### 2019-09-25 sections/acknowledgements.tex
initializes basic MNRAS paper

% !TEX root = ../main.tex
This publication made use of SDSS-I/II data. Funding for the SDSS and SDSS-II was provided by the Alfred P. Sloan Foundation, the Participating Institutions, the National Science Foundation, the U.S. Department of Energy, the National Aeronautics and Space Administration, the Japanese Monbukagakusho, the Max Planck Society, and the Higher Education Funding Council for England. The SDSS Web Site is \url{http://www.sdss.org/}.

This project was partially funded by a Google Faculty Research Award to Karen Masters (\url{https://ai.google/research/outreach/faculty-research-awards/})

### 2019-12-10 sections/intro_reviews.tex
fleshes out intro, begins method

% !TEX root = ../main.tex

\citet{2014PASA...31...35D} introduction:

\begin{itemize}
  \item spirals patterns are present in 2/3 of all massive galaxies (GZ2)
  \item Sites of a majority of a star formation
  \item understanding spirals is essential for understanding star formation and galaxy evolution
  \item hubble classification of spirals are divided by tightness and the presence of a bar
  \item can also include information on bulge size and luminosity, and galaxy gas content
  \item Elmegreen \& Elmegreen proposed another scheme, classifying spirals into 12 types depending on the number and lenght of spiral arms
  \item This was simplified into \textit{flocculent}, \textit{grand design} and \textit{multi-armed} spirals
  \item 60\% of galaxies exhibit some grand design structure
  \item the type of spiral is linked to the mechanism which generated them
  \begin{itemize}
    \item (quasi-stationary) density wave theory (\textit{QSDW})
    \item local instabilities, perturbations or noise which are swing-amplified
    \item tidal interactions
    \item bars may also play a role in inducing spiral arms
  \end{itemize}
  \item \textbf{these mechanisms are not mutually exclusive}
  \item flocculent and multi-arm spirals generally thought to arise from local instabilities
  \item grand designs are thought to have undergone a tidal interaction, have a bar driving arms or be obeying \textit{QSDW}
  \item we can measure spiral number, pitch angle, amplitude, arm shape and lifetime
\end{itemize}

\citet{2019MNRAS.487.1808M} introduction:

\begin{itemize}
  \item Hubble classification is a common technique
  \item Hubble spirals are ordered in a sequence extending away from the ellipticals, and separated depending on the presence of a bar
  \item Extended by de Vaucouleurs to include Sd
  \begin{itemize}
    \item split by spiral arm appearance (how tightly wound and how distinct the arms were)
    \item the prominence of a central bulge
  \end{itemize}
  \item galaxy morphology encodes information on its dynamical history, including its formation and evolution
  \item morphology is known to correlate well with other physical properties
  \item people have taken to using proxies for morphology
  \item this is not a valid approach
  \item this paper explores an updated view of the Hubble sequence with the morphological classifications provided by Willett et al. (2013)
  \item most experts say that classifying based on bulge size vs spiral tightness result in consistent classification
  \item modern automatic galaxy classification conflates bulge size alone with spiral type
  \item differences in arm characteristics is linked to different formation mechanisms (floculent to shearing, grand design to density waves)
\end{itemize}

\citet{2019arXiv190804246D} introduction:

\begin{itemize}
  \item Arms are noticeable and pretty and they host intense star formation, H2 regions and dust.
  \item More prominent in blue, but backbone is comprised of old stars.
  \item 2/3 massive galaxies are Spirals.
  \item 3 types of spiral arm (GD, 18\%; FL, 50\%; MA, 32\%; Elmegreen et al, 2011 and Buta et al., 2015).
  \item GD have two long and well-defined spiral arms, FL have short and fragmented arm sections, MA are a fairly symmetric middle-ground, comprised of central 2-armed section which develops long ramifications in teh outer parts of the optical disc.
  \item Spiral formation is a matter of debate - including \textit{density wave theory}, \textit{tidal triggering} and \textit{swing amplification}.
  \item GD spirals and inner section of MA spirals are seen as density wave driven, whereas FL patterns are seen as swing-amplified regions of local gravitational instability
  \item Numerical models show that FL patterns are transient and reccurent, whereas GD last longer ($\sim1$ Gyr).
  \item pitch angle is a measure of arm tightness. Positive pich angle gives trailing arm, and vice versa.
  \item pitch angle is not necessarily constant, regardless of the class of the spiral arm.
  \item It is claimed that pitch angles depend on central mass concentration and atomic gas density (e.g. Kennicutt 1981; Block et al. 1994; Davis et al. 2015; Yu \& Ho 2019), galactic shear rate (e.g. Seigar et al. 2006; Grand et al. 2013), or on the steepness of the rotation curves (Seigar et al. 2005, 2014).
  \item Spiral arms are more open in galaxies with rising rotation curves (larger absolute pitch angle) and tighter in those with falling rotation curves.
  \item Very tight scaling relation reported by Davis et al. (2017) between supermassive black hole mass and the spiral pitch angle.
  \item Spirals rearrange gas and lead to the formation of disc-like bulges (e.g. Kormendy \& Kennicutt 2004). Making them important agents for the secular evolution of disc galaxies (a process in which bars also play a significant role).
  \item Spirals could be driven by bars, though this is still debated. Manifold theory suggests a strong coupling between bars and spirals (Romero-G\'omez et al. 2006, and many others). Numerical simulations show that material becomes confined to tubes (invariant manifolds) that extend from the two unstable Lagrangian points at the end of the bar ($L_1$ and $L_2$). This theory predicts a dependence of the pitch angle on the bar perturbation strength (Athanassoula et al. 2009a).
\end{itemize}

\citet{2019arXiv190910291P} introduction:

\begin{itemize}
  \item Spirals make up \textbf{one-third} (\textit{this is a typo, two-thirds is correct}) of all massive galaxies
  \item They are the site of a majority of star formation
  \item Dobbs and Baba argue that in unbarred spirals spiral arms are either transient and recurrent due to self-gravity within steller and / or gaseous discs, or are the result of tidal interactions
  \item Supported by lack of relation between pitch angle and central galaxy concentration observed in GZ (Hart 2017)
  \item Shabani et al. (2018) only find evidence of a fixed density wave in a galaxy with a strong bar, however Yu and Ho (2018, 2019) do find correlations between pitch angle and galaxy morphology
\end{itemize}

\subsection*{TL;DR}

Understanding spirals is very useful

Spiral tightness can be used as a measure of useful things

It's difficult to measure spiral tightness at scale

Bars cause predictions of tightness to change measurably

### 2019-12-10 sections/introduction.tex
fleshes out intro, begins method

% !TEX root = ../main.tex

% BEGIN HERE
Spiral structure is present in a majority of massive galaxies \citep{2008MNRAS.389.1179L}, yet the formation mechanisms through which spiral structure originates are still hotly debated. Spirals themselves are as diverse as the theories proposed to govern their evolution; ranging from the quintessential pair of well-defined arcs of the grand design spiral, to the patchy and fragmented arm segments of the flocculent spiral, to the disjointed multi-armed spiral. These variatons on structure account for $18\%$, $50\%$ and $32\%$ of the population respectively (\citealt{2011ApJ...737...32E}, \citealt{2015yCat..22170032B}). Our current understanding of the mechanisms which drive spiral growth and evolution suggest that each of the different forms of spiral galaxy may be triggered primarily by different processes. Grand Design spirals are thought to have undergone a tidal interaction, be driven by a bar (as suggested by Manifold theory, \citealt{2011hsa6.conf..548R}), or be obeying quasi-stationary density wave theory, in which spiral arms are slowly evolving, ever-present structures in the disc \citep{1964ApJ...140..646L}. Flocculent spirals are thought to be formed through swing amplification (shearing of small gravitational instabilities in the disc), and be transient and reccurent in nature \citep{1966ApJ...146..810J}. However it is recognised that no two methods of spiral formation are mutually exclusive.

Arms of spiral galaxies are the source of the vast majority of star formation in the Universe, and spirals rearrange disc gas and can lead to the formation of disc-like bulges (e.g. \citealt{2004ARA&A..42..603K}). Studies of spiral morphology have found interesting correlations between spiral morphology and other galactic properties, such as a correlation between spiral tightness and central mass concentration (\citealt{2019ApJ...871..194Y}, \citealt{2015PhDT........14D}, though \citealt{2017MNRAS.472.2263H} found no such relation) and tightness and rotation curve shape (\citealt{2005MNRAS.359.1065S}, with rising rotation curves creating more open spiral structure). These predictions and observations provide compelling reasons for investigating their underlying rules and dynamics, as doing so is essential for understanding the secular evolution of disc galaxies.

Many methodologies have been proposed and implemented to measure spiral arm properties, including visual inspection (i.e. \citealt{2015A&A...582A..86H}), fourier analysis (i.e. \citealt{2019arXiv190804246D}), more complex automated identification (i.e. SpArcFiRe, \citealt{2014ApJ...790...87D}), and combinations of automated methods and human classifiers \citep{2017MNRAS.472.2263H}. One potentially underused method of obtaining measurements of spirals is through photometric fitting of spiral structure, as possible using tools such as \textsc{GALFIT} \citep{2010AJ....139.2097P} and \textit{Galaxy Builder} \comment{(Lingard et al, in prep)}. These methods attempt to localize light from an image of a galaxy into distinct subcomponents, such as a galaxy disc, bulge, bar and spiral arms, generally finding the optimum solution using computational optimization.

This paper makes use of the photometric models obtained through the \textit{Galaxy Builder} citizen science project \comment{(Lingard et al, in prep)} which measure spiral brightness, thickness and tightness. In this paper we focus on the use of measured spiral tightness (quantified using pitch angle, \citealt{1987gady.book.....B}) as a probe into the dynamical mechanisms governing a spiral galaxy's evolution. We make use of multilevel modelling to test for signs of spiral arm winding using the predictions derived by \citealt{2019arXiv190910291P} (uniformity of galaxy pitch angle in $\cot\;\phi$) and conclude that our sample of galaxies do not show winding of this form. We model galaxy pitch angle as a truncated normal distribution make use of the resulting pitch angles to examine the correlation between pitch angle and bulge size, implied by the Hubble sequence, and bar strength, implied by Manifold theory.

### 2019-12-10 sections/method.tex
fleshes out intro, begins method

% !TEX root = ../main.tex

\subsection{Length-weighted spiral arms}

A common method used to obtain a pitch angle of a galaxy is by identifying all

\subsection{Bayesian modelling of spiral arms in \textit{Galaxy Builder}}

We fit a number of models, including various priors on the global pitch angle distribution. We make use of PYMC3\footnote{\url{https://docs.pymc.io/}}, an open source probabilistic programming framework written in python.

\subsubsection{Spiral arms from \textit{Galaxy Builder}}

Clustered, cleaned points in polar coordinates

$$ r_\mathrm{arm} = \exp\left[\theta\tan(\phi_\mathrm{arm}) + c_\mathrm{arm}\right] + \sigma_r.$$

\subsubsection{Cot-Uniform model}

Assume the pitch angle of a galaxy's spiral arms are drawn from a Normal distribution, truncated between 0 and 90, centred on the galaxy's ``pitch angle'', $\phi_\mathrm{gal}$ with some measure of spread, $\sigma_\mathrm{gal}$, common to all galaxies:

\begin{equation}
\phi_\mathrm{arm} \sim \mathrm{TruncatedNormal}(\phi_\mathrm{gal}, \sigma_\mathrm{gal}, \mathrm{min}=0, \mathrm{max}=90).
\end{equation}

We choose a prior on $\phi_\mathrm{gal}$ such that

\begin{equation}
\cot\left(\phi_\mathrm{gal}\right) \sim \mathrm{Uniform}(\mathrm{min}=1, \mathrm{max}=4),
\end{equation}

and $\sigma_\mathrm{gal}$ of

\begin{equation}
\sigma_\mathrm{gal} \sim \mathrm{InverseGamma}(\alpha=2,\,\beta=20).
\end{equation}

We model point

\begin{equation}
c_\mathrm{arm} \sim \mathrm{Cauchy}(\alpha=0,\,\beta=10).
\end{equation}

\begin{equation}
\sigma_r \sim \mathrm{HalfCauchy}(\beta=0.2)
\end{equation}

\subsubsection{Hierarchial Normal Model}

Instead of the above model, assume

\begin{equation}
\phi_\mathrm{gal} \sim \mathrm{TruncatedNormal}(\mu_\mathrm{global}, \sigma_\mathrm{global}, \mathrm{min}=0, \mathrm{max}=90),
\end{equation}

Hyperpriors used are as follows:

\begin{equation}
\mu_\phi \sim \mathrm{Uniform}(0, 90)
\end{equation}

\begin{equation}
\sigma_\mathrm{global} \sim \mathrm{InverseGamma}(\alpha=1,\,\beta=10)
\end{equation}

### 2019-12-19 sections/introduction.tex
fleshing out method and results now anaylsis is complete

@@
 % BEGIN HERE
-Spiral structure is present in a majority of massive galaxies \citep{2008MNRAS.389.1179L}, yet the formation mechanisms through which spiral structure originates are still hotly debated.
+Spiral structure is present in a majority of massive galaxies (\citealt{1989gadv.book..151B}, \citealt{2008MNRAS.389.1179L}), yet the formation mechanisms through which spiral structure originates are still hotly debated.
 Spirals themselves are as diverse as the theories proposed to govern their evolution; ranging from the quintessential pair of well-defined arcs of the grand design spiral, to the patchy and fragmented arm segments of the flocculent spiral, to the disjointed multi-armed spiral.
@@
 
-Many methodologies have been proposed and implemented to measure spiral arm properties, including visual inspection (i.e.
-\citealt{2015A&A...582A..86H}), fourier analysis (i.e.
-\citealt{2019arXiv190804246D}), more complex automated identification (i.e.
+\subsection{Measuring galaxy pitch angle}
+
+Many methodologies have been proposed and implemented to measure spiral arm properties, including visual inspection (\citealt{2015A&A...582A..86H}), fourier analysis (i.e.
+\textsc{2DFFT}, \citealt{2012ApJS..199...33D}), texture analysis (i.e.
 SpArcFiRe, \citealt{2014ApJ...790...87D}), and combinations of automated methods and human classifiers \citep{2017MNRAS.472.2263H}.
-One potentially underused method of obtaining measurements of spirals is through photometric fitting of spiral structure, as possible using tools such as \textsc{GALFIT} \citep{2010AJ....139.2097P} and \textit{Galaxy Builder} \comment{(Lingard et al, in prep)}.
+One potentially underused method of obtaining measurements of spirals is through photometric fitting of spiral structure, as possible using tools such as \textsc{GALFIT} \citep{2010AJ....139.2097P} and \textit{Galaxy Builder} (\comment{Lingard et al, in prep}).
 These methods attempt to localize light from an image of a galaxy into distinct subcomponents, such as a galaxy disc, bulge, bar and spiral arms, generally finding the optimum solution using computational optimization.
+This optimization process, however, is often not robust for complex, many-component models and requires significant supervision to converge to a physically meaningful result \citep{Gao2017:1709.00746v1}.
+\comment{Lingard et al, in prep} successfully solve this problem through the use of citizen science to provide the starting point for a computational fit.
 
-This paper makes use of the photometric models obtained through the \textit{Galaxy Builder} citizen science project \comment{(Lingard et al, in prep)} which measure spiral brightness, thickness and tightness.
+A common assumption when measuring galaxy pitch angle is that observed spiral arms have a constant pitch angle, these spirals are known as logarithmic spirals and are described by
+
+\begin{equation}
+r = A\,e^{\theta\tan\phi}.
+\end{equation}
+
+One method used to obtain a pitch angle of a galaxy from identified arm segments is by taking the weighted mean of their pitch angles (as used by \textsc{SpArcFiRe}, \citealt{2014ApJ...790...87D}), where weighting is determined by the length of the arc segment, with longer arcs having higher weights, i.e.
+for a galaxy where we have identified $N$ arm segments, each with length $L_i$ and pitch angle $\phi_i$
+
+\begin{equation}
+  \phi_\mathrm{gal} = \left(\sum_{i=1}^{N}L_i\right)^{-1}\sum_{i=1}^{N}L_i \phi_i.
+\end{equation}
+
+The most commonly used measurement of uncertainty of length-weighted pitch angles is the \comment{unweighted?} sample variance between the arm segments identified.
+
+One notable drawback of the use of length-weighted arms is the sensitivity of the end result to the number and quality of the spiral arm segments identified \comment{...}
+
+This paper makes use of the photometric models obtained through the \textit{Galaxy Builder} citizen science project for the 196 spiral galaxies present in \comment{Lingard et al, in prep}.
 In this paper we focus on the use of measured spiral tightness (quantified using pitch angle, \citealt{1987gady.book.....B}) as a probe into the dynamical mechanisms governing a spiral galaxy's evolution.
-We make use of multilevel modelling to test for signs of spiral arm winding using the predictions derived by \citealt{2019arXiv190910291P} (uniformity of galaxy pitch angle in $\cot\;\phi$) and conclude that our sample of galaxies do not show winding of this form.
-We model galaxy pitch angle as a truncated normal distribution make use of the resulting pitch angles to examine the correlation between pitch angle and bulge size, implied by the Hubble sequence, and bar strength, implied by Manifold theory.
+We make use of Bayesian hierarchical modelling to measure galaxy pitch angle from the spiral arm clusters producted by \textit{Galaxy Builder}.
+We test for signs of spiral arm winding using the predictions derived by \citealt{2019arXiv190910291P} (uniformity of galaxy pitch angle in $\cot\;\phi$) and conclude using a marginalized Anderson-Darling test that we \comment{cannot unilaterally reject winding of this form at the 1\% level}.
+Section \ref{section:morphology_comparision} examines the correlation between pitch angle and bulge size, implied by the Hubble sequence, and bar strength, implied by Manifold theory, and find \comment{no significant correlation}.

### 2019-12-19 sections/method.tex
fleshing out method and results now anaylsis is complete

@@
 % !TEX root = ../main.tex
+\subsection{The Galaxy Sample}
+The galaxies selected analysed in this paper are the 198 galaxies from \comment{Lingard et al, in prep.}.
+We combine classifications of galaxies which were repeated in the \textit{validation subset} with the original classifications, and perform clustering and point cleaning as detailed in \comment{Lingard et al, in prep.}, and remove any galaxies for which no arms were identified.
+This results in a hierarchiacal data structure of 109 galaxies and 250 spiral arms, with 307,861 points in polar coordinates, which are scaled such that the radius of each spiral arm has unit maximum.
 
-\subsection{Length-weighted spiral arms}
-
-A common method used to obtain a pitch angle of a galaxy is by identifying all
+\comment{Should I elaborate more on the sample?}
 
@@
 
-We fit a number of models, including various priors on the global pitch angle distribution.
-We make use of PYMC3\footnote{\url{https://docs.pymc.io/}}, an open source probabilistic programming framework written in python.
+Assume we can model spiral arms as a logarithmic spiral, described by
 
-\subsubsection{Spiral arms from \textit{Galaxy Builder}}
+\begin{equation}
+r_\mathrm{arm} = \exp\left[\theta_\mathrm{arm}\tan\phi_\mathrm{arm} + c_\mathrm{arm}\right].
+\end{equation}
 
-Clustered, cleaned points in polar coordinates
-
-$$ r_\mathrm{arm} = \exp\left[\theta\tan(\phi_\mathrm{arm}) + c_\mathrm{arm}\right] + \sigma_r.$$
-
-\subsubsection{Cot-Uniform model}
-
-Assume the pitch angle of a galaxy's spiral arms are drawn from a Normal distribution, truncated between 0 and 90, centred on the galaxy's ``pitch angle'', $\phi_\mathrm{gal}$ with some measure of spread, $\sigma_\mathrm{gal}$, common to all galaxies:
+Assume the pitch angle of a galaxy's spiral arms are drawn from a Normal distribution, truncated between 0 and 90, centred on some value we will call the galaxy's pitch angle, $\phi_\mathrm{gal}$.
+This dispersion in arm pitch angle has some measure of spread, $\sigma_\mathrm{gal}$, which we will assume is the same in galaxies:
 
@@
 
-We choose a prior on $\phi_\mathrm{gal}$ such that
+We choose priors on $\phi_\mathrm{gal}$ and $\sigma_\mathrm{gal}$ of
 
-\begin{equation}
-\cot\left(\phi_\mathrm{gal}\right) \sim \mathrm{Uniform}(\mathrm{min}=1, \mathrm{max}=4),
-\end{equation}
+\begin{align}
+  \phi_\mathrm{gal} &\sim \mathrm{Uniform}(\mathrm{min}=0, \mathrm{max}=90),\\
+  \sigma_\mathrm{gal} &\sim \mathrm{InverseGamma}(\alpha=2,\,\beta=20).
+\end{align}
 
-and $\sigma_\mathrm{gal}$ of
+We also have the offset parameter $c$, and a measure of radial uncertainty $\sigma_r$:
 
-\begin{equation}
-\sigma_\mathrm{gal} \sim \mathrm{InverseGamma}(\alpha=2,\,\beta=20).
-\end{equation}
+\begin{align}
+  c_\mathrm{arm} &\sim \mathrm{Cauchy}(\alpha=0,\,\beta=10),\\
+  \sigma_r &\sim \mathrm{HalfCauchy}(\beta=0.2).
+\end{align}
 
-We model point
+\begin{figure}
+  \includegraphics[width=8cm]{plots/plots_n109d1000t500/model.pdf}
+  \caption{The model used for galaxy pitch angle measurement for the sample.}
+  \label{fig:ad-cot-test}
+\end{figure}
 
-\begin{equation}
-c_\mathrm{arm} \sim \mathrm{Cauchy}(\alpha=0,\,\beta=10).
-\end{equation}
-
-\begin{equation}
-\sigma_r \sim \mathrm{HalfCauchy}(\beta=0.2)
-\end{equation}
-
-\subsubsection{Hierarchial Normal Model}
-
-Instead of the above model, assume
-
-\begin{equation}
-\phi_\mathrm{gal} \sim \mathrm{TruncatedNormal}(\mu_\mathrm{global}, \sigma_\mathrm{global}, \mathrm{min}=0, \mathrm{max}=90),
-\end{equation}
-
-Hyperpriors used are as follows:
-
-\begin{equation}
-\mu_\phi \sim \mathrm{Uniform}(0, 90)
-\end{equation}
-
-\begin{equation}
-\sigma_\mathrm{global} \sim \mathrm{InverseGamma}(\alpha=1,\,\beta=10)
-\end{equation}
+To perform inference, we make use of the No-U-Turn-Sampler (NUTS), implemented in PYMC3\footnote{\url{https://docs.pymc.io/}}, an open source probabilistic programming framework written in python.
+\comment{does PYMC3 have a citation?}

### 2019-12-19 sections/results.tex
fleshing out method and results now anaylsis is complete

% !TEX root = ../main.tex

\subsection{Spiral Winding}

In order to test the possible progenitor distribution of our estimated galaxy pitch angles, we repeatedly perform an Anderson-Darling test over each draw present in the MCMC trace, resulting in a distribution of Anderson-Darling statistics. We will refer to this test as the \textit{marginalized Anderson-Darling test}.

We perform the marginalized Anderson-Darling test on a distribution uniform in $\cot\phi$ between the limits present in \citet{2019arXiv190910291P} ($1.00 < \cot\phi < 4.75$, or roughly $11.9 < \phi < 45.0$). The resulting distribution of Anderson-Darling statistics can be seen in Figure \ref{fig:ad-cot-test}. We observe that we reject the null hypothesis at the 1\% level for only 92\% of the possible realizations of galaxy pitch angle. Therefore we cannot unilaterally reject winding of the kind described by \citet{2019arXiv190910291P}, though .

\begin{figure*}
  \includegraphics[width=17.7cm]{plots/cot_uniform_marginalized_tests.pdf}
  \caption{The results of a marginalized Anderson-Darling test (top panel), with confidence intervals shown, and marginalized Kolmogorov-Smirnov test p-values (lower panel).}
  \label{fig:ad-cot-test}
\end{figure*}

\subsection{Dependence of pitch angle on Galaxy Morphology}
\label{section:morphology_comparision}
\subsubsection{Pitch angle vs. Bulge size}
We see no correlation between galaxy pitch angle derived from the \textit{hierarchial normal model} and GZ2's \textit{pbulge}, which is widely viewed as a good measure of bulge size.

\subsubsection{Pitch angle vs. Bar Strength}
We see no correlation between galaxy pitch angle derived from the \textit{hierarchial normal model} and GZ2's \textit{pbar}, which is widely viewed as a good measure of bar strength, and therefore a measure of the torque applied on the disc gas.

A marginalized Anderson-Darling test does not find that the samples were drawn from different distributions (statistic of -0.91).

### 2020-01-02 sections/discussion.tex
continues work on fleshing out paper. Adds discussion. Almost at 1st draft.

% !TEX root = ../main.tex
This paper presented a new Bayesian approach to estimate galaxy pitch angle, making use of citizen science results to measure spiral arms through photometric modelling. We introduce an adaptation of the Anderson-Darling test to incorporate full Bayesian posterior probabilities and utilize this test to investigate theories governing spiral formation and evolution.

The statistical approach implemented in this paper allows a more thorough approach than calculating length-weighted pitch angles, and better accounts for inter-arm pitch angle variation than fourier analysis. \comment{This is a strong statement, how can we quantify it?}.

Our method and results do not completely reject spiral winding of the form described by \citet{2019arXiv190910291P}, however this result is highly sensitive to the boundaries used; using limits of $10.0^\circ < \phi < 40.0^\circ$ results in us rejecting the null hypothesis at the 1\% level for every MCMC draw. As no physical justification was provided for the limits present in \citet{2019arXiv190910291P}, this is a far from ideal test of arm winding.

We we do not find a relationship between bar strength and pitch angle, as would have been predicted by Manifold theory, and do not find evidence for the relationship between central mass concentration and pitch angle predicted by the Hubble sequence.

This work is primarily held back by the sample used; due to time constraints the \textit{Galaxy Builder} galaxies used are not guaranteed to be representative. A larger and more complete sample is needed to be fully confident in our results, but we are confident that the methodlogy proposed here is a scalable solution to the problems facing investigation of spiral morphology.

### 2020-01-02 sections/introduction.tex
continues work on fleshing out paper. Adds discussion. Almost at 1st draft.

@@
 % BEGIN HERE
-Spiral structure is present in a majority of massive galaxies (\citealt{1989gadv.book..151B}, \citealt{2008MNRAS.389.1179L}), yet the formation mechanisms through which spiral structure originates are still hotly debated.
-Spirals themselves are as diverse as the theories proposed to govern their evolution; ranging from the quintessential pair of well-defined arcs of the grand design spiral, to the patchy and fragmented arm segments of the flocculent spiral, to the disjointed multi-armed spiral.
+Spiral structure is present in a majority of massive galaxies (\citealt{1989gadv.book..151B}, \citealt{2008MNRAS.389.1179L}) yet the formation mechanisms through which spiral structure originates are still hotly debated.
+Spirals are as diverse as the theories proposed to govern their evolution; from the quintessential pair of well-defined arcs of the grand design spiral, to the fragmented arm segments of the flocculent spiral, to the disjointed multi-armed spiral.
 These variatons on structure account for $18\%$, $50\%$ and $32\%$ of the population respectively (\citealt{2011ApJ...737...32E}, \citealt{2015yCat..22170032B}).
 Our current understanding of the mechanisms which drive spiral growth and evolution suggest that each of the different forms of spiral galaxy may be triggered primarily by different processes.
-Grand Design spirals are thought to have undergone a tidal interaction, be driven by a bar (as suggested by Manifold theory, \citealt{2011hsa6.conf..548R}), or be obeying quasi-stationary density wave theory, in which spiral arms are slowly evolving, ever-present structures in the disc \citep{1964ApJ...140..646L}.
+Grand Design spirals are thought to have undergone a tidal interaction, be driven by a bar (as suggested by Manifold theory, \citealt{2011hsa6.conf..548R}), or be obeying quasi-stationary density wave theory (QSDW theory), in which spiral arms are slowly evolving, ever-present structures in the disc \citep{1964ApJ...140..646L}.
 Flocculent spirals are thought to be formed through swing amplification (shearing of small gravitational instabilities in the disc), and be transient and reccurent in nature \citep{1966ApJ...146..810J}.
-However it is recognised that no two methods of spiral formation are mutually exclusive.
+It is recognised that methods of spiral formation are not mutually exclusive.
+There is also debate as to the evolution of spiral structure: whether spirals maintain their shape over many rotational periods (assumed by QSDW theory, \citealt{1964ApJ...140..646L}), or ``wind up'' with the differential rotation of the disc (as found by \citealt{2019MNRAS.487.1808M}).
 
@@
 This optimization process, however, is often not robust for complex, many-component models and requires significant supervision to converge to a physically meaningful result \citep{Gao2017:1709.00746v1}.
-\comment{Lingard et al, in prep} successfully solve this problem through the use of citizen science to provide the starting point for a computational fit.
+\comment{Lingard et al, in prep} propose a solution to this problem through the use of citizen science to provide the starting points for computational fitting.
 
-A common assumption when measuring galaxy pitch angle is that observed spiral arms have a constant pitch angle, these spirals are known as logarithmic spirals and are described by
+A common assumption when measuring galaxy pitch angle is that observed spiral arms have a constant pitch angle. Spirals of this kind are known as logarithmic spirals and are described by
 
 \begin{equation}
-r = A\,e^{\theta\tan\phi}.
+r = A\,e^{\theta\tan\phi},
 \end{equation}
 
-One method used to obtain a pitch angle of a galaxy from identified arm segments is by taking the weighted mean of their pitch angles (as used by \textsc{SpArcFiRe}, \citealt{2014ApJ...790...87D}), where weighting is determined by the length of the arc segment, with longer arcs having higher weights, i.e.
+where $\phi$ is the arm's pitch angle.
+One method used to obtain a pitch angle of a galaxy is to fit logarithmic spirals to individually identified arm segments and take the weighted mean of their pitch angles (which often vary by upwards of $10^\circ$, \citealt{2014ApJ...790...87D}).
+Weighting is determined by the length of the arc segment, with longer being assigned higher weights, i.e.
 for a galaxy where we have identified $N$ arm segments, each with length $L_i$ and pitch angle $\phi_i$
@@
 
-One notable drawback of the use of length-weighted arms is the sensitivity of the end result to the number and quality of the spiral arm segments identified \comment{...}
+A notable drawback of length-weighted pitch angle is sensitivity to the number and quality of the spiral arm segments identified; \citet{2017MNRAS.472.2263H} found that only 15\% of the arm segments identified by a leading algorithm (\textsc{SpArcFiRe}) were identified as ``good'' matches to real spiral arms by citizen science classifiers.
+
+Fourier analysis in one-dimension (as performed by \comment{...}) and two-dimensions (\comment{...}) is another widely used method of computationally obtaining galaxy pitch angles.
+Two-dimensional fourier methods generally decompose a deprojected image of a galaxy into a superpositions of logarithmic spirals between inner and outer annuli \citep{2012ApJS..199...33D}, and reports the pitch angle with the highest amplitude as the galaxy's pitch angle.
+It is uncertain how the large observed variation between pitch angles of individual arms impacts this measurement.
 
@@
 We make use of Bayesian hierarchical modelling to measure galaxy pitch angle from the spiral arm clusters producted by \textit{Galaxy Builder}.
-We test for signs of spiral arm winding using the predictions derived by \citealt{2019arXiv190910291P} (uniformity of galaxy pitch angle in $\cot\;\phi$) and conclude using a marginalized Anderson-Darling test that we \comment{cannot unilaterally reject winding of this form at the 1\% level}.
-Section \ref{section:morphology_comparision} examines the correlation between pitch angle and bulge size, implied by the Hubble sequence, and bar strength, implied by Manifold theory, and find \comment{no significant correlation}.
+
+Section \ref{section:spiral_winding} investigates spiral arm winding using the test derived by \citealt{2019arXiv190910291P} (uniformity of galaxy pitch angle in $\cot\;\phi$) and concludes using a marginalized Anderson-Darling test that we \comment{cannot unilaterally reject winding of this form at the 1\% level}.
+Section \ref{section:morphology_comparision} examines the correlation between pitch angle and bulge size implied by the Hubble sequence and reported by \comment{citation}, and pitch angle and bar strength implied by Manifold theory, and find \comment{no significant correlation}.

### 2020-01-02 sections/method.tex
continues work on fleshing out paper. Adds discussion. Almost at 1st draft.

@@
 The galaxies selected analysed in this paper are the 198 galaxies from \comment{Lingard et al, in prep.}.
-We combine classifications of galaxies which were repeated in the \textit{validation subset} with the original classifications, and perform clustering and point cleaning as detailed in \comment{Lingard et al, in prep.}, and remove any galaxies for which no arms were identified.
-This results in a hierarchiacal data structure of 109 galaxies and 250 spiral arms, with 307,861 points in polar coordinates, which are scaled such that the radius of each spiral arm has unit maximum.
+We combine classifications of galaxies which were repeated in the \textit{validation subset} with the original classifications, and perform clustering and point cleaning as detailed in \comment{Lingard et al, in prep.}.
+We remove any galaxies for which no arms were identified.
+This results in a hierarchiacal data structure of 109 galaxies, 250 spiral arms and 307,861 points in polar coordinates.
+We apply a scaling to the radius to unit variance (which does not impact log spiral pitch angle due to its scale invariance).
 
@@
 
-Assume we can model spiral arms as a logarithmic spiral, described by
+We model a spiral arm as a logarithmic spiral, described by
 
@@
 
-Assume the pitch angle of a galaxy's spiral arms are drawn from a Normal distribution, truncated between 0 and 90, centred on some value we will call the galaxy's pitch angle, $\phi_\mathrm{gal}$.
+Assume the pitch angle of a galaxy's spiral arms are drawn from a Normal distribution, truncated between 0 and 90 degrees, centred on some value we will call the galaxy's pitch angle, $\phi_\mathrm{gal}$.
 This dispersion in arm pitch angle has some measure of spread, $\sigma_\mathrm{gal}$, which we will assume is the same in galaxies:
@@
 
-We choose priors on $\phi_\mathrm{gal}$ and $\sigma_\mathrm{gal}$ of
+We choose hyperpriors on $\phi_\mathrm{gal}$ and $\sigma_\mathrm{gal}$ of
 
@@
   \caption{The model used for galaxy pitch angle measurement for the sample.}
-  \label{fig:ad-cot-test}
+  \label{fig:pymc3-model}
 \end{figure}
 
-To perform inference, we make use of the No-U-Turn-Sampler (NUTS), implemented in PYMC3\footnote{\url{https://docs.pymc.io/}}, an open source probabilistic programming framework written in python.
-\comment{does PYMC3 have a citation?}
+To perform inference, we make use of the No-U-Turn-Sampler (NUTS, \citealt{2011arXiv1111.4246H}), implemented in PYMC3\footnote{\url{https://docs.pymc.io/}}, an open source probabilistic programming framework written in python \citep{pymc3_paper}.

### 2020-01-02 sections/results.tex
continues work on fleshing out paper. Adds discussion. Almost at 1st draft.

@@
 \subsection{Spiral Winding}
-
+\label{section:spiral_winding}
 In order to test the possible progenitor distribution of our estimated galaxy pitch angles, we repeatedly perform an Anderson-Darling test over each draw present in the MCMC trace, resulting in a distribution of Anderson-Darling statistics.
 We will refer to this test as the \textit{marginalized Anderson-Darling test}.
+We make use of the Kolmogorov-Smirnov test in a similar manner for comparison.
 
-We perform the marginalized Anderson-Darling test on a distribution uniform in $\cot\phi$ between the limits present in \citet{2019arXiv190910291P} ($1.00 < \cot\phi < 4.75$, or roughly $11.9 < \phi < 45.0$).
+We perform the marginalized Anderson-Darling test for a potential source distribution uniform in $\cot\phi$ between the limits present in \citet{2019arXiv190910291P} ($1.00 < \cot\phi < 4.75$, or roughly $11.9^\circ < \phi < 45.0^\circ$).
 The resulting distribution of Anderson-Darling statistics can be seen in Figure \ref{fig:ad-cot-test}.
-We observe that we reject the null hypothesis at the 1\% level for only 92\% of the possible realizations of galaxy pitch angle.
-Therefore we cannot unilaterally reject winding of the kind described by \citet{2019arXiv190910291P}, though .
+We observe that we reject the null hypothesis at the 1\% level for only 86\% of the possible realizations of galaxy pitch angle, therefore with our sample and methodology we cannot unilaterally reject winding of the kind described by \citet{2019arXiv190910291P}.
 
@@
   \includegraphics[width=17.7cm]{plots/cot_uniform_marginalized_tests.pdf}
-  \caption{The results of a marginalized Anderson-Darling test (top panel), with confidence intervals shown, and marginalized Kolmogorov-Smirnov test p-values (lower panel).}
+  \caption{The results of a marginalized Anderson-Darling test (top panel), with confidence intervals shown, and marginalized Kolmogorov-Smirnov test p-values (lower panel).
+Moving rightwards on the x-axis implies greater confidence in rejecting the null hypothesis.}
   \label{fig:ad-cot-test}
@@
 \subsubsection{Pitch angle vs. Bulge size}
-We see no correlation between galaxy pitch angle derived from the \textit{hierarchial normal model} and GZ2's \textit{pbulge}, which is widely viewed as a good measure of bulge size.
+We see no correlation between galaxy pitch angle derived from the \textit{hierarchial normal model} and Galaxy Zoo 2's debiased \citep{Willett2013:1308.3496v2} \textit{pbulge}, which has been shown to be a good measure of bulge size.
+
+We separate our sample into ``disc-dominated galaxies'' and ``obvious bulge galaxies'' using the debiased fractions from Galaxy Zoo 2 following \citet{2017MNRAS.469.3363K}, defining \textit{no bulge + just noticeable $>$ obvious + dominant} for the former and the converse for the latter.
+A marginalized two-sample Anderson-Darling test \citep{doi:10.1080/01621459.1987.10478517} does not find strong evidence that the samples were drawn from different distributions; we reject the null hypothesis at the 1\% level for only 8\% of the samples.
 
 \subsubsection{Pitch angle vs. Bar Strength}
-We see no correlation between galaxy pitch angle derived from the \textit{hierarchial normal model} and GZ2's \textit{pbar}, which is widely viewed as a good measure of bar strength, and therefore a measure of the torque applied on the disc gas.
+We see no correlation between galaxy pitch angle derived from the \textit{hierarchial normal model} and Galaxy Zoo 2's debiased \textit{pbar}, which is widely viewed as a good measure of bar strength, and therefore a measure of the torque applied on the disc gas.
 
-A marginalized Anderson-Darling test does not find that the samples were drawn from different distributions (statistic of -0.91).
+Separating the sample based off of $\mathrm{\textit{pbar}} > 0.5$, and restricting to galaxies with more than 10 classifications for \textit{pbar} (as performed by \citealt{2011MNRAS.411.2026M} and \citealt{2017MNRAS.469.3363K}) and performing a marginalized two-sample Anderson-Darling test does not find that the samples were drawn from different distributions.
+
+\begin{figure*}
+  \includegraphics[width=17.7cm]{plots/bulge_bar_test_results.pdf}
+  \caption{The results of marginalized two-sample Anderson-Darling tests examining whether pitch angles for Bulge-dominated and Disc-dominated galaxies are drawn from the same distribution (top panel), and the results of the same test for strongly-barred vs unbarred galaxies (bottom panel).}
+  \label{fig:ad-morphology-test}
+\end{figure*}

### 2020-01-14 sections/discussion.tex
update results

@@
 % !TEX root = ../main.tex
-This paper presented a new Bayesian approach to estimate galaxy pitch angle, making use of citizen science results to measure spiral arms through photometric modelling.
+This paper presents a new Bayesian approach to estimate galaxy pitch angle, making use of citizen science results to measure spiral arms through photometric modelling.
 We introduce an adaptation of the Anderson-Darling test to incorporate full Bayesian posterior probabilities and utilize this test to investigate theories governing spiral formation and evolution.
 
-The statistical approach implemented in this paper allows a more thorough approach than calculating length-weighted pitch angles, and better accounts for inter-arm pitch angle variation than fourier analysis.
-\comment{This is a strong statement, how can we quantify it?}.
+The statistical approach implemented in this paper allows a more thorough examination of pitch angle than simpler methods of logarithmic spiral fitting, and better accounts for inter-arm pitch angle variation than fourier analysis \comment{This is a strong statement, how can we quantify it?}.
 
@@
 
-This work is primarily held back by the sample used; due to time constraints the \textit{Galaxy Builder} galaxies used are not guaranteed to be representative.
-A larger and more complete sample is needed to be fully confident in our results, but we are confident that the methodlogy proposed here is a scalable solution to the problems facing investigation of spiral morphology.
+This work is primarily limited by the sample used; due to time constraints the \textit{Galaxy Builder} galaxies used are not guaranteed to be representative.
+A larger and more complete sample is needed to be completely confident in our results.
+However, we believe that the methodlogy proposed here is a scalable solution to the problems facing investigation of spiral morphology.

### 2020-01-14 sections/method.tex
update results

@@
 The galaxies selected analysed in this paper are the 198 galaxies from \comment{Lingard et al, in prep.}.
-We combine classifications of galaxies which were repeated in the \textit{validation subset} with the original classifications, and perform clustering and point cleaning as detailed in \comment{Lingard et al, in prep.}.
+We combine classifications of galaxies which were repeated in the \textit{validation subset} with the original classifications.
+Clustering of drawn spiral arms and cleaning of points was then performed as detailed in \comment{Lingard et al, in prep.}.
 We remove any galaxies for which no arms were identified.
-This results in a hierarchiacal data structure of 109 galaxies, 250 spiral arms and 307,861 points in polar coordinates.
-We apply a scaling to the radius to unit variance (which does not impact log spiral pitch angle due to its scale invariance).
+This results in a hierarchiacal data structure of 109 galaxies, 250 spiral arms and 307,861 points.
+
+Spiral arm points are deprojected to a face-on orientation using the position angle \textsc{PETRO\_PHI90} and axis ratio \textsc{PETRO\_BA90} keywords from the NASA-Sloan Atlas \comment{citation}.
+We scale the radii to have unit variance (which does not impact log spiral pitch angle due to its scale invariance).
 

### 2020-01-14 sections/results.tex
update results

@@
 % !TEX root = ../main.tex
+\subsection{Constraints on Galaxy Pitch angle}
+Our hierarchical model can predict individual arm pitch angle with a hich degree of certainty, however it returns a large spread of potential values for the pitch angles of arms
 
@@
   \includegraphics[width=17.7cm]{plots/cot_uniform_marginalized_tests.pdf}
-  \caption{The results of a marginalized Anderson-Darling test (top panel), with confidence intervals shown, and marginalized Kolmogorov-Smirnov test p-values (lower panel).
+  \caption{The results of a marginalized Anderson-Darling test (top panel), with values corresponding to various confidence intervals shown, and marginalized Kolmogorov-Smirnov test p-values (lower panel).
 Moving rightwards on the x-axis implies greater confidence in rejecting the null hypothesis.}
@@
 \subsubsection{Pitch angle vs. Bulge size}
-We see no correlation between galaxy pitch angle derived from the \textit{hierarchial normal model} and Galaxy Zoo 2's debiased \citep{Willett2013:1308.3496v2} \textit{pbulge}, which has been shown to be a good measure of bulge size.
+We see no correlation between galaxy pitch angle derived from the \textit{hierarchical normal model} and Galaxy Zoo 2's debiased \citep{Willett2013:1308.3496v2} \textit{pbulge}, which has been shown to be a good measure of bulge size.
 
@@
 \subsubsection{Pitch angle vs. Bar Strength}
-We see no correlation between galaxy pitch angle derived from the \textit{hierarchial normal model} and Galaxy Zoo 2's debiased \textit{pbar}, which is widely viewed as a good measure of bar strength, and therefore a measure of the torque applied on the disc gas.
+We see no correlation between galaxy pitch angle derived from the \textit{hierarchical normal model} and Galaxy Zoo 2's debiased \textit{pbar}, which is widely viewed as a good measure of bar strength, and therefore a measure of the torque applied on the disc gas.
 
 Separating the sample based off of $\mathrm{\textit{pbar}} > 0.5$, and restricting to galaxies with more than 10 classifications for \textit{pbar} (as performed by \citealt{2011MNRAS.411.2026M} and \citealt{2017MNRAS.469.3363K}) and performing a marginalized two-sample Anderson-Darling test does not find that the samples were drawn from different distributions.
+\comment{Talk about number of galaxies in each sub-sample?}
 

### 2020-03-23 sections/discussion.tex
Includes new results (deprojection from GZB fitting, updated hyperparameters on clustering). Greatly expands on intro section

@@
 
-The statistical approach implemented in this paper allows a more thorough examination of pitch angle than simpler methods of logarithmic spiral fitting, and better accounts for inter-arm pitch angle variation than fourier analysis \comment{This is a strong statement, how can we quantify it?}.
+The statistical approach implemented in this paper allows a more thorough examination of pitch angle than simpler methods of logarithmic spiral fitting, and better accounts for inter-arm pitch angle variation than fourier analysis, which assumes all arms in a given mode have the same pitch angle.
 
-Our method and results do not completely reject spiral winding of the form described by \citet{2019arXiv190910291P}, however this result is highly sensitive to the boundaries used; using limits of $10.0^\circ < \phi < 40.0^\circ$ results in us rejecting the null hypothesis at the 1\% level for every MCMC draw.
+Our method and results do not completely reject spiral winding of the form described by \citet{2019arXiv190910291P}, however this result is highly sensitive to the boundaries used; using limits of $10.0^\circ < \phi < 40.0^\circ$ results in us rejecting the null hypothesis at the 1\% level for 99\% of the MCMC draws.
 As no physical justification was provided for the limits present in \citet{2019arXiv190910291P}, this is a far from ideal test of arm winding.
@@
 
-This work is primarily limited by the sample used; due to time constraints the \textit{Galaxy Builder} galaxies used are not guaranteed to be representative.
-A larger and more complete sample is needed to be completely confident in our results.
-However, we believe that the methodlogy proposed here is a scalable solution to the problems facing investigation of spiral morphology.
+As with most analyses, the most impactful improvement it would be possible to make here would be to increase the cleanliness and volume of data analysed; due to time constraints the \textit{Galaxy Builder} galaxies used are not guaranteed to be representative.
+However, we believe that the methodlogy proposed here is a scalable, robust solution to the problems facing investigation of spiral morphology.

### 2020-03-23 sections/introduction.tex
Includes new results (deprojection from GZB fitting, updated hyperparameters on clustering). Greatly expands on intro section

@@
 
-% BEGIN HERE
 Spiral structure is present in a majority of massive galaxies (\citealt{1989gadv.book..151B}, \citealt{2008MNRAS.389.1179L}) yet the formation mechanisms through which spiral structure originates are still hotly debated.
@@
 Our current understanding of the mechanisms which drive spiral growth and evolution suggest that each of the different forms of spiral galaxy may be triggered primarily by different processes.
-Grand Design spirals are thought to have undergone a tidal interaction, be driven by a bar (as suggested by Manifold theory, \citealt{2011hsa6.conf..548R}), or be obeying quasi-stationary density wave theory (QSDW theory), in which spiral arms are slowly evolving, ever-present structures in the disc \citep{1964ApJ...140..646L}.
+Grand Design spirals are thought to have undergone a tidal interaction, be driven by a bar (as seen in gas simulations, \citealt{1976ApJ...209...53S}, \citealt{2008A&A...489..115R}, and suggested for stars by Manifold theory, \citealt{2006A&A...453...39R}, \citealt{2009MNRAS.394...67A}, \citealt{2009MNRAS.400.1706A}), or be obeying (quasi-stationary) density wave theory (QSDW theory), in which spiral arms are slowly evolving, ever-present structures in the disc \citep{1964ApJ...140..646L}.
 Flocculent spirals are thought to be formed through swing amplification (shearing of small gravitational instabilities in the disc), and be transient and reccurent in nature \citep{1966ApJ...146..810J}.
 It is recognised that methods of spiral formation are not mutually exclusive.
-There is also debate as to the evolution of spiral structure: whether spirals maintain their shape over many rotational periods (assumed by QSDW theory, \citealt{1964ApJ...140..646L}), or ``wind up'' with the differential rotation of the disc (as found by \citealt{2019MNRAS.487.1808M}).
+There is also debate as to the evolution of spiral structure: whether spirals maintain their shape over many rotational periods (assumed by QSDW theory), or ``wind up'' with the differential rotation of the disc (as found by \citealt{2003MNRAS.344..358B}, \citealt{2019MNRAS.487.1808M}).
 
@@
 
+\subsection{Secular evolution of spiral arms}
+
+\comment{I'm worried that a lot of this section repeats what was said earlier}
+
+One of the foundational assumptions of early work on spiral formation mechanisms (primarily QSDW) was that the disc of a galaxy, if unstable to spiral perturbations, would create a stable, static wave which would exist unchanging for many rotational periods \citep{1964ApJ...140..646L}.
+The motivation for static waves with small numbers of arms (with a preference for $m=2$) was primarily observational; most galaxies show spiral structure, indicating that spirals exist for a long time or are continually rebuilt.
+
+Many simulations demonstrate that spirals do not maintain a constant pitch angle, and instead wind-up over time due to the differential rotation of the disc \citep{2013ApJ...763...46B}.
+Recent research suggests that spirals are dynamical in nature, and continually dissapate and re-form \citep{2014PASA...31...35D}.
+These spirals can be maintained through the same mechanisms that drive QSDW spirals (i.e.
+WASER, \citealt{1976ApJ...205..363M}, swing amplification, \citealt{1965MNRAS.130..125G}), but do not require the idealistic disc conditions required for the formation and maintenance of a stationary wave.
+The pitch angles of these transient spiral arms will decrease due to the differential rotation of the disk, with the density of the arm peaking at some critical pitch angle and then dissapating to be reformed.
+
+In this dynamic picture of spiral arms, pitch angle monotonically decreases from a spiral arm's formation to its dissapation.
+\citet{2019arXiv190910291P} proposes a simple test of spiral arm winding, assuming the cotangent of the pitch angle evolves linearly with time.
+Finding compelling evidence of this relationship against QSDW theory, in favour of the dynamic spirals now favoured by many simulations.
+
+Spiral evolution also appears to be influenced by the presence and strength of a bar; in barred grand-design spirals the arms often appear to start from the ends of the bar.
+Simulations of gas in barred galaxies often demonstrate that bars can drive long-term spiral evolution \citep{2008A&A...489..115R}, or boost transient spiral structure \citep{2012MNRAS.426..167G}.
+Manifold theory is one attempt to determine the orbits of stars in bar-driven spiral arms: it proposes that stars in the vicinity of the unstable Lagrangian points at either end of the bar tend to escape along predictable orbits, governed by invariant manifolds.
+One of the primary factors influencing the shape of this invariant manifold is the relative strength of the non-axisymmetric forcing caused by the bar, with stronger bars resulting in spirals with larger pitch angles.
+
+Many other systems contribute to spiral morphology, including potential ties to bulge fraction (\citealt{1975A&A....44..363Y}, \citealt{2013MNRAS.436.1074S}, \citealt{2019MNRAS.487.1808M}) and black hole mass (\citealt{2015PhDT........14D}, \citealt{2017MNRAS.471.2187D}), via a series of correlations.
+Stronger bulges and more more massive central black holes are both linked to more tightly wound spiral arms, albeit with large amounts of scatter.
+
 \subsection{Measuring galaxy pitch angle}
@@
 \textsc{2DFFT}, \citealt{2012ApJS..199...33D}), texture analysis (i.e.
-SpArcFiRe, \citealt{2014ApJ...790...87D}), and combinations of automated methods and human classifiers \citep{2017MNRAS.472.2263H}.
-One potentially underused method of obtaining measurements of spirals is through photometric fitting of spiral structure, as possible using tools such as \textsc{GALFIT} \citep{2010AJ....139.2097P} and \textit{Galaxy Builder} (\comment{Lingard et al, in prep}).
+SpArcFiRe, \citealt{2014ApJ...790...87D}), and combinations of automated methods and human classifiers (\citealt{2017MNRAS.472.2263H}, \citealt{2020MNRAS.493.3854H}).
+One potentially underused method of obtaining measurements of spirals is through photometric fitting of spiral structure, as possible using tools such as \textsc{GALFIT} \citep{2010AJ....139.2097P} and \textit{Galaxy Builder} (\Lingard).
 These methods attempt to localize light from an image of a galaxy into distinct subcomponents, such as a galaxy disc, bulge, bar and spiral arms, generally finding the optimum solution using computational optimization.
 This optimization process, however, is often not robust for complex, many-component models and requires significant supervision to converge to a physically meaningful result \citep{Gao2017:1709.00746v1}.
-\comment{Lingard et al, in prep} propose a solution to this problem through the use of citizen science to provide the starting points for computational fitting.
+\Lingard propose a solution to this problem through the use of citizen science to provide priors on parameters used in computational fitting.
 
@@
 
-The most commonly used measurement of uncertainty of length-weighted pitch angles is the \comment{unweighted?} sample variance between the arm segments identified.
+The most commonly used measurement of uncertainty of length-weighted pitch angles is the unweighted sample variance between the arm segments identified.
 
@@
 
-Fourier analysis in one-dimension (as performed by \comment{...}) and two-dimensions (\comment{...}) is another widely used method of computationally obtaining galaxy pitch angles.
+Fourier analysis in one- and two-dimensions (as performed by \citealt{2019arXiv190804246D}, \citealt{2012ApJS..199...33D}, \citealt{2018MNRAS.474.2594M}) is another widely used method of computationally obtaining galaxy pitch angles.
 Two-dimensional fourier methods generally decompose a deprojected image of a galaxy into a superpositions of logarithmic spirals between inner and outer annuli \citep{2012ApJS..199...33D}, and reports the pitch angle with the highest amplitude as the galaxy's pitch angle.
@@
 
-This paper makes use of the photometric models obtained through the \textit{Galaxy Builder} citizen science project for the 196 spiral galaxies present in \comment{Lingard et al, in prep}.
+This paper makes use of the classification data and fitted photometric models obtained through the \textit{Galaxy Builder} citizen science project for the 196 spiral galaxies present in \Lingard.
 In this paper we focus on the use of measured spiral tightness (quantified using pitch angle, \citealt{1987gady.book.....B}) as a probe into the dynamical mechanisms governing a spiral galaxy's evolution.
@@
 
-Section \ref{section:spiral_winding} investigates spiral arm winding using the test derived by \citealt{2019arXiv190910291P} (uniformity of galaxy pitch angle in $\cot\;\phi$) and concludes using a marginalized Anderson-Darling test that we \comment{cannot unilaterally reject winding of this form at the 1\% level}.
-Section \ref{section:morphology_comparision} examines the correlation between pitch angle and bulge size implied by the Hubble sequence and reported by \comment{citation}, and pitch angle and bar strength implied by Manifold theory, and find \comment{no significant correlation}.
+Section \ref{section:spiral_winding} investigates spiral arm winding using the test derived by \cite{2019arXiv190910291P} (uniformity of galaxy pitch angle in $\cot\;\phi$) and concludes using a marginalized Anderson-Darling test that we cannot unilaterally reject winding of this form at the 1\% level.
+Section \ref{section:morphology_comparision} examines the correlation between pitch angle and bulge size implied by the Hubble sequence, and pitch angle and bar strength implied by Manifold theory, and find no significant correlation.

### 2020-03-23 sections/method.tex
Includes new results (deprojection from GZB fitting, updated hyperparameters on clustering). Greatly expands on intro section

@@
 \subsection{The Galaxy Sample}
-The galaxies selected analysed in this paper are the 198 galaxies from \comment{Lingard et al, in prep.}.
+The galaxies selected analysed in this paper are the 198 galaxies from \Lingard.
+These are a subset of the \textit{stellar mass-complete sample} in \citet{2017MNRAS.472.2263H}, a sample of low-redshift face-on spiral galaxies selected using data from the NASA-Sloan Atlas \citep{2011AJ....142...31B} and Galaxy Zoo 2 \citep{Willett2013:1308.3496v2}.
+
 We combine classifications of galaxies which were repeated in the \textit{validation subset} with the original classifications.
-Clustering of drawn spiral arms and cleaning of points was then performed as detailed in \comment{Lingard et al, in prep.}.
-We remove any galaxies for which no arms were identified.
-This results in a hierarchiacal data structure of 109 galaxies, 250 spiral arms and 307,861 points.
+Clustering of drawn spiral arms and cleaning of points was then performed as detailed in \Lingard.
+We remove any galaxies for which fewer than two spiral arms were identified \comment{why exclude one-armed?}.
+This results in a hierarchical data structure of 91 galaxies, 211 spiral arms and 215,678 points.
 
-Spiral arm points are deprojected to a face-on orientation using the position angle \textsc{PETRO\_PHI90} and axis ratio \textsc{PETRO\_BA90} keywords from the NASA-Sloan Atlas \comment{citation}.
-We scale the radii to have unit variance (which does not impact log spiral pitch angle due to its scale invariance).
-
-\comment{Should I elaborate more on the sample?}
+Spiral arm points are deprojected to a face-on orientation using the disk inclination and position angle obtained through photometric model fitting performed in \Lingard.
+We scale the deprojected point radii to have unit variance, to aid chain convergence.
 
@@
 
-\begin{figure}
-  \includegraphics[width=8cm]{plots/plots_n109d1000t500/model.pdf}
-  \caption{The model used for galaxy pitch angle measurement for the sample.}
-  \label{fig:pymc3-model}
-\end{figure}
+% \begin{figure}
+%   \includegraphics[width=8cm]{plots/plots_n109d1000t500/model.pdf}
+%   \caption{The model used for galaxy pitch angle measurement for the sample.}
+%   \label{fig:pymc3-model}
+% \end{figure}
 

### 2020-03-23 sections/results.tex
Includes new results (deprojection from GZB fitting, updated hyperparameters on clustering). Greatly expands on intro section

@@
 \subsection{Constraints on Galaxy Pitch angle}
-Our hierarchical model can predict individual arm pitch angle with a hich degree of certainty, however it returns a large spread of potential values for the pitch angles of arms
+Our hierarchical model identifies arm pitch angle with a hich degree of certainty (uncertainty less than 1.6\degree for 95\% of arms, assuming no error on disc inclination and position angle), however it returns a large spread of potential values for the pitch angles of galaxies, primarily caused by only measuring pitch angles of a small number of arms per galaxy.
+For galaxies with two arms identified in \textit{Galaxy Builder}, we have a mean uncertainty of 7.8\degree, which decreases to 6.7\degree and 5.8\degree for galaxies with three and four arms respectively.
+
+Our measure of inter-arm variability of pitch angle, $\sigma_\mathrm{gal} = 11.02 \pm 0.95$, confirming
 
@@
 The resulting distribution of Anderson-Darling statistics can be seen in Figure \ref{fig:ad-cot-test}.
-We observe that we reject the null hypothesis at the 1\% level for only 86\% of the possible realizations of galaxy pitch angle, therefore with our sample and methodology we cannot unilaterally reject winding of the kind described by \citet{2019arXiv190910291P}.
+We observe that we reject the null hypothesis at the 1\% level for 69\% of the possible realizations of galaxy pitch angle.
+While, with our sample and methodology, we cannot unilaterally reject winding of the kind described by \citet{2019arXiv190910291P}, it suggests this model is not found in our data.
 
@@
 
+Not seeing uniformity in $\cot\phi$ is not evidence against spiral winding, nor is it evidence for static spiral structures.
+Our results simply suggest that a more comprehensive test is needed for observational data of this kind.
+
 \subsection{Dependence of pitch angle on Galaxy Morphology}
@@
 We separate our sample into ``disc-dominated galaxies'' and ``obvious bulge galaxies'' using the debiased fractions from Galaxy Zoo 2 following \citet{2017MNRAS.469.3363K}, defining \textit{no bulge + just noticeable $>$ obvious + dominant} for the former and the converse for the latter.
-A marginalized two-sample Anderson-Darling test \citep{doi:10.1080/01621459.1987.10478517} does not find strong evidence that the samples were drawn from different distributions; we reject the null hypothesis at the 1\% level for only 8\% of the samples.
+A marginalized two-sample Anderson-Darling test \citep{doi:10.1080/01621459.1987.10478517} does not find any evidence that the samples were drawn from different distributions; we do not reject the null hypothesis at the 1\% level for any samples.
+The distribution of Anderson-Darling test statistics is shown in the upper panel of Figure \ref{fig:ad-morphology-test}.
 
@@
 
-Separating the sample based off of $\mathrm{\textit{pbar}} > 0.5$, and restricting to galaxies with more than 10 classifications for \textit{pbar} (as performed by \citealt{2011MNRAS.411.2026M} and \citealt{2017MNRAS.469.3363K}) and performing a marginalized two-sample Anderson-Darling test does not find that the samples were drawn from different distributions.
-\comment{Talk about number of galaxies in each sub-sample?}
+Separating the sample based off of $\mathrm{\textit{pbar}} > 0.5$, and restricting to galaxies with more than 10 classifications for \textit{pbar} (as performed by \citealt{2011MNRAS.411.2026M} and \citealt{2017MNRAS.469.3363K}) and performing a marginalized two-sample Anderson-Darling test does not find that the samples were drawn from different distributions (we reject the null hypothesis at the 1\% level for only 1\% of samples).
+The distribution of Anderson-Darling test statistics is shown in the lower panel of Figure \ref{fig:ad-morphology-test}.
+
+We do not account for variation in bulge size in this work, however predictions from Manifold theory should not be affected as bulges do not provide a non-axisymmetric focing.
 

### 2020-04-01 sections/discussion.tex
addresses most comments in #1

@@
 
-We we do not find a relationship between bar strength and pitch angle, as would have been predicted by Manifold theory, and do not find evidence for the relationship between central mass concentration and pitch angle predicted by the Hubble sequence.
+We do not find a relationship between bar strength and pitch angle, as would have been predicted by Manifold theory, and do not find evidence for the relationship between central mass concentration and pitch angle predicted by the Hubble sequence.
 

### 2020-04-01 sections/introduction.tex
addresses most comments in #1

@@
 These variatons on structure account for $18\%$, $50\%$ and $32\%$ of the population respectively (\citealt{2011ApJ...737...32E}, \citealt{2015yCat..22170032B}).
-Our current understanding of the mechanisms which drive spiral growth and evolution suggest that each of the different forms of spiral galaxy may be triggered primarily by different processes.
-Grand Design spirals are thought to have undergone a tidal interaction, be driven by a bar (as seen in gas simulations, \citealt{1976ApJ...209...53S}, \citealt{2008A&A...489..115R}, and suggested for stars by Manifold theory, \citealt{2006A&A...453...39R}, \citealt{2009MNRAS.394...67A}, \citealt{2009MNRAS.400.1706A}), or be obeying (quasi-stationary) density wave theory (QSDW theory), in which spiral arms are slowly evolving, ever-present structures in the disc \citep{1964ApJ...140..646L}.
-Flocculent spirals are thought to be formed through swing amplification (shearing of small gravitational instabilities in the disc), and be transient and reccurent in nature \citep{1966ApJ...146..810J}.
-It is recognised that methods of spiral formation are not mutually exclusive.
-There is also debate as to the evolution of spiral structure: whether spirals maintain their shape over many rotational periods (assumed by QSDW theory), or ``wind up'' with the differential rotation of the disc (as found by \citealt{2003MNRAS.344..358B}, \citealt{2019MNRAS.487.1808M}).
 
@@
 
-\subsection{Secular evolution of spiral arms}
+\subsection{Formation and Evolution of Spiral Arms}
 
-\comment{I'm worried that a lot of this section repeats what was said earlier}
+Our current understanding of the mechanisms which drive spiral growth and evolution suggest that each of the different forms of spiral galaxy may be triggered primarily by different processes.
+Grand Design spirals are thought to have undergone a tidal interaction, be driven by a bar (as seen in gas simulations, \citealt{1976ApJ...209...53S}, \citealt{2008A&A...489..115R}, and suggested for stars by Manifold theory, \citealt{2006A&A...453...39R}, \citealt{2009MNRAS.394...67A}, \citealt{2009MNRAS.400.1706A}), or be obeying (quasi-stationary) density wave theory (QSDW theory), in which spiral arms are slowly evolving, ever-present structures in the disc \citep{1964ApJ...140..646L}.
+Flocculent spirals are thought to be formed through swing amplification (shearing of small gravitational instabilities in the disc), and be transient and reccurent in nature \citep{1966ApJ...146..810J}.
+It is recognised that methods of spiral formation are not mutually exclusive.
 
-One of the foundational assumptions of early work on spiral formation mechanisms (primarily QSDW) was that the disc of a galaxy, if unstable to spiral perturbations, would create a stable, static wave which would exist unchanging for many rotational periods \citep{1964ApJ...140..646L}.
+One of the fundamental assumptions of early work on spiral formation mechanisms (primarily QSDW) was that the disc of a galaxy, if unstable to spiral perturbations, would create a stable, static wave which would exist unchanging for many rotational periods \citep{1964ApJ...140..646L}.
 The motivation for static waves with small numbers of arms (with a preference for $m=2$) was primarily observational; most galaxies show spiral structure, indicating that spirals exist for a long time or are continually rebuilt.
@@
 Two-dimensional fourier methods generally decompose a deprojected image of a galaxy into a superpositions of logarithmic spirals between inner and outer annuli \citep{2012ApJS..199...33D}, and reports the pitch angle with the highest amplitude as the galaxy's pitch angle.
-It is uncertain how the large observed variation between pitch angles of individual arms impacts this measurement.
+\citet{2020MNRAS.493.3854H} combined fourier analysis of spiral galaxies with human tracing of spiral arms, to great effect.
+It is uncertain how variation between pitch angles of individual arms impacts this measurement.
 

### 2020-04-01 sections/method.tex
addresses most comments in #1

@@
 \subsection{The Galaxy Sample}
-The galaxies selected analysed in this paper are the 198 galaxies from \Lingard.
+The galaxies analysed in this paper are the 198 galaxies from \Lingard.
 These are a subset of the \textit{stellar mass-complete sample} in \citet{2017MNRAS.472.2263H}, a sample of low-redshift face-on spiral galaxies selected using data from the NASA-Sloan Atlas \citep{2011AJ....142...31B} and Galaxy Zoo 2 \citep{Willett2013:1308.3496v2}.
 
-We combine classifications of galaxies which were repeated in the \textit{validation subset} with the original classifications.
+Some galaxies in \Lingard were shown to volunteers again in a repeat \textit{validation subset} in order to create a second aggregate model used to test internal consistency.
+We combine the classifications of galaxies in this \textit{validation subset} with the original classifications.
 Clustering of drawn spiral arms and cleaning of points was then performed as detailed in \Lingard.
@@
 Spiral arm points are deprojected to a face-on orientation using the disk inclination and position angle obtained through photometric model fitting performed in \Lingard.
-We scale the deprojected point radii to have unit variance, to aid chain convergence.
 
@@
 \begin{equation}
-r_\mathrm{arm} = \exp\left[\theta_\mathrm{arm}\tan\phi_\mathrm{arm} + c_\mathrm{arm}\right].
+r_\mathrm{arm} \sim \mathcal{N}\left(\exp\left[\theta_\mathrm{arm}\tan\phi_\mathrm{arm} + c_\mathrm{arm}\right], \sigma_r\right).
 \end{equation}
@@
 Assume the pitch angle of a galaxy's spiral arms are drawn from a Normal distribution, truncated between 0 and 90 degrees, centred on some value we will call the galaxy's pitch angle, $\phi_\mathrm{gal}$.
-This dispersion in arm pitch angle has some measure of spread, $\sigma_\mathrm{gal}$, which we will assume is the same in galaxies:
+This dispersion in arm pitch angle has some measure of spread, $\sigma_\mathrm{gal}$, which we will assume is the same in all galaxies:
 
@@
 
+Our likelihood function is therefore
+
+\begin{equation}
+\mathcal{L} \sim \mathcal{N}\left(\exp\left[\theta_\mathrm{arm}\tan\phi_\mathrm{arm} + c_\mathrm{arm}\right], \sigma_r\right).
+\end{equation}
+
 To perform inference, we make use of the No-U-Turn-Sampler (NUTS, \citealt{2011arXiv1111.4246H}), implemented in PYMC3\footnote{\url{https://docs.pymc.io/}}, an open source probabilistic programming framework written in python \citep{pymc3_paper}.
+We scale the radii of deprojected points to have unit variance, to aid the convergence of MC chains.

### 2020-04-01 sections/results.tex
addresses most comments in #1

@@
 \subsection{Constraints on Galaxy Pitch angle}
-Our hierarchical model identifies arm pitch angle with a hich degree of certainty (uncertainty less than 1.6\degree for 95\% of arms, assuming no error on disc inclination and position angle), however it returns a large spread of potential values for the pitch angles of galaxies, primarily caused by only measuring pitch angles of a small number of arms per galaxy.
+Our hierarchical model identifies arm pitch angle with a high degree of certainty (uncertainty less than 1.6\degree for 95\% of arms, assuming no error on disc inclination and position angle), however it returns a large spread of potential values for the pitch angles of galaxies, primarily caused by only measuring pitch angles of a small number of arms per galaxy.
 For galaxies with two arms identified in \textit{Galaxy Builder}, we have a mean uncertainty of 7.8\degree, which decreases to 6.7\degree and 5.8\degree for galaxies with three and four arms respectively.
 
-Our measure of inter-arm variability of pitch angle, $\sigma_\mathrm{gal} = 11.02 \pm 0.95$, confirming
+Our measure of inter-arm variability of pitch angle, $\sigma_\mathrm{gal}$ has a posterior distribution of $11.02 \pm 0.95$, confirming the finding of \citet{2014ApJ...790...87D} and emphaising the need for fitting algorithms to not assume all arms have the same pitch angle.
 
@@
 \subsubsection{Pitch angle vs. Bulge size}
-We see no correlation between galaxy pitch angle derived from the \textit{hierarchical normal model} and Galaxy Zoo 2's debiased \citep{Willett2013:1308.3496v2} \textit{pbulge}, which has been shown to be a good measure of bulge size.
+We see no correlation between galaxy pitch angle derived from the hierarchical model and Galaxy Zoo 2's debiased \citep{Willett2013:1308.3496v2} \textit{pbulge}, which has been shown to be a good measure of bulge size.
 

### 2020-04-02 sections/acknowledgements.tex
Includes new results (deprojection from GZB fitting, updated hyperparameters on clustering). Greatly expands on intro section
Continues making changes suggested by KLM and CK

@@
 
-This project was partially funded by a Google Faculty Research Award to Karen Masters (\url{https://ai.google/research/outreach/faculty-research-awards/})
+This publication uses data generated via the Zooniverse.org platform, development of which is funded by generous support, including a Global Impact Award from Google, and by a grant from the Alfred P.
+Sloan Foundation.
+We would also like to thank the 2,340 volunteers who have submitted classifications to the \textit{Galaxy Builder} project, especially user EliabethB, whose presence on the \textit{Galaxy Builder} forum on top of a large number of galaxies modelled has been a huge help.
+
+This project was partially funded by a Google Faculty Research Award to Karen Masters (\url{https://ai.google/research/outreach/faculty-research-awards/}), and Timothy Lingard acknowledges studentship funding from the Science and Technology Facilities Council (ST/N504245/1).

### 2020-04-02 sections/introduction.tex
Continues making changes suggested by KLM and CK

@@
 These variatons on structure account for $18\%$, $50\%$ and $32\%$ of the population respectively (\citealt{2011ApJ...737...32E}, \citealt{2015yCat..22170032B}).
+The Hubble classification scheme \citep{1926ApJ....64..321H}, and its revisions and expansions (\citealt{1961hag..book.....S}, \citealt{1991rc3..book.....D}) contain detailed variations of different types of spiral galaxy, divided by the presence of a bar and ordered by how obvious spiral arm patterns are, how tightly they are wound and the prominence of a central bulge.
 
@@
 These predictions and observations provide compelling reasons for investigating their underlying rules and dynamics, as doing so is essential for understanding the secular evolution of disc galaxies.
-
-\subsection{Formation and Evolution of Spiral Arms}
 
@@
 One of the fundamental assumptions of early work on spiral formation mechanisms (primarily QSDW) was that the disc of a galaxy, if unstable to spiral perturbations, would create a stable, static wave which would exist unchanging for many rotational periods \citep{1964ApJ...140..646L}.
-The motivation for static waves with small numbers of arms (with a preference for $m=2$) was primarily observational; most galaxies show spiral structure, indicating that spirals exist for a long time or are continually rebuilt.
+The motivation for static waves with small numbers of arms (with a preference for $m=2$) was primarily observational; most galaxies show spiral structure, suggesting that spirals exist for a long time or are continually rebuilt.
 
@@
 Many other systems contribute to spiral morphology, including potential ties to bulge fraction (\citealt{1975A&A....44..363Y}, \citealt{2013MNRAS.436.1074S}, \citealt{2019MNRAS.487.1808M}) and black hole mass (\citealt{2015PhDT........14D}, \citealt{2017MNRAS.471.2187D}), via a series of correlations.
-Stronger bulges and more more massive central black holes are both linked to more tightly wound spiral arms, albeit with large amounts of scatter.
-
-\subsection{Measuring galaxy pitch angle}
-
-Many methodologies have been proposed and implemented to measure spiral arm properties, including visual inspection (\citealt{2015A&A...582A..86H}), fourier analysis (i.e.
-\textsc{2DFFT}, \citealt{2012ApJS..199...33D}), texture analysis (i.e.
-SpArcFiRe, \citealt{2014ApJ...790...87D}), and combinations of automated methods and human classifiers (\citealt{2017MNRAS.472.2263H}, \citealt{2020MNRAS.493.3854H}).
-One potentially underused method of obtaining measurements of spirals is through photometric fitting of spiral structure, as possible using tools such as \textsc{GALFIT} \citep{2010AJ....139.2097P} and \textit{Galaxy Builder} (\Lingard).
-These methods attempt to localize light from an image of a galaxy into distinct subcomponents, such as a galaxy disc, bulge, bar and spiral arms, generally finding the optimum solution using computational optimization.
-This optimization process, however, is often not robust for complex, many-component models and requires significant supervision to converge to a physically meaningful result \citep{Gao2017:1709.00746v1}.
-\Lingard propose a solution to this problem through the use of citizen science to provide priors on parameters used in computational fitting.
-
-A common assumption when measuring galaxy pitch angle is that observed spiral arms have a constant pitch angle. Spirals of this kind are known as logarithmic spirals and are described by
-
-\begin{equation}
-r = A\,e^{\theta\tan\phi},
-\end{equation}
-
-where $\phi$ is the arm's pitch angle.
-One method used to obtain a pitch angle of a galaxy is to fit logarithmic spirals to individually identified arm segments and take the weighted mean of their pitch angles (which often vary by upwards of $10^\circ$, \citealt{2014ApJ...790...87D}).
-Weighting is determined by the length of the arc segment, with longer being assigned higher weights, i.e.
-for a galaxy where we have identified $N$ arm segments, each with length $L_i$ and pitch angle $\phi_i$
-
-\begin{equation}
-  \phi_\mathrm{gal} = \left(\sum_{i=1}^{N}L_i\right)^{-1}\sum_{i=1}^{N}L_i \phi_i.
-\end{equation}
-
-The most commonly used measurement of uncertainty of length-weighted pitch angles is the unweighted sample variance between the arm segments identified.
-
-A notable drawback of length-weighted pitch angle is sensitivity to the number and quality of the spiral arm segments identified; \citet{2017MNRAS.472.2263H} found that only 15\% of the arm segments identified by a leading algorithm (\textsc{SpArcFiRe}) were identified as ``good'' matches to real spiral arms by citizen science classifiers.
-
-Fourier analysis in one- and two-dimensions (as performed by \citealt{2019arXiv190804246D}, \citealt{2012ApJS..199...33D}, \citealt{2018MNRAS.474.2594M}) is another widely used method of computationally obtaining galaxy pitch angles.
-Two-dimensional fourier methods generally decompose a deprojected image of a galaxy into a superpositions of logarithmic spirals between inner and outer annuli \citep{2012ApJS..199...33D}, and reports the pitch angle with the highest amplitude as the galaxy's pitch angle.
-\citet{2020MNRAS.493.3854H} combined fourier analysis of spiral galaxies with human tracing of spiral arms, to great effect.
-It is uncertain how variation between pitch angles of individual arms impacts this measurement.
+Stronger bulges and more more massive central black holes have both been linked to more tightly wound spiral arms.
 

### 2020-04-02 sections/method.tex
Continues making changes suggested by KLM and CK

@@
 % !TEX root = ../main.tex
+\subsection{Measuring galaxy pitch angle}
+
+Many methodologies have been proposed and implemented to measure spiral arm properties, including visual inspection (\citealt{2015A&A...582A..86H}), fourier analysis (i.e.
+\textsc{2DFFT}, \citealt{2012ApJS..199...33D}), texture analysis (i.e.
+SpArcFiRe, \citealt{2014ApJ...790...87D}), and combinations of automated methods and human classifiers (\citealt{2017MNRAS.472.2263H}, \citealt{2020MNRAS.493.3854H}).
+One potentially underused method of obtaining measurements of spirals is through photometric fitting of spiral structure, as possible using tools such as \textsc{GALFIT} \citep{2010AJ....139.2097P} and \textit{Galaxy Builder} (\Lingard).
+These methods attempt to localize light from an image of a galaxy into distinct subcomponents, such as a galaxy disc, bulge, bar and spiral arms, generally finding the optimum solution using computational optimization.
+This optimization process, however, is often not robust for complex, many-component models and requires significant supervision to converge to a physically meaningful result \citep{Gao2017:1709.00746v1}.
+\Lingard propose a solution to this problem through the use of citizen science to provide priors on parameters used in computational fitting.
+
+A common assumption when measuring galaxy pitch angle is that observed spiral arms have a constant pitch angle. Spirals of this kind are known as logarithmic spirals and are described by
+
+\begin{equation}
+r = A\,e^{\theta\tan\phi},
+\end{equation}
+
+where $\phi$ is the arm's pitch angle.
+One method used to obtain a pitch angle of a galaxy is to fit logarithmic spirals to individually identified arm segments and take the weighted mean of their pitch angles (which often vary by upwards of $10^\circ$, \citealt{2014ApJ...790...87D}).
+Weighting is determined by the length of the arc segment, with longer being assigned higher weights, i.e.
+for a galaxy where we have identified $N$ arm segments, each with length $L_i$ and pitch angle $\phi_i$
+
+\begin{equation}
+  \phi_\mathrm{gal} = \left(\sum_{i=1}^{N}L_i\right)^{-1}\sum_{i=1}^{N}L_i \phi_i.
+\end{equation}
+
+The most commonly used measurement of uncertainty of length-weighted pitch angles is the unweighted sample variance between the arm segments identified.
+
+A notable drawback of length-weighted pitch angle is sensitivity to the number and quality of the spiral arm segments identified; \citet{2017MNRAS.472.2263H} found that only 15\% of the arm segments identified by a leading algorithm (\textsc{SpArcFiRe}) were identified as ``good'' matches to real spiral arms by citizen science classifiers.
+
+Fourier analysis in one- and two-dimensions (as performed by \citealt{2019arXiv190804246D}, \citealt{2012ApJS..199...33D}, \citealt{2018MNRAS.474.2594M}) is another widely used method of computationally obtaining galaxy pitch angles.
+Two-dimensional fourier methods generally decompose a deprojected image of a galaxy into a superpositions of logarithmic spirals between inner and outer annuli \citep{2012ApJS..199...33D}, and reports the pitch angle with the highest amplitude as the galaxy's pitch angle.
+\citet{2020MNRAS.493.3854H} combined fourier analysis of spiral galaxies with human tracing of spiral arms, to great effect.
+It is uncertain how variation between pitch angles of individual arms impacts this measurement.
+
 \subsection{The Galaxy Sample}
@@
 Clustering of drawn spiral arms and cleaning of points was then performed as detailed in \Lingard.
-We remove any galaxies for which fewer than two spiral arms were identified \comment{why exclude one-armed?}.
+We remove any galaxies for which fewer than two spiral arms were identified, as these do not provide our model with any information about inter-arm scatter.
 This results in a hierarchical data structure of 91 galaxies, 211 spiral arms and 215,678 points.
@@
 \begin{equation}
-r_\mathrm{arm} \sim \mathcal{N}\left(\exp\left[\theta_\mathrm{arm}\tan\phi_\mathrm{arm} + c_\mathrm{arm}\right], \sigma_r\right).
+r_\mathrm{arm} \sim \mathrm{Normal}\left(\exp\left[\theta_\mathrm{arm}\tan\phi_\mathrm{arm} + c_\mathrm{arm}\right], \sigma_r\right).
 \end{equation}
@@
 
-% \begin{figure}
-%   \includegraphics[width=8cm]{plots/plots_n109d1000t500/model.pdf}
-%   \caption{The model used for galaxy pitch angle measurement for the sample.}
-%   \label{fig:pymc3-model}
-% \end{figure}
-
+The Cauchy distirbution is equivalent to the Student's t-distribution with one degree of freedom, and was chosen due to its fatter tails than the normal distribution.
 Our likelihood function is therefore
@@
 \begin{equation}
-\mathcal{L} \sim \mathcal{N}\left(\exp\left[\theta_\mathrm{arm}\tan\phi_\mathrm{arm} + c_\mathrm{arm}\right], \sigma_r\right).
+\mathcal{L} \sim \mathrm{Normal}\left(\exp\left[\theta_\mathrm{arm}\tan\phi_\mathrm{arm} + c_\mathrm{arm}\right], \sigma_r\right).
 \end{equation}

### 2020-04-02 sections/results.tex
Continues making changes suggested by KLM and CK

@@
 \subsection{Constraints on Galaxy Pitch angle}
-Our hierarchical model identifies arm pitch angle with a high degree of certainty (uncertainty less than 1.6\degree for 95\% of arms, assuming no error on disc inclination and position angle), however it returns a large spread of potential values for the pitch angles of galaxies, primarily caused by only measuring pitch angles of a small number of arms per galaxy.
-For galaxies with two arms identified in \textit{Galaxy Builder}, we have a mean uncertainty of 7.8\degree, which decreases to 6.7\degree and 5.8\degree for galaxies with three and four arms respectively.
+Our hierarchical model identifies the pitch angle of individual arms ($\phi_\mathrm{arm}$) with a high degree of certainty (uncertainty less than {1.6\degree} for 95\% of arms, assuming no error on disc inclination and position angle), however the pitch angle of a galaxy as a whole ($\phi_\mathrm{gal}$) is not well constrained.
+This is primarily a result of only having pitch angles measurements for a small number of arms per galaxy, and reflects the difficulty in providing a single value for the pitch angle of a galaxy containing individual arms with very different pitch angles.
+For galaxies with two arms identified in \textit{Galaxy Builder}, we have a mean uncertainty of {7.8\degree}, which decreases to {6.7\degree} and {5.8\degree} for galaxies with three and four arms respectively.
 
-Our measure of inter-arm variability of pitch angle, $\sigma_\mathrm{gal}$ has a posterior distribution of $11.02 \pm 0.95$, confirming the finding of \citet{2014ApJ...790...87D} and emphaising the need for fitting algorithms to not assume all arms have the same pitch angle.
+Our measure of inter-arm variability of pitch angle, $\sigma_\mathrm{gal}$ has a posterior distribution of $11.0\pm 1.0$, similar to the finding of \citet{2014ApJ...790...87D} and emphaising the need for fitting algorithms to not assume all arms have the same pitch angle.
 
@@
 We will refer to this test as the \textit{marginalized Anderson-Darling test}.
-We make use of the Kolmogorov-Smirnov test in a similar manner for comparison.
+We also make use of the Kolmogorov-Smirnov test in a similar manner for comparison.
 

### 2020-04-14 sections/discussion.tex
separates analysis of galaxy pitch angle and arm pitch angle

@@
 
-Our method and results do not completely reject spiral winding of the form described by \citet{2019arXiv190910291P}, however this result is highly sensitive to the boundaries used; using limits of $10.0^\circ < \phi < 40.0^\circ$ results in us rejecting the null hypothesis at the 1\% level for 99\% of the MCMC draws.
-As no physical justification was provided for the limits present in \citet{2019arXiv190910291P}, this is a far from ideal test of arm winding.
+We do not find a relationship between bar strength and pitch angle, as would have been predicted by Manifold theory, and do not find evidence for the relationship between central mass concentration and pitch angle predicted by the Hubble sequence.
 
-We do not find a relationship between bar strength and pitch angle, as would have been predicted by Manifold theory, and do not find evidence for the relationship between central mass concentration and pitch angle predicted by the Hubble sequence.
+Our results are consistent with spiral winding of the form described by \citet{2019arXiv190910291P}; providing evidence for transient and recurrent spiral arms, the evolution of which is governed by self-gravity (such as through swing-amplification, \citealt{1965MNRAS.130..125G})) and which wind up over time.
+The assumptions of this model of spiral winding are also highly reductionist, and leaves many unanswered questions: what determines the limits on $\phi$?
+Is the spiral arm equally apparent at all pitch angles, or is a selection effect present?
+This result is not evidence against QSDW, as it is possible that our distribtuion of pitch angles is dictated by other factors such as disk shear and central mass concentration.
 

### 2020-04-14 sections/introduction.tex
separates analysis of galaxy pitch angle and arm pitch angle

@@
 In this dynamic picture of spiral arms, pitch angle monotonically decreases from a spiral arm's formation to its dissapation.
-\citet{2019arXiv190910291P} proposes a simple test of spiral arm winding, assuming the cotangent of the pitch angle evolves linearly with time.
-Finding compelling evidence of this relationship against QSDW theory, in favour of the dynamic spirals now favoured by many simulations.
+\citet{2019arXiv190910291P} proposes a simple test of spiral arm winding, assuming the cotangent of the pitch angle of a spiral arm evolves linearly with time.
+They found that the pitch angles of their sample of 86 galaxies was consistent with this distribution, evidence against QSDW theory in favour of the dynamic spirals produced in many simulations.
 

### 2020-04-14 sections/method.tex
separates analysis of galaxy pitch angle and arm pitch angle

@@
 Spiral arm points are deprojected to a face-on orientation using the disk inclination and position angle obtained through photometric model fitting performed in \Lingard.
+Arms are individually corrected to all have the same chirality (a pitch angle greater than or equal to zero) using the logarithmic spiral fit in \Lingard.
 
@@
 
-We model a spiral arm as a logarithmic spiral, described by
-
-\begin{equation}
-r_\mathrm{arm} \sim \mathrm{Normal}\left(\exp\left[\theta_\mathrm{arm}\tan\phi_\mathrm{arm} + c_\mathrm{arm}\right], \sigma_r\right).
-\end{equation}
-
-Assume the pitch angle of a galaxy's spiral arms are drawn from a Normal distribution, truncated between 0 and 90 degrees, centred on some value we will call the galaxy's pitch angle, $\phi_\mathrm{gal}$.
-This dispersion in arm pitch angle has some measure of spread, $\sigma_\mathrm{gal}$, which we will assume is the same in all galaxies:
+We assume that a galaxy has a single value for pitch angle, $\phi_\mathrm{gal}$, and that the pitch angles of spiral arms in this galaxy, $\phi_\mathrm{arm}$, are constants (giving logarithmic spirals) drawn from a normal distribution centred on $\phi_\mathrm{gal}$, with some spread $\sigma_\mathrm{gal}$ common to all galaxies.
+We truncate this normal distribution between the physical limits of {0\degree} (a ring) and {90\degree} (a ``spoke''), giving
 
@@
 
-We choose hyperpriors on $\phi_\mathrm{gal}$ and $\sigma_\mathrm{gal}$ of
+Furthermore, we assume that the observed points in a \textit{Galaxy Builder} spiral arm, once deprojected, follow a logarithmic spiral with gaussian radial error $\sigma_r$,
+
+\begin{equation}
+\widetilde{r_\mathrm{arm}} = \exp\left(\widevec{\theta_\mathrm{arm}}\tan\phi_\mathrm{arm} + c_\mathrm{arm}\right).
+\end{equation}
+
+Where $\widetilde{r_\mathrm{arm}}$ is the model's predictions for the radii of the deprojected points in a \textit{Galaxy Builder} arm ($\widevec{r_\mathrm{arm}}$), and $\widevec{\theta_\mathrm{arm}}$ is the polar angles of the points.
+
+We choose hyperpriors over $\phi_\mathrm{gal}$, $\sigma_\mathrm{gal}$, $c_\mathrm{arm}$ and $\sigma_\mathrm{r}$ of
 
@@
   \phi_\mathrm{gal} &\sim \mathrm{Uniform}(\mathrm{min}=0, \mathrm{max}=90),\\
-  \sigma_\mathrm{gal} &\sim \mathrm{InverseGamma}(\alpha=2,\,\beta=20).
+  \sigma_\mathrm{gal} &\sim \mathrm{InverseGamma}(\alpha=2,\,\beta=20),\\
+  c_\mathrm{arm} &\sim \mathrm{Cauchy}(\alpha=0,\,\beta=10),\\
+  \sigma_r &\sim \mathrm{InverseGamma}(\alpha=2, \beta=0.5).
 \end{align}
 
-We also have the offset parameter $c$, and a measure of radial uncertainty $\sigma_r$:
-
-\begin{align}
-  c_\mathrm{arm} &\sim \mathrm{Cauchy}(\alpha=0,\,\beta=10),\\
-  \sigma_r &\sim \mathrm{HalfCauchy}(\beta=0.2).
-\end{align}
-
+The inverse gamma distribution is used to aid the convergence of the Hamiltonian Monte Carlo algorithm used (discussed later).
 The Cauchy distirbution is equivalent to the Student's t-distribution with one degree of freedom, and was chosen due to its fatter tails than the normal distribution.
-Our likelihood function is therefore
+Our likelihood function for $N$ arms, each of which with $n_\mathrm{arm}$ points is
 
 \begin{equation}
-\mathcal{L} \sim \mathrm{Normal}\left(\exp\left[\theta_\mathrm{arm}\tan\phi_\mathrm{arm} + c_\mathrm{arm}\right], \sigma_r\right).
+  \mathcal{L} = \prod_{\mathrm{arm}=1}^{N}\left(2\pi\sigma_r^2\right)^{-n_\mathrm{arm}/2}
+  \exp\left(\frac{||\widevec{r_\mathrm{arm}}\,-\,\widetilde{r_\mathrm{arm}}||^2}{2\sigma_r^2}\right).
 \end{equation}
@@
 To perform inference, we make use of the No-U-Turn-Sampler (NUTS, \citealt{2011arXiv1111.4246H}), implemented in PYMC3\footnote{\url{https://docs.pymc.io/}}, an open source probabilistic programming framework written in python \citep{pymc3_paper}.
-We scale the radii of deprojected points to have unit variance, to aid the convergence of MC chains.
+To aid the convergence of MC chains, we scale the radii of deprojected points to have unit variance.

### 2020-04-14 sections/results.tex
separates analysis of galaxy pitch angle and arm pitch angle

@@
 \subsection{Constraints on Galaxy Pitch angle}
-Our hierarchical model identifies the pitch angle of individual arms ($\phi_\mathrm{arm}$) with a high degree of certainty (uncertainty less than {1.6\degree} for 95\% of arms, assuming no error on disc inclination and position angle), however the pitch angle of a galaxy as a whole ($\phi_\mathrm{gal}$) is not well constrained.
+Our hierarchical model identifies the pitch angle of individual arms ($\phi_\mathrm{arm}$) with less than {1.6\degree} of uncertainty for 95\% of arms, assuming no error on disc inclination and position angle.
+The pitch angle of a galaxy as a whole ($\phi_\mathrm{gal}$), however, is not well constrained.
 This is primarily a result of only having pitch angles measurements for a small number of arms per galaxy, and reflects the difficulty in providing a single value for the pitch angle of a galaxy containing individual arms with very different pitch angles.
-For galaxies with two arms identified in \textit{Galaxy Builder}, we have a mean uncertainty of {7.8\degree}, which decreases to {6.7\degree} and {5.8\degree} for galaxies with three and four arms respectively.
+For galaxies with two arms identified in \textit{Galaxy Builder}, we have a mean uncertainty of ($\sigma_{\phi_\mathrm{gal}}$) of  {7.8\degree}, which decreases to {6.7\degree} and {5.8\degree} for galaxies with three and four arms respectively.
+This is consistent with the standard error on the mean for a galaxy with $N$ arms,
 
-Our measure of inter-arm variability of pitch angle, $\sigma_\mathrm{gal}$ has a posterior distribution of $11.0\pm 1.0$, similar to the finding of \citet{2014ApJ...790...87D} and emphaising the need for fitting algorithms to not assume all arms have the same pitch angle.
+\begin{equation}
+  \sigma_{\phi_\mathrm{gal}} = \frac{\sigma_\mathrm{gal}}{\sqrt{N}},
+\end{equation}
 
-\subsection{Spiral Winding}
-\label{section:spiral_winding}
-In order to test the possible progenitor distribution of our estimated galaxy pitch angles, we repeatedly perform an Anderson-Darling test over each draw present in the MCMC trace, resulting in a distribution of Anderson-Darling statistics.
-We will refer to this test as the \textit{marginalized Anderson-Darling test}.
-We also make use of the Kolmogorov-Smirnov test in a similar manner for comparison.
-
-We perform the marginalized Anderson-Darling test for a potential source distribution uniform in $\cot\phi$ between the limits present in \citet{2019arXiv190910291P} ($1.00 < \cot\phi < 4.75$, or roughly $11.9^\circ < \phi < 45.0^\circ$).
-The resulting distribution of Anderson-Darling statistics can be seen in Figure \ref{fig:ad-cot-test}.
-We observe that we reject the null hypothesis at the 1\% level for 69\% of the possible realizations of galaxy pitch angle.
-While, with our sample and methodology, we cannot unilaterally reject winding of the kind described by \citet{2019arXiv190910291P}, it suggests this model is not found in our data.
-
-\begin{figure*}
-  \includegraphics[width=17.7cm]{plots/cot_uniform_marginalized_tests.pdf}
-  \caption{The results of a marginalized Anderson-Darling test (top panel), with values corresponding to various confidence intervals shown, and marginalized Kolmogorov-Smirnov test p-values (lower panel).
-Moving rightwards on the x-axis implies greater confidence in rejecting the null hypothesis.}
-  \label{fig:ad-cot-test}
-\end{figure*}
-
-Not seeing uniformity in $\cot\phi$ is not evidence against spiral winding, nor is it evidence for static spiral structures.
-Our results simply suggest that a more comprehensive test is needed for observational data of this kind.
+where $\sigma_\mathrm{gal}$ is our measure of inter-arm variability of pitch angle and has a posterior distribution of $11.0^\circ\pm 1.0^\circ$.
+This variability is similar to the finding of \citet{2014ApJ...790...87D} and emphaises the need for fitting algorithms to not assume all arms have the same pitch angle.
 
@@
 \label{section:morphology_comparision}
+In order to test the possible progenitor distribution of our estimated arm pitch angles, we repeatedly perform an Anderson-Darling test \citep{10.2307/2286009} over each draw present in the MC trace, resulting in a distribution of Anderson-Darling statistics.
+We will refer to this test as the \textit{marginalized Anderson-Darling test}.
+We also make use of the two-sample Anderson-Darling \citep{doi:10.1080/01621459.1987.10478517} and Kolmogorov-Smirnov tests in a similar manner for comparison.
+
 \subsubsection{Pitch angle vs. Bulge size}
-We see no correlation between galaxy pitch angle derived from the hierarchical model and Galaxy Zoo 2's debiased \citep{Willett2013:1308.3496v2} \textit{pbulge}, which has been shown to be a good measure of bulge size.
+We define a bulge prominence from Galaxy Zoo 2 as Equation 3 in \citet{2019MNRAS.487.1808M}:
 
-We separate our sample into ``disc-dominated galaxies'' and ``obvious bulge galaxies'' using the debiased fractions from Galaxy Zoo 2 following \citet{2017MNRAS.469.3363K}, defining \textit{no bulge + just noticeable $>$ obvious + dominant} for the former and the converse for the latter.
-A marginalized two-sample Anderson-Darling test \citep{doi:10.1080/01621459.1987.10478517} does not find any evidence that the samples were drawn from different distributions; we do not reject the null hypothesis at the 1\% level for any samples.
+\begin{equation}
+  B_\mathrm{avg} = 0.2p_\mathrm{just noticeable} + 0.8p_\mathrm{obvious} + 1.0p_\mathrm{dominant}.
+\end{equation}
+
+We see no correlation between galaxy pitch angle derived from the hierarchical model and $B_\mathrm{avg}$.
+The Pearson correlation coefficient between the expectation value of galaxy pitch angle ($E(\phi_\mathrm{gal})$) and $B_\mathrm{avg}$ is -0.09.
+We see a slight correlation between $E(\phi_\mathrm{gal})$ and the fraction of volunteers who said a galaxy did not have a bulge (Pearson correlation coefficient of 0.19).
+
+\comment{Do we need to do an Anderson-Darling test here?}
+
+We separate our sample into ``weaker bulged galaxies'' and ``stronger bulged galaxies'' using , in order to examine whether they could be drawn from significantly different distributions.
+A marginalized two-sample Anderson-Darling test does not find any evidence that the samples were drawn from different distributions; we do not reject the null hypothesis at the 1\% level for any samples.
 The distribution of Anderson-Darling test statistics is shown in the upper panel of Figure \ref{fig:ad-morphology-test}.
@@
 \end{figure*}
+
+\subsection{Spiral Winding}
+\label{section:spiral_winding}
+
+For transient and reccurent spiral arms driven by self-gravity, \citet{2019arXiv190910291P} suggest that spiral patterns form at some maximum pitch angle ($\phi_\mathrm{max}$), continually wind up over time and finally dissapate at some minimum pitch angle ($\phi_\mathrm{min}$).
+They propose that, under a set of very simple assumptions, the evolution of pitch angle would be governed by
+
+\begin{equation}
+  \label{eq:winding}
+  \cot{\phi} = \left[R\frac{\mathrm{d}\Omega_p}{\mathrm{d}R}\right](t - t_0) + \cot{\phi_\mathrm{max}},
+\end{equation}
+
+where $\Omega_p$ is the radially dependant pattern speed of the spiral arm and $t_0$ is the initial time at which it formed.
+
+In QSDW theory, the pattern speed $\Omega_p$ is a constant in R, as spiral arms obey rigid-body rotation.
+If $\Omega_p$ instead varies with radius we would expect $\cot{\phi}$ to be uniformly distributed between $\cot{\phi_\mathrm{max}}$ and $\cot{\phi_\mathrm{min}}$.
+
+In order to test this theory, \citet{2019arXiv190910291P} used a Kolmogorov-Smirnov test to examine the consistency of a sample of observed galaxy pitch angles with one uniform in cot.
+Pitch angles were measured using discrete fourier transformations in one- and two-dimensions, and as such do not account for inter-arm variations.
+They chose limits of $\cot{\phi} \in [1.00, 4.75]$ (roughly $11.9^\circ < \phi < 45.0^\circ$), motivated by examination of their data.
+
+We aim to replicate their work here, using our sample and methods.
+We will make use of the marginalized tests described above, and examine winding on a per-arm basis, as well as a per-galaxy basis.
+Observation of the distribution of arm pitch angles in our sample suggests limits of $15^\circ < \phi < 50.0^\circ$.
+
+\subsubsection{Galaxy Pitch angle}
+
+Testing the uniformity in cot of $\phi_\mathrm{gal}$, between the limits specified above, using a marginalized Anderson-Darling test, results in rejecting the null hypothesis at the 1\% level for 68\% of samples.
+This suggests that a cot uniform model is not a good fit to galaxy pitch angle, but is not conclusive enough for us to unilaterally reject winding governed by Equation \ref{eq:winding}.
+The full distribution of Anderson-Darling statistics can be seen in the upper panel of Figure \ref{fig:ad-cot-test}.
+
+This inconclusive result is perhaps unsurprising; were we to assume that spiral arms are transient and reccurent instabilities, there is little reason for all of the arms to be at precisely the same evolutionary stage at the same time.
+This is supported by the large spread in inter-arm pitch angles.
+
+\subsubsection{Arm Pitch angle}
+If we assume that spirals form and wind independantly in a galaxy and that their evolution over time can be described by Equation \ref{eq:winding}, the distribution of pitch angles of individual arms should be uniform in cot between our limits.
+
+Using the marginalized Anderson-Darling test we cannot reject the null hypothesis at the 10\% level for any of the possible realizations of arm pitch angle.
+The resulting distribution of Anderson-Darling statistics in the lower panel of Figure \ref{fig:ad-cot-test}.
+This result is highly consistent with the model for spiral winding proposed by \citet{2019arXiv190910291P}, and can be seen as evidence that spirals are formed through local disc perturbations, and are primarily governed by local forces.
+
+\begin{figure*}
+  \includegraphics[width=17.7cm]{plots/combined_cot_uniform_marginalized_tests.pdf}
+  \caption{The results of a marginalized Anderson-Darling test for uniformity in $\cot$ for $\phi_\mathrm{gal}$ (top panel) and $\phi_\mathrm{arm}$ (bottom panel), with values corresponding to various confidence intervals shown.
+Moving rightwards on the x-axis implies greater confidence in rejecting the null hypothesis.}
+  \label{fig:ad-cot-test}
+\end{figure*}
+
+This result is highly sensitive to the lower limit of $\phi_\mathrm{gal}$ chosen; decreasing it to 10\degree results in us rejecting the cot-uniform progenitor at greater than the 0.1\% level for every realization of the posterior.
+As we have no information available on the biases present in \textit{Galaxy Builder} spiral arm classification, we should use the higher limits.
+
+\comment{We do not account for observation effects, instead assuming that galaxy builder spiral arms are equally likely to be identified and recovered at all pitch angles.
+We also assume that the galaxy builder is representative of the general spiral population, which is not necessarily the case}

### 2020-04-17 sections/discussion.tex
redoes analysis including all spirals present in galaxy builder. Includes small paragraphs on motivation behind tests in results section

@@
 
-The statistical approach implemented in this paper allows a more thorough examination of pitch angle than simpler methods of logarithmic spiral fitting, and better accounts for inter-arm pitch angle variation than fourier analysis, which assumes all arms in a given mode have the same pitch angle.
+The statistical approach implemented in this paper allows a more thorough examination of pitch angle than length-weighted pitch angle calculation, and better accounts for the large varitations observed in inter-arm pitch angle than fourier analysis, which assumes all arms in a given mode have the same pitch angle.
 
-We do not find a relationship between bar strength and pitch angle, as would have been predicted by Manifold theory, and do not find evidence for the relationship between central mass concentration and pitch angle predicted by the Hubble sequence.
+We do not find a relationship between bar strength and pitch angle, as would have been predicted by Manifold theory, and do not find evidence for the relationship between bulge size and pitch angle predicted by the Hubble sequence.
+\citet{2007ApJ...655...77G} propose that bulge S\'ersic index is correlated with black hole mas, which has in turn been shown to correlate with spiral pitch angle, we do not investigate this here as \textit{Galaxy Builder} could not accurately recover bulge S\'ersic indices.
 
 Our results are consistent with spiral winding of the form described by \citet{2019arXiv190910291P}; providing evidence for transient and recurrent spiral arms, the evolution of which is governed by self-gravity (such as through swing-amplification, \citealt{1965MNRAS.130..125G})) and which wind up over time.
-The assumptions of this model of spiral winding are also highly reductionist, and leaves many unanswered questions: what determines the limits on $\phi$?
+The assumptions of this model of spiral winding are highly reductionist, and it leaves many unanswered questions: what determines the limits on $\phi$?
 Is the spiral arm equally apparent at all pitch angles, or is a selection effect present?
-This result is not evidence against QSDW, as it is possible that our distribtuion of pitch angles is dictated by other factors such as disk shear and central mass concentration.
+This result is also not evidence against QSDW, as it is possible that our distribtuion of pitch angles is dictated by other factors such as disk shear and central mass concentration.
 
+We do not account for observation effects, instead assuming that galaxy builder spiral arms are equally likely to be identified and recovered at all pitch angles within the limits specified above.
+We also assume that the galaxy sample is representative of the general spiral population, which is not necessarily the case.
 As with most analyses, the most impactful improvement it would be possible to make here would be to increase the cleanliness and volume of data analysed; due to time constraints the \textit{Galaxy Builder} galaxies used are not guaranteed to be representative.
-However, we believe that the methodlogy proposed here is a scalable, robust solution to the problems facing investigation of spiral morphology.
+
+The methodlogy proposed here is a robust solution to the problems facing investigation of spiral morphology, namely that of reliably identifying spiral arms, and properly accounting for the spread in pitch angles of arms within a galaxy.

### 2020-04-17 sections/introduction.tex
redoes analysis including all spirals present in galaxy builder. Includes small paragraphs on motivation behind tests in results section

@@
 \citealt{2004ARA&A..42..603K}).
-Studies of spiral morphology have found interesting correlations between spiral morphology and other galactic properties, such as a correlation between spiral tightness and central mass concentration (\citealt{2019ApJ...871..194Y}, \citealt{2015PhDT........14D}, though \citealt{2017MNRAS.472.2263H} found no such relation) and tightness and rotation curve shape (\citealt{2005MNRAS.359.1065S}, with rising rotation curves creating more open spiral structure).
+Studies of spiral morphology have found interesting correlations between spiral morphology and other galactic properties, such as a correlation between spiral tightness and central mass concentration (\citealt{2019ApJ...871..194Y}, though \citealt{2017MNRAS.472.2263H} found no such relation) and tightness and rotation curve shape (\citealt{2005MNRAS.359.1065S}, with rising rotation curves creating more open spiral structure).
 These predictions and observations provide compelling reasons for investigating their underlying rules and dynamics, as doing so is essential for understanding the secular evolution of disc galaxies.
@@
 Many simulations demonstrate that spirals do not maintain a constant pitch angle, and instead wind-up over time due to the differential rotation of the disc \citep{2013ApJ...763...46B}.
-Recent research suggests that spirals are dynamical in nature, and continually dissapate and re-form \citep{2014PASA...31...35D}.
+Recent research suggests that spirals arms are transient in nature, and continually dissapate and re-form \citep{2014PASA...31...35D}.
 These spirals can be maintained through the same mechanisms that drive QSDW spirals (i.e.
 WASER, \citealt{1976ApJ...205..363M}, swing amplification, \citealt{1965MNRAS.130..125G}), but do not require the idealistic disc conditions required for the formation and maintenance of a stationary wave.
-The pitch angles of these transient spiral arms will decrease due to the differential rotation of the disk, with the density of the arm peaking at some critical pitch angle and then dissapating to be reformed.
+The pitch angles of these transient spiral arms will decrease due to the differential rotation of the disk, with the density of the arm peaking at some critical pitch angle, before dissapating to be reformed.
 
@@
 \citet{2019arXiv190910291P} proposes a simple test of spiral arm winding, assuming the cotangent of the pitch angle of a spiral arm evolves linearly with time.
-They found that the pitch angles of their sample of 86 galaxies was consistent with this distribution, evidence against QSDW theory in favour of the dynamic spirals produced in many simulations.
+They found that the distribution of pitch angles of their sample of 86 galaxies was consistent with this prediction, evidence against QSDW theory in favour of the dynamic spirals produced in many simulations.
 
@@
 
-Many other systems contribute to spiral morphology, including potential ties to bulge fraction (\citealt{1975A&A....44..363Y}, \citealt{2013MNRAS.436.1074S}, \citealt{2019MNRAS.487.1808M}) and black hole mass (\citealt{2015PhDT........14D}, \citealt{2017MNRAS.471.2187D}), via a series of correlations.
+Many other systems contribute to spiral morphology, including potential ties to bulge fraction (\citealt{1975A&A....44..363Y}, \citealt{2013MNRAS.436.1074S}, \citealt{2019MNRAS.487.1808M}) and black hole mass (\citealt{2008ApJ...678L..93S}, \citealt{2017MNRAS.471.2187D}, \citealt{2019MS&E..571a2118A}).
 Stronger bulges and more more massive central black holes have both been linked to more tightly wound spiral arms.

### 2020-04-17 sections/method.tex
redoes analysis including all spirals present in galaxy builder. Includes small paragraphs on motivation behind tests in results section

@@
 Clustering of drawn spiral arms and cleaning of points was then performed as detailed in \Lingard.
-We remove any galaxies for which fewer than two spiral arms were identified, as these do not provide our model with any information about inter-arm scatter.
-This results in a hierarchical data structure of 91 galaxies, 211 spiral arms and 215,678 points.
+We remove any galaxies for which no spiral arms were identified, resulting in a hierarchical data structure of 139 galaxies, 261 spiral arms and 239,947 points.
 
@@
 
-We assume that a galaxy has a single value for pitch angle, $\phi_\mathrm{gal}$, and that the pitch angles of spiral arms in this galaxy, $\phi_\mathrm{arm}$, are constants (giving logarithmic spirals) drawn from a normal distribution centred on $\phi_\mathrm{gal}$, with some spread $\sigma_\mathrm{gal}$ common to all galaxies.
+We assume that a galaxy has a single value for pitch angle, $\phi_\mathrm{gal}$, and that the pitch angles of spiral arms in the galaxy, $\phi_\mathrm{arm}$, are constant (giving logarithmic spirals) and drawn from a normal distribution centred on $\phi_\mathrm{gal}$, with some spread $\sigma_\mathrm{gal}$ common to all galaxies.
 We truncate this normal distribution between the physical limits of {0\degree} (a ring) and {90\degree} (a ``spoke''), giving
@@
 
-The inverse gamma distribution is used to aid the convergence of the Hamiltonian Monte Carlo algorithm used (discussed later).
+The inverse gamma distribution is used to aid the convergence of the Hamiltonian Monte Carlo (HMC) algorithm used (discussed later).
 The Cauchy distirbution is equivalent to the Student's t-distribution with one degree of freedom, and was chosen due to its fatter tails than the normal distribution.

### 2020-04-17 sections/results.tex
redoes analysis including all spirals present in galaxy builder. Includes small paragraphs on motivation behind tests in results section

@@
 This is primarily a result of only having pitch angles measurements for a small number of arms per galaxy, and reflects the difficulty in providing a single value for the pitch angle of a galaxy containing individual arms with very different pitch angles.
-For galaxies with two arms identified in \textit{Galaxy Builder}, we have a mean uncertainty of ($\sigma_{\phi_\mathrm{gal}}$) of  {7.8\degree}, which decreases to {6.7\degree} and {5.8\degree} for galaxies with three and four arms respectively.
-This is consistent with the standard error on the mean for a galaxy with $N$ arms,
+For galaxies with two arms identified in \textit{Galaxy Builder}, we have a mean uncertainty of ($\sigma_{\phi_\mathrm{gal}}$) of  {7.9\degree}, which decreases to {6.8\degree} and {6.0\degree} for galaxies with three and four arms respectively.
+This is roughly consistent with the standard error on the mean for a galaxy with $N$ arms,
 
@@
 
-where $\sigma_\mathrm{gal}$ is our measure of inter-arm variability of pitch angle and has a posterior distribution of $11.0^\circ\pm 1.0^\circ$.
+where $\sigma_\mathrm{gal}$ is our measure of inter-arm variability of pitch angle and has a posterior distribution of $11.0^\circ\pm 0.9^\circ$.
 This variability is similar to the finding of \citet{2014ApJ...790...87D} and emphaises the need for fitting algorithms to not assume all arms have the same pitch angle.
+
+\comment{what about for one-armed galaxies? (A: 9.86\degree)}
 
@@
 \label{section:morphology_comparision}
-In order to test the possible progenitor distribution of our estimated arm pitch angles, we repeatedly perform an Anderson-Darling test \citep{10.2307/2286009} over each draw present in the MC trace, resulting in a distribution of Anderson-Darling statistics.
+In order to test the possible progenitor distribution of our estimated arm pitch angles, we repeatedly perform an Anderson-Darling test (\citealt{10.2307/2286009}, implemented in \textsc{Scipy}, \citealt{scipy-paper}) over each draw present in the MC trace, resulting in a distribution of Anderson-Darling statistics.
 We will refer to this test as the \textit{marginalized Anderson-Darling test}.
-We also make use of the two-sample Anderson-Darling \citep{doi:10.1080/01621459.1987.10478517} and Kolmogorov-Smirnov tests in a similar manner for comparison.
+We also make use of the two-sample Anderson-Darling \citep{doi:10.1080/01621459.1987.10478517} test in a similar manner.
 
 \subsubsection{Pitch angle vs. Bulge size}
-We define a bulge prominence from Galaxy Zoo 2 as Equation 3 in \citet{2019MNRAS.487.1808M}:
+
+Morphological classification commonly links bulge size to spiral tightness, and such a link is implied by the Hubble Sequence (\citealt{2005ARA&A..43..581S}, \citealt{2009MNRAS.393.1531G}, \citealt{2013seg..book..155B}).
+Some studies have indeed reported a link between measured spiral galaxy pitch angle and bulge size (i.e.
+\citealt{2017MNRAS.472.2263H}, \citealt{2019ApJ...873...85D}), while others have not found any significant correlation \citep{2019MNRAS.487.1808M}.
+We investigate this relationship here using a measure of bulge prominence from Galaxy Zoo 2, as Equation 3 in \citet{2019MNRAS.487.1808M}:
 
 \begin{equation}
-  B_\mathrm{avg} = 0.2p_\mathrm{just noticeable} + 0.8p_\mathrm{obvious} + 1.0p_\mathrm{dominant}.
+  B_\mathrm{avg} = 0.2\times p_\mathrm{just\ noticeable} + 0.8\times p_\mathrm{obvious} + 1.0\times p_\mathrm{dominant},
 \end{equation}
 
+Where $p_\mathrm{just\ noticeable}$, $p_\mathrm{obvious}$ and $p_\mathrm{dominant}$ are the fractions of classifications indicating the galaxy's bulge was ``just noticeable'', ``obvious'' or ``dominant'' respectively.
+
 We see no correlation between galaxy pitch angle derived from the hierarchical model and $B_\mathrm{avg}$.
-The Pearson correlation coefficient between the expectation value of galaxy pitch angle ($E(\phi_\mathrm{gal})$) and $B_\mathrm{avg}$ is -0.09.
-We see a slight correlation between $E(\phi_\mathrm{gal})$ and the fraction of volunteers who said a galaxy did not have a bulge (Pearson correlation coefficient of 0.19).
+The Pearson correlation coefficient between the expectation value of galaxy pitch angle ($E[\phi_\mathrm{gal}]$) and $B_\mathrm{avg}$ is 0.00.
 
-\comment{Do we need to do an Anderson-Darling test here?}
+We separate our sample into galaxies with weaker bulges ($B_\mathrm{avg} < 0.28$, 83 galaxies) and those with stronger bulges ($B_\mathrm{avg} \ge 0.28$, 54 galaxies), in order to test whether their pitch angles could be drawn from significantly different distributions.
+A marginalized two-sample Anderson-Darling test comparing the distributions of $\phi_\mathrm{gal}$ for the samples does not find evidence that galaxy pitch angles were drawn from different distributions: we reject the null hypothesis at the 1\% level for only 1\% of the samples.
+Similarly comparing arm pitch angles for galaxies in the different samples results in not rejecting the null hypothesis at the 1\% level for any of the samples.
+The distributions of Anderson-Darling test statistic for $\phi_\mathrm{gal}$ and $\phi_\mathrm{arm}$ are shown in the upper panel of Figure \ref{fig:ad-morphology-test} in blue and orange respectively.
 
-We separate our sample into ``weaker bulged galaxies'' and ``stronger bulged galaxies'' using , in order to examine whether they could be drawn from significantly different distributions.
-A marginalized two-sample Anderson-Darling test does not find any evidence that the samples were drawn from different distributions; we do not reject the null hypothesis at the 1\% level for any samples.
-The distribution of Anderson-Darling test statistics is shown in the upper panel of Figure \ref{fig:ad-morphology-test}.
+These results may be caused by the galaxy sample not containing many galaxies with dominant bulges: $B_\mathrm{avg}$ only varied from 0.09 to 0.75 (the allowed maximum being 1.0), with only four galaxies having $B_\mathrm{avg} > 0.5$.
+The split point of 0.28 was also chosen to produce evenly sized comparison samples rather than from some physical motivation.
+However, the lack of any form of correlation strongly suggests that there is no evidence in our data for the link between bulge size and pitch angle observed in other studies.
 
 \subsubsection{Pitch angle vs. Bar Strength}
-We see no correlation between galaxy pitch angle derived from the \textit{hierarchical normal model} and Galaxy Zoo 2's debiased \textit{pbar}, which is widely viewed as a good measure of bar strength, and therefore a measure of the torque applied on the disc gas.
 
-Separating the sample based off of $\mathrm{\textit{pbar}} > 0.5$, and restricting to galaxies with more than 10 classifications for \textit{pbar} (as performed by \citealt{2011MNRAS.411.2026M} and \citealt{2017MNRAS.469.3363K}) and performing a marginalized two-sample Anderson-Darling test does not find that the samples were drawn from different distributions (we reject the null hypothesis at the 1\% level for only 1\% of samples).
-The distribution of Anderson-Darling test statistics is shown in the lower panel of Figure \ref{fig:ad-morphology-test}.
+One of the predictions of Manifold theory is that pitch angle increases with bar strength \citep{2009MNRAS.400.1706A}.
+In order to investigate this relationship in our data, we make use of Galaxy Zoo 2's bar fraction (\textit{pbar}), which is widely viewed as a good measure of bar strength (\citealt{2012MNRAS.423.1485S}, \citealt{2012MNRAS.424.2180M}) and therefore a measure of the torque applied on the disc gas.
 
-We do not account for variation in bulge size in this work, however predictions from Manifold theory should not be affected as bulges do not provide a non-axisymmetric focing.
+We do not observe a correlation between \textit{pbar} and $E[\phi_\mathrm{gal}]$ (Pearson correlation coefficient of -0.05).
+Following \citet{2012MNRAS.424.2180M} and \citet{2012MNRAS.423.1485S}, we separate the sample into galaxies without a bar ($\mathrm{\textit{pbar}} < 0.2$), weakly-barred galaxies ($0.2 \le \mathrm{\textit{pbar}} \le 0.5$) and strongly-barred galaxies ($\mathrm{\textit{pbar}} > 0.5$).
+
+A marginalized k-sample Anderson-Darling tests on the distributions of $\phi_\mathrm{gal}$ and $\phi_\mathrm{arm}$ does not find that pitch angles ($\phi_\mathrm{gal}$ or $\phi_\mathrm{arm}$) of galaxies with different bar strengths were drawn from different distributions; we do not reject the null hypothesis at the 1\% level for any of samples for either the test of $\phi_\mathrm{gal}$ or the test of $\phi_\mathrm{arm}$.
+The distributions of Anderson-Darling test statistic is shown in the lower panel of Figure \ref{fig:ad-morphology-test}.
+
+\comment{this paragraph is quite clunky, can we rephrase it?}
 
@@
   \includegraphics[width=17.7cm]{plots/bulge_bar_test_results.pdf}
-  \caption{The results of marginalized two-sample Anderson-Darling tests examining whether pitch angles for Bulge-dominated and Disc-dominated galaxies are drawn from the same distribution (top panel), and the results of the same test for strongly-barred vs unbarred galaxies (bottom panel).}
+  \caption{The results of marginalized two-sample Anderson-Darling tests examining whether pitch angles ($\phi_\mathrm{gal}$ in blue and $\phi_\mathrm{gal}$ in orange) for galaxies with $B_\mathrm{avg} < 0.28$ and $B_\mathrm{avg} \ge 0.28$ are drawn from the same distribution (top panel), and the results for strongly-barred vs unbarred galaxies (bottom panel).}
   \label{fig:ad-morphology-test}
 \end{figure*}
+
+We do not correct for variation in bulge size in this work, however predictions from Manifold theory should not be affected as bulges do not provide a strong non-axisymmetric focing.
+The fact that we do not find any link between bar strength and pitch angle suggests that the primary mechanism driving the evolution of the spirals in our sample is not Manifold theory.
 
@@
 We aim to replicate their work here, using our sample and methods.
-We will make use of the marginalized tests described above, and examine winding on a per-arm basis, as well as a per-galaxy basis.
+We will make use of the marginalized Anderson-Darling test described above, and examine winding on a per-arm basis, as well as a per-galaxy basis.
 Observation of the distribution of arm pitch angles in our sample suggests limits of $15^\circ < \phi < 50.0^\circ$.
@@
 
-Testing the uniformity in cot of $\phi_\mathrm{gal}$, between the limits specified above, using a marginalized Anderson-Darling test, results in rejecting the null hypothesis at the 1\% level for 68\% of samples.
-This suggests that a cot uniform model is not a good fit to galaxy pitch angle, but is not conclusive enough for us to unilaterally reject winding governed by Equation \ref{eq:winding}.
+Testing the uniformity of $\cot\phi_\mathrm{gal}$ between {15\degree} and {50\degree} using a marginalized Anderson-Darling test results in rejecting the null hypothesis at the 1\% level for just 5\% of samples, with a large spread in observed test values.
 The full distribution of Anderson-Darling statistics can be seen in the upper panel of Figure \ref{fig:ad-cot-test}.
 
-This inconclusive result is perhaps unsurprising; were we to assume that spiral arms are transient and reccurent instabilities, there is little reason for all of the arms to be at precisely the same evolutionary stage at the same time.
-This is supported by the large spread in inter-arm pitch angles.
+This result suggests that we cannot rule out a cot-uniform source distribution for galaxy pitch angle, but the large uncertainty in $\phi_\mathrm{gal}$ makes it difficult to make any conclusive statements.
+This result is also highly sensitive to the lower limit of $\phi$: decreasing it to {10\degree} results in us rejecting the cot-uniform model at greater than the 0.1\% level for 96\% of the posterior samples.
+As we have no information available on the biases present in \textit{Galaxy Builder} spiral arm classification, we choose to keep the less strict limit of {15\degree}.
 
 \subsubsection{Arm Pitch angle}
-If we assume that spirals form and wind independantly in a galaxy and that their evolution over time can be described by Equation \ref{eq:winding}, the distribution of pitch angles of individual arms should be uniform in cot between our limits.
+The inconclusive result for $\phi_\mathrm{gal}$ is perhaps unsurprising: were we to assume that spiral arms are transient and reccurent instabilities, there is little reason for all of the arms to be at precisely the same evolutionary stage at the same time.
+This is supported by the large observed spread in inter-arm pitch angles.
 
-Using the marginalized Anderson-Darling test we cannot reject the null hypothesis at the 10\% level for any of the possible realizations of arm pitch angle.
-The resulting distribution of Anderson-Darling statistics in the lower panel of Figure \ref{fig:ad-cot-test}.
+If we assume instead that spirals form and wind independantly in a galaxy, and that their evolution over time can be described by Equation \ref{eq:winding}, the distribution of pitch angles of individual arms should be uniform in cot between our limits, rather than that of the galaxy's pitch angle as a whole.
+
+Using the marginalized Anderson-Darling test we cannot reject the null hypothesis at even the 5\% level for any of the possible realizations of arm pitch angle.
+The resulting distribution of Anderson-Darling statistics is shown in in the lower panel of Figure \ref{fig:ad-cot-test}.
 This result is highly consistent with the model for spiral winding proposed by \citet{2019arXiv190910291P}, and can be seen as evidence that spirals are formed through local disc perturbations, and are primarily governed by local forces.
@@
 \end{figure*}
-
-This result is highly sensitive to the lower limit of $\phi_\mathrm{gal}$ chosen; decreasing it to 10\degree results in us rejecting the cot-uniform progenitor at greater than the 0.1\% level for every realization of the posterior.
-As we have no information available on the biases present in \textit{Galaxy Builder} spiral arm classification, we should use the higher limits.
-
-\comment{We do not account for observation effects, instead assuming that galaxy builder spiral arms are equally likely to be identified and recovered at all pitch angles.
-We also assume that the galaxy builder is representative of the general spiral population, which is not necessarily the case}

### 2020-04-17 sections/discussion.tex
cleanup

@@
 This paper presents a new Bayesian approach to estimate galaxy pitch angle, making use of citizen science results to measure spiral arms through photometric modelling.
-We introduce an adaptation of the Anderson-Darling test to incorporate full Bayesian posterior probabilities and utilize this test to investigate theories governing spiral formation and evolution.
+We introduce an adaptation of the Anderson-Darling test, which we name the \textit{marginalized Anderson-Darling test}, to incorporate full Bayesian posterior probabilities and utilize this test to investigate theories governing spiral formation and evolution.
 
@@
 
-We do not find a relationship between bar strength and pitch angle, as would have been predicted by Manifold theory, and do not find evidence for the relationship between bulge size and pitch angle predicted by the Hubble sequence.
-\citet{2007ApJ...655...77G} propose that bulge S\'ersic index is correlated with black hole mas, which has in turn been shown to correlate with spiral pitch angle, we do not investigate this here as \textit{Galaxy Builder} could not accurately recover bulge S\'ersic indices.
+We do not see a relationship between bar strength and pitch angle in our data, as would have been predicted by Manifold theory, and do not find evidence for the relationship between bulge size and pitch angle predicted by the Hubble sequence.
+\citet{2007ApJ...655...77G} propose that bulge S\'ersic index is correlated with black hole mass, which has in turn been shown to correlate with spiral pitch angle, we do not investigate this here as \textit{Galaxy Builder} could not accurately recover bulge S\'ersic indices.
 
-Our results are consistent with spiral winding of the form described by \citet{2019arXiv190910291P}; providing evidence for transient and recurrent spiral arms, the evolution of which is governed by self-gravity (such as through swing-amplification, \citealt{1965MNRAS.130..125G})) and which wind up over time.
-The assumptions of this model of spiral winding are highly reductionist, and it leaves many unanswered questions: what determines the limits on $\phi$?
+Our results are consistent with spiral winding of the form described by \citet{2019arXiv190910291P}, in which spiral arms are transient and reccurent, evolve through mechanisms such as swing-amplification \citep{1965MNRAS.130..125G} and which wind up over time.
+However, the assumptions of this model of spiral winding are highly simplistic, and it leaves many unanswered questions: what determines the limits on $\phi$?
 Is the spiral arm equally apparent at all pitch angles, or is a selection effect present?

### 2020-04-17 sections/results.tex
cleanup

@@
 
-\comment{what about for one-armed galaxies? (A: 9.86\degree)}
+% for one-armed galaxies = {9.86\degree}
 
@@
 The split point of 0.28 was also chosen to produce evenly sized comparison samples rather than from some physical motivation.
-However, the lack of any form of correlation strongly suggests that there is no evidence in our data for the link between bulge size and pitch angle observed in other studies.
+However, the lack of any form of correlation implies that there is no evidence in our data for the link between bulge size and pitch angle predicted by the Hubble sequence and observed in other studies.
 
@@
 We do not observe a correlation between \textit{pbar} and $E[\phi_\mathrm{gal}]$ (Pearson correlation coefficient of -0.05).
-Following \citet{2012MNRAS.424.2180M} and \citet{2012MNRAS.423.1485S}, we separate the sample into galaxies without a bar ($\mathrm{\textit{pbar}} < 0.2$), weakly-barred galaxies ($0.2 \le \mathrm{\textit{pbar}} \le 0.5$) and strongly-barred galaxies ($\mathrm{\textit{pbar}} > 0.5$).
-
-A marginalized k-sample Anderson-Darling tests on the distributions of $\phi_\mathrm{gal}$ and $\phi_\mathrm{arm}$ does not find that pitch angles ($\phi_\mathrm{gal}$ or $\phi_\mathrm{arm}$) of galaxies with different bar strengths were drawn from different distributions; we do not reject the null hypothesis at the 1\% level for any of samples for either the test of $\phi_\mathrm{gal}$ or the test of $\phi_\mathrm{arm}$.
+Following \citet{2012MNRAS.424.2180M} and \citet{2012MNRAS.423.1485S}, we separate the sample into galaxies without a bar ($\mathrm{\textit{pbar}} < 0.2$), with a weak bar ($0.2 \le \mathrm{\textit{pbar}} \le 0.5$) and with a strong bar ($\mathrm{\textit{pbar}} > 0.5$).
+Performing marginalized three-sample Anderson-Darling tests does not find that pitch angles ($\phi_\mathrm{gal}$ or $\phi_\mathrm{arm}$) of galaxies with different bar strengths were drawn from different distributions; we do not reject the null hypothesis at the 1\% level for any samples for the test of $\phi_\mathrm{gal}$, and at the 10\% level for the test of $\phi_\mathrm{arm}$.
 The distributions of Anderson-Darling test statistic is shown in the lower panel of Figure \ref{fig:ad-morphology-test}.
-
-\comment{this paragraph is quite clunky, can we rephrase it?}
 
@@
   \includegraphics[width=17.7cm]{plots/bulge_bar_test_results.pdf}
-  \caption{The results of marginalized two-sample Anderson-Darling tests examining whether pitch angles ($\phi_\mathrm{gal}$ in blue and $\phi_\mathrm{gal}$ in orange) for galaxies with $B_\mathrm{avg} < 0.28$ and $B_\mathrm{avg} \ge 0.28$ are drawn from the same distribution (top panel), and the results for strongly-barred vs unbarred galaxies (bottom panel).}
+  \caption{The results of marginalized two-sample Anderson-Darling tests examining whether pitch angles ($\phi_\mathrm{gal}$ in blue and $\phi_\mathrm{gal}$ in orange) for galaxies with $B_\mathrm{avg} < 0.28$ and $B_\mathrm{avg} \ge 0.28$ are drawn from the same distribution (top panel), and the results of marginalized three-sample Anderson-Darling tests for galaxies with no bar ($\mathrm{\textit{pbar}} < 0.2$), a weak bar ($0.2 \le \mathrm{\textit{pbar}} \le 0.5$) and a strong bar ($\mathrm{\textit{pbar}} > 0.5$) (bottom panel).
+Confidence intervals are shown, with moving rightwards indicating more confidence in rejecting the null hypothesis.}
   \label{fig:ad-morphology-test}
@@
 The fact that we do not find any link between bar strength and pitch angle suggests that the primary mechanism driving the evolution of the spirals in our sample is not Manifold theory.
+\comment{This is possibly too strong a statement?}
 

### 2020-04-22 sections/discussion.tex
Adds figures suggested by KLM. Makes changes to figures suggested by CMK. Changes some content as suggested by KLM.

@@
 The statistical approach implemented in this paper allows a more thorough examination of pitch angle than length-weighted pitch angle calculation, and better accounts for the large varitations observed in inter-arm pitch angle than fourier analysis, which assumes all arms in a given mode have the same pitch angle.
+In this work we find that the mean inter-arm difference in pitch angle is $11.0^\circ\pm 0.9^\circ$.
 
-We do not see a relationship between bar strength and pitch angle in our data, as would have been predicted by Manifold theory, and do not find evidence for the relationship between bulge size and pitch angle predicted by the Hubble sequence.
-\citet{2007ApJ...655...77G} propose that bulge S\'ersic index is correlated with black hole mass, which has in turn been shown to correlate with spiral pitch angle, we do not investigate this here as \textit{Galaxy Builder} could not accurately recover bulge S\'ersic indices.
+There is no evidence in our data for the link between bulge size and pitch angle predicted by the Hubble sequence and observed in other studies.
+
+We do not find any link between bar strength and pitch angle suggests that the primary mechanism driving the evolution of the spirals in our sample is not Manifold theory.
 
@@
 Is the spiral arm equally apparent at all pitch angles, or is a selection effect present?
-This result is also not evidence against QSDW, as it is possible that our distribtuion of pitch angles is dictated by other factors such as disk shear and central mass concentration.
+This result is also not evidence against QSDW, as it is possible that our distribtuion of pitch angles is dictated by other factors such as disk shear.
 
-We do not account for observation effects, instead assuming that galaxy builder spiral arms are equally likely to be identified and recovered at all pitch angles within the limits specified above.
-We also assume that the galaxy sample is representative of the general spiral population, which is not necessarily the case.
-As with most analyses, the most impactful improvement it would be possible to make here would be to increase the cleanliness and volume of data analysed; due to time constraints the \textit{Galaxy Builder} galaxies used are not guaranteed to be representative.
+In this work we assume that spiral arms are equally likely to be identified and recovered at all pitch angles.
+This is not an unfair assumption given the amount of human effort that went into obtaining spiral arms (more so than any other pitch angle measurement method, with each galaxy receiving 30 human classifications).
+The galaxy sample used is not guaranteed to be representative of the general spiral population, but is comparable in size to those used in other similar studies (\citealt{2013MNRAS.436.1074S}, \citealt{2019ApJ...871..194Y}, \citealt{2019arXiv190910291P}).
 
 The methodlogy proposed here is a robust solution to the problems facing investigation of spiral morphology, namely that of reliably identifying spiral arms, and properly accounting for the spread in pitch angles of arms within a galaxy.
+As with most analyses, the most impactful improvement it would be possible to make here would be to increase the cleanliness and volume of data analysed.
+
+The processes governing the formation and evolution of spiral arms are immensely complicated, but the prevalence of spiral galaxies in the Universe, and the spiral nature of our own Milky Way, makes investigating their dynamics of fundamental importance to the scientific aims of understanding, predicting and explaining the nature of the cosmos.

### 2020-04-22 sections/introduction.tex
Adds figures suggested by KLM. Makes changes to figures suggested by CMK. Changes some content as suggested by KLM.

@@
 Spirals are as diverse as the theories proposed to govern their evolution; from the quintessential pair of well-defined arcs of the grand design spiral, to the fragmented arm segments of the flocculent spiral, to the disjointed multi-armed spiral.
-These variatons on structure account for $18\%$, $50\%$ and $32\%$ of the population respectively (\citealt{2011ApJ...737...32E}, \citealt{2015yCat..22170032B}).
+These variatons on structure account for $18\%$, $50\%$ and $32\%$ of the population respectively (\citealt{2011ApJ...737...32E}, \citealt{2015yCat..22170032B}, examples of each type are shown in Figure \ref{fig:spiral-galaxy-types}).
 The Hubble classification scheme \citep{1926ApJ....64..321H}, and its revisions and expansions (\citealt{1961hag..book.....S}, \citealt{1991rc3..book.....D}) contain detailed variations of different types of spiral galaxy, divided by the presence of a bar and ordered by how obvious spiral arm patterns are, how tightly they are wound and the prominence of a central bulge.
 
-Arms of spiral galaxies are the source of the vast majority of star formation in the Universe, and spirals rearrange disc gas and can lead to the formation of disc-like bulges (e.g.
+\begin{figure*}
+  \includegraphics[width=13cm]{plots/galaxy_types.jpg}
+  \caption{Examples of the different types of spiral galaxy present in the sky.
+The left column shows grand design spirals, the middle shows many-armed spirals and the right shows flocculent spirals.
+Images were taken with the Sloan Digital Sky Survey Telescope.}
+  \label{fig:spiral-galaxy-types}
+\end{figure*}
+
+A majority of the population of young stars in a galaxy are located in its spiral arms \citep{2011EAS....51...19E}, and it is possible that their passage triggers star formation \citep{2013A&A...560A..59C}, though their main role may instead be to sweep up material, promoting the growth of Giant Molecular Clouds.
+\citep{2014IAUS..298..221D}.
+This rearrangement of disc gas can lead to the formation of disc-like bulges (e.g.
 \citealt{2004ARA&A..42..603K}).

### 2020-04-22 sections/method.tex
cleanup
Adds figures suggested by KLM. Makes changes to figures suggested by CMK. Changes some content as suggested by KLM.

@@
 
+{\comment{KLM: Need to talk about individual arms}}
+
 \subsection{Bayesian modelling of spiral arms in \textit{Galaxy Builder}}
 
-We assume that a galaxy has a single value for pitch angle, $\phi_\mathrm{gal}$, and that the pitch angles of spiral arms in the galaxy, $\phi_\mathrm{arm}$, are constant (giving logarithmic spirals) and drawn from a normal distribution centred on $\phi_\mathrm{gal}$, with some spread $\sigma_\mathrm{gal}$ common to all galaxies.
+We assume that a galaxy has a single value for pitch angle, $\phi_\mathrm{gal}$, and that the pitch angles of spiral arms in the galaxy, $\phi_\mathrm{arm}$, are constant with radius (giving logarithmic spirals) and drawn from a normal distribution centred on $\phi_\mathrm{gal}$, with some spread $\sigma_\mathrm{gal}$ common to all galaxies.
 We truncate this normal distribution between the physical limits of {0\degree} (a ring) and {90\degree} (a ``spoke''), giving
@@
 
+We assume that the radial error is Gaussian for simplicity of analysis, Shapiro-Wilk tests on the residuals of the logarithmic spirals fit in {\Lingard} suggest that this is not a good assumption, and a more robust likelihood (such as the Student's t-distirbution) would possibly more appropriate.
+
 To perform inference, we make use of the No-U-Turn-Sampler (NUTS, \citealt{2011arXiv1111.4246H}), implemented in PYMC3\footnote{\url{https://docs.pymc.io/}}, an open source probabilistic programming framework written in python \citep{pymc3_paper}.

### 2020-04-22 sections/results.tex
Adds figures suggested by KLM. Makes changes to figures suggested by CMK. Changes some content as suggested by KLM.

@@
 This variability is similar to the finding of \citet{2014ApJ...790...87D} and emphaises the need for fitting algorithms to not assume all arms have the same pitch angle.
+The spread of arm pitch angle from the mean galaxy pitch angle can be seen in Figure \ref{fig:arm-pa-spread}, with points colour-coded by the number of arms measured for a galaxy.
+We see a slight drop in $E[\phi_\mathrm{gal}]$ compared to $E[\phi_\mathrm{arm}]$ at small galaxy pitch angles, due to the truncation of $\phi_\mathrm{gal}$ at {0\degree}.
+
+\begin{figure*}
+  \includegraphics[width=17.7cm]{plots/arm_pa_spread.pdf}
+  \caption{Scatter plot showing how arm pitch angle compares to galaxy pitch angle for galaxies with different pitch angles and number of arms.
+The top panel shows a Gaussian KDE for $E[\phi_\mathrm{gal}]$, and the right panel shows a Gaussian KDE for $E[\phi_\mathrm{arm} - \phi_\mathrm{gal}]$.}
+  \label{fig:arm-pa-spread}
+\end{figure*}
 
@@
 \label{section:morphology_comparision}
+
 In order to test the possible progenitor distribution of our estimated arm pitch angles, we repeatedly perform an Anderson-Darling test (\citealt{10.2307/2286009}, implemented in \textsc{Scipy}, \citealt{scipy-paper}) over each draw present in the MC trace, resulting in a distribution of Anderson-Darling statistics.
@@
 
-These results may be caused by the galaxy sample not containing many galaxies with dominant bulges: $B_\mathrm{avg}$ only varied from 0.09 to 0.75 (the allowed maximum being 1.0), with only four galaxies having $B_\mathrm{avg} > 0.5$.
+One limitation of this result is that our sample does not contain many galaxies with dominant bulges: $B_\mathrm{avg}$ only varied from 0.09 to 0.75 (the allowed maximum being 1.0), with only four galaxies having $B_\mathrm{avg} > 0.5$.
 The split point of 0.28 was also chosen to produce evenly sized comparison samples rather than from some physical motivation.
-However, the lack of any form of correlation implies that there is no evidence in our data for the link between bulge size and pitch angle predicted by the Hubble sequence and observed in other studies.
+However, the lack of any form of correlation implies that \textbf{there is no evidence in our data for the link between bulge size and pitch angle predicted by the Hubble sequence and observed in other studies}.
 
@@
 One of the predictions of Manifold theory is that pitch angle increases with bar strength \citep{2009MNRAS.400.1706A}.
-In order to investigate this relationship in our data, we make use of Galaxy Zoo 2's bar fraction (\textit{pbar}), which is widely viewed as a good measure of bar strength (\citealt{2012MNRAS.423.1485S}, \citealt{2012MNRAS.424.2180M}) and therefore a measure of the torque applied on the disc gas.
+In order to investigate this relationship in our data, we make use of Galaxy Zoo 2's bar fraction (\textit{pbar}), which has been demonstrated to be a good measure of bar strength (\citealt{2012MNRAS.423.1485S}, \citealt{2012MNRAS.424.2180M}) and therefore a measure of the torque applied on the disc gas.
 
@@
   \caption{The results of marginalized two-sample Anderson-Darling tests examining whether pitch angles ($\phi_\mathrm{gal}$ in blue and $\phi_\mathrm{gal}$ in orange) for galaxies with $B_\mathrm{avg} < 0.28$ and $B_\mathrm{avg} \ge 0.28$ are drawn from the same distribution (top panel), and the results of marginalized three-sample Anderson-Darling tests for galaxies with no bar ($\mathrm{\textit{pbar}} < 0.2$), a weak bar ($0.2 \le \mathrm{\textit{pbar}} \le 0.5$) and a strong bar ($\mathrm{\textit{pbar}} > 0.5$) (bottom panel).
-Confidence intervals are shown, with moving rightwards indicating more confidence in rejecting the null hypothesis.}
+Confidence intervals are shown, with moving rightwards indicating more confidence in rejecting the null hypothesis that the compared values were drawn from the same parent distribution.}
   \label{fig:ad-morphology-test}
@@
 We do not correct for variation in bulge size in this work, however predictions from Manifold theory should not be affected as bulges do not provide a strong non-axisymmetric focing.
-The fact that we do not find any link between bar strength and pitch angle suggests that the primary mechanism driving the evolution of the spirals in our sample is not Manifold theory.
-\comment{This is possibly too strong a statement?}
+The fact that we do not find any link between bar strength and pitch angle suggests that \textbf{the primary mechanism driving the evolution of the spirals in our sample is not Manifold theory}.
 
@@
 The full distribution of Anderson-Darling statistics can be seen in the upper panel of Figure \ref{fig:ad-cot-test}.
+The large spread in results is caused by the large uncertainties in $\phi_\mathrm{gal}$.
 
@@
 The resulting distribution of Anderson-Darling statistics is shown in in the lower panel of Figure \ref{fig:ad-cot-test}.
-This result is highly consistent with the model for spiral winding proposed by \citet{2019arXiv190910291P}, and can be seen as evidence that spirals are formed through local disc perturbations, and are primarily governed by local forces.
+\textbf{This result is highly consistent with the model for spiral winding proposed by \citet{2019arXiv190910291P}}, and can be seen as evidence that spirals are formed through local disc perturbations, and are primarily governed by local forces.
 
@@
   \caption{The results of a marginalized Anderson-Darling test for uniformity in $\cot$ for $\phi_\mathrm{gal}$ (top panel) and $\phi_\mathrm{arm}$ (bottom panel), with values corresponding to various confidence intervals shown.
-Moving rightwards on the x-axis implies greater confidence in rejecting the null hypothesis.}
+Moving rightwards on the x-axis implies greater confidence in rejecting the null hypothesis that the sample was drawn from a distribution uniform in $\cot$ between $15^\circ < \phi < 50.0^\circ$.}
   \label{fig:ad-cot-test}

### 2020-05-12 sections/abstract.tex
final changes before being sent to co-authors

% !TEX root = ../main.tex
Spiral structure is endemic in the Universe, and the pitch angle of arms in spiral galaxies is an important observable in efforts to discriminate between different mechanisms of spiral arm evolution. In this paper we present a hierarchical Bayesian approach to galaxy pitch angle determination, using spiral arm data obtained through the \textit{Galaxy Builder} citizen science project {\Lingard}. This approach better deals with the large variations in pitch angle between arms in a galaxy than existing fourier-based methods, and obtains full posterior distributions on parameters. We utilise fit pitch angles to examine the links to morphology reported in other work, finding no link in our data, and examine a simple model of spiral winding, with which our results are consistent.

### 2020-05-12 sections/discussion.tex
final changes before being sent to co-authors

@@
 
-The statistical approach implemented in this paper allows a more thorough examination of pitch angle than length-weighted pitch angle calculation, and better accounts for the large varitations observed in inter-arm pitch angle than fourier analysis, which assumes all arms in a given mode have the same pitch angle.
+The hierarchical Bayesian approach implemented in this paper allows a more thorough examination of pitch angle than length-weighted pitch angle calculation; obtaining posterior distributions of measured parameters.
+It better accounts for the large varitations observed in inter-arm pitch angle than fourier analysis, which assumes all arms in a given mode have the same pitch angle.
 In this work we find that the mean inter-arm difference in pitch angle is $11.0^\circ\pm 0.9^\circ$.

### 2020-05-12 sections/method.tex
final changes before being sent to co-authors

@@
 
-{\comment{KLM: Need to talk about individual arms}}
-
 \subsection{Bayesian modelling of spiral arms in \textit{Galaxy Builder}}
 
-We assume that a galaxy has a single value for pitch angle, $\phi_\mathrm{gal}$, and that the pitch angles of spiral arms in the galaxy, $\phi_\mathrm{arm}$, are constant with radius (giving logarithmic spirals) and drawn from a normal distribution centred on $\phi_\mathrm{gal}$, with some spread $\sigma_\mathrm{gal}$ common to all galaxies.
+In this section we lay out our bayesian hierarchical model for galaxy pitch angle.
+We fit directly to clustered, cleaned points from polylines drawn in \textit{Galaxy Builder}, deprojected and unwrapped to polar coordinates.
+We fit a logarithmic spiral to each clustered arm, with the pitch angles of multiple arms in a single galaxy being drawn from a single parent distribution.
+
+We assume that a galaxy has some value for pitch angle, $\phi_\mathrm{gal}$, and that the pitch angles of spiral arms in that galaxy, $\phi_\mathrm{arm}$, are constant with radius (giving logarithmic spirals) and drawn from a normal distribution centred on $\phi_\mathrm{gal}$, with some spread $\sigma_\mathrm{gal}$ common to all galaxies.
 We truncate this normal distribution between the physical limits of {0\degree} (a ring) and {90\degree} (a ``spoke''), giving

### 2020-05-21 sections/results.tex
Adds paper with measure of variability of arm pitch angles

@@
 where $\sigma_\mathrm{gal}$ is our measure of inter-arm variability of pitch angle and has a posterior distribution of $11.0^\circ\pm 0.9^\circ$.
-This variability is similar to the finding of \citet{2014ApJ...790...87D} and emphaises the need for fitting algorithms to not assume all arms have the same pitch angle.
+This variability is similar to that found by \citet{1981AJ.....86.1847K} and \citet{2014ApJ...790...87D}, and emphaises the need for fitting algorithms to not assume all arms have the same pitch angle.
 The spread of arm pitch angle from the mean galaxy pitch angle can be seen in Figure \ref{fig:arm-pa-spread}, with points colour-coded by the number of arms measured for a galaxy.

### 2020-06-02 sections/abstract.tex
Adds changes suggested by Chris

@@
 In this paper we present a hierarchical Bayesian approach to galaxy pitch angle determination, using spiral arm data obtained through the \textit{Galaxy Builder} citizen science project {\Lingard}.
-This approach better deals with the large variations in pitch angle between arms in a galaxy than existing fourier-based methods, and obtains full posterior distributions on parameters.
+This approach better deals with the large variations in pitch angle between arms in a galaxy than existing methods, and obtains full posterior distributions on parameters.
 We utilise fit pitch angles to examine the links to morphology reported in other work, finding no link in our data, and examine a simple model of spiral winding, with which our results are consistent.

### 2020-06-02 sections/discussion.tex
Adds changes suggested by Chris

@@
 
-The methodlogy proposed here is a robust solution to the problems facing investigation of spiral morphology, namely that of reliably identifying spiral arms, and properly accounting for the spread in pitch angles of arms within a galaxy.
+The methodology proposed here is a robust solution to the problems facing investigation of spiral morphology, namely that of reliably identifying spiral arms, and properly accounting for the spread in pitch angles of arms within a galaxy.
 As with most analyses, the most impactful improvement it would be possible to make here would be to increase the cleanliness and volume of data analysed.
+A larger sample would make possible further comparisons, such as splitting galaxyies into spiral type (grand design / many-armed / flocculent) and examining the differences between populations.
 

### 2020-06-02 sections/introduction.tex
Adds changes suggested by Chris

@@
 \citep{2014IAUS..298..221D}.
-This rearrangement of disc gas can lead to the formation of disc-like bulges (e.g.
+This rearrangement of disc gas can lead to the formation of disc-like bulges (commonly called ``pseudobulges''; e.g.
 \citealt{2004ARA&A..42..603K}).
@@
 Flocculent spirals are thought to be formed through swing amplification (shearing of small gravitational instabilities in the disc), and be transient and reccurent in nature \citep{1966ApJ...146..810J}.
-It is recognised that methods of spiral formation are not mutually exclusive.
 
 One of the fundamental assumptions of early work on spiral formation mechanisms (primarily QSDW) was that the disc of a galaxy, if unstable to spiral perturbations, would create a stable, static wave which would exist unchanging for many rotational periods \citep{1964ApJ...140..646L}.
-The motivation for static waves with small numbers of arms (with a preference for $m=2$) was primarily observational; most galaxies show spiral structure, suggesting that spirals exist for a long time or are continually rebuilt.
+The motivation for static waves with small numbers of arms (with a preference for $m=2$) was primarily observational; most galaxies show spiral structure with low spiral arm numbers, suggesting that spirals exist for a long time or are continually rebuilt.
 
@@
 In this dynamic picture of spiral arms, pitch angle monotonically decreases from a spiral arm's formation to its dissapation.
-\citet{2019arXiv190910291P} proposes a simple test of spiral arm winding, assuming the cotangent of the pitch angle of a spiral arm evolves linearly with time.
+\citet{2019arXiv190910291P} propose a simple test of spiral arm winding, assuming the cotangent of the pitch angle of a spiral arm evolves linearly with time.
 They found that the distribution of pitch angles of their sample of 86 galaxies was consistent with this prediction, evidence against QSDW theory in favour of the dynamic spirals produced in many simulations.
 
-Spiral evolution also appears to be influenced by the presence and strength of a bar; in barred grand-design spirals the arms often appear to start from the ends of the bar.
-Simulations of gas in barred galaxies often demonstrate that bars can drive long-term spiral evolution \citep{2008A&A...489..115R}, or boost transient spiral structure \citep{2012MNRAS.426..167G}.
+We aim to test this idea of spiral winding using data from the \textit{Galaxy Builder} citizen science project for the 196 spiral galaxies present in \Lingard.
+We make use of Bayesian hierarchical modelling to measure galaxy pitch angle from the spiral arm clusters producted by \textit{Galaxy Builder}.
+This methodology allows us to quantify the differences in pitch angles between arms in a single galaxy, as well as investigate the distribution of pitch angles in the galaxy population and investigate relationships between pitch angle and other galaxy properties such as stellar mass or morphology.
+
+Using Galaxy Zoo 2 data \citep{Willett2013:1308.3496v2} allows the separation of galaxies by the presence and strength of bars and bulges; simulations of gas in barred galaxies often demonstrate that bars can drive long-term spiral evolution \citep{2008A&A...489..115R}, or boost transient spiral structure \citep{2012MNRAS.426..167G}.
 Manifold theory is one attempt to determine the orbits of stars in bar-driven spiral arms: it proposes that stars in the vicinity of the unstable Lagrangian points at either end of the bar tend to escape along predictable orbits, governed by invariant manifolds.
@@
 
-Many other systems contribute to spiral morphology, including potential ties to bulge fraction (\citealt{1975A&A....44..363Y}, \citealt{2013MNRAS.436.1074S}, \citealt{2019MNRAS.487.1808M}) and black hole mass (\citealt{2008ApJ...678L..93S}, \citealt{2017MNRAS.471.2187D}, \citealt{2019MS&E..571a2118A}).
-Stronger bulges and more more massive central black holes have both been linked to more tightly wound spiral arms.
-
-This paper makes use of the classification data and fitted photometric models obtained through the \textit{Galaxy Builder} citizen science project for the 196 spiral galaxies present in \Lingard.
-In this paper we focus on the use of measured spiral tightness (quantified using pitch angle, \citealt{1987gady.book.....B}) as a probe into the dynamical mechanisms governing a spiral galaxy's evolution.
-We make use of Bayesian hierarchical modelling to measure galaxy pitch angle from the spiral arm clusters producted by \textit{Galaxy Builder}.
+Many other systems correlate with spiral morphology, including potential ties to bulge fraction (\citealt{1975A&A....44..363Y}, \citealt{2013MNRAS.436.1074S}, \citealt{2019MNRAS.487.1808M}) and black hole mass (\citealt{2008ApJ...678L..93S}, \citealt{2017MNRAS.471.2187D}, \citealt{2019MS&E..571a2118A}).
+Stronger bulges and more more massive central black holes have both correlated with more tightly wound spiral arms.
 

### 2020-06-02 sections/method.tex
Adds changes suggested by Chris

@@
 One potentially underused method of obtaining measurements of spirals is through photometric fitting of spiral structure, as possible using tools such as \textsc{GALFIT} \citep{2010AJ....139.2097P} and \textit{Galaxy Builder} (\Lingard).
-These methods attempt to localize light from an image of a galaxy into distinct subcomponents, such as a galaxy disc, bulge, bar and spiral arms, generally finding the optimum solution using computational optimization.
+These methods attempt to separate light from an image of a galaxy into distinct subcomponents, such as a galaxy disc, bulge, bar and spiral arms, generally finding the optimum solution using computational optimization.
 This optimization process, however, is often not robust for complex, many-component models and requires significant supervision to converge to a physically meaningful result \citep{Gao2017:1709.00746v1}.
@@
 Two-dimensional fourier methods generally decompose a deprojected image of a galaxy into a superpositions of logarithmic spirals between inner and outer annuli \citep{2012ApJS..199...33D}, and reports the pitch angle with the highest amplitude as the galaxy's pitch angle.
-\citet{2020MNRAS.493.3854H} combined fourier analysis of spiral galaxies with human tracing of spiral arms, to great effect.
+\citet{2020MNRAS.493.3854H} combined fourier analysis of spiral galaxies with visual tracing of spiral arms, to great effect.
 It is uncertain how variation between pitch angles of individual arms impacts this measurement.
@@
 
-Some galaxies in \Lingard were shown to volunteers again in a repeat \textit{validation subset} in order to create a second aggregate model used to test internal consistency.
+Some galaxies in \Lingard were shown to volunteers a second time in a repeat \textit{validation subset} in order to create a second aggregate model used to test internal consistency.
 We combine the classifications of galaxies in this \textit{validation subset} with the original classifications.
@@
 We remove any galaxies for which no spiral arms were identified, resulting in a hierarchical data structure of 139 galaxies, 261 spiral arms and 239,947 points.
+
+As we wish to utilize the logarithmic spiral's desirable properties of a constant pitch angle and small number of fittable parameters, we make use of it here without an explicit comparison to other models; a simple visual inspection of the fitted logarithmic spirals suggests that it is an appropriate model.
+A comparison of a logarithmic spiral profile to other spiral forms (i.e.
+Archimedian or polynomial) is another important piece of work, outside of the scope of this research, as it has been reported that galaxy arms do not have constant pitch angles (\citealt{1981AJ.....86.1847K}; \citealt{2009MNRAS.397..164R}).
 
@@
 
-Furthermore, we assume that the observed points in a \textit{Galaxy Builder} spiral arm, once deprojected, follow a logarithmic spiral with gaussian radial error $\sigma_r$,
+The choice to assume all galaxies show the same inter-arm variation in pitch angle (represented by a common value of $\sigma_\mathrm{gal}$ across all galaxies) was motivated by our small sample size and the low number of arms measured per galaxy.
+
+We assume that the observed points in a \textit{Galaxy Builder} spiral arm, once deprojected, follow a logarithmic spiral with gaussian radial error $\sigma_r$,
 

### 2020-06-02 sections/results.tex
Adds changes suggested by Chris

@@
 One of the predictions of Manifold theory is that pitch angle increases with bar strength \citep{2009MNRAS.400.1706A}.
-In order to investigate this relationship in our data, we make use of Galaxy Zoo 2's bar fraction (\textit{pbar}), which has been demonstrated to be a good measure of bar strength (\citealt{2012MNRAS.423.1485S}, \citealt{2012MNRAS.424.2180M}) and therefore a measure of the torque applied on the disc gas.
+In order to investigate this relationship in our data, we make use of Galaxy Zoo 2's bar fraction (\textit{pbar}), which has been demonstrated to be a good measure of bar length \citep{Willett2013:1308.3496v2} and bar strength (\citealt{2012MNRAS.423.1485S}, \citealt{2012MNRAS.424.2180M}) and therefore a good measure of the torque applied on the disc gas.
 

### 2020-06-22 sections/abstract.tex
implements brooke's suggested changes

@@
 Spiral structure is endemic in the Universe, and the pitch angle of arms in spiral galaxies is an important observable in efforts to discriminate between different mechanisms of spiral arm evolution.
-In this paper we present a hierarchical Bayesian approach to galaxy pitch angle determination, using spiral arm data obtained through the \textit{Galaxy Builder} citizen science project {\Lingard}.
+In this paper we present a hierarchical Bayesian approach to galaxy pitch angle determination, using spiral arm data obtained through the \textit{Galaxy Builder} citizen science project \citep{2020arXiv200610450L}.
 This approach better deals with the large variations in pitch angle between arms in a galaxy than existing methods, and obtains full posterior distributions on parameters.
-We utilise fit pitch angles to examine the links to morphology reported in other work, finding no link in our data, and examine a simple model of spiral winding, with which our results are consistent.
+We utilise fit pitch angles to examine links between bulge and bar strength and pitch angle reported in other work, finding no link in our data.
+We similarly test the simple model of spiral winding proposed by \citet{2019arXiv190910291P}, which assumes the uniformity of the cotangent of pitch angle between some unknown upper and lower limits, with results supporting this model of transient and reccurent spiral pitch angle.

### 2020-06-22 sections/introduction.tex
implements brooke's suggested changes

@@
 Spirals are as diverse as the theories proposed to govern their evolution; from the quintessential pair of well-defined arcs of the grand design spiral, to the fragmented arm segments of the flocculent spiral, to the disjointed multi-armed spiral.
-These variatons on structure account for $18\%$, $50\%$ and $32\%$ of the population respectively (\citealt{2011ApJ...737...32E}, \citealt{2015yCat..22170032B}, examples of each type are shown in Figure \ref{fig:spiral-galaxy-types}).
+These variations on structure account for $18\%$, $50\%$ and $32\%$ of the population respectively (\citealt{2011ApJ...737...32E}, \citealt{2015yCat..22170032B}, examples of each type are shown in Figure \ref{fig:spiral-galaxy-types}).
 The Hubble classification scheme \citep{1926ApJ....64..321H}, and its revisions and expansions (\citealt{1961hag..book.....S}, \citealt{1991rc3..book.....D}) contain detailed variations of different types of spiral galaxy, divided by the presence of a bar and ordered by how obvious spiral arm patterns are, how tightly they are wound and the prominence of a central bulge.
@@
 \begin{figure*}
-  \includegraphics[width=13cm]{plots/galaxy_types.jpg}
+  \includegraphics[width=13cm]{plots/galaxy_types.png}
   \caption{Examples of the different types of spiral galaxy present in the sky.
@@
 
-Our current understanding of the mechanisms which drive spiral growth and evolution suggest that each of the different forms of spiral galaxy may be triggered primarily by different processes.
+Our current understanding of the mechanisms which drive spiral growth and evolution suggests that each of the different forms of spiral galaxy may be triggered primarily by different processes.
 Grand Design spirals are thought to have undergone a tidal interaction, be driven by a bar (as seen in gas simulations, \citealt{1976ApJ...209...53S}, \citealt{2008A&A...489..115R}, and suggested for stars by Manifold theory, \citealt{2006A&A...453...39R}, \citealt{2009MNRAS.394...67A}, \citealt{2009MNRAS.400.1706A}), or be obeying (quasi-stationary) density wave theory (QSDW theory), in which spiral arms are slowly evolving, ever-present structures in the disc \citep{1964ApJ...140..646L}.
-Flocculent spirals are thought to be formed through swing amplification (shearing of small gravitational instabilities in the disc), and be transient and reccurent in nature \citep{1966ApJ...146..810J}.
+Flocculent spirals are thought to be formed through swing amplification (shearing of small gravitational instabilities in the disc), and be transient and recurrent in nature \citep{1966ApJ...146..810J}.
 
 One of the fundamental assumptions of early work on spiral formation mechanisms (primarily QSDW) was that the disc of a galaxy, if unstable to spiral perturbations, would create a stable, static wave which would exist unchanging for many rotational periods \citep{1964ApJ...140..646L}.
-The motivation for static waves with small numbers of arms (with a preference for $m=2$) was primarily observational; most galaxies show spiral structure with low spiral arm numbers, suggesting that spirals exist for a long time or are continually rebuilt.
+The motivation for static waves with small numbers of arms (with a preference for $m=2$) was primarily observational; most disc galaxies show spiral structure with low spiral arm numbers, suggesting that spirals exist for a long time or are continually rebuilt.
 
 Many simulations demonstrate that spirals do not maintain a constant pitch angle, and instead wind-up over time due to the differential rotation of the disc \citep{2013ApJ...763...46B}.
-Recent research suggests that spirals arms are transient in nature, and continually dissapate and re-form \citep{2014PASA...31...35D}.
+Recent research suggests that spirals arms are transient in nature, and continually dissipate and re-form \citep{2014PASA...31...35D}.
 These spirals can be maintained through the same mechanisms that drive QSDW spirals (i.e.
 WASER, \citealt{1976ApJ...205..363M}, swing amplification, \citealt{1965MNRAS.130..125G}), but do not require the idealistic disc conditions required for the formation and maintenance of a stationary wave.
-The pitch angles of these transient spiral arms will decrease due to the differential rotation of the disk, with the density of the arm peaking at some critical pitch angle, before dissapating to be reformed.
+The pitch angles of these transient spiral arms will decrease due to the differential rotation of the disk, with the density of the arm peaking at some critical pitch angle, before dissipating to be reformed.
 
-In this dynamic picture of spiral arms, pitch angle monotonically decreases from a spiral arm's formation to its dissapation.
+In this dynamic picture of spiral arms, pitch angle monotonically decreases from a spiral arm's formation to its dissipation.
 \citet{2019arXiv190910291P} propose a simple test of spiral arm winding, assuming the cotangent of the pitch angle of a spiral arm evolves linearly with time.
@@
 
-We aim to test this idea of spiral winding using data from the \textit{Galaxy Builder} citizen science project for the 196 spiral galaxies present in \Lingard.
-We make use of Bayesian hierarchical modelling to measure galaxy pitch angle from the spiral arm clusters producted by \textit{Galaxy Builder}.
-This methodology allows us to quantify the differences in pitch angles between arms in a single galaxy, as well as investigate the distribution of pitch angles in the galaxy population and investigate relationships between pitch angle and other galaxy properties such as stellar mass or morphology.
+We aim to test this idea of spiral winding using data from the \textit{Galaxy Builder} citizen science project for the spiral galaxies present in \citet{2020arXiv200610450L}.
+We make use of Bayesian hierarchical modelling to measure galaxy pitch angle from the spiral arm clusters produced by \textit{Galaxy Builder}.
+This methodology allows us to quantify the differences in pitch angles between arms in a single galaxy, as well as investigate the distribution of pitch angles in the galaxy population and investigate relationships between pitch angle and galaxy morphology.
 
-Using Galaxy Zoo 2 data \citep{Willett2013:1308.3496v2} allows the separation of galaxies by the presence and strength of bars and bulges; simulations of gas in barred galaxies often demonstrate that bars can drive long-term spiral evolution \citep{2008A&A...489..115R}, or boost transient spiral structure \citep{2012MNRAS.426..167G}.
+Using Galaxy Zoo 2 data \citep{Willett2013:1308.3496v2} allows the separation of galaxies by the presence and strength of a stellar bar; simulations of gas in barred galaxies often demonstrate that bars can drive long-term spiral evolution \citep{2008A&A...489..115R}, or boost transient spiral structure \citep{2012MNRAS.426..167G}.
 Manifold theory is one attempt to determine the orbits of stars in bar-driven spiral arms: it proposes that stars in the vicinity of the unstable Lagrangian points at either end of the bar tend to escape along predictable orbits, governed by invariant manifolds.
@@
 Many other systems correlate with spiral morphology, including potential ties to bulge fraction (\citealt{1975A&A....44..363Y}, \citealt{2013MNRAS.436.1074S}, \citealt{2019MNRAS.487.1808M}) and black hole mass (\citealt{2008ApJ...678L..93S}, \citealt{2017MNRAS.471.2187D}, \citealt{2019MS&E..571a2118A}).
-Stronger bulges and more more massive central black holes have both correlated with more tightly wound spiral arms.
+Stronger bulges and more massive central black holes have both correlated with more tightly wound spiral arms.
 
-Section \ref{section:spiral_winding} investigates spiral arm winding using the test derived by \cite{2019arXiv190910291P} (uniformity of galaxy pitch angle in $\cot\;\phi$) and concludes using a marginalized Anderson-Darling test that we cannot unilaterally reject winding of this form at the 1\% level.
-Section \ref{section:morphology_comparision} examines the correlation between pitch angle and bulge size implied by the Hubble sequence, and pitch angle and bar strength implied by Manifold theory, and find no significant correlation.
+Section \ref{section:spiral_winding} investigates spiral arm winding using the test derived by \cite{2019arXiv190910291P} (uniformity of galaxy pitch angle in $\cot\;\phi$).
+Section \ref{section:morphology_comparision} examines the correlation between pitch angle and bulge size implied by the Hubble sequence, and pitch angle and bar strength implied by Manifold theory.
+
+Where necessary, we make use of $H_0 = 70\ \text{km}\ \text{s}^{-1}\ \text{Mpc}^{-1}$.

### 2020-06-22 sections/method.tex
implements brooke's suggested changes

@@
 
-Many methodologies have been proposed and implemented to measure spiral arm properties, including visual inspection (\citealt{2015A&A...582A..86H}), fourier analysis (i.e.
+Many methodologies have been proposed and implemented to measure spiral arm properties, including visual inspection (\citealt{2015A&A...582A..86H}), Fourier analysis (i.e.
 \textsc{2DFFT}, \citealt{2012ApJS..199...33D}), texture analysis (i.e.
 SpArcFiRe, \citealt{2014ApJ...790...87D}), and combinations of automated methods and human classifiers (\citealt{2017MNRAS.472.2263H}, \citealt{2020MNRAS.493.3854H}).
-One potentially underused method of obtaining measurements of spirals is through photometric fitting of spiral structure, as possible using tools such as \textsc{GALFIT} \citep{2010AJ....139.2097P} and \textit{Galaxy Builder} (\Lingard).
+One potentially underused method of obtaining measurements of spirals is through photometric fitting of spiral structure, as possible using tools such as \textsc{GALFIT} \citep{2010AJ....139.2097P} and \textit{Galaxy Builder} \citep{2020arXiv200610450L}.
 These methods attempt to separate light from an image of a galaxy into distinct subcomponents, such as a galaxy disc, bulge, bar and spiral arms, generally finding the optimum solution using computational optimization.
 This optimization process, however, is often not robust for complex, many-component models and requires significant supervision to converge to a physically meaningful result \citep{Gao2017:1709.00746v1}.
-\Lingard propose a solution to this problem through the use of citizen science to provide priors on parameters used in computational fitting.
+\citet{2020arXiv200610450L} propose a solution to this problem through the use of citizen science to provide priors on parameters used in computational fitting.
 
@@
 \end{equation}
-
+%
 where $\phi$ is the arm's pitch angle.
@@
 
-A notable drawback of length-weighted pitch angle is sensitivity to the number and quality of the spiral arm segments identified; \citet{2017MNRAS.472.2263H} found that only 15\% of the arm segments identified by a leading algorithm (\textsc{SpArcFiRe}) were identified as ``good'' matches to real spiral arms by citizen science classifiers.
+A notable drawback of length-weighted pitch angle is sensitivity to the number and quality of the spiral arm segments identified; \citet{2017MNRAS.472.2263H} found that only 15\% of the arm segments identified using \textsc{SpArcFiRe} were identified as ``good'' matches to real spiral arms by citizen science classifiers.
 
 Fourier analysis in one- and two-dimensions (as performed by \citealt{2019arXiv190804246D}, \citealt{2012ApJS..199...33D}, \citealt{2018MNRAS.474.2594M}) is another widely used method of computationally obtaining galaxy pitch angles.
-Two-dimensional fourier methods generally decompose a deprojected image of a galaxy into a superpositions of logarithmic spirals between inner and outer annuli \citep{2012ApJS..199...33D}, and reports the pitch angle with the highest amplitude as the galaxy's pitch angle.
-\citet{2020MNRAS.493.3854H} combined fourier analysis of spiral galaxies with visual tracing of spiral arms, to great effect.
-It is uncertain how variation between pitch angles of individual arms impacts this measurement.
+Two-dimensional Fourier methods generally decompose a deprojected image of a galaxy into a superposition of logarithmic spirals between inner and outer annuli \citep{2012ApJS..199...33D} and reports the pitch angle with the highest amplitude as the galaxy's pitch angle.
+\citet{2020MNRAS.493.3854H} combined Fourier analysis of spiral galaxies with visual tracing of spiral arms, to great effect.
+It is uncertain how the variation between pitch angles of individual arms impacts this measurement.
 
 \subsection{The Galaxy Sample}
-The galaxies analysed in this paper are the 198 galaxies from \Lingard.
+The galaxies analysed in this paper are those for which photometric models were obtained in \citet{2020arXiv200610450L}.
 These are a subset of the \textit{stellar mass-complete sample} in \citet{2017MNRAS.472.2263H}, a sample of low-redshift face-on spiral galaxies selected using data from the NASA-Sloan Atlas \citep{2011AJ....142...31B} and Galaxy Zoo 2 \citep{Willett2013:1308.3496v2}.
 
-Some galaxies in \Lingard were shown to volunteers a second time in a repeat \textit{validation subset} in order to create a second aggregate model used to test internal consistency.
+Some galaxies in \citet{2020arXiv200610450L} were shown to volunteers a second time in a repeat \textit{validation subset} in order to create a second aggregate model used to test internal consistency.
 We combine the classifications of galaxies in this \textit{validation subset} with the original classifications.
-Clustering of drawn spiral arms and cleaning of points was then performed as detailed in \Lingard.
+Clustering of drawn spiral arms and cleaning of points was then performed as detailed in \citet{2020arXiv200610450L}.
 We remove any galaxies for which no spiral arms were identified, resulting in a hierarchical data structure of 139 galaxies, 261 spiral arms and 239,947 points.
 
-As we wish to utilize the logarithmic spiral's desirable properties of a constant pitch angle and small number of fittable parameters, we make use of it here without an explicit comparison to other models; a simple visual inspection of the fitted logarithmic spirals suggests that it is an appropriate model.
+As we wish to utilize the logarithmic spiral's desirable properties of a constant pitch angle and a small number of free parameters, we make use of it here without an explicit comparison to other models; a simple visual inspection of the fitted logarithmic spirals suggests that it is an appropriate model.
 A comparison of a logarithmic spiral profile to other spiral forms (i.e.
@@
 
-Spiral arm points are deprojected to a face-on orientation using the disk inclination and position angle obtained through photometric model fitting performed in \Lingard.
-Arms are individually corrected to all have the same chirality (a pitch angle greater than or equal to zero) using the logarithmic spiral fit in \Lingard.
+Spiral arm points are deprojected to a face-on orientation using the disk inclination and position angle obtained through photometric model fitting performed in \citet{2020arXiv200610450L}.
+Arms are individually corrected to all have the same chirality (a pitch angle greater than or equal to zero) using the logarithmic spiral fit in \citet{2020arXiv200610450L}.
 
@@
 
-In this section we lay out our bayesian hierarchical model for galaxy pitch angle.
+In this Section, we lay out our bayesian hierarchical model for galaxy pitch angle.
 We fit directly to clustered, cleaned points from polylines drawn in \textit{Galaxy Builder}, deprojected and unwrapped to polar coordinates.
@@
 The inverse gamma distribution is used to aid the convergence of the Hamiltonian Monte Carlo (HMC) algorithm used (discussed later).
-The Cauchy distirbution is equivalent to the Student's t-distribution with one degree of freedom, and was chosen due to its fatter tails than the normal distribution.
-Our likelihood function for $N$ arms, each of which with $n_\mathrm{arm}$ points is
+The Cauchy distribution is equivalent to the Student's t-distribution with one degree of freedom, and was chosen due to its fatter tails than the normal distribution.
+Our likelihood function for $N$ arms, each with $n_\mathrm{arm}$ points, is
 
@@
 
-We assume that the radial error is Gaussian for simplicity of analysis, Shapiro-Wilk tests on the residuals of the logarithmic spirals fit in {\Lingard} suggest that this is not a good assumption, and a more robust likelihood (such as the Student's t-distirbution) would possibly more appropriate.
+We assume that the radial error is Gaussian for simplicity of analysis, however, Shapiro-Wilk tests on the residuals of the logarithmic spirals fit in \citet{2020arXiv200610450L} suggest that this is not a good assumption, and a more robust likelihood (such as the Student's t-distribution) would possibly more appropriate.
 
-To perform inference, we make use of the No-U-Turn-Sampler (NUTS, \citealt{2011arXiv1111.4246H}), implemented in PYMC3\footnote{\url{https://docs.pymc.io/}}, an open source probabilistic programming framework written in python \citep{pymc3_paper}.
+To perform inference, we make use of the No-U-Turn-Sampler (NUTS, \citealt{2011arXiv1111.4246H}), implemented in PYMC3\footnote{\url{https://docs.pymc.io/}}, an open-source probabilistic programming framework written in Python \citep{pymc3_paper}.
 To aid the convergence of MC chains, we scale the radii of deprojected points to have unit variance.

### 2020-06-22 sections/results.tex
implements brooke's suggested changes

--- 2020-04-22 sections/results.tex
@@
 \end{equation}
-
 where $\sigma_\mathrm{gal}$ is our measure of inter-arm variability of pitch angle and has a posterior distribution of $11.0^\circ\pm 0.9^\circ$.
-This variability is similar to the finding of \citet{2014ApJ...790...87D} and emphaises the need for fitting algorithms to not assume all arms have the same pitch angle.
+This variability is similar to that found by \citet{1981AJ.....86.1847K} and \citet{2014ApJ...790...87D} and emphasises the need for fitting algorithms to not assume all arms have the same pitch angle.
 The spread of arm pitch angle from the mean galaxy pitch angle can be seen in Figure \ref{fig:arm-pa-spread}, with points colour-coded by the number of arms measured for a galaxy.
@@
   \caption{Scatter plot showing how arm pitch angle compares to galaxy pitch angle for galaxies with different pitch angles and number of arms.
-The top panel shows a Gaussian KDE for $E[\phi_\mathrm{gal}]$, and the right panel shows a Gaussian KDE for $E[\phi_\mathrm{arm} - \phi_\mathrm{gal}]$.}
+The top panel shows a Gaussian KDE for $E[\phi_\mathrm{gal}]$, and the right panel shows a Gaussian KDE for $E[\phi_\mathrm{arm} - \phi_\mathrm{gal}]$.
+The galaxy pitch angle is consistent with the mean of its arms, with large scatter and a slight bias against values near the lower bound of $0$.}
   \label{fig:arm-pa-spread}
@@
 \end{equation}
-
-Where $p_\mathrm{just\ noticeable}$, $p_\mathrm{obvious}$ and $p_\mathrm{dominant}$ are the fractions of classifications indicating the galaxy's bulge was ``just noticeable'', ``obvious'' or ``dominant'' respectively.
+where $p_\mathrm{just\ noticeable}$, $p_\mathrm{obvious}$ and $p_\mathrm{dominant}$ are the fractions of classifications indicating the galaxy's bulge was ``just noticeable'', ``obvious'' or ``dominant'' respectively.
 
@@
 Similarly comparing arm pitch angles for galaxies in the different samples results in not rejecting the null hypothesis at the 1\% level for any of the samples.
-The distributions of Anderson-Darling test statistic for $\phi_\mathrm{gal}$ and $\phi_\mathrm{arm}$ are shown in the upper panel of Figure \ref{fig:ad-morphology-test} in blue and orange respectively.
+The distributions of the Anderson-Darling test statistic for $\phi_\mathrm{gal}$ and $\phi_\mathrm{arm}$ are shown in the upper panel of Figure \ref{fig:ad-morphology-test} in blue and orange respectively.
 
@@
 The split point of 0.28 was also chosen to produce evenly sized comparison samples rather than from some physical motivation.
-However, the lack of any form of correlation implies that \textbf{there is no evidence in our data for the link between bulge size and pitch angle predicted by the Hubble sequence and observed in other studies}.
+However, the lack of any form of correlation implies that there is no evidence in our data for the link between bulge size and pitch angle predicted by the Hubble sequence and observed in other studies.
 
@@
 One of the predictions of Manifold theory is that pitch angle increases with bar strength \citep{2009MNRAS.400.1706A}.
-In order to investigate this relationship in our data, we make use of Galaxy Zoo 2's bar fraction (\textit{pbar}), which has been demonstrated to be a good measure of bar strength (\citealt{2012MNRAS.423.1485S}, \citealt{2012MNRAS.424.2180M}) and therefore a measure of the torque applied on the disc gas.
+In order to investigate this relationship in our data, we make use of Galaxy Zoo 2's bar fraction ($p_\mathrm{bar}$), which has been demonstrated to be a good measure of bar length \citep{Willett2013:1308.3496v2} and bar strength (\citealt{2012MNRAS.423.1485S}, \citealt{2012MNRAS.424.2180M}) and therefore a good measure of the torque applied on the disc gas.
 
-We do not observe a correlation between \textit{pbar} and $E[\phi_\mathrm{gal}]$ (Pearson correlation coefficient of -0.05).
-Following \citet{2012MNRAS.424.2180M} and \citet{2012MNRAS.423.1485S}, we separate the sample into galaxies without a bar ($\mathrm{\textit{pbar}} < 0.2$), with a weak bar ($0.2 \le \mathrm{\textit{pbar}} \le 0.5$) and with a strong bar ($\mathrm{\textit{pbar}} > 0.5$).
+We do not observe a correlation between $p_\mathrm{bar}$ and $E[\phi_\mathrm{gal}]$ (Pearson correlation coefficient of -0.05).
+Following \citet{2012MNRAS.424.2180M} and \citet{2012MNRAS.423.1485S}, we separate the sample into galaxies without a bar ($p_\mathrm{bar} < 0.2$), with a weak bar ($0.2 \le p_\mathrm{bar} \le 0.5$) and with a strong bar ($p_\mathrm{bar} > 0.5$).
 Performing marginalized three-sample Anderson-Darling tests does not find that pitch angles ($\phi_\mathrm{gal}$ or $\phi_\mathrm{arm}$) of galaxies with different bar strengths were drawn from different distributions; we do not reject the null hypothesis at the 1\% level for any samples for the test of $\phi_\mathrm{gal}$, and at the 10\% level for the test of $\phi_\mathrm{arm}$.
@@
   \includegraphics[width=17.7cm]{plots/bulge_bar_test_results.pdf}
-  \caption{The results of marginalized two-sample Anderson-Darling tests examining whether pitch angles ($\phi_\mathrm{gal}$ in blue and $\phi_\mathrm{gal}$ in orange) for galaxies with $B_\mathrm{avg} < 0.28$ and $B_\mathrm{avg} \ge 0.28$ are drawn from the same distribution (top panel), and the results of marginalized three-sample Anderson-Darling tests for galaxies with no bar ($\mathrm{\textit{pbar}} < 0.2$), a weak bar ($0.2 \le \mathrm{\textit{pbar}} \le 0.5$) and a strong bar ($\mathrm{\textit{pbar}} > 0.5$) (bottom panel).
-Confidence intervals are shown, with moving rightwards indicating more confidence in rejecting the null hypothesis that the compared values were drawn from the same parent distribution.}
+  \caption{The results of marginalized two-sample Anderson-Darling tests examining whether pitch angles ($\phi_\mathrm{gal}$ in blue and $\phi_\mathrm{gal}$ in orange) for galaxies with $B_\mathrm{avg} < 0.28$ and $B_\mathrm{avg} \ge 0.28$ are drawn from the same distribution (top panel), and the results of marginalized three-sample Anderson-Darling tests for galaxies with no bar ($p_\mathrm{bar} < 0.2$), a weak bar ($0.2 \le p_\mathrm{bar} \le 0.5$) and a strong bar ($p_\mathrm{bar} > 0.5$) (bottom panel).
+Confidence intervals are shown, with moving rightwards indicating more confidence in rejecting the null hypothesis that the compared values were drawn from the same parent distribution.
+We cannot reject the null hypotheis at the 1\% level for any of the tests conducted.}
   \label{fig:ad-morphology-test}
@@
 
-We do not correct for variation in bulge size in this work, however predictions from Manifold theory should not be affected as bulges do not provide a strong non-axisymmetric focing.
-The fact that we do not find any link between bar strength and pitch angle suggests that \textbf{the primary mechanism driving the evolution of the spirals in our sample is not Manifold theory}.
+We do not correct for variation in bulge size in this work, however, predictions from Manifold theory should not be affected as bulges do not provide a strong non-axisymmetric forcing.
+The fact that we do not find any link between bar strength and pitch angle suggests that the primary mechanism driving the evolution of the spirals in our sample is not Manifold theory.
 
@@
 
-For transient and reccurent spiral arms driven by self-gravity, \citet{2019arXiv190910291P} suggest that spiral patterns form at some maximum pitch angle ($\phi_\mathrm{max}$), continually wind up over time and finally dissapate at some minimum pitch angle ($\phi_\mathrm{min}$).
+For transient and recurrent spiral arms driven by self-gravity, \citet{2019arXiv190910291P} suggest that spiral patterns form at some maximum pitch angle ($\phi_\mathrm{max}$), continually wind up over time and finally dissipate at some minimum pitch angle ($\phi_\mathrm{min}$).
 They propose that, under a set of very simple assumptions, the evolution of pitch angle would be governed by
@@
 \end{equation}
-
 where $\Omega_p$ is the radially dependant pattern speed of the spiral arm and $t_0$ is the initial time at which it formed.
@@
 
-In order to test this theory, \citet{2019arXiv190910291P} used a Kolmogorov-Smirnov test to examine the consistency of a sample of observed galaxy pitch angles with one uniform in cot.
-Pitch angles were measured using discrete fourier transformations in one- and two-dimensions, and as such do not account for inter-arm variations.
+In order to test this theory, \citet{2019arXiv190910291P} used a Kolmogorov-Smirnov test to examine whether a sample of observed galaxy pitch angles was likely to have been drawn from a distribution uniform in its cotangent.
+Pitch angles were measured using discrete Fourier transformations in one- and two-dimensions, and as such do not account for inter-arm variations.
 They chose limits of $\cot{\phi} \in [1.00, 4.75]$ (roughly $11.9^\circ < \phi < 45.0^\circ$), motivated by examination of their data.
@@
 \subsubsection{Arm Pitch angle}
-The inconclusive result for $\phi_\mathrm{gal}$ is perhaps unsurprising: were we to assume that spiral arms are transient and reccurent instabilities, there is little reason for all of the arms to be at precisely the same evolutionary stage at the same time.
+The inconclusive result for $\phi_\mathrm{gal}$ is perhaps unsurprising: were we to assume that spiral arms are transient and recurrent instabilities, there is little reason for all of the arms to be at precisely the same evolutionary stage at the same time.
 This is supported by the large observed spread in inter-arm pitch angles.
 
-If we assume instead that spirals form and wind independantly in a galaxy, and that their evolution over time can be described by Equation \ref{eq:winding}, the distribution of pitch angles of individual arms should be uniform in cot between our limits, rather than that of the galaxy's pitch angle as a whole.
+If we assume instead that spirals form and wind independently in a galaxy, and that their evolution over time can be described by Equation \ref{eq:winding}, the distribution of the cotangent of pitch angles of individual arms should be uniform between our limits, rather than that of the galaxy's pitch angle as a whole.
 
@@
 The resulting distribution of Anderson-Darling statistics is shown in in the lower panel of Figure \ref{fig:ad-cot-test}.
-\textbf{This result is highly consistent with the model for spiral winding proposed by \citet{2019arXiv190910291P}}, and can be seen as evidence that spirals are formed through local disc perturbations, and are primarily governed by local forces.
+This result is highly consistent with the model for spiral winding proposed by \citet{2019arXiv190910291P} and can be seen as evidence that spirals are formed through local disc perturbations, and are primarily governed by local forces.
+
+\begin{figure*}
+  \includegraphics[width=17.7cm]{plots/phi_distribution_comparison.pdf}
+  \caption{The distributions of pitch angles relative to one uniform in $\cot\phi$.
+Histograms have been normalised by the area between the limits such that they are comparable.
+The histogram was recalculated with identical bins for each posterior sample of $\phi_\mathrm{gal}$ and $\phi_\mathrm{arm}$, we plot the mean value of each bin, with the sample standard deviation shown as errorbars.
+It is evident that the distributions are very similar between the chosen limits.}
+  \label{fig:pa-cot-distributions}
+\end{figure*}
 
@@
   \includegraphics[width=17.7cm]{plots/combined_cot_uniform_marginalized_tests.pdf}
-  \caption{The results of a marginalized Anderson-Darling test for uniformity in $\cot$ for $\phi_\mathrm{gal}$ (top panel) and $\phi_\mathrm{arm}$ (bottom panel), with values corresponding to various confidence intervals shown.
-Moving rightwards on the x-axis implies greater confidence in rejecting the null hypothesis that the sample was drawn from a distribution uniform in $\cot$ between $15^\circ < \phi < 50.0^\circ$.}
+  \caption{The results of a marginalized Anderson-Darling test for uniformity in $\cot$ for $\phi_\mathrm{gal}$ (blue) and $\phi_\mathrm{arm}$ (orange), with values corresponding to various confidence intervals shown.
+Moving rightwards on the x-axis implies greater confidence in rejecting the null hypothesis that the sample was drawn from a distribution uniform in $\cot$ between $15^\circ < \phi < 50.0^\circ$.
+In this instance, we would not be able to reject the null hypothesis at the 1\% level for either $\phi_\mathrm{gal}$ or $\phi_\mathrm{arm}$.}
   \label{fig:ad-cot-test}

### 2020-06-22 sections/summary.tex
implements brooke's suggested changes

--- 2020-06-02 sections/discussion.tex
@@
 The hierarchical Bayesian approach implemented in this paper allows a more thorough examination of pitch angle than length-weighted pitch angle calculation; obtaining posterior distributions of measured parameters.
-It better accounts for the large varitations observed in inter-arm pitch angle than fourier analysis, which assumes all arms in a given mode have the same pitch angle.
-In this work we find that the mean inter-arm difference in pitch angle is $11.0^\circ\pm 0.9^\circ$.
+It better accounts for the large variations observed in inter-arm pitch angle than Fourier analysis, which assumes all arms in a given mode have the same pitch angle.
+In this work, we find that the mean inter-arm difference in pitch angle is $11.0^\circ\pm 0.9^\circ$.
 
@@
 
-Our results are consistent with spiral winding of the form described by \citet{2019arXiv190910291P}, in which spiral arms are transient and reccurent, evolve through mechanisms such as swing-amplification \citep{1965MNRAS.130..125G} and which wind up over time.
+Our results are consistent with spiral winding of the form described by \citet{2019arXiv190910291P}, in which spiral arms are transient and recurrent, evolve through mechanisms such as swing-amplification \citep{1965MNRAS.130..125G} and which wind up over time.
 However, the assumptions of this model of spiral winding are highly simplistic, and it leaves many unanswered questions: what determines the limits on $\phi$?
 Is the spiral arm equally apparent at all pitch angles, or is a selection effect present?
-This result is also not evidence against QSDW, as it is possible that our distribtuion of pitch angles is dictated by other factors such as disk shear.
+This result is also not evidenced against QSDW, as it is possible that our distribution of pitch angles is dictated by other factors such as disk shear.
 
-In this work we assume that spiral arms are equally likely to be identified and recovered at all pitch angles.
-This is not an unfair assumption given the amount of human effort that went into obtaining spiral arms (more so than any other pitch angle measurement method, with each galaxy receiving 30 human classifications).
+In this work, we assume that spiral arms are equally likely to be identified and recovered at all pitch angles.
+This is not an unfair assumption given the amount of human effort that went into obtaining spiral arm measurements (more so than any other pitch angle measurement method, with each galaxy receiving 30 human classifications).
 The galaxy sample used is not guaranteed to be representative of the general spiral population, but is comparable in size to those used in other similar studies (\citealt{2013MNRAS.436.1074S}, \citealt{2019ApJ...871..194Y}, \citealt{2019arXiv190910291P}).
@@
 As with most analyses, the most impactful improvement it would be possible to make here would be to increase the cleanliness and volume of data analysed.
-A larger sample would make possible further comparisons, such as splitting galaxyies into spiral type (grand design / many-armed / flocculent) and examining the differences between populations.
+A larger sample would make possible further comparisons, such as splitting galaxies into spiral type (grand design / many-armed / flocculent) and examining the differences between populations.
 

### 2020-06-23 sections/introduction.tex
makes changes suggested by SK

@@
 Spiral structure is present in a majority of massive galaxies (\citealt{1989gadv.book..151B}, \citealt{2008MNRAS.389.1179L}) yet the formation mechanisms through which spiral structure originates are still hotly debated.
-Spirals are as diverse as the theories proposed to govern their evolution; from the quintessential pair of well-defined arcs of the grand design spiral, to the fragmented arm segments of the flocculent spiral, to the disjointed multi-armed spiral.
+Spirals are as diverse as the theories proposed to govern their evolution, from the quintessential pair of well-defined arcs of the grand design spiral, to the fragmented arm segments of the flocculent spiral, to the disjointed multi-armed spiral.
 These variations on structure account for $18\%$, $50\%$ and $32\%$ of the population respectively (\citealt{2011ApJ...737...32E}, \citealt{2015yCat..22170032B}, examples of each type are shown in Figure \ref{fig:spiral-galaxy-types}).
-The Hubble classification scheme \citep{1926ApJ....64..321H}, and its revisions and expansions (\citealt{1961hag..book.....S}, \citealt{1991rc3..book.....D}) contain detailed variations of different types of spiral galaxy, divided by the presence of a bar and ordered by how obvious spiral arm patterns are, how tightly they are wound and the prominence of a central bulge.
+The Hubble classification scheme \citep{1926ApJ....64..321H} and its revisions and expansions (\citealt{1961hag..book.....S}, \citealt{1991rc3..book.....D}) contain detailed variations of different types of spiral galaxy, divided by the presence of a bar and ordered by how obvious spiral arm patterns are, how tightly they are wound and the prominence of a central bulge.
 
@@
 
-A majority of the population of young stars in a galaxy are located in its spiral arms \citep{2011EAS....51...19E}, and it is possible that their passage triggers star formation \citep{2013A&A...560A..59C}, though their main role may instead be to sweep up material, promoting the growth of Giant Molecular Clouds.
+A majority of the population of young stars in a galaxy are located in its spiral arms \citep{2011EAS....51...19E}, and it is possible that they trigger star formation \citep{2013A&A...560A..59C}, though their main role may instead be to sweep up material, promoting the growth of Giant Molecular Clouds.
 \citep{2014IAUS..298..221D}.
 This rearrangement of disc gas can lead to the formation of disc-like bulges (commonly called ``pseudobulges''; e.g.
-\citealt{2004ARA&A..42..603K}).
-Studies of spiral morphology have found interesting correlations between spiral morphology and other galactic properties, such as a correlation between spiral tightness and central mass concentration (\citealt{2019ApJ...871..194Y}, though \citealt{2017MNRAS.472.2263H} found no such relation) and tightness and rotation curve shape (\citealt{2005MNRAS.359.1065S}, with rising rotation curves creating more open spiral structure).
-These predictions and observations provide compelling reasons for investigating their underlying rules and dynamics, as doing so is essential for understanding the secular evolution of disc galaxies.
+\citealt{2004ARA&A..42..603K}), which are prevalent in most spiral galaxies \citep{2018MNRAS.473.4731K}.
+Studies of spiral morphology have found interesting correlations with other galactic properties, such as a correlation between the tightness of spiral arms and central mass concentration (\citealt{2019ApJ...871..194Y}, though neither \citealt{2017MNRAS.472.2263H} not \citealt{2019MNRAS.487.1808M} found such a relation in large samples); and spiral tightness and rotation curve shape (\citealt{2005MNRAS.359.1065S}, with rising rotation curves creating more open spiral structure).
+These predictions and observations provide compelling reasons for investigating the underlying rules and dynamics of spiral stucture, as doing so is essential for understanding the secular evolution of disc galaxies.
 
-Our current understanding of the mechanisms which drive spiral growth and evolution suggests that each of the different forms of spiral galaxy may be triggered primarily by different processes.
-Grand Design spirals are thought to have undergone a tidal interaction, be driven by a bar (as seen in gas simulations, \citealt{1976ApJ...209...53S}, \citealt{2008A&A...489..115R}, and suggested for stars by Manifold theory, \citealt{2006A&A...453...39R}, \citealt{2009MNRAS.394...67A}, \citealt{2009MNRAS.400.1706A}), or be obeying (quasi-stationary) density wave theory (QSDW theory), in which spiral arms are slowly evolving, ever-present structures in the disc \citep{1964ApJ...140..646L}.
+Our current understanding of the mechanisms which drive spiral growth and evolution suggests that each of the different forms of spiral galaxy may be triggered by different processes.
+Grand Design spirals are thought to have undergone a tidal interaction \citep{2010MNRAS.403..625D,2017ApJ...834....7S}, be driven by a bar (as seen in gas simulations, \citealt{1976ApJ...209...53S,2008A&A...489..115R}, and suggested for stars by Manifold theory, \citealt{2006A&A...453...39R,2009MNRAS.394...67A,2009MNRAS.400.1706A}), or be obeying (quasi-stationary) density wave theory (QSDW theory), in which spiral arms are slowly evolving, ever-present structures in the disc \citep{1964ApJ...140..646L}.
 Flocculent spirals are thought to be formed through swing amplification (shearing of small gravitational instabilities in the disc), and be transient and recurrent in nature \citep{1966ApJ...146..810J}.
 
-One of the fundamental assumptions of early work on spiral formation mechanisms (primarily QSDW) was that the disc of a galaxy, if unstable to spiral perturbations, would create a stable, static wave which would exist unchanging for many rotational periods \citep{1964ApJ...140..646L}.
-The motivation for static waves with small numbers of arms (with a preference for $m=2$) was primarily observational; most disc galaxies show spiral structure with low spiral arm numbers, suggesting that spirals exist for a long time or are continually rebuilt.
+One of the fundamental assumptions of early work on spiral formation mechanisms (primarily QSDW) was that the disc of a galaxy, if unstable to spiral perturbations, would create a stable, static wave which would exist unchanged for many rotational periods \citep{1964ApJ...140..646L}.
+The motivation for static waves with small numbers of arms (with a preference for $m=2$) was primarily observational: most disc galaxies show spiral structure with low spiral arm numbers, suggesting that spirals exist for a long time or are continually rebuilt.
 
-Many simulations demonstrate that spirals do not maintain a constant pitch angle, and instead wind-up over time due to the differential rotation of the disc \citep{2013ApJ...763...46B}.
+Many simulations demonstrate that spirals do not maintain a constant tightness (often quantified by pitch angle, the angle between the spiral and the tangent to a circle centred on the galaxy, \citealt{1987gady.book.....B}, illustrated in Figure \ref{fig:pitch-angle-example}), and instead wind-up over time due to the differential rotation of the disc \citep{2013ApJ...763...46B}.
 Recent research suggests that spirals arms are transient in nature, and continually dissipate and re-form \citep{2014PASA...31...35D}.
 These spirals can be maintained through the same mechanisms that drive QSDW spirals (i.e.
-WASER, \citealt{1976ApJ...205..363M}, swing amplification, \citealt{1965MNRAS.130..125G}), but do not require the idealistic disc conditions required for the formation and maintenance of a stationary wave.
+``wave amplification by stimulated emission of radiation'', \citealt{1976ApJ...205..363M}; swing amplification, \citealt{1965MNRAS.130..125G}), but do not require the idealistic disc conditions required for the formation and maintenance of a stationary wave.
 The pitch angles of these transient spiral arms will decrease due to the differential rotation of the disk, with the density of the arm peaking at some critical pitch angle, before dissipating to be reformed.
 
+\begin{figure}
+  \includegraphics[width=8.4cm]{plots/pitch-angle-explanation.pdf}
+  \caption{Illustration of the definition of pitch angle.
+It is given as $\phi = \tan^{-1}\left(\frac{\mathrm{d}r}{\mathrm{d}\theta}\ /\ r\right)$, or the angle between the spiral (red) and the tangent to a circle centred on the galaxy (blue).}
+  \label{fig:example-spiral-fits}
+\end{figure}
 In this dynamic picture of spiral arms, pitch angle monotonically decreases from a spiral arm's formation to its dissipation.
@@
 We aim to test this idea of spiral winding using data from the \textit{Galaxy Builder} citizen science project for the spiral galaxies present in \citet{2020arXiv200610450L}.
-We make use of Bayesian hierarchical modelling to measure galaxy pitch angle from the spiral arm clusters produced by \textit{Galaxy Builder}.
+We make use of Bayesian hierarchical modelling to measure galaxy pitch angle from the spiral arms produced by \textit{Galaxy Builder}.
 This methodology allows us to quantify the differences in pitch angles between arms in a single galaxy, as well as investigate the distribution of pitch angles in the galaxy population and investigate relationships between pitch angle and galaxy morphology.
 
-Using Galaxy Zoo 2 data \citep{Willett2013:1308.3496v2} allows the separation of galaxies by the presence and strength of a stellar bar; simulations of gas in barred galaxies often demonstrate that bars can drive long-term spiral evolution \citep{2008A&A...489..115R}, or boost transient spiral structure \citep{2012MNRAS.426..167G}.
+Using Galaxy Zoo 2 data \citep{Willett2013:1308.3496v2} allows the separation of galaxies by the presence and strength of a stellar bar, simulations of gas in barred galaxies often demonstrate that bars can drive long-term spiral evolution \citep{2008A&A...489..115R}, or boost transient spiral structure \citep{2012MNRAS.426..167G}.
 Manifold theory is one attempt to determine the orbits of stars in bar-driven spiral arms: it proposes that stars in the vicinity of the unstable Lagrangian points at either end of the bar tend to escape along predictable orbits, governed by invariant manifolds.
@@
 
-Many other systems correlate with spiral morphology, including potential ties to bulge fraction (\citealt{1975A&A....44..363Y}, \citealt{2013MNRAS.436.1074S}, \citealt{2019MNRAS.487.1808M}) and black hole mass (\citealt{2008ApJ...678L..93S}, \citealt{2017MNRAS.471.2187D}, \citealt{2019MS&E..571a2118A}).
-Stronger bulges and more massive central black holes have both correlated with more tightly wound spiral arms.
+Many other components correlate with spiral morphology, including potential ties to bulge fraction (\citealt{1975A&A....44..363Y}, \citealt{2013MNRAS.436.1074S}, \citealt{2019MNRAS.487.1808M}) and black hole mass (\citealt{2008ApJ...678L..93S}, \citealt{2017MNRAS.471.2187D}, \citealt{2019MS&E..571a2118A}).
+Larger bulges and more massive central black holes have both correlated with more tightly wound spiral arms.
 

### 2020-06-23 sections/method.tex
makes changes suggested by SK

@@
 \subsection{Measuring galaxy pitch angle}
-
 Many methodologies have been proposed and implemented to measure spiral arm properties, including visual inspection (\citealt{2015A&A...582A..86H}), Fourier analysis (i.e.
@@
 \begin{equation}
+  \label{eq:log-spiral}
 r = A\,e^{\theta\tan\phi},
@@
 %
-where $\phi$ is the arm's pitch angle.
+where $\phi$ is the arm's pitch angle, $A$ is an amplitude coefficient and $\theta$ is the polar coordinate.
 One method used to obtain a pitch angle of a galaxy is to fit logarithmic spirals to individually identified arm segments and take the weighted mean of their pitch angles (which often vary by upwards of $10^\circ$, \citealt{2014ApJ...790...87D}).
@@
 
-A notable drawback of length-weighted pitch angle is sensitivity to the number and quality of the spiral arm segments identified; \citet{2017MNRAS.472.2263H} found that only 15\% of the arm segments identified using \textsc{SpArcFiRe} were identified as ``good'' matches to real spiral arms by citizen science classifiers.
+A notable drawback of length-weighted pitch angle is sensitivity to the number and quality of the spiral arm segments identified; \citet{2017MNRAS.472.2263H} found that only 15\% of the arm segments identified using \textsc{SpArcFiRe} \citep{2014ApJ...790...87D} were identified as ``good'' matches to real spiral arms by citizen science classifiers.
 
@@
 Two-dimensional Fourier methods generally decompose a deprojected image of a galaxy into a superposition of logarithmic spirals between inner and outer annuli \citep{2012ApJS..199...33D} and reports the pitch angle with the highest amplitude as the galaxy's pitch angle.
-\citet{2020MNRAS.493.3854H} combined Fourier analysis of spiral galaxies with visual tracing of spiral arms, to great effect.
+\citet{2020MNRAS.493.3854H} combined Fourier analysis of spiral galaxies with visual tracing of spiral arms, succesfully eliminating observed bias in a sample of toy images of galaxies.
 It is uncertain how the variation between pitch angles of individual arms impacts this measurement.
@@
 Some galaxies in \citet{2020arXiv200610450L} were shown to volunteers a second time in a repeat \textit{validation subset} in order to create a second aggregate model used to test internal consistency.
-We combine the classifications of galaxies in this \textit{validation subset} with the original classifications.
+We combine the 30 classifications of galaxies in this \textit{validation subset} with the 30 original classifications.
 Clustering of drawn spiral arms and cleaning of points was then performed as detailed in \citet{2020arXiv200610450L}.
@@
 Arms are individually corrected to all have the same chirality (a pitch angle greater than or equal to zero) using the logarithmic spiral fit in \citet{2020arXiv200610450L}.
+This was achieved by multiplying the polar coordinate $\theta$ by $-1$ for arms identified as winding counter-clockwise.
 
 \subsection{Bayesian modelling of spiral arms in \textit{Galaxy Builder}}
-
-In this Section, we lay out our bayesian hierarchical model for galaxy pitch angle.
+\label{section:bhsm-model}
+In this Section, we lay out our Bayesian hierarchical model for galaxy pitch angle.
 We fit directly to clustered, cleaned points from polylines drawn in \textit{Galaxy Builder}, deprojected and unwrapped to polar coordinates.
@@
 
-Where $\widetilde{r_\mathrm{arm}}$ is the model's predictions for the radii of the deprojected points in a \textit{Galaxy Builder} arm ($\widevec{r_\mathrm{arm}}$), and $\widevec{\theta_\mathrm{arm}}$ is the polar angles of the points.
+Where $\widetilde{r_\mathrm{arm}}$ is the model's predictions for the radii of the deprojected points in a \textit{Galaxy Builder} arm ($\widevec{r_\mathrm{arm}}$), $c_\mathrm{arm}$ is the amplitude parameter (equivalent to $A$ in Equation \ref{eq:log-spiral}), and $\widevec{\theta_\mathrm{arm}}$ is the polar angles of the points.
 

### 2020-06-23 sections/results.tex
makes changes suggested by SK

@@
 \subsection{Constraints on Galaxy Pitch angle}
-Our hierarchical model identifies the pitch angle of individual arms ($\phi_\mathrm{arm}$) with less than {1.6\degree} of uncertainty for 95\% of arms, assuming no error on disc inclination and position angle.
+\label{section:constraints-on-galaxy-phi}
+Our hierarchical model identifies the pitch angle of individual arms ($\phi_\mathrm{arm}$) with posterior standard deviations less than {1.6\degree} for 95\% of arms, assuming no error on disc inclination and position angle.
+This is illustrated well by the small uncertainties on fit spiral arms in Figure \ref{fig:example-spiral-fits}.
 The pitch angle of a galaxy as a whole ($\phi_\mathrm{gal}$), however, is not well constrained.
@@
 This is roughly consistent with the standard error on the mean for a galaxy with $N$ arms,
+
+\begin{figure*}
+  \includegraphics[width=17.7cm]{plots/example-spiral-fits.pdf}
+  \caption{Examples of spiral profiles fit using the hierarchical model described in Section \ref{section:bhsm-model}.
+Deprojected points from \textit{Galaxy Builder} clustered, cleaned spiral arms are shown in black; fit logarithmic spiral arms are shown in red, with the width of the line corresponding to the $2\sigma$ interval on predicted values of $\widetilde{r_\mathrm{arm}}$.}
+  \label{fig:example-spiral-fits}
+\end{figure*}
 
@@
 The spread of arm pitch angle from the mean galaxy pitch angle can be seen in Figure \ref{fig:arm-pa-spread}, with points colour-coded by the number of arms measured for a galaxy.
-We see a slight drop in $E[\phi_\mathrm{gal}]$ compared to $E[\phi_\mathrm{arm}]$ at small galaxy pitch angles, due to the truncation of $\phi_\mathrm{gal}$ at {0\degree}.
+We see a slight drop in the expectation values of galaxy pitch angle ($E[\phi_\mathrm{gal}]$) compared to the expectation of arm pitch angles ($E[\phi_\mathrm{arm}]$) at small galaxy pitch angles, due to the truncation of $\phi_\mathrm{gal}$ at {0\degree}.
 
@@
 We see no correlation between galaxy pitch angle derived from the hierarchical model and $B_\mathrm{avg}$.
-The Pearson correlation coefficient between the expectation value of galaxy pitch angle ($E[\phi_\mathrm{gal}]$) and $B_\mathrm{avg}$ is 0.00.
+The Pearson correlation coefficient between the expectation value of galaxy pitch angle ($E[\phi_\mathrm{gal}]$) and $B_\mathrm{avg}$ is 0.00 (with a p-value of 0.95).
+
+\begin{figure*}
+  \includegraphics[width=17.7cm]{plots/bulge_bar_phigal_distribution.pdf}
+  \caption{Density plot showing bulge strength ($B_\mathrm{avg}$; left, orange) and bar strength ($p_\mathrm{bar}$; right, green) against galaxy pitch angle ($\phi_\mathrm{gal}$).
+Split points for the marginalized Anderson-Darling tests are labelled.
+There is no statistically significant relationship for either bulge or bar strength.}
+  \label{fig:bulge-bar-pa-hist}
+\end{figure*}
 
@@
 One of the predictions of Manifold theory is that pitch angle increases with bar strength \citep{2009MNRAS.400.1706A}.
-In order to investigate this relationship in our data, we make use of Galaxy Zoo 2's bar fraction ($p_\mathrm{bar}$), which has been demonstrated to be a good measure of bar length \citep{Willett2013:1308.3496v2} and bar strength (\citealt{2012MNRAS.423.1485S}, \citealt{2012MNRAS.424.2180M}) and therefore a good measure of the torque applied on the disc gas.
+In order to investigate this relationship in our data, we make use of Galaxy Zoo 2's bar fraction ($p_\mathrm{bar}$), which has been demonstrated to be a good measure of bar length \citep{Willett2013:1308.3496v2} and bar strength \citep{2012MNRAS.423.1485S,2012MNRAS.424.2180M,2018MNRAS.473.4731K} and therefore a good measure of the torque applied on the disc gas.
 
-We do not observe a correlation between $p_\mathrm{bar}$ and $E[\phi_\mathrm{gal}]$ (Pearson correlation coefficient of -0.05).
+We do not observe a correlation between $p_\mathrm{bar}$ and $E[\phi_\mathrm{gal}]$ (Pearson correlation coefficient of -0.05, with a p-value of 0.54).
 Following \citet{2012MNRAS.424.2180M} and \citet{2012MNRAS.423.1485S}, we separate the sample into galaxies without a bar ($p_\mathrm{bar} < 0.2$), with a weak bar ($0.2 \le p_\mathrm{bar} \le 0.5$) and with a strong bar ($p_\mathrm{bar} > 0.5$).
@@
 
-We do not correct for variation in bulge size in this work, however, predictions from Manifold theory should not be affected as bulges do not provide a strong non-axisymmetric forcing.
 The fact that we do not find any link between bar strength and pitch angle suggests that the primary mechanism driving the evolution of the spirals in our sample is not Manifold theory.
@@
 
-We aim to replicate their work here, using our sample and methods.
+We perform a similar test in this work, using our sample and methods.
 We will make use of the marginalized Anderson-Darling test described above, and examine winding on a per-arm basis, as well as a per-galaxy basis.
-Observation of the distribution of arm pitch angles in our sample suggests limits of $15^\circ < \phi < 50.0^\circ$.
+Observation of the distribution of arm pitch angles in our sample (Figure \ref{fig:pa-cot-distributions}) suggests limits of $15^\circ < \phi < 50.0^\circ$.
 
@@
 This result is also highly sensitive to the lower limit of $\phi$: decreasing it to {10\degree} results in us rejecting the cot-uniform model at greater than the 0.1\% level for 96\% of the posterior samples.
-As we have no information available on the biases present in \textit{Galaxy Builder} spiral arm classification, we choose to keep the less strict limit of {15\degree}.
+As we have no information available on the selection biases present for classification of extremely loose or tight spiral arms in \textit{Galaxy Builder}, we choose to keep the less strict limit of {15\degree}.
 
@@
 The inconclusive result for $\phi_\mathrm{gal}$ is perhaps unsurprising: were we to assume that spiral arms are transient and recurrent instabilities, there is little reason for all of the arms to be at precisely the same evolutionary stage at the same time.
-This is supported by the large observed spread in inter-arm pitch angles.
+This is supported by the large observed spread in inter-arm pitch angles (Section \ref{section:constraints-on-galaxy-phi}).
 
-If we assume instead that spirals form and wind independently in a galaxy, and that their evolution over time can be described by Equation \ref{eq:winding}, the distribution of the cotangent of pitch angles of individual arms should be uniform between our limits, rather than that of the galaxy's pitch angle as a whole.
+If we assume instead that spirals form and wind independently inside a galaxy, and that their evolution over time can be described by Equation \ref{eq:winding}, the distribution of the cotangent of pitch angles of individual arms should be uniform between our limits, rather than that of the galaxy's pitch angle as a whole.
 

### 2020-06-23 sections/summary.tex
makes changes suggested by SK

@@
 The methodology proposed here is a robust solution to the problems facing investigation of spiral morphology, namely that of reliably identifying spiral arms, and properly accounting for the spread in pitch angles of arms within a galaxy.
-As with most analyses, the most impactful improvement it would be possible to make here would be to increase the cleanliness and volume of data analysed.
+As with most analyses, the most impactful improvement it would be possible to make here would be to increase the quality and volume of data analysed.
 A larger sample would make possible further comparisons, such as splitting galaxies into spiral type (grand design / many-armed / flocculent) and examining the differences between populations.

### 2020-06-29 sections/abstract.tex
makes KLM suggested changes

@@
 % !TEX root = ../main.tex
-Spiral structure is endemic in the Universe, and the pitch angle of arms in spiral galaxies is an important observable in efforts to discriminate between different mechanisms of spiral arm evolution.
+Spiral structure is ubiquitous in the Universe, and the pitch angle of arms in spiral galaxies is an important observable in efforts to discriminate between different mechanisms of spiral arm formation and evolution.
 In this paper we present a hierarchical Bayesian approach to galaxy pitch angle determination, using spiral arm data obtained through the \textit{Galaxy Builder} citizen science project \citep{2020arXiv200610450L}.

### 2020-06-29 sections/introduction.tex
makes KLM suggested changes

@@
 
-Spiral structure is present in a majority of massive galaxies (\citealt{1989gadv.book..151B}, \citealt{2008MNRAS.389.1179L}) yet the formation mechanisms through which spiral structure originates are still hotly debated.
+Spiral structure is present in a majority of massive galaxies (e.g.
+\citealt{1989gadv.book..151B}, \citealt{2008MNRAS.389.1179L}) yet the formation mechanisms through which spiral structure originates are still hotly debated (e.g.
+\citealt{2014PASA...31...35D}).
 Spirals are as diverse as the theories proposed to govern their evolution, from the quintessential pair of well-defined arcs of the grand design spiral, to the fragmented arm segments of the flocculent spiral, to the disjointed multi-armed spiral.
-These variations on structure account for $18\%$, $50\%$ and $32\%$ of the population respectively (\citealt{2011ApJ...737...32E}, \citealt{2015yCat..22170032B}, examples of each type are shown in Figure \ref{fig:spiral-galaxy-types}).
-The Hubble classification scheme \citep{1926ApJ....64..321H} and its revisions and expansions (\citealt{1961hag..book.....S}, \citealt{1991rc3..book.....D}) contain detailed variations of different types of spiral galaxy, divided by the presence of a bar and ordered by how obvious spiral arm patterns are, how tightly they are wound and the prominence of a central bulge.
+These variations on structure account for $18\%$, $50\%$ and $32\%$ of the population respectively (\citealt{2011ApJ...737...32E}; examples of each type are shown in Figure \ref{fig:spiral-galaxy-types}).
+The Hubble classification scheme \citep{1926ApJ....64..321H} and its revisions and expansions \citep{1961hag..book.....S,1991rc3..book.....D} contain detailed variations of different types of spiral galaxy, divided by the presence of a bar and ordered by how obvious spiral arm patterns are, how tightly they are wound and the prominence of a central bulge.
 
@@
 This rearrangement of disc gas can lead to the formation of disc-like bulges (commonly called ``pseudobulges''; e.g.
-\citealt{2004ARA&A..42..603K}), which are prevalent in most spiral galaxies \citep{2018MNRAS.473.4731K}.
+\citealt{2004ARA&A..42..603K}), which are prevalent in most spiral galaxies \citep{2018MNRAS.473.4731K,2010ApJ...716..942F}.
 Studies of spiral morphology have found interesting correlations with other galactic properties, such as a correlation between the tightness of spiral arms and central mass concentration (\citealt{2019ApJ...871..194Y}, though neither \citealt{2017MNRAS.472.2263H} not \citealt{2019MNRAS.487.1808M} found such a relation in large samples); and spiral tightness and rotation curve shape (\citealt{2005MNRAS.359.1065S}, with rising rotation curves creating more open spiral structure).
-These predictions and observations provide compelling reasons for investigating the underlying rules and dynamics of spiral stucture, as doing so is essential for understanding the secular evolution of disc galaxies.
+These predictions and observations provide compelling reasons for continued investigation of the underlying rules and dynamics of spiral stucture, as doing so is essential for understanding the secular evolution of disc galaxies.
 
 Our current understanding of the mechanisms which drive spiral growth and evolution suggests that each of the different forms of spiral galaxy may be triggered by different processes.
-Grand Design spirals are thought to have undergone a tidal interaction \citep{2010MNRAS.403..625D,2017ApJ...834....7S}, be driven by a bar (as seen in gas simulations, \citealt{1976ApJ...209...53S,2008A&A...489..115R}, and suggested for stars by Manifold theory, \citealt{2006A&A...453...39R,2009MNRAS.394...67A,2009MNRAS.400.1706A}), or be obeying (quasi-stationary) density wave theory (QSDW theory), in which spiral arms are slowly evolving, ever-present structures in the disc \citep{1964ApJ...140..646L}.
+Grand design spirals are thought to have undergone a tidal interaction \citep{2010MNRAS.403..625D,2017ApJ...834....7S}, be driven by a bar (as seen in gas simulations, \citealt{1976ApJ...209...53S,2008A&A...489..115R}, and suggested for stars by Manifold theory, \citealt{2006A&A...453...39R,2009MNRAS.394...67A,2009MNRAS.400.1706A}), or be obeying (quasi-stationary) density wave theory (QSDW theory), in which spiral arms are slowly evolving, ever-present structures in the disc (as first proposed by \citealt{1964ApJ...140..646L}).
 Flocculent spirals are thought to be formed through swing amplification (shearing of small gravitational instabilities in the disc), and be transient and recurrent in nature \citep{1966ApJ...146..810J}.
@@
 It is given as $\phi = \tan^{-1}\left(\frac{\mathrm{d}r}{\mathrm{d}\theta}\ /\ r\right)$, or the angle between the spiral (red) and the tangent to a circle centred on the galaxy (blue).}
-  \label{fig:example-spiral-fits}
+  \label{fig:pitch-angle-example}
 \end{figure}
+
 In this dynamic picture of spiral arms, pitch angle monotonically decreases from a spiral arm's formation to its dissipation.
@@
 Using Galaxy Zoo 2 data \citep{Willett2013:1308.3496v2} allows the separation of galaxies by the presence and strength of a stellar bar, simulations of gas in barred galaxies often demonstrate that bars can drive long-term spiral evolution \citep{2008A&A...489..115R}, or boost transient spiral structure \citep{2012MNRAS.426..167G}.
-Manifold theory is one attempt to determine the orbits of stars in bar-driven spiral arms: it proposes that stars in the vicinity of the unstable Lagrangian points at either end of the bar tend to escape along predictable orbits, governed by invariant manifolds.
+Manifold theory \citealt{2006A&A...453...39R,2009MNRAS.394...67A,2009MNRAS.400.1706A} is one attempt to determine the orbits of stars in bar-driven spiral arms: it proposes that stars in the vicinity of the unstable Lagrangian points at either end of the bar tend to escape along predictable orbits, governed by invariant manifolds.
 One of the primary factors influencing the shape of this invariant manifold is the relative strength of the non-axisymmetric forcing caused by the bar, with stronger bars resulting in spirals with larger pitch angles.
@@
 
+Section \ref{section:morphology-comparison} examines the correlation between pitch angle and bulge size implied by the Hubble sequence, and pitch angle and bar strength implied by Manifold theory.
 Section \ref{section:spiral_winding} investigates spiral arm winding using the test derived by \cite{2019arXiv190910291P} (uniformity of galaxy pitch angle in $\cot\;\phi$).
-Section \ref{section:morphology_comparision} examines the correlation between pitch angle and bulge size implied by the Hubble sequence, and pitch angle and bar strength implied by Manifold theory.
 

### 2020-06-29 sections/method.tex
makes KLM suggested changes

@@
 This optimization process, however, is often not robust for complex, many-component models and requires significant supervision to converge to a physically meaningful result \citep{Gao2017:1709.00746v1}.
-\citet{2020arXiv200610450L} propose a solution to this problem through the use of citizen science to provide priors on parameters used in computational fitting.
+\citet{2020arXiv200610450L} proposed a solution to this problem through the use of citizen science to provide priors on parameters used in computational fitting.
 
-A common assumption when measuring galaxy pitch angle is that observed spiral arms have a constant pitch angle. Spirals of this kind are known as logarithmic spirals and are described by
+A common assumption when measuring galaxy pitch angle is that observed spiral arms have a constant pitch angle with radius (e.g.
+\citealt{2012ApJS..199...33D,2013MNRAS.436.1074S,2014ApJ...790...87D}.
+Spirals of this kind are known as logarithmic spirals and are described by
 
@@
 
-The most commonly used measurement of uncertainty of length-weighted pitch angles is the unweighted sample variance between the arm segments identified.
+The most commonly used measurement of uncertainty of length-weighted pitch angles is the unweighted sample variance between the arm segments which were identified.
 
-A notable drawback of length-weighted pitch angle is sensitivity to the number and quality of the spiral arm segments identified; \citet{2017MNRAS.472.2263H} found that only 15\% of the arm segments identified using \textsc{SpArcFiRe} \citep{2014ApJ...790...87D} were identified as ``good'' matches to real spiral arms by citizen science classifiers.
+A notable drawback of length-weighted pitch angle is sensitivity to the number and quality of the spiral arm segments; \citet{2017MNRAS.472.2263H} found that only 15\% of the arm segments which were identified using \textsc{SpArcFiRe} \citep{2014ApJ...790...87D} were identified as ``good'' matches to real spiral arms by citizen science classifiers.
 
@@
 \citet{2020MNRAS.493.3854H} combined Fourier analysis of spiral galaxies with visual tracing of spiral arms, succesfully eliminating observed bias in a sample of toy images of galaxies.
-It is uncertain how the variation between pitch angles of individual arms impacts this measurement.
+It is unclear how the variation between pitch angles of individual arms impacts this measurement.
 
@@
 The galaxies analysed in this paper are those for which photometric models were obtained in \citet{2020arXiv200610450L}.
-These are a subset of the \textit{stellar mass-complete sample} in \citet{2017MNRAS.472.2263H}, a sample of low-redshift face-on spiral galaxies selected using data from the NASA-Sloan Atlas \citep{2011AJ....142...31B} and Galaxy Zoo 2 \citep{Willett2013:1308.3496v2}.
+These are a subset of the \textit{stellar mass-complete sample} in \citet{2017MNRAS.472.2263H}, a sample of low-redshift ($0.02 < z < 0.055$) face-on spiral galaxies selected using data from the NASA-Sloan Atlas \citep{2011AJ....142...31B} and Galaxy Zoo 2 \citep{Willett2013:1308.3496v2}.
+The \textit{stellar mass-complete sample} ranged in stellar mass from $9.45 < \log(M_* / M_\odot) < 11.05$.
 
@@
 We remove any galaxies for which no spiral arms were identified, resulting in a hierarchical data structure of 139 galaxies, 261 spiral arms and 239,947 points.
-
-As we wish to utilize the logarithmic spiral's desirable properties of a constant pitch angle and a small number of free parameters, we make use of it here without an explicit comparison to other models; a simple visual inspection of the fitted logarithmic spirals suggests that it is an appropriate model.
-A comparison of a logarithmic spiral profile to other spiral forms (i.e.
-Archimedian or polynomial) is another important piece of work, outside of the scope of this research, as it has been reported that galaxy arms do not have constant pitch angles (\citealt{1981AJ.....86.1847K}; \citealt{2009MNRAS.397..164R}).
 
@@
 We fit directly to clustered, cleaned points from polylines drawn in \textit{Galaxy Builder}, deprojected and unwrapped to polar coordinates.
-We fit a logarithmic spiral to each clustered arm, with the pitch angles of multiple arms in a single galaxy being drawn from a single parent distribution.
+We fit a logarithmic spiral to each clustered arm (examples are shown in Figure \ref{fig:example-spiral-fits}), with the pitch angles of multiple arms in a single galaxy being drawn from a single parent distribution.
+
+We wish to utilize the logarithmic spiral's desirable properties of a constant pitch angle and a small number of free parameters, therefore, we make use of it here without an explicit comparison to other models.
+A simple visual inspection of the fitted logarithmic spirals suggests that it is an appropriate model, however, a comparison of a logarithmic spiral profile to other spiral forms (i.e.
+Archimedian or polynomial) is another important piece of work, outside of the scope of this research, as it has been reported that galaxy arms do not have constant pitch angles (\citealt{1981AJ.....86.1847K}; \citealt{2009MNRAS.397..164R}).
 
@@
   \mathcal{L} = \prod_{\mathrm{arm}=1}^{N}\left(2\pi\sigma_r^2\right)^{-n_\mathrm{arm}/2}
-  \exp\left(\frac{||\widevec{r_\mathrm{arm}}\,-\,\widetilde{r_\mathrm{arm}}||^2}{2\sigma_r^2}\right).
+  \exp\left(-\frac{||\widevec{r_\mathrm{arm}}\,-\,\widetilde{r_\mathrm{arm}}||^2}{2\sigma_r^2}\right).
 \end{equation}

### 2020-06-29 sections/results.tex
makes KLM suggested changes

@@
 where $\sigma_\mathrm{gal}$ is our measure of inter-arm variability of pitch angle and has a posterior distribution of $11.0^\circ\pm 0.9^\circ$.
-This variability is similar to that found by \citet{1981AJ.....86.1847K} and \citet{2014ApJ...790...87D} and emphasises the need for fitting algorithms to not assume all arms have the same pitch angle.
+This inter-arm variability is similar to that found by \citet{1981AJ.....86.1847K} and \citet{2014ApJ...790...87D} and emphasises the need for fitting algorithms to not assume all arms have the same pitch angle.
 The spread of arm pitch angle from the mean galaxy pitch angle can be seen in Figure \ref{fig:arm-pa-spread}, with points colour-coded by the number of arms measured for a galaxy.
@@
 The top panel shows a Gaussian KDE for $E[\phi_\mathrm{gal}]$, and the right panel shows a Gaussian KDE for $E[\phi_\mathrm{arm} - \phi_\mathrm{gal}]$.
-The galaxy pitch angle is consistent with the mean of its arms, with large scatter and a slight bias against values near the lower bound of $0$.}
+The galaxy pitch angle is consistent with the mean of its arms, with large scatter and a slight bias against values near the lower bound of $0$ due to the lower limit applied.}
   \label{fig:arm-pa-spread}
@@
 \subsection{Dependence of pitch angle on Galaxy Morphology}
-\label{section:morphology_comparision}
+\label{section:morphology-comparison}
 
@@
 \subsubsection{Pitch angle vs. Bulge size}
+\label{section:morphology-comparision-bulge}
 
@@
 \subsubsection{Pitch angle vs. Bar Strength}
+\label{section:morphology-comparision-bar}
 
@@
 Confidence intervals are shown, with moving rightwards indicating more confidence in rejecting the null hypothesis that the compared values were drawn from the same parent distribution.
-We cannot reject the null hypotheis at the 1\% level for any of the tests conducted.}
+We cannot reject the null hypotheis at the 1\% level for any of the tests conducted, meaning there is no evidence in this sample that bulge size or bar strength impacts pitch angle.}
   \label{fig:ad-morphology-test}
@@
 The resulting distribution of Anderson-Darling statistics is shown in in the lower panel of Figure \ref{fig:ad-cot-test}.
-This result is highly consistent with the model for spiral winding proposed by \citet{2019arXiv190910291P} and can be seen as evidence that spirals are formed through local disc perturbations, and are primarily governed by local forces.
+This result is highly consistent with the model for spiral winding proposed by \citet{2019arXiv190910291P} and can be interpreted as evidence that spirals are formed through local disc perturbations, and are primarily governed by local forces.
 
@@
   \includegraphics[width=17.7cm]{plots/phi_distribution_comparison.pdf}
-  \caption{The distributions of pitch angles relative to one uniform in $\cot\phi$.
+  \caption{The distributions of pitch angles (orange and blue) relative to one uniform in $\cot\phi$ (black).
 Histograms have been normalised by the area between the limits such that they are comparable.
@@
 Moving rightwards on the x-axis implies greater confidence in rejecting the null hypothesis that the sample was drawn from a distribution uniform in $\cot$ between $15^\circ < \phi < 50.0^\circ$.
-In this instance, we would not be able to reject the null hypothesis at the 1\% level for either $\phi_\mathrm{gal}$ or $\phi_\mathrm{arm}$.}
+In this instance, we would not be able to reject the null hypothesis at the 1\% level for either $\phi_\mathrm{gal}$ or $\phi_\mathrm{arm}$, meaning our sample is consistent with a cot uniform distribution.}
   \label{fig:ad-cot-test}

### 2020-06-29 sections/summary.tex
makes KLM suggested changes

@@
 
-There is no evidence in our data for the link between bulge size and pitch angle predicted by the Hubble sequence and observed in other studies.
+There is no evidence in our data for the link between bulge size and pitch angle predicted by the Hubble sequence and observed in other studies (see Section \ref{section:morphology-comparision-bulge}).
 
-We do not find any link between bar strength and pitch angle suggests that the primary mechanism driving the evolution of the spirals in our sample is not Manifold theory.
+We do not find any link between bar strength and pitch angle suggests that the primary mechanism driving the evolution of the spirals in our sample is not Manifold theory (see Section \ref{section:morphology-comparision-bar}).
 
@@
 Is the spiral arm equally apparent at all pitch angles, or is a selection effect present?
-This result is also not evidenced against QSDW, as it is possible that our distribution of pitch angles is dictated by other factors such as disk shear.
+This result is also not evidence against QSDW, as it is possible that our distribution of pitch angles is dictated by other factors such as disk shear.
 
 In this work, we assume that spiral arms are equally likely to be identified and recovered at all pitch angles.
-This is not an unfair assumption given the amount of human effort that went into obtaining spiral arm measurements (more so than any other pitch angle measurement method, with each galaxy receiving 30 human classifications).
-The galaxy sample used is not guaranteed to be representative of the general spiral population, but is comparable in size to those used in other similar studies (\citealt{2013MNRAS.436.1074S}, \citealt{2019ApJ...871..194Y}, \citealt{2019arXiv190910291P}).
+This is not an unfair assumption given the amount of human effort that went into obtaining spiral arm measurements (more so than any other pitch angle measurement method, with each galaxy receiving at least 30 human classifications).
+The galaxy sample used is not selected to be representative of the general spiral population, but is comparable in size to those used in other similar studies (\citealt{2013MNRAS.436.1074S}, \citealt{2019ApJ...871..194Y}, \citealt{2019arXiv190910291P}) and covers a range of masses and spiral types.
 
 The methodology proposed here is a robust solution to the problems facing investigation of spiral morphology, namely that of reliably identifying spiral arms, and properly accounting for the spread in pitch angles of arms within a galaxy.
-As with most analyses, the most impactful improvement it would be possible to make here would be to increase the quality and volume of data analysed.
-A larger sample would make possible further comparisons, such as splitting galaxies into spiral type (grand design / many-armed / flocculent) and examining the differences between populations.
+This is one of the largest samples for which this test has been done; a larger sample would make possible further comparisons, such as splitting galaxies into spiral type (grand design / many-armed / flocculent) and examining the differences between populations.
 
-The processes governing the formation and evolution of spiral arms are immensely complicated, but the prevalence of spiral galaxies in the Universe, and the spiral nature of our own Milky Way, makes investigating their dynamics of fundamental importance to the scientific aims of understanding, predicting and explaining the nature of the cosmos.
+The processes governing the formation and evolution of spiral arms are complicated, but the prevalence of spiral galaxies in the Universe, their impact for understanding star formation, and the spiral nature of our own Milky Way, makes investigating their dynamics of fundamental importance to the scientific aims of understanding, predicting and explaining the nature of the cosmos.

### 2020-06-29 sections/abstract.tex
final proofread before submission to MNRAS

@@
 Spiral structure is ubiquitous in the Universe, and the pitch angle of arms in spiral galaxies is an important observable in efforts to discriminate between different mechanisms of spiral arm formation and evolution.
-In this paper we present a hierarchical Bayesian approach to galaxy pitch angle determination, using spiral arm data obtained through the \textit{Galaxy Builder} citizen science project \citep{2020arXiv200610450L}.
+In this paper, we present a hierarchical Bayesian approach to galaxy pitch angle determination, using spiral arm data obtained through the \textit{Galaxy Builder} citizen science project \citep{2020arXiv200610450L}.
 This approach better deals with the large variations in pitch angle between arms in a galaxy than existing methods, and obtains full posterior distributions on parameters.
-We utilise fit pitch angles to examine links between bulge and bar strength and pitch angle reported in other work, finding no link in our data.
-We similarly test the simple model of spiral winding proposed by \citet{2019arXiv190910291P}, which assumes the uniformity of the cotangent of pitch angle between some unknown upper and lower limits, with results supporting this model of transient and reccurent spiral pitch angle.
+We utilise fitted pitch angles to examine links between bulge and bar strength and pitch angle reported in other work, finding no correlation in our data.
+We similarly test the simple model of spiral winding proposed by \citet{2019arXiv190910291P}, which assumes the uniformity of the cotangent of pitch angle between some unknown upper and lower limits, with results supporting this model of transient and recurrent spiral pitch angle.

### 2020-06-29 sections/introduction.tex
final proofread before submission to MNRAS

@@
 
-A majority of the population of young stars in a galaxy are located in its spiral arms \citep{2011EAS....51...19E}, and it is possible that they trigger star formation \citep{2013A&A...560A..59C}, though their main role may instead be to sweep up material, promoting the growth of Giant Molecular Clouds.
+A majority of the population of young stars in a galaxy are located in its spiral arms \citep{2011EAS....51...19E}, and they may trigger star formation \citep{2013A&A...560A..59C}, though their main role may instead be to sweep up material, promoting the growth of Giant Molecular Clouds.
 \citep{2014IAUS..298..221D}.
@@
 Studies of spiral morphology have found interesting correlations with other galactic properties, such as a correlation between the tightness of spiral arms and central mass concentration (\citealt{2019ApJ...871..194Y}, though neither \citealt{2017MNRAS.472.2263H} not \citealt{2019MNRAS.487.1808M} found such a relation in large samples); and spiral tightness and rotation curve shape (\citealt{2005MNRAS.359.1065S}, with rising rotation curves creating more open spiral structure).
-These predictions and observations provide compelling reasons for continued investigation of the underlying rules and dynamics of spiral stucture, as doing so is essential for understanding the secular evolution of disc galaxies.
+These predictions and observations provide compelling reasons for continued investigation of the underlying rules and dynamics of spiral structure, as doing so is essential for understanding the secular evolution of disc galaxies.
 
@@
 Many simulations demonstrate that spirals do not maintain a constant tightness (often quantified by pitch angle, the angle between the spiral and the tangent to a circle centred on the galaxy, \citealt{1987gady.book.....B}, illustrated in Figure \ref{fig:pitch-angle-example}), and instead wind-up over time due to the differential rotation of the disc \citep{2013ApJ...763...46B}.
-Recent research suggests that spirals arms are transient in nature, and continually dissipate and re-form \citep{2014PASA...31...35D}.
+Recent research suggests that spirals arms are transient, and continually dissipate and re-form \citep{2014PASA...31...35D}.
 These spirals can be maintained through the same mechanisms that drive QSDW spirals (i.e.
@@
 In this dynamic picture of spiral arms, pitch angle monotonically decreases from a spiral arm's formation to its dissipation.
-\citet{2019arXiv190910291P} propose a simple test of spiral arm winding, assuming the cotangent of the pitch angle of a spiral arm evolves linearly with time.
+\citet{2019arXiv190910291P} propose a simple test of the winding of spiral arms, assuming the cotangent of the pitch angle of a spiral arm evolves linearly with time.
 They found that the distribution of pitch angles of their sample of 86 galaxies was consistent with this prediction, evidence against QSDW theory in favour of the dynamic spirals produced in many simulations.

### 2020-06-29 sections/method.tex
final proofread before submission to MNRAS

@@
 Two-dimensional Fourier methods generally decompose a deprojected image of a galaxy into a superposition of logarithmic spirals between inner and outer annuli \citep{2012ApJS..199...33D} and reports the pitch angle with the highest amplitude as the galaxy's pitch angle.
-\citet{2020MNRAS.493.3854H} combined Fourier analysis of spiral galaxies with visual tracing of spiral arms, succesfully eliminating observed bias in a sample of toy images of galaxies.
+\citet{2020MNRAS.493.3854H} combined Fourier analysis of spiral galaxies with a visual tracing of spiral arms, successfully eliminating observed bias in a sample of toy images of galaxies.
 It is unclear how the variation between pitch angles of individual arms impacts this measurement.
@@
 
-Some galaxies in \citet{2020arXiv200610450L} were shown to volunteers a second time in a repeat \textit{validation subset} in order to create a second aggregate model used to test internal consistency.
-We combine the 30 classifications of galaxies in this \textit{validation subset} with the 30 original classifications.
+Some galaxies in \citet{2020arXiv200610450L} were shown to volunteers a second time in a repeat validation subset to create a second aggregate model used to test internal consistency.
+We combine the 30 classifications of galaxies in this validation subset with the 30 original classifications.
 Clustering of drawn spiral arms and cleaning of points was then performed as detailed in \citet{2020arXiv200610450L}.
@@
 \label{section:bhsm-model}
-In this Section, we lay out our Bayesian hierarchical model for galaxy pitch angle.
+In this section, we lay out our Bayesian hierarchical model for galaxy pitch angle.
 We fit directly to clustered, cleaned points from polylines drawn in \textit{Galaxy Builder}, deprojected and unwrapped to polar coordinates.

### 2020-06-29 sections/results.tex
final proofread before submission to MNRAS

@@
 
-In order to test the possible progenitor distribution of our estimated arm pitch angles, we repeatedly perform an Anderson-Darling test (\citealt{10.2307/2286009}, implemented in \textsc{Scipy}, \citealt{scipy-paper}) over each draw present in the MC trace, resulting in a distribution of Anderson-Darling statistics.
+To test the possible progenitor distribution of our estimated arm pitch angles, we repeatedly perform an Anderson-Darling test (\citealt{10.2307/2286009}, implemented in \textsc{Scipy}, \citealt{scipy-paper}) over each draw present in the MC trace, resulting in a distribution of Anderson-Darling statistics.
 We will refer to this test as the \textit{marginalized Anderson-Darling test}.
@@
 \subsubsection{Pitch angle vs. Bulge size}
-\label{section:morphology-comparision-bulge}
+\label{section:morphology-comparison-bulge}
 
@@
 
-We separate our sample into galaxies with weaker bulges ($B_\mathrm{avg} < 0.28$, 83 galaxies) and those with stronger bulges ($B_\mathrm{avg} \ge 0.28$, 54 galaxies), in order to test whether their pitch angles could be drawn from significantly different distributions.
+We separate our sample into galaxies with weaker bulges ($B_\mathrm{avg} < 0.28$, 83 galaxies) and those with stronger bulges ($B_\mathrm{avg} \ge 0.28$, 54 galaxies), to test whether their pitch angles could be drawn from significantly different distributions.
 A marginalized two-sample Anderson-Darling test comparing the distributions of $\phi_\mathrm{gal}$ for the samples does not find evidence that galaxy pitch angles were drawn from different distributions: we reject the null hypothesis at the 1\% level for only 1\% of the samples.
@@
 \subsubsection{Pitch angle vs. Bar Strength}
-\label{section:morphology-comparision-bar}
+\label{section:morphology-comparison-bar}
 
 One of the predictions of Manifold theory is that pitch angle increases with bar strength \citep{2009MNRAS.400.1706A}.
-In order to investigate this relationship in our data, we make use of Galaxy Zoo 2's bar fraction ($p_\mathrm{bar}$), which has been demonstrated to be a good measure of bar length \citep{Willett2013:1308.3496v2} and bar strength \citep{2012MNRAS.423.1485S,2012MNRAS.424.2180M,2018MNRAS.473.4731K} and therefore a good measure of the torque applied on the disc gas.
+To investigate this relationship in our data, we make use of Galaxy Zoo 2's bar fraction ($p_\mathrm{bar}$), which has been demonstrated to be a good measure of bar length \citep{Willett2013:1308.3496v2} and bar strength \citep{2012MNRAS.423.1485S,2012MNRAS.424.2180M,2018MNRAS.473.4731K} and therefore a good measure of the torque applied on the disc gas.
 
@@
 
-In order to test this theory, \citet{2019arXiv190910291P} used a Kolmogorov-Smirnov test to examine whether a sample of observed galaxy pitch angles was likely to have been drawn from a distribution uniform in its cotangent.
+To test this theory, \citet{2019arXiv190910291P} used a Kolmogorov-Smirnov test to examine whether a sample of observed galaxy pitch angles was likely to have been drawn from a distribution uniform in its cotangent.
 Pitch angles were measured using discrete Fourier transformations in one- and two-dimensions, and as such do not account for inter-arm variations.
@@
 Histograms have been normalised by the area between the limits such that they are comparable.
-The histogram was recalculated with identical bins for each posterior sample of $\phi_\mathrm{gal}$ and $\phi_\mathrm{arm}$, we plot the mean value of each bin, with the sample standard deviation shown as errorbars.
+The histogram was recalculated with identical bins for each posterior sample of $\phi_\mathrm{gal}$ and $\phi_\mathrm{arm}$, we plot the mean value of each bin, with the sample standard deviation shown as error bars.
 It is evident that the distributions are very similar between the chosen limits.}

### 2020-06-29 sections/summary.tex
final proofread before submission to MNRAS

@@
 
-There is no evidence in our data for the link between bulge size and pitch angle predicted by the Hubble sequence and observed in other studies (see Section \ref{section:morphology-comparision-bulge}).
+There is no evidence in our data for the link between bulge size and pitch angle predicted by the Hubble sequence and observed in other studies (see Section \ref{section:morphology-comparison-bulge}).
 
-We do not find any link between bar strength and pitch angle suggests that the primary mechanism driving the evolution of the spirals in our sample is not Manifold theory (see Section \ref{section:morphology-comparision-bar}).
+We do not find any link between bar strength and pitch angle suggests that the primary mechanism driving the evolution of the spirals in our sample is not Manifold theory (see Section \ref{section:morphology-comparison-bar}).
 
@@
 Is the spiral arm equally apparent at all pitch angles, or is a selection effect present?
-This result is also not evidence against QSDW, as it is possible that our distribution of pitch angles is dictated by other factors such as disk shear.
+This result is also not evidence against QSDW, as our distribution of pitch angles may be dictated by other factors such as disk shear.
 

### 2021-01-19 sections/introduction.tex
adds plots, starts work on tracking changes

@@
 \citealt{2014PASA...31...35D}).
-Spirals are as diverse as the theories proposed to govern their evolution, from the quintessential pair of well-defined arcs of the grand design spiral, to the fragmented arm segments of the flocculent spiral, to the disjointed multi-armed spiral.
-These variations on structure account for $18\%$, $50\%$ and $32\%$ of the population respectively (\citealt{2011ApJ...737...32E}; examples of each type are shown in Figure \ref{fig:spiral-galaxy-types}).
-The Hubble classification scheme \citep{1926ApJ....64..321H} and its revisions and expansions \citep{1961hag..book.....S,1991rc3..book.....D} contain detailed variations of different types of spiral galaxy, divided by the presence of a bar and ordered by how obvious spiral arm patterns are, how tightly they are wound and the prominence of a central bulge.
+Spirals are as diverse as the theories proposed to govern their evolution, from the quintessential pair of well-defined arcs of the grand design spiral, to the fragmented arm segments of the flocculent spiral, to the disjointed multi-armed spiral
+\sout{. These variations on structure account for $18\%$, $50\%$ and $32\%$ of the population respectively}
+(\citealt{2011ApJ...737...32E}; examples of each type are shown in Figure \ref{fig:spiral-galaxy-types}).
+The Hubble classification scheme \citep{1926ApJ....64..321H} and its revisions and expansions \citep{1961hag..book.....S,1991rc3..book.....D} contain detailed variations of different types of spiral galaxy, divided by the presence of a bar and ordered by \textbf{the degree of resolution of the spiral arms (how ``patchy” they are)}, how tightly wound they are and the prominence of a central bulge.
 
 \begin{figure*}
-  \includegraphics[width=13cm]{plots/galaxy_types.png}
+  \includegraphics[width=15cm]{plots/galaxy_types.jpg}
   \caption{Examples of the different types of spiral galaxy present in the sky.
-The left column shows grand design spirals, the middle shows many-armed spirals and the right shows flocculent spirals.
+The left column shows the \textbf{grand design spiral NGC 5248}, the middle shows the \textbf{many-armed spiral NGC-3184} and the right shows the \textbf{flocculent spiral NGC-4689}.
 Images were taken with the Sloan Digital Sky Survey Telescope.}
@@
 
-A majority of the population of young stars in a galaxy are located in its spiral arms \citep{2011EAS....51...19E}, and they may trigger star formation \citep{2013A&A...560A..59C}, though their main role may instead be to sweep up material, promoting the growth of Giant Molecular Clouds.
-\citep{2014IAUS..298..221D}.
-This rearrangement of disc gas can lead to the formation of disc-like bulges (commonly called ``pseudobulges''; e.g.
-\citealt{2004ARA&A..42..603K}), which are prevalent in most spiral galaxies \citep{2018MNRAS.473.4731K,2010ApJ...716..942F}.
-Studies of spiral morphology have found interesting correlations with other galactic properties, such as a correlation between the tightness of spiral arms and central mass concentration (\citealt{2019ApJ...871..194Y}, though neither \citealt{2017MNRAS.472.2263H} not \citealt{2019MNRAS.487.1808M} found such a relation in large samples); and spiral tightness and rotation curve shape (\citealt{2005MNRAS.359.1065S}, with rising rotation curves creating more open spiral structure).
+\textbf{
+Whatever kind of spiral is present in a disc galaxy, there is plenty of evidence to suggest that they have a significant role on the overall evolution of that galaxy.
+For example, a ajority of the population of young stars in a galaxy are located in its spiral arms \citep{2011EAS....51...19E}, and there is evidence that spiral arms may trigger star formation \citep{2013A&A...560A..59C} perhaps via their ability to promote the growth of Giant Molecular Clouds \citep{2014IAUS..298..221D}.
+The rearrangement of disc gas and stars driven by spiral arms (e.g.
+\citealt{2018MNRAS.476.1561D}) may lead to the formation of disc-like bulges (commonly called ``pseudobulges''; e.g.
+\citealt{2004ARA&A..42..603K}), which are prevalent in most spiral galaxies, including those without bars \citep{2010ApJ...716..942F}.
+Studies of spiral morphology have also found interesting correlations with other galactic properties, such as a correlation between the tightness of spiral arms and central mass concentration (\citealt{2019ApJ...871..194Y}, though neither \citealt{2017MNRAS.472.2263H} nor \citealt{2019MNRAS.487.1808M} found such a relation in large samples).
+Spiral tightness is also observed to correlate with with rotation curve shape (\citealt{2005MNRAS.359.1065S}, with rising rotation curves creating more open spiral structure).
 These predictions and observations provide compelling reasons for continued investigation of the underlying rules and dynamics of spiral structure, as doing so is essential for understanding the secular evolution of disc galaxies.
+}
 
@@
 One of the fundamental assumptions of early work on spiral formation mechanisms (primarily QSDW) was that the disc of a galaxy, if unstable to spiral perturbations, would create a stable, static wave which would exist unchanged for many rotational periods \citep{1964ApJ...140..646L}.
-The motivation for static waves with small numbers of arms (with a preference for $m=2$) was primarily observational: most disc galaxies show spiral structure with low spiral arm numbers, suggesting that spirals exist for a long time or are continually rebuilt.
+The motivation for static waves with small numbers of arms (with a preference for $m=2$) was primarily observational: most disc galaxies observed at the time showed spiral structure with low spiral arm numbers, suggesting that spirals exist for a long time or are continually rebuilt.
+\textbf{\citep{1982MNRAS.201.1021E} found that flocculent spirals are more prevalent in unbarred, isolated galaxies.
+The presence of a bar, a binary companion or group membership result in a higher fraction of observed grand design spiral patterns.}
 

### 2021-01-19 referee_response.md
Saving submit version, adding referee response file

# Referee Response

> Assistant Editor's Comments:
> Please remove the references from your abstract.

Reviewer's Comments:

Report on the paper "Galaxy Zoo Builder: Morphological Dependence of Spiral
Galaxy Pitch Angle, by Timothy Lingard, Karen L. Masters, Coleman Krawczyk,
Chris Lintott, Sandor Kruk, Brooke SImmons, William Keel; Robert C. Nichol,
and Elisabeth Baeten.

This paper uses data from the Galaxy Zoo project to address the problem of
the pitch angle of spiral arms in disc galaxies. It uses citizen science data
products of the Galaxy Builder project, in particular those on the pitch
angle of spiral galaxies. Interesting results are obtained from data supplied
by volunteers, which eventually should be published. However, the paper is not
yet very satisfactory, and I recommend that a major revision should be made.
I am willing to see the revised version again.

1. Introduction

The introduction shows already some problems: in Figure 1 examples are shown
of grand design spirals, multi-arm spirals and flocculent spirals. However,
it is easy to show that the spirals shown at the right column of Figure 1
are not flocculent, by rotating their image by 180°, and overlaying them on
the original with 50% transparency: this shows that the top-right spiral has
two arms in the central parts, and is multi-armed in the outer parts, while
the bottom-right image is that of a barred spiral with predominantly two arms.

TODO:  change to caption: 
Examples of the different types of spiral galaxy present in the sky. The left column shows grand design spirals with two smooth arms, the middle and right show many-armed spirals - either with smooth arms, or  more flocculent arms. Images were taken with the Sloan Digital Sky Survey Telescope. 
Or just use example galaxies from Elmegreen et al. 2011 (bonus you can put the whirlpool galaxy in your paper). Did you do that already? If so that’s the response to the ref (and put the names of the galaxies in the caption).

Another problem concerns the Hubble classification, which for spirals is aptly
summarized by Sandage (2005), who writes "It was simplicity itself with wide
classification bins that merged seamlessly into one another according to three
criteria of size of the central bulge, the degree of resolution into
condensations (in the language of Reynolds 1920b, 1927a,b), and the openness
of the spiral arms." It is very problematic that the authors think they can
replace the criterion of "the degree of resolution into condensations" by
"how obvious spiral patterns are". Sandage was consistent in his write-ups,
and the same criterion of the degree of resolution can be found in his 1961
"Hubble Atlas of Galaxies". He mentions in his 2005 paper that the Elmegreen
& Elmegreen spiral arm classes are "important for the physics, but need not be
made general in the [classification] notation".

Just change language on Hubble classification. That was a long paragraph to ask for a simple language change:  
OLD: The Hubble classification scheme (Hubble 1926) and its revisions and expansions (Sandage 1961; de Vaucouleurs et al. 1991) contain detailed variations of different types of spiral galaxy, divided by the presence of a bar and ordered by how obvious spiral arm pat- terns are, how tightly they are wound and the prominence of a central bulge. 
NEW: The Hubble classification scheme (Hubble 1926) and its revisions and expansions (Sandage 1961; de Vaucouleurs et al. 1991) contain detailed variations of different types of spiral galaxy, divided by the presence of a bar and ordered by the degree of resolution of the spiral arms (how ``patchy” they are), how tightly wound they are and the prominence of a central bulge.

OK, this was the 1st paragraph in the Introduction. The second paragraph is
hardly better: A first sentence discusses "a majority of the population of
young stars ..." but switches subject to spiral arms: they may trigger star
formation, though their main role maybe sweeping up material ... The second
sentence asserts that "this rearrangement of) disc gas can lead to the formation
of disc-like bulges (... pseudobulges ...), which are prevalent in most spiral
galaxies (Kruk et al. 2018). However, in the latter paper this rearrangement
of gas is attributed to the bar, and not to the spiral arms…

TODO: edit this paragraph. Both bars and spiral arms do restribute gas and stars in galaxies. Kruk 2018 attributes pseudobulge formation to the bar not the spiral arms.

Let me point out that the motivation for the Lin & Shu (1964) paper was to
overcome the 'winding dilemma' posed by differentially rotating spirals such
as M31 and NGC 5055 (Prendergast & Burbidge 1960, Oort 1962), and that the
statement of preferentially m = 2 armed spirals is in direct contradiction
with the numbers given in Figure 1.

Furthermore, the mechanisms of wave amplification are not the same as those
behind the assumption of quasi stationary density wave theory. The term
"swing amplification" was coined by Toomre (1981), even though he relied on
the Goldreich and Lynden-Bell (1965) work.

I **meant** that one way to counteract the short lifetime of stellar density waves is through amplification mechanisms such as WASER and swing amplification, but yes that is incorrect.

What to make of the sentence "Larger bulges and more massive central black
holes have both correlated with more tightly wound spiral arms" (p3, left
column, line 47). They don't correlate anymore?

This was a grammatical error. It has been fixed.

The next paragraph then jumps to "Section 3.2 examines ...", and "Section 3.3
investigates ...", without explaining what is to come in Section 2 and 3.1…

We have fixed this omission.

Section 2.1

The authors say: "We assume that a galaxy has some value for pitch angle,
\phi_{gal}, and that the pitch angles of spiral arms in that galaxy, \phi_{arm},
are constant with radius (giving logarithmic spirals)" This is very unclear.
If the pitch angle is constant with radius, then in all the segments \phi_{i} in
eq. (2) will be the same and \phi_{gal}=\phi_{arm}. So why introduce \phi_{gal}?
Why introduce two different definitions of the pitch angle, if there is only one
value? How does this hypothesis influence the results in this paper?

Be more explicit that different arms can have different pitch angles

On page 3 it is written: "A common assumption  when  measuring  galaxy pitch
angle is that observed spiral arms have a constant pitch angle with radius (e.g.
Davis et al. 2012;.... " This is repated a number of times in other parts of the
paper. Yet this is wrong, as first duscussed by Considere and Athanassoula (who
introduced the 2DFFT method) and as stressed further in Davis et al 2012.

Davis et al. 2012 say that the inclusion of non-logarithmic spirals (with kinks) can negatively impact measurements using 2D FFT.

Davis et al say in their paper: "As pointed out by Considere &  Athanassoula
(1988), this  method does  not  assume  that observed spiral structures are
logarithmic. It only decomposes the observed distributions into a superposition
of logarithmic spirals of different pitch angles and number of arms, which can
be thought of as building blocks." The authors should read the Considere &
Athanassoula paper carefully, as well as the Davis et al (2012) paper and change
accordingly all parts of the paper where they make use of this assumption, or
find a different way to justify it.

The first line of the abstract of Davis et al (2012) is *"A logarithmic spiral is a prominent feature appearing in a majority of observed galaxies"*. What more justification for their use would you want?

Some of the assumptions made in this paper come without much/any justification.
The authors should discuss the effects of each of the assumptions for which they
offer no justification. E.g. is it reasonable to assume that the spread sigma_[gal]
is constant? What do results from 2DFFT give? Some *quantitative* analysis of
results from previous work would be useful here. If this assumption is simply
based on lack of any better knowledge, the effect of this assumption should be
tested numerically and discussed. But the authors could first check the literature
whether there is better knowledge from previous 2DFFT work.

An ANTI-TODO. I don’t think we should do this (apply P2DFFT). “this is definitely valid, and wasn't done originally due to time constraints. I'm not sure how easy it is to apply P2DFFT, but will investigate:

Section 2.2

This section is rather sparse in information. However, there are 198 galaxies
considered, of which 98 galaxies were shown twice to volunteers. In Lingard
et al. 2020, the Galaxy Builder paper - GBpaper for short - the stellar masses
are shown in Figure 4. For 120 galaxies the stellar mass log M_* is between
9.45 and 10.03, and there after there are fewer galaxies up until the upper
mass range of 11.0. Please provide a Figure with the histogram of stellar
masses of the galaxies studied here. This sample is skewed towards galaxies
with lower stellar mass: a quick analysis of the galaxies in Kennicutt (1981)
which are also in S4G shows very few galaxies in the range of log M_* between
9.45 and 10.03. Please indicate what the consequences are of this selection
effect for the general conclusions.

Histogram added (hoping Tim has done this).

Since a sample of 98 galaxies were shown twice to (different?) volunteers, it
is of interest to compare the results for the same galaxies: did the resulting
pitch angles, number of arms, etc. all agree ? One of the data files available
on the GitHub website indicates that this is not the case, but that website is
poorly documented.

TODO: Compare spiral arms for validation subset

From Figure 10 of the GBpaper, it follows that 16 out of 98 galaxies have
p_bar .ge. 0.5, of which 8 out 98 galaxies have p_bar .ge. 0.55, with none of
them exceeding p_bar = 0.75. Thus there are few really barred galaxies in this
sample. This statistics can be repeated for the other sample of 100 galaxies.
This information is necessary to be able to judge how many barred galaxies are
considered. If only 16% of the galaxies are barred, there is not much you can
do to test hypotheses concerning barred galaxies.
 We agree. Our statistical analysis incorporates this (which is one of the reasons we don't find anything significant). No changes made (or asked for).

3. Section 3.1

You report a final sample of 139 galaxies with 261 spiral arms, i.e. less than
2 arms per galaxy. Please give details of why this is so. Please also list how
many of the galaxies are multiarmed, and state the number of arms.

Tim working on this.

In Figure 4 I count about 40 1-armed spirals, 22 3-armed spirals, 14 4-armed
spirals, and about 63 2-armed spirals (more difficult to count). Why are there
so many 1-armed spirals, and why for these, for pitch angles below about 14
degrees, is PHI_arm not equal to PHI_gal ?

Tim says: Probably due to Aggregation. **Need to check**

4. Section 3.2.1

I quote: "Morphological classification commonly links bulge size to spiral
tightness, and such a link is implied by the Hubble Sequence (Sandage 2005,
Gadotti 2009, Buta 2013). Some studies have indeed reported a link between
measured spiral galaxy pitch angle and bulge size (i.e. Hart et al. 2017,
Davis et al. 2019), while others have not found any significant correlation
(Masters et al. 2019).

Now there is a large overlap in authors between Hart et al. (2017) and Masters
et al. (2019): 7 people, 5 of which are also co-authors of this paper. Surely
these 5 people should be able to explain WHY they found no link between pitch
angle and bulge size in one paper, and found a link in another one ??
Answering this is not in the scope of this paper.

Note to point out this is one of the ethical issues with this report.

Sandage (2005) reiterates that there are Sa galaxies with small bulges, so he
knew that there was a tension between the pitch angle criterion and the one
concerning the importance of the bulge. This is well explained in his 1961
Hubble Atlas of Galaxies. Please take this into account in your explanations.

We have added a citation.

5. Section 3.2.2

Here it is asserted that "To investigate this relationship in our data, we make
use of Galaxy Zoo 2's bar fraction (pbar), which has been demonstrated to be a
good measure of bar length (Willett et al. 2013) and bar strength (Skibba et
al. 2012; Masters et al. 2012; Kruk et al. 2018) and therefore a good measure
of the torque applied on the disc gas." In Athanassoula et al. (2009, MNRAS
394, 67) the bar strength is evaluated as the torque at the Lagrangian point:
Q_t,L1. Their figure 7 shows a sketch of the parameter R_D, which is the ratio
of the radius of the orbits in the crest in the outer ring to the radius of the
curve of zero velocity. A proxy could have been constructed from this Figure
for a measurement by citizen scientists by having them measure the radius of
the outer (pseudo-) ring to the radius of the L_4 point in the dark "banana"
part of the image. In any case in their figure 8 a correlation is found between
R_D and the bar strength. This figure is also shown in a different way as
Figure 5 of Athanassoula et al. (2009, MNRAS 400, 1706), the paper which the
authors cite.

Nothing asked for?

The manifold theory could be tested if Q_t,L1 can be suitably either mesured or
replaced by an appropriate proxy. As R_D can not be used (see discussion above)
the authors argue that Galaxy Zoo 2's bar fraction (pbar) is a good measure of
the bar strength. This, however, is not sufficient for the problem in question,
because not any measure of the bar strength may do, as Athanassoula et al. (2009,
MNRAS, 400, 1706) show in their figures 7 and 8, where a correlation is shown
with Qt,L1, but is practically lost if Qb is used instead. This could be because
Qt,L1 contains information on the bar strength, as well as on the bar pattern
speed which is also linked to the bar strength. Since pbar presumably does not
include information on the pattern speed it is possibly a better proxy for Qb
than for Qt,L1. This, however, together with a number of other points discussed
above remains to be shown. The authors could have consulted the volunteers with
in a more appropriate way so as to obtain answers for the question under
consideration. But, as it is, section 3.2.2 can not reach any conclusion. Thus,
either the whole of 3.2.2 should be omitted, or the last paragraph
should be reworded to explain that no conclusion could be reached.

We are reporting what we have asked the volunteers. We thank the referee for these suggestions for any potential future version of the site.  We can either make this more explicit, or try and find a better measurement. This is fair, and we should not be making specific claims regarding Manifold theory (it is still valid that we find no link between pitch angle and our measure of bar strength, which should be reported).

6. Section 3.3.3

Figure 8 has only one panel, so the sentences referring to the upper and lower
panels of this figure should be rephrased.

Corrected.

The range of values of PHI_gal is from 8 to about 50 degrees, and there is no
indication that the lower limit in the pitch angle is poorly determined. Note
that Kennicutt (1981) find a lower limit of 0 degrees. Setting the lower limit
to 15 degrees, so as just to be able to conclude that a uniform cot PHI_gal or
cot PHI_arm distribution cannot be excluded is thus completely arbitrary. It
would be best to display in a second panel the distribution of cot PHI_gal or
cot PHI_arm with a lower limit of 8 degrees, so as to alert the reader to the
strong influence of this assumption.

We believe the supporting text is sufficient to explain this. And have added some directly to the caption.

Obviously, the abstract and the conclusions should be toned down as well,
since a hurried reader would only retain your (possibly erroneous) conclusion.

We have toned down the abstract and conclusions.

7. Section 4

I quote: "We do not find any link between bar strength and pitch angle suggests
that the primary mechanism driving the evolution of the spirals in our sample
is not Manifold theory (see Section 3.2.2). Just in case you reach a conclusion
in that section, and want to mention it in this conclusion section, could you
please formulate this in proper English? With so many native English/American
speakers on the list of authors this ought not to be too much asked, and would
help the reader understand better, as well as save considerable work for the
copy editor.

We have reworded this section. 
This seems unnecessarily rude.

Another quote: "In this work, we assume that spiral arms are equally likely to
be identified and recovered at all pitch angles." OK, but then you should NOT
arbitrarely discard low pitch angle values when testing the Pringle & Dobbs
2019 hypothesis.

We assume no selection bias in galaxy builder results, which is not necessarily valid, however the requirement for a lower limit of pitch angle is needed for the test.

The last two paragraphs contain more advertisements for the work, rather than
Conclusions.

We have reworded this section.

8. Data availibility

The GitHub references refer to poorly documented working directories, and an
interested researcher needs to make a considerable effort to start finding the
basic information. It would be much more helpful to provide the most relevant
information about the only several 100s of galaxies in machine readable tables,
which can be stored at e.g. the CDS.

Tim TODO - checking with him.

To sum up: There are a number of problems with this paper, and a major revision
should be made, so as to rescue the valuable results provided by the volunteers,
and present them in a proper context.

We are happy the referee is convinced our experimental design is sound, and are positive about the annotations obtained. We also agree it needs to be published and presented to fully make use of the valuable results the volunteers provided.

### 2021-01-19 referee_response.md
Finished edits to intro

@@
 
+The authors thank the referee for a careful reading, with useful suggestions for additional citations.
+We respond in line below to specific points.
+Major additions to the text are hi-lighted in bold, while major removals are shown with strike-through.
+
 > Assistant Editor's Comments:
+
 > Please remove the references from your abstract.
 
-Reviewer's Comments:
+This has been done.
 
-Report on the paper "Galaxy Zoo Builder: Morphological Dependence of Spiral
-Galaxy Pitch Angle, by Timothy Lingard, Karen L. Masters, Coleman Krawczyk,
-Chris Lintott, Sandor Kruk, Brooke SImmons, William Keel; Robert C. Nichol,
-and Elisabeth Baeten.
+> Reviewer's Comments:
 
-This paper uses data from the Galaxy Zoo project to address the problem of
-the pitch angle of spiral arms in disc galaxies. It uses citizen science data
-products of the Galaxy Builder project, in particular those on the pitch
-angle of spiral galaxies. Interesting results are obtained from data supplied
-by volunteers, which eventually should be published. However, the paper is not
-yet very satisfactory, and I recommend that a major revision should be made.
+> This paper uses data from the Galaxy Zoo project to address the problem of the pitch angle of spiral arms in disc galaxies.
+It uses citizen science data products of the Galaxy Builder project, in particular those on the pitch angle of spiral galaxies.
+Interesting results are obtained from data supplied by volunteers, which eventually should be published.
+However, the paper is not yet very satisfactory, and I recommend that a major revision should be made.
 I am willing to see the revised version again.
 
-1. Introduction
+We are pleased the referee thinks our results are interesting and should be published.
 
-The introduction shows already some problems: in Figure 1 examples are shown
-of grand design spirals, multi-arm spirals and flocculent spirals. However,
-it is easy to show that the spirals shown at the right column of Figure 1
-are not flocculent, by rotating their image by 180°, and overlaying them on
-the original with 50% transparency: this shows that the top-right spiral has
-two arms in the central parts, and is multi-armed in the outer parts, while
-the bottom-right image is that of a barred spiral with predominantly two arms.
+> 1. Introduction
 
-TODO:  change to caption: 
-Examples of the different types of spiral galaxy present in the sky.
-The left column shows grand design spirals with two smooth arms, the middle and right show many-armed spirals - either with smooth arms, or  more flocculent arms.
-Images were taken with the Sloan Digital Sky Survey Telescope.
-Or just use example galaxies from Elmegreen et al.
-2011 (bonus you can put the whirlpool galaxy in your paper).
-Did you do that already?
-If so that’s the response to the ref (and put the names of the galaxies in the caption).
+> The introduction shows already some problems: in Figure 1 examples are shown of grand design spirals, multi-arm spirals and flocculent spirals.
+However, it is easy to show that the spirals shown at the right column of Figure 1 are not flocculent, by rotating their image by 180°, and overlaying them on the original with 50% transparency: this shows that the top-right spiral has two arms in the central parts, and is multi-armed in the outer parts, while the bottom-right image is that of a barred spiral with predominantly two arms.
 
-Another problem concerns the Hubble classification, which for spirals is aptly
-summarized by Sandage (2005), who writes "It was simplicity itself with wide
-classification bins that merged seamlessly into one another according to three
-criteria of size of the central bulge, the degree of resolution into
-condensations (in the language of Reynolds 1920b, 1927a,b), and the openness
-of the spiral arms." It is very problematic that the authors think they can
-replace the criterion of "the degree of resolution into condensations" by
-"how obvious spiral patterns are". Sandage was consistent in his write-ups,
-and the same criterion of the degree of resolution can be found in his 1961
-"Hubble Atlas of Galaxies". He mentions in his 2005 paper that the Elmegreen
-& Elmegreen spiral arm classes are "important for the physics, but need not be
-made general in the [classification] notation".
+The figure and caption has been updated.
 
-Just change language on Hubble classification. That was a long paragraph to ask for a simple language change:  
-OLD: The Hubble classification scheme (Hubble 1926) and its revisions and expansions (Sandage 1961; de Vaucouleurs et al.
-1991) contain detailed variations of different types of spiral galaxy, divided by the presence of a bar and ordered by how obvious spiral arm pat- terns are, how tightly they are wound and the prominence of a central bulge.
-NEW: The Hubble classification scheme (Hubble 1926) and its revisions and expansions (Sandage 1961; de Vaucouleurs et al.
-1991) contain detailed variations of different types of spiral galaxy, divided by the presence of a bar and ordered by the degree of resolution of the spiral arms (how ``patchy” they are), how tightly wound they are and the prominence of a central bulge.
+> Another problem concerns the Hubble classification, which for spirals is aptly summarized by Sandage (2005), who writes "It was simplicity itself with wide classification bins that merged seamlessly into one another according to three criteria of size of the central bulge, the degree of resolution into condensations (in the language of Reynolds 1920b, 1927a,b), and the openness of the spiral arms." It is very problematic that the authors think they can replace the criterion of "the degree of resolution into condensations" by "how obvious spiral patterns are".
+Sandage was consistent in his write-ups, and the same criterion of the degree of resolution can be found in his 1961 "Hubble Atlas of Galaxies".
+He mentions in his 2005 paper that the Elmegreen & Elmegreen spiral arm classes are "important for the physics, but need not be made general in the [classification] notation".
 
-OK, this was the 1st paragraph in the Introduction. The second paragraph is
-hardly better: A first sentence discusses "a majority of the population of
-young stars ..." but switches subject to spiral arms: they may trigger star
-formation, though their main role maybe sweeping up material ... The second
-sentence asserts that "this rearrangement of) disc gas can lead to the formation
-of disc-like bulges (... pseudobulges ...), which are prevalent in most spiral
-galaxies (Kruk et al. 2018). However, in the latter paper this rearrangement
-of gas is attributed to the bar, and not to the spiral arms…
+We have scaled down the mention of Elmegreen & Elmegreen spiral classes, and now use the traditional language of "degree of resolution" in the description the Hubble spiral sequence.
 
-TODO: edit this paragraph. Both bars and spiral arms do restribute gas and stars in galaxies. Kruk 2018 attributes pseudobulge formation to the bar not the spiral arms.
+> OK, this was the 1st paragraph in the Introduction.
+The second paragraph is hardly better: A first sentence discusses "a majority of the population of young stars ..." but switches subject to spiral arms: they may trigger star formation, though their main role maybe sweeping up material ...
+The second sentence asserts that "this rearrangement of) disc gas can lead to the formation of disc-like bulges (...
+pseudobulges ...), which are prevalent in most spiral galaxies (Kruk et al.
+2018).
+However, in the latter paper this rearrangement of gas is attributed to the bar, and not to the spiral arms…
 
-Let me point out that the motivation for the Lin & Shu (1964) paper was to
-overcome the 'winding dilemma' posed by differentially rotating spirals such
-as M31 and NGC 5055 (Prendergast & Burbidge 1960, Oort 1962), and that the
-statement of preferentially m = 2 armed spirals is in direct contradiction
-with the numbers given in Figure 1.
+There is evidence that both bars and spirals do rearrange disc material - this paragraph is intended to act as a short summary of the impact spirals have on galaxies. It has been reworded.
 
-Furthermore, the mechanisms of wave amplification are not the same as those
-behind the assumption of quasi stationary density wave theory. The term
-"swing amplification" was coined by Toomre (1981), even though he relied on
-the Goldreich and Lynden-Bell (1965) work.
+> Let me point out that the motivation for the Lin & Shu (1964) paper was to overcome the 'winding dilemma' posed by differentially rotating spirals such as M31 and NGC 5055 (Prendergast & Burbidge 1960, Oort 1962), and that the statement of preferentially m = 2 armed spirals is in direct contradiction with the numbers given in Figure 1.
 
-I **meant** that one way to counteract the short lifetime of stellar density waves is through amplification mechanisms such as WASER and swing amplification, but yes that is incorrect.
+We have removed both the mention of m=2, and the numbers from Figure 1. Neither of these are central to the point of our paper.
 
-What to make of the sentence "Larger bulges and more massive central black
-holes have both correlated with more tightly wound spiral arms" (p3, left
-column, line 47). They don't correlate anymore?
+> Furthermore, the mechanisms of wave amplification are not the same as those behind the assumption of quasi stationary density wave theory.
+The term "swing amplification" was coined by Toomre (1981), even though he relied on the Goldreich and Lynden-Bell (1965) work.
 
-This was a grammatical error. It has been fixed.
+We have added a clarification to this point.
 
-The next paragraph then jumps to "Section 3.2 examines ...", and "Section 3.3
-investigates ...", without explaining what is to come in Section 2 and 3.1…
+> What to make of the sentence "Larger bulges and more massive central black holes have both correlated with more tightly wound spiral arms" (p3, left column, line 47). They don't correlate anymore?
+
+This was a grammatical error. Thank you for bring it to our attention. It has been fixed.
+
+> The next paragraph then jumps to "Section 3.2 examines ...", and "Section 3.3 investigates ...", without explaining what is to come in Section 2 and 3.1…
 
@@
 
-Section 2.1
+> Section 2.1
 
-The authors say: "We assume that a galaxy has some value for pitch angle,
-\phi_{gal}, and that the pitch angles of spiral arms in that galaxy, \phi_{arm},
+> The authors say: "We assume that a galaxy has some value for pitch angle, \phi_{gal}, and that the pitch angles of spiral arms in that galaxy, \phi_{arm},
 are constant with radius (giving logarithmic spirals)" This is very unclear.
-If the pitch angle is constant with radius, then in all the segments \phi_{i} in
-eq. (2) will be the same and \phi_{gal}=\phi_{arm}. So why introduce \phi_{gal}?
-Why introduce two different definitions of the pitch angle, if there is only one
-value? How does this hypothesis influence the results in this paper?
+If the pitch angle is constant with radius, then in all the segments \phi_{i} in eq.
+(2) will be the same and \phi_{gal}=\phi_{arm}.
+So why introduce \phi_{gal}?
+Why introduce two different definitions of the pitch angle, if there is only one value?
+How does this hypothesis influence the results in this paper?
 
-Be more explicit that different arms can have different pitch angles
+We have added text to make it clearer that different arms can have different pitch angles, even if in a single arm pitch angle may be constant with radius.
 
-On page 3 it is written: "A common assumption  when  measuring  galaxy pitch
-angle is that observed spiral arms have a constant pitch angle with radius (e.g.
-Davis et al. 2012;.... " This is repated a number of times in other parts of the
-paper. Yet this is wrong, as first duscussed by Considere and Athanassoula (who
-introduced the 2DFFT method) and as stressed further in Davis et al 2012.
+> On page 3 it is written: "A common assumption  when  measuring  galaxy pitch angle is that observed spiral arms have a constant pitch angle with radius (e.g.
+Davis et al.
+2012;....
+" This is repated a number of times in other parts of the paper.
+Yet this is wrong, as first duscussed by Considere and Athanassoula (who introduced the 2DFFT method) and as stressed further in Davis et al 2012.
 

### 2021-01-19 sections/abstract.tex
Finished edits to intro

--- 2020-06-29 sections/abstract.tex
@@
 Spiral structure is ubiquitous in the Universe, and the pitch angle of arms in spiral galaxies is an important observable in efforts to discriminate between different mechanisms of spiral arm formation and evolution.
-In this paper we present a hierarchical Bayesian approach to galaxy pitch angle determination, using spiral arm data obtained through the \textit{Galaxy Builder} citizen science project \citep{2020arXiv200610450L}.
-This approach better deals with the large variations in pitch angle between arms in a galaxy than existing methods, and obtains full posterior distributions on parameters.
-We utilise fit pitch angles to examine links between bulge and bar strength and pitch angle reported in other work, finding no link in our data.
-We similarly test the simple model of spiral winding proposed by \citet{2019arXiv190910291P}, which assumes the uniformity of the cotangent of pitch angle between some unknown upper and lower limits, with results supporting this model of transient and reccurent spiral pitch angle.
+In this paper, we present a hierarchical Bayesian approach to galaxy pitch angle determination, using spiral arm data obtained through the \textit{Galaxy Builder} citizen science project.
+We present a new approach to deal with observed large variations in pitch angle between different arms in a single galaxy, which and obtains full posterior distributions on parameters.
+We make use of our fitted pitch angles to examine previously reported links between bulge and bar strength and pitch angle, finding no correlation in our data.
+We test a recent model for spiral arm winding, which predicts uniformity of the cotangent of pitch angle between some unknown upper and lower limits, with results supporting this model of transient and recurrent spiral pitch angle.

### 2021-01-19 sections/introduction.tex
Finished edits to intro

@@
 The Hubble classification scheme \citep{1926ApJ....64..321H} and its revisions and expansions \citep{1961hag..book.....S,1991rc3..book.....D} contain detailed variations of different types of spiral galaxy, divided by the presence of a bar and ordered by \textbf{the degree of resolution of the spiral arms (how ``patchy” they are)}, how tightly wound they are and the prominence of a central bulge.
+\textbf{Building on this, \citet{1982MNRAS.201.1021E} found that flocculent spirals are more prevalent in unbarred, isolated galaxies.
+The presence of a bar, a binary companion or group membership result in a higher fraction of observed grand design spiral patterns.}
 
@@
 Whatever kind of spiral is present in a disc galaxy, there is plenty of evidence to suggest that they have a significant role on the overall evolution of that galaxy.
-For example, a ajority of the population of young stars in a galaxy are located in its spiral arms \citep{2011EAS....51...19E}, and there is evidence that spiral arms may trigger star formation \citep{2013A&A...560A..59C} perhaps via their ability to promote the growth of Giant Molecular Clouds \citep{2014IAUS..298..221D}.
+For example, a majority of the population of young stars in a galaxy are located in its spiral arms \citep{2011EAS....51...19E}, and there is evidence that spiral arms may trigger star formation \citep{2013A&A...560A..59C} perhaps via their ability to promote the growth of Giant Molecular Clouds \citep{2014IAUS..298..221D}.
 The rearrangement of disc gas and stars driven by spiral arms (e.g.
@@
 Studies of spiral morphology have also found interesting correlations with other galactic properties, such as a correlation between the tightness of spiral arms and central mass concentration (\citealt{2019ApJ...871..194Y}, though neither \citealt{2017MNRAS.472.2263H} nor \citealt{2019MNRAS.487.1808M} found such a relation in large samples).
-Spiral tightness is also observed to correlate with with rotation curve shape (\citealt{2005MNRAS.359.1065S}, with rising rotation curves creating more open spiral structure).
+Spiral tightness is also observed to correlate with with rotation curve shape (\citealt{2005MNRAS.359.1065S}), with rising rotation curves creating more open spiral structure.
 These predictions and observations provide compelling reasons for continued investigation of the underlying rules and dynamics of spiral structure, as doing so is essential for understanding the secular evolution of disc galaxies.
@@
 
-Our current understanding of the mechanisms which drive spiral growth and evolution suggests that each of the different forms of spiral galaxy may be triggered by different processes.
+Our current understanding of the mechanisms which drive spiral growth and evolution suggests that different forms of spiral arms in a galaxy may be triggered by different processes.
 Grand design spirals are thought to have undergone a tidal interaction \citep{2010MNRAS.403..625D,2017ApJ...834....7S}, be driven by a bar (as seen in gas simulations, \citealt{1976ApJ...209...53S,2008A&A...489..115R}, and suggested for stars by Manifold theory, \citealt{2006A&A...453...39R,2009MNRAS.394...67A,2009MNRAS.400.1706A}), or be obeying (quasi-stationary) density wave theory (QSDW theory), in which spiral arms are slowly evolving, ever-present structures in the disc (as first proposed by \citealt{1964ApJ...140..646L}).
@@
 One of the fundamental assumptions of early work on spiral formation mechanisms (primarily QSDW) was that the disc of a galaxy, if unstable to spiral perturbations, would create a stable, static wave which would exist unchanged for many rotational periods \citep{1964ApJ...140..646L}.
-The motivation for static waves with small numbers of arms (with a preference for $m=2$) was primarily observational: most disc galaxies observed at the time showed spiral structure with low spiral arm numbers, suggesting that spirals exist for a long time or are continually rebuilt.
-\textbf{\citep{1982MNRAS.201.1021E} found that flocculent spirals are more prevalent in unbarred, isolated galaxies.
-The presence of a bar, a binary companion or group membership result in a higher fraction of observed grand design spiral patterns.}
+The motivation for static waves with small numbers of arms \sout{(with a preference for $m=2$)} was primarily observational: most disc galaxies observed at the time showed spiral structure with low spiral arm numbers, suggesting that spirals exist for a long time or are continually rebuilt.
+\textbf{This, in combination with theoretical arguments about the ``winding problem", motivated the original static density waves of \citet{1964ApJ...140..646L}, to which swing amplification was added by \citet{Toomre1981}  to provide a way to counteract the short lifetime of stellar density waves.}
 
-Many simulations demonstrate that spirals do not maintain a constant tightness (often quantified by pitch angle, the angle between the spiral and the tangent to a circle centred on the galaxy, \citealt{1987gady.book.....B}, illustrated in Figure \ref{fig:pitch-angle-example}), and instead wind-up over time due to the differential rotation of the disc \citep{2013ApJ...763...46B}.
+More recently, simulations demonstrate that spirals do not maintain a constant tightness (often quantified by pitch angle, the angle between the spiral and the tangent to a circle centred on the galaxy, \citealt{1987gady.book.....B}, illustrated in Figure \ref{fig:pitch-angle-example}), and instead wind-up over time due to the differential rotation of the disc \citep{2013ApJ...763...46B}.
 Recent research suggests that spirals arms are transient, and continually dissipate and re-form \citep{2014PASA...31...35D}.
@@
 In this dynamic picture of spiral arms, pitch angle monotonically decreases from a spiral arm's formation to its dissipation.
-\citet{2019arXiv190910291P} propose a simple test of the winding of spiral arms, assuming the cotangent of the pitch angle of a spiral arm evolves linearly with time.
-They found that the distribution of pitch angles of their sample of 86 galaxies was consistent with this prediction, evidence against QSDW theory in favour of the dynamic spirals produced in many simulations.
+As a particular example of this, \citet{2019arXiv190910291P} propose a simple test of the winding of spiral arms, predicting that the cotangent of the pitch angle of a spiral arm ($\cot \phi$) evolves linearly with time.
+They found that the distribution of pitch angles of their sample of 86 galaxies was consistent with this prediction, which they present as evidence against QSDW theory in favour of the dynamic spirals produced in many simulations.
 
@@
 
-Using Galaxy Zoo 2 data \citep{Willett2013:1308.3496v2} allows the separation of galaxies by the presence and strength of a stellar bar, simulations of gas in barred galaxies often demonstrate that bars can drive long-term spiral evolution \citep{2008A&A...489..115R}, or boost transient spiral structure \citep{2012MNRAS.426..167G}.
+Using Galaxy Zoo 2 data \citep{Willett2013:1308.3496v2} we further separate of galaxies by the presence and strength of a stellar bar.
+This will allow us to test simulations of gas in barred galaxies, which often demonstrate that bars can drive long-term spiral evolution \citep{2008A&A...489..115R}, or boost transient spiral structure \citep{2012MNRAS.426..167G}.
 Manifold theory \citealt{2006A&A...453...39R,2009MNRAS.394...67A,2009MNRAS.400.1706A} is one attempt to determine the orbits of stars in bar-driven spiral arms: it proposes that stars in the vicinity of the unstable Lagrangian points at either end of the bar tend to escape along predictable orbits, governed by invariant manifolds.
@@
 
-Many other components correlate with spiral morphology, including potential ties to bulge fraction (\citealt{1975A&A....44..363Y}, \citealt{2013MNRAS.436.1074S}, \citealt{2019MNRAS.487.1808M}) and black hole mass (\citealt{2008ApJ...678L..93S}, \citealt{2017MNRAS.471.2187D}, \citealt{2019MS&E..571a2118A}).
-Larger bulges and more massive central black holes have both correlated with more tightly wound spiral arms.
+Many other galactic components may correlate with spiral morphology, including bulge fraction (\citealt{1975A&A....44..363Y}, \citealt{2013MNRAS.436.1074S}, \citealt{2019MNRAS.487.1808M}) and black hole mass (\citealt{2008ApJ...678L..93S}, \citealt{2017MNRAS.471.2187D}, \citealt{2019MS&E..571a2118A}).
+Larger bulges and more massive central black holes have both \textbf{been observed to} correlate with more tightly wound spiral arms.
+\textbf{We can test this with \textit{Galaxy Builder}.}
 
-Section \ref{section:morphology-comparison} examines the correlation between pitch angle and bulge size implied by the Hubble sequence, and pitch angle and bar strength implied by Manifold theory.
+\textbf{This paper is structured as follows: In Section \ref{section:method} we introduce methods to measure galaxy pitch angle, and present our sample, and our Bayseian hierarchical modeling method making use of \textit{Galaxy Builder} to estimate galaxy and population pitch angles.
+Section \ref{section:constraints-on-galaxy-phi} presents our general constraints on pitch angles in our section, while} Section \ref{section:morphology-comparison} examines the correlation between pitch angle and bulge size implied by the Hubble sequence, and pitch angle and bar strength implied by Manifold theory.
 Section \ref{section:spiral_winding} investigates spiral arm winding using the test derived by \cite{2019arXiv190910291P} (uniformity of galaxy pitch angle in $\cot\;\phi$).
-
-Where necessary, we make use of $H_0 = 70\ \text{km}\ \text{s}^{-1}\ \text{Mpc}^{-1}$.
+\textbf{We provide a summary and conclusions in Section \ref{section:summary}.}  Where necessary, we make use of $H_0 = 70\ \text{km}\ \text{s}^{-1}\ \text{Mpc}^{-1}$.

### 2021-01-19 sections/summary.tex
Finished edits to intro

@@
 % !TEX root = ../main.tex
+\label{section:summary}
 This paper presents a new Bayesian approach to estimate galaxy pitch angle, making use of citizen science results to measure spiral arms through photometric modelling.

### 2021-01-19 referee_response.md
Text edits to most of the rest of the referees points

@@
 
-Davis et al. 2012 say that the inclusion of non-logarithmic spirals (with kinks) can negatively impact measurements using 2D FFT.
+> Davis et al say in their paper: "As pointed out by Considere &  Athanassoula (1988), this  method does  not  assume  that observed spiral structures are logarithmic.
+It only decomposes the observed distributions into a superposition of logarithmic spirals of different pitch angles and number of arms, which can be thought of as building blocks." The authors should read the Considere & Athanassoula paper carefully, as well as the Davis et al (2012) paper and change accordingly all parts of the paper where they make use of this assumption, or find a different way to justify it.
 
-Davis et al say in their paper: "As pointed out by Considere &  Athanassoula
-(1988), this  method does  not  assume  that observed spiral structures are
-logarithmic. It only decomposes the observed distributions into a superposition
-of logarithmic spirals of different pitch angles and number of arms, which can
-be thought of as building blocks." The authors should read the Considere &
-Athanassoula paper carefully, as well as the Davis et al (2012) paper and change
-accordingly all parts of the paper where they make use of this assumption, or
-find a different way to justify it.
+We thank the referee for bringing Considere & Athanassoula 1988 back to our attention.
+We have reviewed the paper again, and added a citation.
+In that paper it is very clear the method does not assume logarithmic spirals.
+However the first line of the abstract of Davis et al (2012) is *"A logarithmic spiral is a prominent feature appearing in a majority of observed galaxies"*, and they clearly say (Section 4.3.2) that the inclusion of non-logarithmic spirals (with significant variation of pitch angle with radius) can negatively impact measurements using 2DFFT.
+They also say "In general, logarithmic spirals are good approximations of the shape of galactic spiral arms".
+We think there is evidence in the literature that the assumption of logarithmic spirals is widespread, even if it's clear this method can in principle discover radial variation in phi.
 
-The first line of the abstract of Davis et al (2012) is *"A logarithmic spiral is a prominent feature appearing in a majority of observed galaxies"*.
-What more justification for their use would you want?
+> Some of the assumptions made in this paper come without much/any justification.
+The authors should discuss the effects of each of the assumptions for which they offer no justification.
+E.g.
+is it reasonable to assume that the spread sigma_[gal] is constant?
+What do results from 2DFFT give?
+Some *quantitative* analysis of results from previous work would be useful here.
+If this assumption is simply based on lack of any better knowledge, the effect of this assumption should be tested numerically and discussed.
+But the authors could first check the literature whether there is better knowledge from previous 2DFFT work.
 
-Some of the assumptions made in this paper come without much/any justification.
-The authors should discuss the effects of each of the assumptions for which they
-offer no justification. E.g. is it reasonable to assume that the spread sigma_[gal]
-is constant? What do results from 2DFFT give? Some *quantitative* analysis of
-results from previous work would be useful here. If this assumption is simply
-based on lack of any better knowledge, the effect of this assumption should be
-tested numerically and discussed. But the authors could first check the literature
-whether there is better knowledge from previous 2DFFT work.
+We have added more discussion of the assumptions which were made.
+The specific question about sigma_[gal] is interesting.
+We have found two previous attempts to measure this, which did not note any significant variation between galaxies, however sample sizes are small.
+We note this as an interesting followup question for larger samples.
 
-An ANTI-TODO.
-I don’t think we should do this (apply P2DFFT).
-“this is definitely valid, and wasn't done originally due to time constraints.
-I'm not sure how easy it is to apply P2DFFT, but will investigate:
+*** TODO: Check more if anyone has measured typical inter-arm variation in pitch angle ***
 
-Section 2.2
+> Section 2.2
 
-This section is rather sparse in information. However, there are 198 galaxies
-considered, of which 98 galaxies were shown twice to volunteers. In Lingard
-et al. 2020, the Galaxy Builder paper - GBpaper for short - the stellar masses
-are shown in Figure 4. For 120 galaxies the stellar mass log M_* is between
-9.45 and 10.03, and there after there are fewer galaxies up until the upper
-mass range of 11.0. Please provide a Figure with the histogram of stellar
-masses of the galaxies studied here. This sample is skewed towards galaxies
-with lower stellar mass: a quick analysis of the galaxies in Kennicutt (1981)
-which are also in S4G shows very few galaxies in the range of log M_* between
-9.45 and 10.03. Please indicate what the consequences are of this selection
-effect for the general conclusions.
+> This section is rather sparse in information.
+However, there are 198 galaxies considered, of which 98 galaxies were shown twice to volunteers.
+In Lingard et al.
+2020, the Galaxy Builder paper - GBpaper for short - the stellar masses are shown in Figure 4.
+For 120 galaxies the stellar mass log M_* is between 9.45 and 10.03, and there after there are fewer galaxies up until the upper mass range of 11.0.
+Please provide a Figure with the histogram of stellar masses of the galaxies studied here.
+This sample is skewed towards galaxies with lower stellar mass: a quick analysis of the galaxies in Kennicutt (1981) which are also in S4G shows very few galaxies in the range of log M_* between 9.45 and 10.03.
+Please indicate what the consequences are of this selection effect for the general conclusions.
 
-Histogram added (hoping Tim has done this).
+We have added a histogram, and an investigation of how pitch angle varies with stellar mass in Section 3.1.
 
-Since a sample of 98 galaxies were shown twice to (different?) volunteers, it
-is of interest to compare the results for the same galaxies: did the resulting
-pitch angles, number of arms, etc. all agree ? One of the data files available
-on the GitHub website indicates that this is not the case, but that website is
-poorly documented.
+> Since a sample of 98 galaxies were shown twice to (different?) volunteers, it is of interest to compare the results for the same galaxies: did the resulting pitch angles, number of arms, etc.
+all agree ?
+One of the data files available on the GitHub website indicates that this is not the case, but that website is poorly documented.
 
-TODO: Compare spiral arms for validation subset
+*** TODO: Compare spiral arms for validation subset ***
 
-From Figure 10 of the GBpaper, it follows that 16 out of 98 galaxies have
-p_bar .ge. 0.5, of which 8 out 98 galaxies have p_bar .ge. 0.55, with none of
-them exceeding p_bar = 0.75. Thus there are few really barred galaxies in this
-sample. This statistics can be repeated for the other sample of 100 galaxies.
-This information is necessary to be able to judge how many barred galaxies are
-considered. If only 16% of the galaxies are barred, there is not much you can
-do to test hypotheses concerning barred galaxies.
+> From Figure 10 of the GBpaper, it follows that 16 out of 98 galaxies have p_bar .ge.
+0.5, of which 8 out 98 galaxies have p_bar .ge.
+0.55, with none of them exceeding p_bar = 0.75.
+Thus there are few really barred galaxies in this sample.
+This statistics can be repeated for the other sample of 100 galaxies.
+This information is necessary to be able to judge how many barred galaxies are considered.
+If only 16% of the galaxies are barred, there is not much you can do to test hypotheses concerning barred galaxies.
+
 We agree. Our statistical analysis incorporates this (which is one of the reasons we don't find anything significant). No changes made (or asked for).
 
-3. Section 3.1
+> 3. Section 3.1
 
-You report a final sample of 139 galaxies with 261 spiral arms, i.e. less than
-2 arms per galaxy. Please give details of why this is so. Please also list how
-many of the galaxies are multiarmed, and state the number of arms.
+> You report a final sample of 139 galaxies with 261 spiral arms, i.e.
+less than 2 arms per galaxy.
+Please give details of why this is so.
+Please also list how many of the galaxies are multiarmed, and state the number of arms.
 
-Tim working on this.
+** Tim working on this. **
 
-In Figure 4 I count about 40 1-armed spirals, 22 3-armed spirals, 14 4-armed
-spirals, and about 63 2-armed spirals (more difficult to count). Why are there
-so many 1-armed spirals, and why for these, for pitch angles below about 14
-degrees, is PHI_arm not equal to PHI_gal ?
+> In Figure 4 I count about 40 1-armed spirals, 22 3-armed spirals, 14 4-armed spirals, and about 63 2-armed spirals (more difficult to count).
+Why are there so many 1-armed spirals, and why for these, for pitch angles below about 14 degrees, is PHI_arm not equal to PHI_gal ?
 
-Tim says: Probably due to Aggregation. **Need to check**
+** Tim says: Probably due to Aggregation. Need to check this **
 
-4. Section 3.2.1
+> 4. Section 3.2.1
 
-I quote: "Morphological classification commonly links bulge size to spiral
-tightness, and such a link is implied by the Hubble Sequence (Sandage 2005,
-Gadotti 2009, Buta 2013). Some studies have indeed reported a link between
-measured spiral galaxy pitch angle and bulge size (i.e. Hart et al. 2017,
-Davis et al. 2019), while others have not found any significant correlation
-(Masters et al. 2019).
+> I quote: "Morphological classification commonly links bulge size to spiral tightness, and such a link is implied by the Hubble Sequence (Sandage 2005, Gadotti 2009, Buta 2013).
+Some studies have indeed reported a link between measured spiral galaxy pitch angle and bulge size (i.e.
+Hart et al.
+2017, Davis et al.
+2019), while others have not found any significant correlation (Masters et al.
+2019).
 
-Now there is a large overlap in authors between Hart et al. (2017) and Masters
-et al. (2019): 7 people, 5 of which are also co-authors of this paper. Surely
-these 5 people should be able to explain WHY they found no link between pitch
-angle and bulge size in one paper, and found a link in another one ??
-Answering this is not in the scope of this paper.
+> Now there is a large overlap in authors between Hart et al.
+(2017) and Masters et al.
+(2019): 7 people, 5 of which are also co-authors of this paper.
+Surely these 5 people should be able to explain WHY they found no link between pitch angle and bulge size in one paper, and found a link in another one ??
 
-Note to point out this is one of the ethical issues with this report.
+This isn't really in the scope of this paper.
+The Hart et al.
+(2017) result really showed that pitch angle depended on disc mass more than bulge size (but more massive discs tend to have larger bulges.
+In the interest of keeping to the point of this paper, we have removed that citation, but also note that these correlation results can depend on both how bulge size and pitch angles are measured.
 
-Sandage (2005) reiterates that there are Sa galaxies with small bulges, so he
-knew that there was a tension between the pitch angle criterion and the one
-concerning the importance of the bulge. This is well explained in his 1961
-Hubble Atlas of Galaxies. Please take this into account in your explanations.
+> Sandage (2005) reiterates that there are Sa galaxies with small bulges, so he knew that there was a tension between the pitch angle criterion and the one concerning the importance of the bulge.
+This is well explained in his 1961 Hubble Atlas of Galaxies.
+Please take this into account in your explanations.
 
-We have added a citation.
+We have added a citation. Thanks.
 
-5. Section 3.2.2
+> 5. Section 3.2.2
 
-Here it is asserted that "To investigate this relationship in our data, we make
-use of Galaxy Zoo 2's bar fraction (pbar), which has been demonstrated to be a
-good measure of bar length (Willett et al. 2013) and bar strength (Skibba et
-al. 2012; Masters et al. 2012; Kruk et al. 2018) and therefore a good measure
-of the torque applied on the disc gas." In Athanassoula et al. (2009, MNRAS
-394, 67) the bar strength is evaluated as the torque at the Lagrangian point:
-Q_t,L1. Their figure 7 shows a sketch of the parameter R_D, which is the ratio
-of the radius of the orbits in the crest in the outer ring to the radius of the
-curve of zero velocity. A proxy could have been constructed from this Figure
-for a measurement by citizen scientists by having them measure the radius of
-the outer (pseudo-) ring to the radius of the L_4 point in the dark "banana"
-part of the image. In any case in their figure 8 a correlation is found between
-R_D and the bar strength. This figure is also shown in a different way as
-Figure 5 of Athanassoula et al. (2009, MNRAS 400, 1706), the paper which the
-authors cite.
+> Here it is asserted that "To investigate this relationship in our data, we make use of Galaxy Zoo 2's bar fraction (pbar), which has been demonstrated to be a good measure of bar length (Willett et al.
+2013) and bar strength (Skibba et al.
+2012; Masters et al.
+2012; Kruk et al.
+2018) and therefore a good measure of the torque applied on the disc gas." In Athanassoula et al.
+(2009, MNRAS 394, 67) the bar strength is evaluated as the torque at the Lagrangian point: Q_t,L1.
+Their figure 7 shows a sketch of the parameter R_D, which is the ratio of the radius of the orbits in the crest in the outer ring to the radius of the curve of zero velocity.
+A proxy could have been constructed from this Figure for a measurement by citizen scientists by having them measure the radius of the outer (pseudo-) ring to the radius of the L_4 point in the dark "banana" part of the image.
+In any case in their figure 8 a correlation is found between R_D and the bar strength.
+This figure is also shown in a different way as Figure 5 of Athanassoula et al.
+(2009, MNRAS 400, 1706), the paper which the authors cite.
 
-Nothing asked for?
-
-The manifold theory could be tested if Q_t,L1 can be suitably either mesured or
-replaced by an appropriate proxy. As R_D can not be used (see discussion above)
-the authors argue that Galaxy Zoo 2's bar fraction (pbar) is a good measure of
-the bar strength. This, however, is not sufficient for the problem in question,
-because not any measure of the bar strength may do, as Athanassoula et al. (2009,
-MNRAS, 400, 1706) show in their figures 7 and 8, where a correlation is shown
-with Qt,L1, but is practically lost if Qb is used instead. This could be because
-Qt,L1 contains information on the bar strength, as well as on the bar pattern
-speed which is also linked to the bar strength. Since pbar presumably does not
-include information on the pattern speed it is possibly a better proxy for Qb
-than for Qt,L1. This, however, together with a number of other points discussed
-above remains to be shown. The authors could have consulted the volunteers with
-in a more appropriate way so as to obtain answers for the question under
-consideration. But, as it is, section 3.2.2 can not reach any conclusion. Thus,
-either the whole of 3.2.2 should be omitted, or the last paragraph
-should be reworded to explain that no conclusion could be reached.
+> The manifold theory could be tested if Q_t,L1 can be suitably either mesured or replaced by an appropriate proxy.
+As R_D can not be used (see discussion above) the authors argue that Galaxy Zoo 2's bar fraction (pbar) is a good measure of the bar strength.
+This, however, is not sufficient for the problem in question, because not any measure of the bar strength may do, as Athanassoula et al.
+(2009, MNRAS, 400, 1706) show in their figures 7 and 8, where a correlation is shown with Qt,L1, but is practically lost if Qb is used instead.
+This could be because Qt,L1 contains  nformation on the bar strength, as well as on the bar pattern speed which is also linked to the bar strength.
+Since pbar presumably does not include information on the pattern speed it is possibly a better proxy for Qb than for Qt,L1.
+This, however, together with a number of other points discussed above remains to be shown.
+The authors could have consulted the volunteers with in a more appropriate way so as to obtain answers for the question under consideration.
+But, as it is, section 3.2.2 can not reach any conclusion.
+Thus, either the whole of 3.2.2 should be omitted, or the last paragraph should be reworded to explain that no conclusion could be reached.
 
@@
 We thank the referee for these suggestions for any potential future version of the site.
-We can either make this more explicit, or try and find a better measurement.
-This is fair, and we should not be making specific claims regarding Manifold theory (it is still valid that we find no link between pitch angle and our measure of bar strength, which should be reported).
+We agree we cannot make any strong claims regarding Manifold theory, however it is still valid that we find no link between pitch angle and our measure of bar strength, which we feel should be reported.
+Edits have been made to the text.
 
-6. Section 3.3.3
+> 6. Section 3.3.3
 
-Figure 8 has only one panel, so the sentences referring to the upper and lower
-panels of this figure should be rephrased.
+> Figure 8 has only one panel, so the sentences referring to the upper and lower panels of this figure should be rephrased.
 
-Corrected.
+This error has been corrected.
 
-The range of values of PHI_gal is from 8 to about 50 degrees, and there is no
-indication that the lower limit in the pitch angle is poorly determined. Note
-that Kennicutt (1981) find a lower limit of 0 degrees. Setting the lower limit
-to 15 degrees, so as just to be able to conclude that a uniform cot PHI_gal or
-cot PHI_arm distribution cannot be excluded is thus completely arbitrary. It
-would be best to display in a second panel the distribution of cot PHI_gal or
-cot PHI_arm with a lower limit of 8 degrees, so as to alert the reader to the
-strong influence of this assumption.
+> The range of values of PHI_gal is from 8 to about 50 degrees, and there is no indication that the lower limit in the pitch angle is poorly determined.
+Note that Kennicutt (1981) find a lower limit of 0 degrees.
+Setting the lower limit to 15 degrees, so as just to be able to conclude that a uniform cot PHI_gal or cot PHI_arm distribution cannot be excluded is thus completely arbitrary.
+It would be best to display in a second panel the distribution of cot PHI_gal or cot PHI_arm with a lower limit of 8 degrees, so as to alert the reader to the strong influence of this assumption.
 
-We believe the supporting text is sufficient to explain this. And have added some directly to the caption.
+*** Do we though??? The should think about this point *** We believe the supporting text is sufficient to explain this. And have added some directly to the caption.
 
-Obviously, the abstract and the conclusions should be toned down as well,
-since a hurried reader would only retain your (possibly erroneous) conclusion.
+> Obviously, the abstract and the conclusions should be toned down as well, since a hurried reader would only retain your (possibly erroneous) conclusion.
 
-We have toned down the abstract and conclusions.
+** TDOD: *** We have toned down the abstract and conclusions.
 
-7. Section 4
+> 7. Section 4
 
-I quote: "We do not find any link between bar strength and pitch angle suggests
-that the primary mechanism driving the evolution of the spirals in our sample
-is not Manifold theory (see Section 3.2.2). Just in case you reach a conclusion
-in that section, and want to mention it in this conclusion section, could you
-please formulate this in proper English? With so many native English/American
-speakers on the list of authors this ought not to be too much asked, and would
-help the reader understand better, as well as save considerable work for the
-copy editor.
+> I quote: "We do not find any link between bar strength and pitch angle suggests that the primary mechanism driving the evolution of the spirals in our sample is not Manifold theory (see Section 3.2.2).
+Just in case you reach a conclusion in that section, and want to mention it in this conclusion section, could you please formulate this in proper English?
+With so many native English/American speakers on the list of authors this ought not to be too much asked, and would help the reader understand better, as well as save considerable work for the copy editor.
 
 We have reworded this section.
-This seems unnecessarily rude.
 
-Another quote: "In this work, we assume that spiral arms are equally likely to
-be identified and recovered at all pitch angles." OK, but then you should NOT
-arbitrarely discard low pitch angle values when testing the Pringle & Dobbs
-2019 hypothesis.
+> Another quote: "In this work, we assume that spiral arms are equally likely to be identified and recovered at all pitch angles." OK, but then you should NOT arbitrarely discard low pitch angle values when testing the Pringle & Dobbs 2019 hypothesis.
 
@@
 
-The last two paragraphs contain more advertisements for the work, rather than
-Conclusions.
+> The last two paragraphs contain more advertisements for the work, rather than Conclusions.
 
@@
 
-8. Data availibility
+> 8. Data availibility
 
-The GitHub references refer to poorly documented working directories, and an
-interested researcher needs to make a considerable effort to start finding the
-basic information. It would be much more helpful to provide the most relevant
-information about the only several 100s of galaxies in machine readable tables,
-which can be stored at e.g. the CDS.
+> The GitHub references refer to poorly documented working directories, and an interested researcher needs to make a considerable effort to start finding the basic information.
+It would be much more helpful to provide the most relevant information about the only several 100s of galaxies in machine readable tables, which can be stored at e.g.
+the CDS.
 
-Tim TODO - checking with him.
+*** Tim TODO - might just need a better README ***
 
-To sum up: There are a number of problems with this paper, and a major revision
-should be made, so as to rescue the valuable results provided by the volunteers,
-and present them in a proper context.
+>To sum up: There are a number of problems with this paper, and a major revision should be made, so as to rescue the valuable results provided by the volunteers, and present them in a proper context.
 
 We are happy the referee is convinced our experimental design is sound, and are positive about the annotations obtained.
-We also agree it needs to be published and presented to fully make use of the valuable results the volunteers provided.
+We are happy to agree this work should be published and presented to fully make use of the valuable results the volunteers provided.

### 2021-01-19 sections/method.tex
Finished edits to intro
Text edits to most of the rest of the referees points

@@
 % !TEX root = ../main.tex
+\label{section:method}
+
 \subsection{Measuring galaxy pitch angle} 
@@
 A common assumption when measuring galaxy pitch angle is that observed spiral arms have a constant pitch angle with radius (e.g.
-\citealt{2012ApJS..199...33D,2013MNRAS.436.1074S,2014ApJ...790...87D}.
+\citealt{2012ApJS..199...33D,2013MNRAS.436.1074S,2014ApJ...790...87D}).
 Spirals of this kind are known as logarithmic spirals and are described by
@@
 where $\phi$ is the arm's pitch angle, $A$ is an amplitude coefficient and $\theta$ is the polar coordinate.
+\textbf{Different arms in a galaxy could have different values of $\phi$, however for each arm, $\phi$ is assumed to constant with radius}.
 One method used to obtain a pitch angle of a galaxy is to fit logarithmic spirals to individually identified arm segments and take the weighted mean of their pitch angles (which often vary by upwards of $10^\circ$, \citealt{2014ApJ...790...87D}).
@@
 
-Fourier analysis in one- and two-dimensions (as performed by \citealt{2019arXiv190804246D}, \citealt{2012ApJS..199...33D}, \citealt{2018MNRAS.474.2594M}) is another widely used method of computationally obtaining galaxy pitch angles.
+Fourier analysis in one- and two-dimensions (as performed by \citealt{2019arXiv190804246D}, \citealt{2012ApJS..199...33D}, \citealt{2018MNRAS.474.2594M}, \textbf{and dating back to the seminal work of \citealt{ConsidereAthanassoula1988}}) is another widely used method of computationally obtaining galaxy pitch angles.
 Two-dimensional Fourier methods generally decompose a deprojected image of a galaxy into a superposition of logarithmic spirals between inner and outer annuli \citep{2012ApJS..199...33D} and reports the pitch angle with the highest amplitude as the galaxy's pitch angle.
@@
 It is unclear how the variation between pitch angles of individual arms impacts this measurement.
+\textbf{We note that while this method is able to model non-logarithmic spirals - as a sum of logarithmic spirals with differing pitch angles, most applications of the model favour picking regions of the galaxy in which the pitch angle is constant with radius - e.g.
+see Section 4.3.2 of \citealt{2012ApJS..199...33D}.}
 
@@
 These are a subset of the \textit{stellar mass-complete sample} in \citet{2017MNRAS.472.2263H}, a sample of low-redshift ($0.02 < z < 0.055$) face-on spiral galaxies selected using data from the NASA-Sloan Atlas \citep{2011AJ....142...31B} and Galaxy Zoo 2 \citep{Willett2013:1308.3496v2}.
-The \textit{stellar mass-complete sample} ranged in stellar mass from $9.45 < \log(M_* / M_\odot) < 11.05$.
+The \textit{stellar mass-complete sample} ranged in stellar mass from $9.45 < \log(M_* / M_\odot) < 11.05$, with most of the sample between $9.5 < \log(M_* / M_\odot) < 10.0$.
+A histogram of stellar masses can be see in Section \ref{section:constraints-on-galaxy-phi}, where variation with stellar mass, to check the impact of this limited mass range, is also investigated.
 
@@
 
-We wish to utilize the logarithmic spiral's desirable properties of a constant pitch angle and a small number of free parameters, therefore, we make use of it here without an explicit comparison to other models.
+Logarithmic spirals have the desirable properties of a constant pitch angle and a small number of free parameters.
+\textbf{For this first analysis of the \textit{Galaxy Builder} models we choose to} make use of it here without an explicit comparison to other models.
 A simple visual inspection of the fitted logarithmic spirals suggests that it is an appropriate model, however, a comparison of a logarithmic spiral profile to other spiral forms (i.e.
@@
 
-We assume that a galaxy has some value for pitch angle, $\phi_\mathrm{gal}$, and that the pitch angles of spiral arms in that galaxy, $\phi_\mathrm{arm}$, are constant with radius (giving logarithmic spirals) and drawn from a normal distribution centred on $\phi_\mathrm{gal}$, with some spread $\sigma_\mathrm{gal}$ common to all galaxies.
-We truncate this normal distribution between the physical limits of {0\degree} (a ring) and {90\degree} (a ``spoke''), giving
+\textbf{As suggested by spiral formation models which correlate a galaxy wide pitch angle, with galaxy wide properties,} we will assume that a give galaxy has some preferred value for arm pitch angle, $\phi_\mathrm{gal}$, and that the pitch angles of spiral arms in that galaxy, $\phi_\mathrm{arm}$, are constant with radius (giving logarithmic spirals) and drawn from a normal distribution centred on $\phi_\mathrm{gal}$, with some spread $\sigma_\mathrm{gal}$ common to all galaxies.
+We truncate the normal distribution of spiral arm pitch angles in a single galaxy between the physical limits of {0\degree} (a ring) and {90\degree} (a ``spoke''), giving
 
@@
 
-The choice to assume all galaxies show the same inter-arm variation in pitch angle (represented by a common value of $\sigma_\mathrm{gal}$ across all galaxies) was motivated by our small sample size and the low number of arms measured per galaxy.
+\textbf{The choice to assume all galaxies show the same inter-arm variation in pitch angle (represented by a common value of $\sigma_\mathrm{gal}$ across all galaxies) was motivated by our small sample size and the low number of arms measured per galaxy.
+With this sample size we do not find, nor expect to be sensitive to variations in this parameter.
+It is possible that it does vary between galaxies, and that this variation is physically interesting.
+At least \citet{1981AJ.....86.1847K} and \citet{2014ApJ...790...87D} have previously made attempts to measure this parameter.
+\citet{1981AJ.....86.1847K} fit logarithmic spirals to 113 nearby (NGC) galaxies, and note the dominant error in average galaxy pitch angle comes from interarm variation, which they measure to have an average value of 5$^\circ$.
+\citet{2014ApJ...790...87D} is primarily a machine learning method paper, and details on the galaxies investigated are not clear, but their Table 1 presents the median difference in pitch angle between pairs of arms with different lengths, which varies from 14.5$^\circ$ in very short arms, to 2.6$^\circ$ in the longest traced arms.
+It is unclear how much this encodes error in their method versus real variation in the galaxy population.
+Futher investigation of this issue in the \textit{Galaxy Builder} sample would be an interesting follow-up project.}
 
@@
 
-The inverse gamma distribution is used to aid the convergence of the Hamiltonian Monte Carlo (HMC) algorithm used (discussed later).
+\textbf{These are conservative priors, which are not expected to have significant impact of the results.} The inverse gamma distribution is used to aid the convergence of the Hamiltonian Monte Carlo (HMC) algorithm used (discussed later).
 The Cauchy distribution is equivalent to the Student's t-distribution with one degree of freedom, and was chosen due to its fatter tails than the normal distribution.

### 2021-01-19 sections/results.tex
Text edits to most of the rest of the referees points

@@
 
+\textbf{In Figure \ref{fig:stellar-mass-phigal} we present the stellar mass distribution of the sample, and investigate how the global galaxy pitch angle depends on this parameter.
+The majority of our sample has stellar masses $9.5 < \log(M_* / M_\odot) < 10.0$, and galaxy average pitch angles 10$^\circ$--20$^\circ$.
+This should be remembered in our physical interpretation of other results.
+We observe no significant trend of the global pitch angle with mass, although there is a hint that more massive spirals may on average have more tightly wound arms.
+}
+
+\begin{figure}
+  \includegraphics[width=9cm]{plots/stellar_mass_phigal_distribution.pdf}
+  \caption{The stellar mass distribution of the sample and the galaxy average pitch angle distribution shown as a 2D histogram (centre) and also projected along each axis. }
+  \label{fig:stellar-mass-phigal}
+\end{figure}
+
 % for one-armed galaxies = {9.86\degree}
@@
 
-Morphological classification commonly links bulge size to spiral tightness, and such a link is implied by the Hubble Sequence (\citealt{2005ARA&A..43..581S}, \citealt{2009MNRAS.393.1531G}, \citealt{2013seg..book..155B}).
-Some studies have indeed reported a link between measured spiral galaxy pitch angle and bulge size (i.e.
-\citealt{2017MNRAS.472.2263H}, \citealt{2019ApJ...873...85D}), while others have not found any significant correlation \citep{2019MNRAS.487.1808M}.
-We investigate this relationship here using a measure of bulge prominence from Galaxy Zoo 2, as Equation 3 in \citet{2019MNRAS.487.1808M}:
+Morphological classification commonly links bulge size to spiral tightness, and such a link is implied by the Hubble Sequence (\citealt{2005ARA&A..43..581S}, \citealt{2009MNRAS.393.1531G}, \citealt{2013seg..book..155B}), \textbf{although small bulge Sa galaxies have been noted for decades (e.g.
+for a review see \citet{2005ARA&A..43..581S}; this is also noted in \citealt{2019MNRAS.487.1808M})}.
+Some recent studies have indeed reported a link between measured spiral galaxy pitch angle and bulge size (i.e.
+\citealt{2019ApJ...873...85D}), while others have not found any significant correlation \citep{2019MNRAS.487.1808M}.
+\textbf{The differing results of this may depend on the details of how both the bulge size, and spiral pitch angles are measured and suggest further investigation is needed.} We investigate this relationship here using a measure of bulge prominence from Galaxy Zoo 2, as Equation 3 in \citet{2019MNRAS.487.1808M}:
 
@@
 One of the predictions of Manifold theory is that pitch angle increases with bar strength \citep{2009MNRAS.400.1706A}.
-To investigate this relationship in our data, we make use of Galaxy Zoo 2's bar fraction ($p_\mathrm{bar}$), which has been demonstrated to be a good measure of bar length \citep{Willett2013:1308.3496v2} and bar strength \citep{2012MNRAS.423.1485S,2012MNRAS.424.2180M,2018MNRAS.473.4731K} and therefore a good measure of the torque applied on the disc gas.
+\textbf{In this work we do not have any direct measurement of bar strength; nor are there significant numbers of barred galaxies in our sample.
+However in an attempt to} investigate this relationship in our data, we make use of Galaxy Zoo 2's bar fraction ($p_\mathrm{bar}$), which has been demonstrated to be a good measure of bar length \citep{Willett2013:1308.3496v2} and bar strength \citep{2012MNRAS.423.1485S,2012MNRAS.424.2180M,2018MNRAS.473.4731K} and therefore a good measure of the torque applied on the disc gas.
 
@@
 
-The fact that we do not find any link between bar strength and pitch angle suggests that the primary mechanism driving the evolution of the spirals in our sample is not Manifold theory.
+The fact that we do not find any link between \textbf{our measure of bar strength (based on the prominance of the bar in Galaxy Zoo 2) and pitch angle is suggestive that their is actually no link between bar strength and pitch angle, which would exclude Manifold theory as the primary mechanism driving the evolution of the spirals in our sample.
+However since we use only a proxy for bar strength, which has not been well tested, this is not conclusive.}
 
@@
 Testing the uniformity of $\cot\phi_\mathrm{gal}$ between {15\degree} and {50\degree} using a marginalized Anderson-Darling test results in rejecting the null hypothesis at the 1\% level for just 5\% of samples, with a large spread in observed test values.
-The full distribution of Anderson-Darling statistics can be seen in the upper panel of Figure \ref{fig:ad-cot-test}.
+The full distribution of Anderson-Darling statistics can be seen in Figure \ref{fig:ad-cot-test}.
 The large spread in results is caused by the large uncertainties in $\phi_\mathrm{gal}$.

### 2021-01-19 sections/summary.tex
Text edits to most of the rest of the referees points

@@
 The hierarchical Bayesian approach implemented in this paper allows a more thorough examination of pitch angle than length-weighted pitch angle calculation; obtaining posterior distributions of measured parameters.
-It better accounts for the large variations observed in inter-arm pitch angle than Fourier analysis, which assumes all arms in a given mode have the same pitch angle.
+It better accounts for the large variations observed in inter-arm pitch angle than Fourier analysis, which assumes all arms in a given \textbf{symmetric} mode have the same pitch angle.
 In this work, we find that the mean inter-arm difference in pitch angle is $11.0^\circ\pm 0.9^\circ$.
@@
 
-We do not find any link between bar strength and pitch angle suggests that the primary mechanism driving the evolution of the spirals in our sample is not Manifold theory (see Section \ref{section:morphology-comparison-bar}).
+We do not find any link between bar strength and pitch angle \textbf{in our sample.
+We make use of a novel (not tested) measure of bar strength, so at best this observation} suggests that the primary mechanism driving the evolution of the spirals in our sample is not Manifold theory (see Section \ref{section:morphology-comparison-bar}), \textbf{but cannot completely rule it out.}
 
 Our results are consistent with spiral winding of the form described by \citet{2019arXiv190910291P}, in which spiral arms are transient and recurrent, evolve through mechanisms such as swing-amplification \citep{1965MNRAS.130..125G} and which wind up over time.
-However, the assumptions of this model of spiral winding are highly simplistic, and it leaves many unanswered questions: what determines the limits on $\phi$?
+However, \textbf{this result depends on the range of pitch angles included in this test}.
+The assumptions of this model of spiral winding are also highly simplistic, and it leaves many unanswered questions: what determines the limits on $\phi$?
 Is the spiral arm equally apparent at all pitch angles, or is a selection effect present?
@@
 
-In this work, we assume that spiral arms are equally likely to be identified and recovered at all pitch angles.
+In this work, we assume that spiral arms are equally likely to be identified and recovered at all pitch angles, \textbf{which suggests the absence of galaxies at low pitch angles is not a selection effect}.
 This is not an unfair assumption given the amount of human effort that went into obtaining spiral arm measurements (more so than any other pitch angle measurement method, with each galaxy receiving at least 30 human classifications).
-The galaxy sample used is not selected to be representative of the general spiral population, but is comparable in size to those used in other similar studies (\citealt{2013MNRAS.436.1074S}, \citealt{2019ApJ...871..194Y}, \citealt{2019arXiv190910291P}) and covers a range of masses and spiral types.
+The galaxy sample used is not selected to be representative of the general spiral population, but is comparable in size to those used in other similar studies (\citealt{2013MNRAS.436.1074S}, \citealt{2019ApJ...871..194Y}, \citealt{2019arXiv190910291P}) and covers a range of masses ($9.45 < \log(M_* / M_\odot) < 11.05$ and spiral types.
 
-The methodology proposed here is a robust solution to the problems facing investigation of spiral morphology, namely that of reliably identifying spiral arms, and properly accounting for the spread in pitch angles of arms within a galaxy.
-This is one of the largest samples for which this test has been done; a larger sample would make possible further comparisons, such as splitting galaxies into spiral type (grand design / many-armed / flocculent) and examining the differences between populations.
+\textbf{We have presented evidence that} the methodology proposed here is a robust solution to the problems facing investigation of spiral morphology, namely that of reliably identifying spiral arms, and properly accounting for the spread in pitch angles of arms within a galaxy.
+This is one of the largest samples for which this test has been done \textbf{and is scaleable to larger samples}; such a sample would make possible further comparisons, such as splitting galaxies into spiral type (grand design / many-armed / flocculent), examining the differences between populations, \textbf{investigating if the interarm spread depends on other galaxy properties.}
 

### 2021-02-02 referee_response.md
Updated which github respository is linked.
Added stellar mass histogram, and more discussion of pitch angle variation

@@
 They also say "In general, logarithmic spirals are good approximations of the shape of galactic spiral arms".
-We think there is evidence in the literature that the assumption of logarithmic spirals is widespread, even if it's clear this method can in principle discover radial variation in phi.
+We think there is evidence in the literature that the assumption of logarithmic spirals is widespread, even if it's clear the 2DFFT method can in principle discover radial variation in phi.
 
@@
 The specific question about sigma_[gal] is interesting.
-We have found two previous attempts to measure this, which did not note any significant variation between galaxies, however sample sizes are small.
-We note this as an interesting followup question for larger samples.
-
-*** TODO: Check more if anyone has measured typical inter-arm variation in pitch angle ***
+We have added citations to work where arms in spirals are fit separately.
+All of them comment that pitch angles do vary between arms in spirals, but there is no discussion of how the variation changes between galaxies.
+In general sample sizes are small.
+We note this as an interesting followup question for larger samples, including potentially a large sample run through Galaxy Builder.
 
@@
 
-We have added a histogram, and an investigation of how pitch angle varies with stellar mass in Section 3.1.
+We have added a histogram (which was previously published in Lingard et al.
+2020) as a new Figure 3 in Section 2.2.
+There is also a histogram, and an investigation of how pitch angle varies with stellar mass in Section 3.1.
 
@@
 
+We apologise, we mistakenly linked a working respository, rather than the planned public one, which is available at https://github.com/tingard/hierarchical-modelling-of-spiral-pitch-angle
+
 >To sum up: There are a number of problems with this paper, and a major revision should be made, so as to rescue the valuable results provided by the volunteers, and present them in a proper context.

### 2021-02-02 sections/method.tex
Added stellar mass histogram, and more discussion of pitch angle variation

@@
 The \textit{stellar mass-complete sample} ranged in stellar mass from $9.45 < \log(M_* / M_\odot) < 11.05$, with most of the sample between $9.5 < \log(M_* / M_\odot) < 10.0$.
-A histogram of stellar masses can be see in Section \ref{section:constraints-on-galaxy-phi}, where variation with stellar mass, to check the impact of this limited mass range, is also investigated.
+A histogram of stellar masses for our subset can be see in Section \ref{section:constraints-on-galaxy-phi}, where variation with stellar mass, to check the impact of this limited mass range, is also investigated.
+{\bf For the reader's convenience we also reproduce Figure 4 from \citet{2020arXiv200610450L} here (see Figure \ref{fig:stellarmass}) which shows the redshift and stellar mass distribution of our analysis sample compared to the full \textit{stellar mass-complete sample} from \citet{2017MNRAS.472.2263H}.
+Our choice (see \citet{2020arXiv200610450L} for details) to prefer lower redshift galaxies for {\it Galaxy Builder} analysis is clear in the mass distribution which which results in a sample which favours galaxies $9.5 < \log(M_* / M_\odot) < 10.0$, and includes a smaller number of spirals with masses up to  $\log(M_* / M_\odot) = 11.05$}
+
+\begin{figure*}
+  \includegraphics[width=17.7cm]{plots/stellarmass.png}
+  \caption{A plot of redshift against stellar mass for the \textit{stellar mass-complete sample} from \citet{2017MNRAS.472.2263H}; the subset we use for analysis in Galaxy Builder are shown in red.
+At right we show a histogram of the stellar masses.
+This Figure is identical to Figure 4 from \citet{2020arXiv200610450L} who use the same sample.}
+  \label{fig:stellarmass}
+\end{figure*}
 
@@
 It is possible that it does vary between galaxies, and that this variation is physically interesting.
-At least \citet{1981AJ.....86.1847K} and \citet{2014ApJ...790...87D} have previously made attempts to measure this parameter.
-\citet{1981AJ.....86.1847K} fit logarithmic spirals to 113 nearby (NGC) galaxies, and note the dominant error in average galaxy pitch angle comes from interarm variation, which they measure to have an average value of 5$^\circ$.
-\citet{2014ApJ...790...87D} is primarily a machine learning method paper, and details on the galaxies investigated are not clear, but their Table 1 presents the median difference in pitch angle between pairs of arms with different lengths, which varies from 14.5$^\circ$ in very short arms, to 2.6$^\circ$ in the longest traced arms.
+Several authors have previously made attempts to measure this parameter.
+In a seminal work, \citet{1981AJ.....86.1847K} fit logarithmic spirals to 113 nearby (NGC) galaxies, and note the dominant error in average galaxy pitch angle comes from interarm variation, which they measure to have an average value of 5$^\circ$.
+\citet{2014ApJ...790...87D} is primarily a machine learning method paper, and while details on the galaxies investigated are not clear, their Table 1 presents the median difference in pitch angle between pairs of arms with different lengths, which varies from 14.5$^\circ$ in very short arms, to 2.6$^\circ$ in the longest traced arms.
 It is unclear how much this encodes error in their method versus real variation in the galaxy population.
-Futher investigation of this issue in the \textit{Galaxy Builder} sample would be an interesting follow-up project.}
+Within our own Milky Way, \citet{Vallee2015} did a meta analysis and comparison of several technique to conclude a range of 12-14$^\circ$ was reasonable for all Milky Way spiral arms.
+In a very detailed study of four very nearby spirals, \citet{HonigRead2015} conclude there are large variations of pitch angles  between spirals, and among arms in a given spiral, but made no comments as to if the variation was consistent with being constant.
+Further investigation of this issue in a larger \textit{Galaxy Builder} sample would be an interesting follow-up project.}
 

### 2021-02-09 sections/method.tex
Updates from Tim

@@
 Clustering of drawn spiral arms and cleaning of points was then performed as detailed in \citet{2020arXiv200610450L}.
-We remove any galaxies for which no spiral arms were identified, resulting in a hierarchical data structure of 139 galaxies, 261 spiral arms and 239,947 points.
+We remove any galaxies for which no spiral arms were identified, resulting in a hierarchical data structure of {\bf 129 galaxies, 247 spiral arms and 238,433 points}.
 

### 2021-02-09 referee_response.md
Updates from Tim
Updates on phi_gal and phi_arm for single armed galaxies

@@
 
-*** TODO: Compare spiral arms for validation subset ***
+ In this set of 98 galaxies, we find that 93% have the number of arms the same within +/-1 (53% have exactly the same number of arms).
+
+*** TODO: Compare pitch angles for validation subset ***
 
@@
 
-** Tim working on this. **
+When we looked into this we found this needed to be updated to 129 galaxies with 247 spiral arms which is closer to two arms per galaxy (1.9).
+We are unsure of the reason for this discrepancy, it may have been a typo, or left over from an earlier run.
+
+** TODO - list how many of the galaxies are multiarmed, and state the number of arms.
+
+ We have found that in some cases, while volunteers drew two arms, the clustering/aggregation code was unable to cluster them, so the final method reported a single arm.
+This happened in **TODO** cases, and is expected behaviour when clustering noisy data.
+The presence of this fraction of single armed galaxies drives the average number of arms per galaxy down below the real value.
 
@@
 
-** Tim says: Probably due to Aggregation. Need to check this **
+** TO ANSWER: Why are there so many 1-armed spirals ***
+
+We have added to the explanation of why PHI_gal is consistently below PHI_gal at low values of pitch angle for a single arm at the end of section 3.1 to make clearer the point that it's related to finding the mean of a broad, but skewed (truncated at zero) distribution.
 

### 2021-02-09 sections/method.tex
Updates on phi_gal and phi_arm for single armed galaxies

@@
 It is unclear how much this encodes error in their method versus real variation in the galaxy population.
-Within our own Milky Way, \citet{Vallee2015} did a meta analysis and comparison of several technique to conclude a range of 12-14$^\circ$ was reasonable for all Milky Way spiral arms.
+Within our own Milky Way, \citet{Vallee2015} did a meta analysis and comparison of several technique to conclude a range of 12-14$^\circ$ (i.e.
+2$^\circ$) was reasonable for all Milky Way spiral arms.
 In a very detailed study of four very nearby spirals, \citet{HonigRead2015} conclude there are large variations of pitch angles  between spirals, and among arms in a given spiral, but made no comments as to if the variation was consistent with being constant.

### 2021-02-09 sections/results.tex
Updates on phi_gal and phi_arm for single armed galaxies

@@
 The spread of arm pitch angle from the mean galaxy pitch angle can be seen in Figure \ref{fig:arm-pa-spread}, with points colour-coded by the number of arms measured for a galaxy.
-We see a slight drop in the expectation values of galaxy pitch angle ($E[\phi_\mathrm{gal}]$) compared to the expectation of arm pitch angles ($E[\phi_\mathrm{arm}]$) at small galaxy pitch angles, due to the truncation of $\phi_\mathrm{gal}$ at {0\degree}.
+We see a slight drop in the expectation values of galaxy pitch angle ($E[\phi_\mathrm{gal}]$) compared to the expectation of arm pitch angles ($E[\phi_\mathrm{arm}]$) at small galaxy pitch angles, {\bf which is caused by the combination of} truncation of $\phi_\mathrm{gal}$ at {0\degree} {\bf and the large spread (so the mean value differs from the mode as the distribution is highly skewed).}
 

### 2021-02-09 sections/method.tex
Added some comments on validation checks

@@
 Some galaxies in \citet{2020arXiv200610450L} were shown to volunteers a second time in a repeat validation subset to create a second aggregate model used to test internal consistency.
-We combine the 30 classifications of galaxies in this validation subset with the 30 original classifications.
+{\bf This sample of 98  galaxies demonstrates that arm number is highly reliable to $\pm1$ (93\% galaxies have arm number counts which agree in this range; 53\% have exactly identical arm number counts).} We combine the 30 classifications of galaxies in this validation subset with the 30 original classifications.
 Clustering of drawn spiral arms and cleaning of points was then performed as detailed in \citet{2020arXiv200610450L}.

### 2021-02-16 referee_response.md
Added some comments on validation checks
Few more numbers.

@@
 
-*** TODO: Compare pitch angles for validation subset ***
+ We have expanded the reasons why we sometimes miss arms (generally in the clustering part of the process) below.
+
+ Because of the Bayesian method used to obtain the final pitch angles, running this twice for the validation set, while potentially interesting, is not something we are able to do at this time.
+We are confident that the errors we report on the pitch angles capture this potential variation.
+
+** TODO CONFIRM THE ABOVE **
 
@@
 
-** TODO - list how many of the galaxies are multiarmed, and state the number of arms.
+Here is the complete breakdown of number of arms per galaxy:
 
- We have found that in some cases, while volunteers drew two arms, the clustering/aggregation code was unable to cluster them, so the final method reported a single arm.
-This happened in **TODO** cases, and is expected behaviour when clustering noisy data.
-The presence of this fraction of single armed galaxies drives the average number of arms per galaxy down below the real value.
+159 total galaxies, 129 of which with identified spiral arms
+247 total spiral arms
+238,433 total points
+----------------------------------------------------------------------
+ 30  galaxies had 0 arm(s)
+ 38  galaxies had 1 arm(s)
+ 68  galaxies had 2 arm(s)
+ 19  galaxies had 3 arm(s)
+ 4   galaxies had 4 arm(s)
+
+This has been entered into the text.
 
@@
 
-** TO ANSWER: Why are there so many 1-armed spirals ***
+ We have found that in some cases, while volunteers drew two arms, the clustering/aggregation code was unable to cluster them, so the final method reported a single arm.
+This happened in **TODO** cases, and is expected behaviour when clustering noisy data.
+The presence of this fraction of single armed galaxies drives the average number of arms per galaxy down below the real value, and because the most common number of actual arms is two, it is particularly notable in resulting in reports of one-armed galaxies.
 
-We have added to the explanation of why PHI_gal is consistently below PHI_gal at low values of pitch angle for a single arm at the end of section 3.1 to make clearer the point that it's related to finding the mean of a broad, but skewed (truncated at zero) distribution.
+We have added to the previously existing explanation of why PHI_gal is consistently below PHI_gal at low values of pitch angle for a single arm at the end of section 3.1 to make clearer the point that it's related to finding the mean of a broad, but skewed (truncated at zero) distribution.
 

### 2021-02-16 sections/method.tex
Few more numbers.

--- 2021-02-09 sections/method.tex
@@
 Some galaxies in \citet{2020arXiv200610450L} were shown to volunteers a second time in a repeat validation subset to create a second aggregate model used to test internal consistency.
-We combine the 30 classifications of galaxies in this validation subset with the 30 original classifications.
+{\bf Section 3.2 of \citet{2020arXiv200610450L} presents a comparison if these classifications to investigate volunteer consistency.
+We can see that arm number is highly reliable within $\pm1$ (93\% galaxies have arm number counts which agree in this range; 53\% have exactly identical arm number counts).} In this work, we combine the 30 classifications of galaxies in this validation subset with the 30 original classifications.
 Clustering of drawn spiral arms and cleaning of points was then performed as detailed in \citet{2020arXiv200610450L}.
-We remove any galaxies for which no spiral arms were identified, resulting in a hierarchical data structure of {\bf 129 galaxies, 247 spiral arms and 238,433 points}.
+We remove any galaxies for which no spiral arms were identified, resulting in a hierarchical data structure of {\bf 129 galaxies, 247 spiral arms and 238,433 points.
+This breaks down further to 68 galaxies (53\% of the sample) having two arms identified (meaning marked by the users, and clustered into an arm), 19 (15\%) with three such arms, 4 (3\%) with four, and the remainder (38 or 29\%) with a single identified arm.
+We believe that our reported number of arms per galaxy is in many cases an underestimate, and since the most common number of arms is two, this most often results in a single arm being measured when two are present.
+The most common way we miss an arm is if the clustering did not converge when users have appeared to identify a second arm.
+This happens in XX\% of galaxies, so we don't recommend using these statistics to make general conclusions about the number of arms per galaxy.}
 

### 2021-02-16 referee_response.md
Formatting

@@
 
-159 total galaxies, 129 of which with identified spiral arms
+```159 total galaxies, 129 of which with identified spiral arms
 247 total spiral arms
 238,433 total points
-----------------------------------------------------------------------
  30  galaxies had 0 arm(s)
@@
  19  galaxies had 3 arm(s)
- 4   galaxies had 4 arm(s)
+ 4   galaxies had 4 arm(s)```
 

### 2021-02-16 referee_response.md
formatting

@@
  19  galaxies had 3 arm(s)
- 4   galaxies had 4 arm(s)```
+ 4   galaxies had 4 arm(s)
 
-This has been entered into the text.
+This has been entered into the text.```
 

### 2021-02-16 referee_response.md
format

--- 2021-02-16 referee_response.md
@@
 
-```159 total galaxies, 129 of which with identified spiral arms
+159 total galaxies, 129 of which with identified spiral arms
 247 total spiral arms
@@
  19  galaxies had 3 arm(s)
- 4   galaxies had 4 arm(s)```
+ 4   galaxies had 4 arm(s)
 

### 2021-02-16 referee_response.md
Edits about the minimum pitch angle in the cot-uniform model

@@
 
-*** Do we though??? The should think about this point *** We believe the supporting text is sufficient to explain this. And have added some directly to the caption.
+ The issue here is that the model itself includes an arbitrary lower pitch angle - Pringle & Dobbs (2019) make no attempt to give any physical motivation for what that minimum angle should be, and in their test against data chose 11.5 degrees as the limit - largely because that's where inspection of the distribution of cot(pitch angle) in the sample they investigate deviates from constant.
+
+ We think the referee makes and excellent point that the previous explanation was a little confusing.
+What we have really done is to say this model fits the data IF the minimum pitch angle for winding spirals is around 15 degrees.
+It cannot be as low as 10 degrees.
+That's a specific observations that any more detailed model based on the simple ideas of Pringle & Dobbs (2019) would have to explain.
+We have made edits to the text to make this point.
 
@@
 
-** TDOD: *** We have toned down the abstract and conclusions.
+We have made edits in both the abstract and conclusions.
 

### 2021-02-16 sections/abstract.tex
Edits about the minimum pitch angle in the cot-uniform model

@@
 We present a new approach to deal with observed large variations in pitch angle between different arms in a single galaxy, which and obtains full posterior distributions on parameters.
-We make use of our fitted pitch angles to examine previously reported links between bulge and bar strength and pitch angle, finding no correlation in our data.
-We test a recent model for spiral arm winding, which predicts uniformity of the cotangent of pitch angle between some unknown upper and lower limits, with results supporting this model of transient and recurrent spiral pitch angle.
+We make use of our fitted pitch angles to examine previously reported links between bulge and bar strength and pitch angle, finding no correlation in our data {\bf (with a caveat that we use observations proxies for both bulge size and bar strength which differ from other work)}.
+We test a recent model for spiral arm winding, which predicts uniformity of the cotangent of pitch angle between some unknown upper and lower limits, {\bf finding our observations are consistent with} this model of transient and recurrent spiral pitch angle {\bf as long as the pitch angle at which most winding spirals dissipate/disappear is larger than 10$^\circ$.}

### 2021-02-16 sections/results.tex
Edits about the minimum pitch angle in the cot-uniform model

@@
 If $\Omega_p$ instead varies with radius we would expect $\cot{\phi}$ to be uniformly distributed between $\cot{\phi_\mathrm{max}}$ and $\cot{\phi_\mathrm{min}}$.
+{\bf The model presented in \citet{2019arXiv190910291P} does not give any physical justification for what $\cot{\phi_\mathrm{max}}$ and $\cot{\phi_\mathrm{min}}$ should be.
+}
 
-To test this theory, \citet{2019arXiv190910291P} used a Kolmogorov-Smirnov test to examine whether a sample of observed galaxy pitch angles was likely to have been drawn from a distribution uniform in its cotangent.
-Pitch angles were measured using discrete Fourier transformations in one- and two-dimensions, and as such do not account for inter-arm variations.
-They chose limits of $\cot{\phi} \in [1.00, 4.75]$ (roughly $11.9^\circ < \phi < 45.0^\circ$), motivated by examination of their data.
+To test this theory, \citet{2019arXiv190910291P} used a Kolmogorov-Smirnov test to examine whether {\bf a sample of 113} galaxies with measured pitch angles was likely to have been drawn from a distribution uniform in its cotangent.
+Pitch angles were measured {\bf by \citet{2019ApJ...871..194Y}} using discrete Fourier transformations in one- and two-dimensions, and as such do not account for inter-arm variations.
+They {\bf conclude the model works within limits of} $\cot{\phi} \in [1.00, 4.75]$ (roughly $11.9^\circ < \phi < 45.0^\circ$), motivated by examination of the data.
 
@@
 We will make use of the marginalized Anderson-Darling test described above, and examine winding on a per-arm basis, as well as a per-galaxy basis.
-Observation of the distribution of arm pitch angles in our sample (Figure \ref{fig:pa-cot-distributions}) suggests limits of $15^\circ < \phi < 50.0^\circ$.
+Observation of the distribution of arm pitch angles in our sample (Figure \ref{fig:pa-cot-distributions}) suggests {\bf they are close to uniform in cotangent} within limits of $15^\circ < \phi < 50.0^\circ$.
 
@@
 This result is also highly sensitive to the lower limit of $\phi$: decreasing it to {10\degree} results in us rejecting the cot-uniform model at greater than the 0.1\% level for 96\% of the posterior samples.
-As we have no information available on the selection biases present for classification of extremely loose or tight spiral arms in \textit{Galaxy Builder}, we choose to keep the less strict limit of {15\degree}.
+{\bf We can conclude from this that the \citet{2019arXiv190910291P} cot-uniform model is a good fit to the data, as long as the minimum pitch angle, $\cot{\phi_\mathrm{min}}$, at which the majority of winding dissipate and/or disappear is  $\phi_\mathrm{min} > $ {10\degree}, and more confidently $\phi_\mathrm{min} = $ {15\degree}.} \sout{As we have no information available on the selection biases present for classification of extremely loose or tight spiral arms in \textit{Galaxy Builder}, we choose to keep the less strict limit of {15\degree}.}
 
@@
 Moving rightwards on the x-axis implies greater confidence in rejecting the null hypothesis that the sample was drawn from a distribution uniform in $\cot$ between $15^\circ < \phi < 50.0^\circ$.
-In this instance, we would not be able to reject the null hypothesis at the 1\% level for either $\phi_\mathrm{gal}$ or $\phi_\mathrm{arm}$, meaning our sample is consistent with a cot uniform distribution.}
+In this instance, we would not be able to reject the null hypothesis at the 1\% level for either $\phi_\mathrm{gal}$ or $\phi_\mathrm{arm}$, meaning our sample is consistent with a cot uniform distribution.
+{\bf The larger error in $\phi_\mathrm{gal}$  means that this result is more significant for $\phi_\mathrm{arm}$, which is also physically motivated, as arms can wind independently.}}
   \label{fig:ad-cot-test}

### 2021-02-16 sections/summary.tex
Edits about the minimum pitch angle in the cot-uniform model

@@
 
-Our results are consistent with spiral winding of the form described by \citet{2019arXiv190910291P}, in which spiral arms are transient and recurrent, evolve through mechanisms such as swing-amplification \citep{1965MNRAS.130..125G} and which wind up over time.
-However, \textbf{this result depends on the range of pitch angles included in this test}.
-The assumptions of this model of spiral winding are also highly simplistic, and it leaves many unanswered questions: what determines the limits on $\phi$?
+Our results are consistent with spiral winding of the form described by \citet{2019arXiv190910291P}, in which spiral arms are transient and recurrent, evolve through mechanisms such as swing-amplification \citep{1965MNRAS.130..125G} and which wind up over time, \textbf{as long as the minimum pitch angle before the majority of arms dissipate or disappear is $\phi_\mathrm{min} = $ {15\degree}}.
+The assumptions of this model of spiral winding are highly simplistic, and it leaves many unanswered questions: what determines the limits on $\phi$?
 Is the spiral arm equally apparent at all pitch angles, or is a selection effect present?
-This result is also not evidence against QSDW, as our distribution of pitch angles may be dictated by other factors such as disk shear.
+\textbf{Our observations suggest that any further development of this model needs to predict that the minimum pitch angle, $\phi_\mathrm{min}>${10\degree}.} This result is also not evidence against QSDW, as our distribution of pitch angles may be dictated by other factors such as disk shear.
 

### 2021-02-22 referee_response.md
Edits about missing arms from clustering.

@@
 
-** TODO CONFIRM THE ABOVE **
-
 > From Figure 10 of the GBpaper, it follows that 16 out of 98 galaxies have p_bar .ge.
@@
  We have found that in some cases, while volunteers drew two arms, the clustering/aggregation code was unable to cluster them, so the final method reported a single arm.
-This happened in **TODO** cases, and is expected behaviour when clustering noisy data.
+This happened in 35 cases (the majority of the 1 arm galaxies), and is expected behaviour when clustering noisy data.
 The presence of this fraction of single armed galaxies drives the average number of arms per galaxy down below the real value, and because the most common number of actual arms is two, it is particularly notable in resulting in reports of one-armed galaxies.
+We have added some text explaining this, and cautioning against over interpreting our reports on the number of arms.
+We believe this should not significantly impact the results on pitch angle.
 

### 2021-02-22 sections/method.tex
Edits about missing arms from clustering.

@@
 The most common way we miss an arm is if the clustering did not converge when users have appeared to identify a second arm.
-This happens in XX\% of galaxies, so we don't recommend using these statistics to make general conclusions about the number of arms per galaxy.}
+For example, we find by visual inspection that just three galaxies in the sample are truly one-armed spirals, while in the remaining 35 galaxies for which only one arm was identified following clustering, the other arms were simply not recovered due to noise in the dataset.
+Give this, we do not recommend using these statistics to make general conclusions about the number of arms per galaxy.
+However, as our hierarchical model incorporates the uncertainty involved with missing spiral arms, the results related to pitch angle should not be significantly affected.}
 

### 2021-02-22 sections/method.tex
Ran tex (and some spell checking)

@@
 One potentially underused method of obtaining measurements of spirals is through photometric fitting of spiral structure, as possible using tools such as \textsc{GALFIT} \citep{2010AJ....139.2097P} and \textit{Galaxy Builder} \citep{2020arXiv200610450L}.
-These methods attempt to separate light from an image of a galaxy into distinct subcomponents, such as a galaxy disc, bulge, bar and spiral arms, generally finding the optimum solution using computational optimization.
-This optimization process, however, is often not robust for complex, many-component models and requires significant supervision to converge to a physically meaningful result \citep{Gao2017:1709.00746v1}.
+These methods attempt to separate light from an image of a galaxy into distinct subcomponents, such as a galaxy disc, bulge, bar and spiral arms, generally finding the optimum solution using computational optimisation.
+This optimisation process, however, is often not robust for complex, many-component models and requires significant supervision to converge to a physically meaningful result \citep{Gao2017:1709.00746v1}.
 \citet{2020arXiv200610450L} proposed a solution to this problem through the use of citizen science to provide priors on parameters used in computational fitting.
@@
 For example, we find by visual inspection that just three galaxies in the sample are truly one-armed spirals, while in the remaining 35 galaxies for which only one arm was identified following clustering, the other arms were simply not recovered due to noise in the dataset.
-Give this, we do not recommend using these statistics to make general conclusions about the number of arms per galaxy.
+Given this, we do not recommend using these statistics to make general conclusions about the number of arms per galaxy.
 However, as our hierarchical model incorporates the uncertainty involved with missing spiral arms, the results related to pitch angle should not be significantly affected.}
@@
 A simple visual inspection of the fitted logarithmic spirals suggests that it is an appropriate model, however, a comparison of a logarithmic spiral profile to other spiral forms (i.e.
-Archimedian or polynomial) is another important piece of work, outside of the scope of this research, as it has been reported that galaxy arms do not have constant pitch angles (\citealt{1981AJ.....86.1847K}; \citealt{2009MNRAS.397..164R}).
+Archimedean or polynomial) is another important piece of work, outside of the scope of this research, as it has been reported that galaxy arms do not have constant pitch angles (\citealt{1981AJ.....86.1847K}; \citealt{2009MNRAS.397..164R}).
 
@@
 Several authors have previously made attempts to measure this parameter.
-In a seminal work, \citet{1981AJ.....86.1847K} fit logarithmic spirals to 113 nearby (NGC) galaxies, and note the dominant error in average galaxy pitch angle comes from interarm variation, which they measure to have an average value of 5$^\circ$.
+In a seminal work, \citet{1981AJ.....86.1847K} fit logarithmic spirals to 113 nearby (NGC) galaxies, and note the dominant error in average galaxy pitch angle comes from inter-arm variation, which they measure to have an average value of 5$^\circ$.
 \citet{2014ApJ...790...87D} is primarily a machine learning method paper, and while details on the galaxies investigated are not clear, their Table 1 presents the median difference in pitch angle between pairs of arms with different lengths, which varies from 14.5$^\circ$ in very short arms, to 2.6$^\circ$ in the longest traced arms.

### 2021-02-25 referee_response.md
Working on referee reponse updates

@@
 However, it is easy to show that the spirals shown at the right column of Figure 1 are not flocculent, by rotating their image by 180°, and overlaying them on the original with 50% transparency: this shows that the top-right spiral has two arms in the central parts, and is multi-armed in the outer parts, while the bottom-right image is that of a barred spiral with predominantly two arms.
-
-The figure and caption has been updated.
-
 > Another problem concerns the Hubble classification, which for spirals is aptly summarized by Sandage (2005), who writes "It was simplicity itself with wide classification bins that merged seamlessly into one another according to three criteria of size of the central bulge, the degree of resolution into condensations (in the language of Reynolds 1920b, 1927a,b), and the openness of the spiral arms." It is very problematic that the authors think they can replace the criterion of "the degree of resolution into condensations" by "how obvious spiral patterns are".
@@
 He mentions in his 2005 paper that the Elmegreen & Elmegreen spiral arm classes are "important for the physics, but need not be made general in the [classification] notation".
-
-We have scaled down the mention of Elmegreen & Elmegreen spiral classes, and now use the traditional language of "degree of resolution" in the description the Hubble spiral sequence.
-
 > OK, this was the 1st paragraph in the Introduction.
@@
 However, in the latter paper this rearrangement of gas is attributed to the bar, and not to the spiral arms…
-
-There is evidence that both bars and spirals do rearrange disc material - this paragraph is intended to act as a short summary of the impact spirals have on galaxies. It has been reworded.
-
 > Let me point out that the motivation for the Lin & Shu (1964) paper was to overcome the 'winding dilemma' posed by differentially rotating spirals such as M31 and NGC 5055 (Prendergast & Burbidge 1960, Oort 1962), and that the statement of preferentially m = 2 armed spirals is in direct contradiction with the numbers given in Figure 1.
-
-We have removed both the mention of m=2, and the numbers from Figure 1. Neither of these are central to the point of our paper.
-
 > Furthermore, the mechanisms of wave amplification are not the same as those behind the assumption of quasi stationary density wave theory.
 The term "swing amplification" was coined by Toomre (1981), even though he relied on the Goldreich and Lynden-Bell (1965) work.
-
-We have added a clarification to this point.
-
 > What to make of the sentence "Larger bulges and more massive central black holes have both correlated with more tightly wound spiral arms" (p3, left column, line 47). They don't correlate anymore?
-
-This was a grammatical error. Thank you for bring it to our attention. It has been fixed.
-
 > The next paragraph then jumps to "Section 3.2 examines ...", and "Section 3.3 investigates ...", without explaining what is to come in Section 2 and 3.1…
 
-We have fixed this omission.
+We have rewritten the first two paragraphs to more carefully reflect the state of the field.
+These changes do not alter the scientific substance of our paper, but they do make the paper clearer and we thank the referee for their corrections.
+Specifically, we have:
+
+1. Updated Figure 1 and caption,
+2. Scaled down the mention of Elmegreen & Elmegreen spiral classes, and now use the traditional language of "degree of resolution" in the description the Hubble spiral sequence. 
+3. Reworded the second paragraph, which was intended to act as a short summary of the impact spirals have on galaxies. There is evidence that both bars and spirals do rearrange disc material 
+4. Removed both the mention of m=2, and the numbers from Figure 1. Neither of these are central to the point of our paper. 
+5. Added a clarification about quasi stationary density wave theory and swing amplification. 
+6. Corrected the grammatical error the referee noted on p3, left column, line 47
+7. Added an explanation of Section 2 and 3.1.
 
@@
 
-We have added text to make it clearer that different arms can have different pitch angles, even if in a single arm pitch angle may be constant with radius.
+We have added text to make it clearer that different arms can have different pitch angles, even if in a single arm, pitch angle may be constant with radius.
 
@@
 They also say "In general, logarithmic spirals are good approximations of the shape of galactic spiral arms".
-We think there is evidence in the literature that the assumption of logarithmic spirals is widespread, even if it's clear the 2DFFT method can in principle discover radial variation in phi.
+In the face of these comments, which show that the assumption of logarithmic spirals is made despite the more general applicability of the method of Considere & Athanassoula (1988) we think there is evidence in the literature that the assumption of logarithmic spirals is widespread, even if it's clear the 2DFFT method can in principle discover radial variation in phi.
 
@@
 2020) as a new Figure 3 in Section 2.2.
-There is also a histogram, and an investigation of how pitch angle varies with stellar mass in Section 3.1.
+We have also added a histogram, and an investigation of how pitch angle varies with stellar mass in Section 3.1 in response to another request below.
 
@@
 
- Because of the Bayesian method used to obtain the final pitch angles, running this twice for the validation set, while potentially interesting, is not something we are able to do at this time.
+ Because of the Bayesian method used to obtain the final pitch angles, running this twice for the validation set, while potentially interesting for a future project, is not something we are able to do at this time.
 We are confident that the errors we report on the pitch angles capture this potential variation.
@@
 
-We agree. Our statistical analysis incorporates this (which is one of the reasons we don't find anything significant). No changes made (or asked for).
+We agree. Our statistical analysis incorporates this (which is one of the reasons we don't find anything significant). No changes have been made (or asked for).
 
@@
 When we looked into this we found this needed to be updated to 129 galaxies with 247 spiral arms which is closer to two arms per galaxy (1.9).
-We are unsure of the reason for this discrepancy, it may have been a typo, or left over from an earlier run.
+We are unsure of the reason for this discrepancy, it may have been as simple as a typo.
 
@@
 
-159 total galaxies, 129 of which with identified spiral arms
-247 total spiral arms
-238,433 total points
- 30  galaxies had 0 arm(s)
- 38  galaxies had 1 arm(s)
- 68  galaxies had 2 arm(s)
- 19  galaxies had 3 arm(s)
- 4   galaxies had 4 arm(s)
+* 159 total galaxies, 129 of which with identified spiral arms
+* 247 total spiral arms
+* 238,433 total points
+  *  30  galaxies had 0 arm(s)
+  *  38  galaxies had 1 arm(s)
+  *  68  galaxies had 2 arm(s)
+  *  19  galaxies had 3 arm(s)
+  *  4   galaxies had 4 arm(s)
 
@@
  We have found that in some cases, while volunteers drew two arms, the clustering/aggregation code was unable to cluster them, so the final method reported a single arm.
-This happened in 35 cases (the majority of the 1 arm galaxies), and is expected behaviour when clustering noisy data.
-The presence of this fraction of single armed galaxies drives the average number of arms per galaxy down below the real value, and because the most common number of actual arms is two, it is particularly notable in resulting in reports of one-armed galaxies.
+This happened in 35 cases (the majority of the 1 arm galaxies), and is a consequence of clustering noisy data.
+Where arms are detected the pitch angles are reliable, and our method of calculating galaxy averages accounts for this possible bias.
+However the presence of this fraction of single-identified spiral arm galaxies drives the average number of arms per galaxy down below the real value, and because the most common number of actual arms is two, it is particularly notable in resulting in reports of one-armed galaxies.
 We have added some text explaining this, and cautioning against over interpreting our reports on the number of arms.
-We believe this should not significantly impact the results on pitch angle.
+We reiterate that this should not significantly impact the results on pitch angle.
 

### 2021-02-25 sections/abstract.tex
Updates plots to remove erroneous phi_gal. Corrects some typos. Reruns LaTeX

@@
 In this paper, we present a hierarchical Bayesian approach to galaxy pitch angle determination, using spiral arm data obtained through the \textit{Galaxy Builder} citizen science project.
-We present a new approach to deal with observed large variations in pitch angle between different arms in a single galaxy, which and obtains full posterior distributions on parameters.
+We present a new approach to deal with observed large variations in pitch angle between different arms in a single galaxy, which obtains full posterior distributions on parameters.
 We make use of our fitted pitch angles to examine previously reported links between bulge and bar strength and pitch angle, finding no correlation in our data {\bf (with a caveat that we use observations proxies for both bulge size and bar strength which differ from other work)}.

### 2021-02-25 sections/introduction.tex
Edits about the minimum pitch angle in the cot-uniform model
Updates plots to remove erroneous phi_gal. Corrects some typos. Reruns LaTeX

@@
 This will allow us to test simulations of gas in barred galaxies, which often demonstrate that bars can drive long-term spiral evolution \citep{2008A&A...489..115R}, or boost transient spiral structure \citep{2012MNRAS.426..167G}.
-Manifold theory \citealt{2006A&A...453...39R,2009MNRAS.394...67A,2009MNRAS.400.1706A} is one attempt to determine the orbits of stars in bar-driven spiral arms: it proposes that stars in the vicinity of the unstable Lagrangian points at either end of the bar tend to escape along predictable orbits, governed by invariant manifolds.
+Manifold theory \citep{2006A&A...453...39R,2009MNRAS.394...67A,2009MNRAS.400.1706A} is one attempt to determine the orbits of stars in bar-driven spiral arms: it proposes that stars in the vicinity of the unstable Lagrangian points at either end of the bar tend to escape along predictable orbits, governed by invariant manifolds.
 One of the primary factors influencing the shape of this invariant manifold is the relative strength of the non-axisymmetric forcing caused by the bar, with stronger bars resulting in spirals with larger pitch angles.
@@
 Larger bulges and more massive central black holes have both \textbf{been observed to} correlate with more tightly wound spiral arms.
-\textbf{We can test this with \textit{Galaxy Builder}.}
+\textbf{We can also test this with the data presented in this paper.}
 

### 2021-02-25 sections/method.tex
Updates plots to remove erroneous phi_gal. Corrects some typos. Reruns LaTeX

@@
 
-\textbf{As suggested by spiral formation models which correlate a galaxy wide pitch angle, with galaxy wide properties,} we will assume that a give galaxy has some preferred value for arm pitch angle, $\phi_\mathrm{gal}$, and that the pitch angles of spiral arms in that galaxy, $\phi_\mathrm{arm}$, are constant with radius (giving logarithmic spirals) and drawn from a normal distribution centred on $\phi_\mathrm{gal}$, with some spread $\sigma_\mathrm{gal}$ common to all galaxies.
+\textbf{As suggested by spiral formation models which correlate a galaxy wide pitch angle with galaxy wide properties,} we will assume that a given galaxy has some preferred value for arm pitch angle, $\phi_\mathrm{gal}$, and that the pitch angles of spiral arms in that galaxy, $\phi_\mathrm{arm}$, are constant with radius (giving logarithmic spirals) and drawn from a normal distribution centred on $\phi_\mathrm{gal}$, with some spread $\sigma_\mathrm{gal}$ common to all galaxies.
 We truncate the normal distribution of spiral arm pitch angles in a single galaxy between the physical limits of {0\degree} (a ring) and {90\degree} (a ``spoke''), giving
@@
 
-Where $\widetilde{r_\mathrm{arm}}$ is the model's predictions for the radii of the deprojected points in a \textit{Galaxy Builder} arm ($\widevec{r_\mathrm{arm}}$), $c_\mathrm{arm}$ is the amplitude parameter (equivalent to $A$ in Equation \ref{eq:log-spiral}), and $\widevec{\theta_\mathrm{arm}}$ is the polar angles of the points.
+Where $\widetilde{r_\mathrm{arm}}$ is the model's prediction for the radii of the deprojected points in a \textit{Galaxy Builder} arm ($\widevec{r_\mathrm{arm}}$), $c_\mathrm{arm}$ is the amplitude parameter (equivalent to $A$ in Equation \ref{eq:log-spiral}), and $\widevec{\theta_\mathrm{arm}}$ is the polar angles of the points.
 

### 2021-02-25 sections/results.tex
Updates plots to remove erroneous phi_gal. Corrects some typos. Reruns LaTeX

@@
 The spread of arm pitch angle from the mean galaxy pitch angle can be seen in Figure \ref{fig:arm-pa-spread}, with points colour-coded by the number of arms measured for a galaxy.
-We see a slight drop in the expectation values of galaxy pitch angle ($E[\phi_\mathrm{gal}]$) compared to the expectation of arm pitch angles ($E[\phi_\mathrm{arm}]$) at small galaxy pitch angles, {\bf which is caused by the combination of} truncation of $\phi_\mathrm{gal}$ at {0\degree} {\bf and the large spread (so the mean value differs from the mode as the distribution is highly skewed).}
+We see a slight drop in the expectation values of galaxy pitch angle ($E[\phi_\mathrm{gal}]$) compared to the expectation of arm pitch angles ($E[\phi_\mathrm{arm}]$) at small galaxy pitch angles, {\bf which is caused by a combination of the} truncation of $\phi_\mathrm{gal}$ at {0\degree} {\bf and the large spread (so the mean value differs from the mode as the distribution is highly skewed).}
 
@@
 
-The fact that we do not find any link between \textbf{our measure of bar strength (based on the prominance of the bar in Galaxy Zoo 2) and pitch angle is suggestive that their is actually no link between bar strength and pitch angle, which would exclude Manifold theory as the primary mechanism driving the evolution of the spirals in our sample.
+The fact that we do not find any link between \textbf{our measure of bar strength (based on the prominance of the bar in Galaxy Zoo 2) and pitch angle is suggestive that there is actually no link between bar strength and pitch angle, which would exclude Manifold theory as the primary mechanism driving the evolution of the spirals in our sample.
 However since we use only a proxy for bar strength, which has not been well tested, this is not conclusive.}
@@
 This result is also highly sensitive to the lower limit of $\phi$: decreasing it to {10\degree} results in us rejecting the cot-uniform model at greater than the 0.1\% level for 96\% of the posterior samples.
-{\bf We can conclude from this that the \citet{2019arXiv190910291P} cot-uniform model is a good fit to the data, as long as the minimum pitch angle, $\cot{\phi_\mathrm{min}}$, at which the majority of winding dissipate and/or disappear is  $\phi_\mathrm{min} > $ {10\degree}, and more confidently $\phi_\mathrm{min} = $ {15\degree}.} \sout{As we have no information available on the selection biases present for classification of extremely loose or tight spiral arms in \textit{Galaxy Builder}, we choose to keep the less strict limit of {15\degree}.}
+{\bf We can conclude from this that the \citet{2019arXiv190910291P} cot-uniform model is an adequate fit to the data, as long as the minimum pitch angle, $\cot{\phi_\mathrm{min}}$, at which the majority of winding dissipate and/or disappear is  $\phi_\mathrm{min} > $ {10\degree}, and more confidently $\phi_\mathrm{min} = $ {15\degree}.} \sout{As we have no information available on the selection biases present for classification of extremely loose or tight spiral arms in \textit{Galaxy Builder}, we choose to keep the less strict limit of {15\degree}.}
 

### 2021-02-25 referee_response.md
Minor edits based on Chris L comments

@@
 However, it is easy to show that the spirals shown at the right column of Figure 1 are not flocculent, by rotating their image by 180°, and overlaying them on the original with 50% transparency: this shows that the top-right spiral has two arms in the central parts, and is multi-armed in the outer parts, while the bottom-right image is that of a barred spiral with predominantly two arms.
+
 > Another problem concerns the Hubble classification, which for spirals is aptly summarized by Sandage (2005), who writes "It was simplicity itself with wide classification bins that merged seamlessly into one another according to three criteria of size of the central bulge, the degree of resolution into condensations (in the language of Reynolds 1920b, 1927a,b), and the openness of the spiral arms." It is very problematic that the authors think they can replace the criterion of "the degree of resolution into condensations" by "how obvious spiral patterns are".
@@
 He mentions in his 2005 paper that the Elmegreen & Elmegreen spiral arm classes are "important for the physics, but need not be made general in the [classification] notation".
+
 > OK, this was the 1st paragraph in the Introduction.
@@
 However, in the latter paper this rearrangement of gas is attributed to the bar, and not to the spiral arms…
+
 > Let me point out that the motivation for the Lin & Shu (1964) paper was to overcome the 'winding dilemma' posed by differentially rotating spirals such as M31 and NGC 5055 (Prendergast & Burbidge 1960, Oort 1962), and that the statement of preferentially m = 2 armed spirals is in direct contradiction with the numbers given in Figure 1.
+
 > Furthermore, the mechanisms of wave amplification are not the same as those behind the assumption of quasi stationary density wave theory.
 The term "swing amplification" was coined by Toomre (1981), even though he relied on the Goldreich and Lynden-Bell (1965) work.
+
 > What to make of the sentence "Larger bulges and more massive central black holes have both correlated with more tightly wound spiral arms" (p3, left column, line 47). They don't correlate anymore?
+
 > The next paragraph then jumps to "Section 3.2 examines ...", and "Section 3.3 investigates ...", without explaining what is to come in Section 2 and 3.1…
@@
 This happened in 35 cases (the majority of the 1 arm galaxies), and is a consequence of clustering noisy data.
-Where arms are detected the pitch angles are reliable, and our method of calculating galaxy averages accounts for this possible bias.
+Where arms are detected the pitch angles are reliable, and our method of calculating galaxy averages accounts for this possible bias of error caused by missing arms.
 However the presence of this fraction of single-identified spiral arm galaxies drives the average number of arms per galaxy down below the real value, and because the most common number of actual arms is two, it is particularly notable in resulting in reports of one-armed galaxies.
@@
 
- We think the referee makes and excellent point that the previous explanation was a little confusing.
-What we have really done is to say this model fits the data IF the minimum pitch angle for winding spirals is around 15 degrees.
-It cannot be as low as 10 degrees.
-That's a specific observations that any more detailed model based on the simple ideas of Pringle & Dobbs (2019) would have to explain.
+ We think the referee makes an excellent point that the previous explanation was a little confusing.
+What we have really determined in our study here is that the cot(phi) uniform "model" fits the data well when the minimum pitch angle for winding spirals is around 15 degrees.
+We also determine that if the minimum pitch angle is 10 degrees the model cannot fit our data.
+This suggests that a minimum pitch angle for this model must be above 10 degrees,(e.g.
+15 degrees).
+That's a specific observation that a more detailed model based on the simple model of Pringle & Dobbs (2019) would have to explain.
 We have made edits to the text to make this point.

### 2021-02-25 sections/abstract.tex
Minor edits based on Chris L comments

--- 2021-01-19 sections/abstract.tex
@@
 % !TEX root = ../main.tex
-Spiral structure is ubiquitous in the Universe, and the pitch angle of arms in spiral galaxies is an important observable in efforts to discriminate between different mechanisms of spiral arm formation and evolution.
+Spiral structure is ubiquitous in the Universe, and the pitch angle of arms in spiral galaxies provide an important observable in efforts to discriminate between different mechanisms of spiral arm formation and evolution.
 In this paper, we present a hierarchical Bayesian approach to galaxy pitch angle determination, using spiral arm data obtained through the \textit{Galaxy Builder} citizen science project.
-We present a new approach to deal with observed large variations in pitch angle between different arms in a single galaxy, which and obtains full posterior distributions on parameters.
-We make use of our fitted pitch angles to examine previously reported links between bulge and bar strength and pitch angle, finding no correlation in our data.
-We test a recent model for spiral arm winding, which predicts uniformity of the cotangent of pitch angle between some unknown upper and lower limits, with results supporting this model of transient and recurrent spiral pitch angle.
+We present a new approach to deal with the large variations in pitch angle between different arms in a single galaxy, which obtains full posterior distributions on parameters.
+We make use of our pitch angles to examine previously reported links between bulge and bar strength and pitch angle, finding no correlation in our data {\bf (with a caveat that we use observational proxies for both bulge size and bar strength which differ from other work)}.
+We test a recent model for spiral arm winding, which predicts uniformity of the cotangent of pitch angle between some unknown upper and lower limits, {\bf finding our observations are consistent with} this model of transient and recurrent spiral pitch angle {\bf as long as the pitch angle at which most winding spirals dissipate or disappear is larger than 10$^\circ$.}

### 2021-02-25 sections/method.tex
Minor edits based on Chris L comments

@@
 where $\phi$ is the arm's pitch angle, $A$ is an amplitude coefficient and $\theta$ is the polar coordinate.
-\textbf{Different arms in a galaxy could have different values of $\phi$, however for each arm, $\phi$ is assumed to constant with radius}.
+\textbf{Different arms in a galaxy could have different values of $\phi$, however for each arm, $\phi$ is assumed to be constant with radius}.
 One method used to obtain a pitch angle of a galaxy is to fit logarithmic spirals to individually identified arm segments and take the weighted mean of their pitch angles (which often vary by upwards of $10^\circ$, \citealt{2014ApJ...790...87D}).
-Weighting is determined by the length of the arc segment, with longer being assigned higher weights, i.e.
+Weighting is determined by the length of the arc segment, with longer arms being assigned higher weights, i.e.
 for a galaxy where we have identified $N$ arm segments, each with length $L_i$ and pitch angle $\phi_i$
@@
 It is unclear how the variation between pitch angles of individual arms impacts this measurement.
-\textbf{We note that while this method is able to model non-logarithmic spirals - as a sum of logarithmic spirals with differing pitch angles, most applications of the model favour picking regions of the galaxy in which the pitch angle is constant with radius - e.g.
+\textbf{We note that while this method is able to model non-logarithmic spirals -- as a sum of logarithmic spirals with differing pitch angles, most applications use models which assume that the pitch angle is constant with radius, in some cases picking regions of a galaxy in which this is true - e.g.
 see Section 4.3.2 of \citealt{2012ApJS..199...33D}.}

### 2021-02-25 sections/summary.tex
Minor edits based on Chris L comments

@@
 We do not find any link between bar strength and pitch angle \textbf{in our sample.
-We make use of a novel (not tested) measure of bar strength, so at best this observation} suggests that the primary mechanism driving the evolution of the spirals in our sample is not Manifold theory (see Section \ref{section:morphology-comparison-bar}), \textbf{but cannot completely rule it out.}
+However, rather that a direct measure of bar strength, we make use of an available parameter which correlates with bar strength, so at best this observation} suggests that the primary mechanism driving the evolution of the spirals in our sample is not Manifold theory (see Section \ref{section:morphology-comparison-bar}), \textbf{but cannot completely rule it out.}
 
-Our results are consistent with spiral winding of the form described by \citet{2019arXiv190910291P}, in which spiral arms are transient and recurrent, evolve through mechanisms such as swing-amplification \citep{1965MNRAS.130..125G} and which wind up over time, \textbf{as long as the minimum pitch angle before the majority of arms dissipate or disappear is $\phi_\mathrm{min} = $ {15\degree}}.
+Our results are consistent with spiral winding of the form described by \citet{2019arXiv190910291P}, in which spiral arms are transient and recurrent, evolve through mechanisms such as swing-amplification \citep{1965MNRAS.130..125G} and which wind up over time, \textbf{if the minimum pitch angle before the majority of arms dissipate or disappear is $\phi_\mathrm{min} = $ {15\degree}, but rule it out if the minimum pitch angle is $\phi_\mathrm{min} = $ {10\degree}}.
 The assumptions of this model of spiral winding are highly simplistic, and it leaves many unanswered questions: what determines the limits on $\phi$?

### 2021-02-25 referee_response.md
Update referee_response.md

@@
 
-*** Tim TODO - might just need a better README ***
+We apologise, we mistakenly linked a working respository, rather than the planned public one, which is available at https://github.com/tingard/hierarchical-modelling-of-spiral-pitch-angle.
 
-We apologise, we mistakenly linked a working respository, rather than the planned public one, which is available at https://github.com/tingard/hierarchical-modelling-of-spiral-pitch-angle
+Following this feedback, we have added (see at end of the README) a description of the contents of each of the Python Notebooks.
 
@@
 We are happy to agree this work should be published and presented to fully make use of the valuable results the volunteers provided.
+We thank the referee for detailed comments which have improved the paper.

### 2021-02-25 sections/results.tex
Edits for clarification of minimum pitch angle argument

@@
 One of the predictions of Manifold theory is that pitch angle increases with bar strength \citep{2009MNRAS.400.1706A}.
-\textbf{In this work we do not have any direct measurement of bar strength; nor are there significant numbers of barred galaxies in our sample.
+\textbf{In this work we do not have any direct measurement of bar strength; nor are there significant numbers of strongly barred galaxies in our sample.
 However in an attempt to} investigate this relationship in our data, we make use of Galaxy Zoo 2's bar fraction ($p_\mathrm{bar}$), which has been demonstrated to be a good measure of bar length \citep{Willett2013:1308.3496v2} and bar strength \citep{2012MNRAS.423.1485S,2012MNRAS.424.2180M,2018MNRAS.473.4731K} and therefore a good measure of the torque applied on the disc gas.
@@
 
-This result suggests that we cannot rule out a cot-uniform source distribution for galaxy pitch angle, but the large uncertainty in $\phi_\mathrm{gal}$ makes it difficult to make any conclusive statements.
+This result suggests that \sout{we cannot rule out} {\bf our data are consistent with} a cot-uniform source distribution for galaxy pitch angle, but the large uncertainty in $\phi_\mathrm{gal}$ makes it difficult to make any conclusive statements.
 This result is also highly sensitive to the lower limit of $\phi$: decreasing it to {10\degree} results in us rejecting the cot-uniform model at greater than the 0.1\% level for 96\% of the posterior samples.
-{\bf We can conclude from this that the \citet{2019arXiv190910291P} cot-uniform model is an adequate fit to the data, as long as the minimum pitch angle, $\cot{\phi_\mathrm{min}}$, at which the majority of winding dissipate and/or disappear is  $\phi_\mathrm{min} > $ {10\degree}, and more confidently $\phi_\mathrm{min} = $ {15\degree}.} \sout{As we have no information available on the selection biases present for classification of extremely loose or tight spiral arms in \textit{Galaxy Builder}, we choose to keep the less strict limit of {15\degree}.}
+{\bf We can conclude from this that the \citet{2019arXiv190910291P} cot-uniform model is an adequate fit to the data, as long as the minimum pitch angle, $\cot{\phi_\mathrm{min}}$, at which the majority of winding dissipate or disappear is  $\phi_\mathrm{min} > $ {10\degree}, and more confidently $\phi_\mathrm{min} = $ {15\degree}.
+We reiterate that there is no prediction in \citet{2019arXiv190910291P} as to what this minimum pitch angle should be, so our observation constrains the allowed range.} \sout{As we have no information available on the selection biases present for classification of extremely loose or tight spiral arms in \textit{Galaxy Builder}, we choose to keep the less strict limit of {15\degree}.}
 
@@
 The resulting distribution of Anderson-Darling statistics is shown in in the lower panel of Figure \ref{fig:ad-cot-test}.
-This result is highly consistent with the model for spiral winding proposed by \citet{2019arXiv190910291P} and can be interpreted as evidence that spirals are formed through local disc perturbations, and are primarily governed by local forces.
+This result is highly consistent with the model for spiral winding proposed by \citet{2019arXiv190910291P} {\bf with a $\cot{\phi_\mathrm{min}} = $ {15\degree} } and can be interpreted as evidence that spirals are formed through local disc perturbations, and are primarily governed by local forces.
 

### 2021-02-25 sections/summary.tex
Edits for clarification of minimum pitch angle argument

@@
 
-Our results are consistent with spiral winding of the form described by \citet{2019arXiv190910291P}, in which spiral arms are transient and recurrent, evolve through mechanisms such as swing-amplification \citep{1965MNRAS.130..125G} and which wind up over time, \textbf{if the minimum pitch angle before the majority of arms dissipate or disappear is $\phi_\mathrm{min} = $ {15\degree}, but rule it out if the minimum pitch angle is $\phi_\mathrm{min} = $ {10\degree}}.
+Our results are consistent with spiral winding of the form described by \citet{2019arXiv190910291P}, in which spiral arms are transient and recurrent, evolve through mechanisms such as swing-amplification \citep{1965MNRAS.130..125G} and which wind up over time.
+\textbf{This model predict a distribution of pitch angles that is uniform in cotangent space across some range.
+No prediction is provided as to what that range should be.
+Our data are consistent with this model, if the minimum pitch angle is $\phi_\mathrm{min} = $ {15\degree}, but rule it out if the minimum pitch angle is $\phi_\mathrm{min} = $ {10\degree}}.
 The assumptions of this model of spiral winding are highly simplistic, and it leaves many unanswered questions: what determines the limits on $\phi$?

### 2021-02-26 sections/method.tex
Minor edits.

@@
 We remove any galaxies for which no spiral arms were identified, resulting in a hierarchical data structure of {\bf 129 galaxies, 247 spiral arms and 238,433 points.
-This breaks down further to 68 galaxies (53\% of the sample) having two arms identified (meaning marked by the users, and clustered into an arm), 19 (15\%) with three such arms, 4 (3\%) with four, and the remainder (38 or 29\%) with a single identified arm.
+This breaks down further to 68 galaxies (53\% of the sample) having two arms identified (meaning that they were marked by enough users to cluster into an arm), 19 (15\%) with three such arms, 4 (3\%) with four, and the remainder (38 or 29\%) with a single identified arm.
 We believe that our reported number of arms per galaxy is in many cases an underestimate, and since the most common number of arms is two, this most often results in a single arm being measured when two are present.
@@
 
-\textbf{These are conservative priors, which are not expected to have significant impact of the results.} The inverse gamma distribution is used to aid the convergence of the Hamiltonian Monte Carlo (HMC) algorithm used (discussed later).
+\textbf{These are conservative priors, which are not expected to have significant impact on the results.} The inverse gamma distribution is used to aid the convergence of the Hamiltonian Monte Carlo (HMC) algorithm used (discussed later).
 The Cauchy distribution is equivalent to the Student's t-distribution with one degree of freedom, and was chosen due to its fatter tails than the normal distribution.

### 2021-02-27 sections/method.tex
deleted unused plots (still available in data repo). Made some small changes to numbers to reflect most recent results

@@
 \begin{figure*}
-  \includegraphics[width=17.7cm]{plots/stellarmass.png}
+  \includegraphics[width=17.7cm]{plots/stellar_mass_selection_plot.pdf}
   \caption{A plot of redshift against stellar mass for the \textit{stellar mass-complete sample} from \citet{2017MNRAS.472.2263H}; the subset we use for analysis in Galaxy Builder are shown in red.
@@
 We believe that our reported number of arms per galaxy is in many cases an underestimate, and since the most common number of arms is two, this most often results in a single arm being measured when two are present.
-The most common way we miss an arm is if the clustering did not converge when users have appeared to identify a second arm.
+The most common way we miss an arm is if the spiral clustering algorithm did not identify a cluster when users appeared to identify a second arm.
 For example, we find by visual inspection that just three galaxies in the sample are truly one-armed spirals, while in the remaining 35 galaxies for which only one arm was identified following clustering, the other arms were simply not recovered due to noise in the dataset.

### 2021-02-27 sections/results.tex
deleted unused plots (still available in data repo). Made some small changes to numbers to reflect most recent results

@@
   \caption{Examples of spiral profiles fit using the hierarchical model described in Section \ref{section:bhsm-model}.
-Deprojected points from \textit{Galaxy Builder} clustered, cleaned spiral arms are shown in black; fit logarithmic spiral arms are shown in red, with the width of the line corresponding to the $2\sigma$ interval on predicted values of $\widetilde{r_\mathrm{arm}}$.}
+Deprojected points from \textit{Galaxy Builder} clustered, cleaned spiral arms are shown in black; fit logarithmic spiral arms are shown in red, with the width of the line corresponding to the $2\sigma$ interval on predicted values of $\widetilde{r_\mathrm{arm}}$.
+The two one-armed spirals in the top left panels are instances where the spiral clustering algorithm failed to identify all spiral arms present in the galaxy.}
   \label{fig:example-spiral-fits}
@@
 
+We make use of Galaxy Zoo 2 data for morhpological comparison.
+Two of the galaxies in our sample could not be matched to Galaxy Zoo 2 data, and as such have been dropped from this comparison (leaving 127 galaxies).
+
 \subsubsection{Pitch angle vs. Bulge size}
@@
 
-We separate our sample into galaxies with weaker bulges ($B_\mathrm{avg} < 0.28$, 83 galaxies) and those with stronger bulges ($B_\mathrm{avg} \ge 0.28$, 54 galaxies), to test whether their pitch angles could be drawn from significantly different distributions.
+We separate our sample into galaxies with weaker bulges ($B_\mathrm{avg} < 0.28$, 79 galaxies) and those with stronger bulges ($B_\mathrm{avg} \ge 0.28$, 48 galaxies), to test whether their pitch angles could be drawn from significantly different distributions.
 A marginalized two-sample Anderson-Darling test comparing the distributions of $\phi_\mathrm{gal}$ for the samples does not find evidence that galaxy pitch angles were drawn from different distributions: we reject the null hypothesis at the 1\% level for only 1\% of the samples.
@@
 We do not observe a correlation between $p_\mathrm{bar}$ and $E[\phi_\mathrm{gal}]$ (Pearson correlation coefficient of -0.05, with a p-value of 0.54).
-Following \citet{2012MNRAS.424.2180M} and \citet{2012MNRAS.423.1485S}, we separate the sample into galaxies without a bar ($p_\mathrm{bar} < 0.2$), with a weak bar ($0.2 \le p_\mathrm{bar} \le 0.5$) and with a strong bar ($p_\mathrm{bar} > 0.5$).
+Following \citet{2012MNRAS.424.2180M} and \citet{2012MNRAS.423.1485S}, we separate the sample into galaxies without a bar ($p_\mathrm{bar} < 0.2$, 50 galaxies), with a weak bar ($0.2 \le p_\mathrm{bar} \le 0.5$, 44 galaxies) and with a strong bar ($p_\mathrm{bar} > 0.5$, 33 galaxies).
 Performing marginalized three-sample Anderson-Darling tests does not find that pitch angles ($\phi_\mathrm{gal}$ or $\phi_\mathrm{arm}$) of galaxies with different bar strengths were drawn from different distributions; we do not reject the null hypothesis at the 1\% level for any samples for the test of $\phi_\mathrm{gal}$, and at the 10\% level for the test of $\phi_\mathrm{arm}$.
@@
 The resulting distribution of Anderson-Darling statistics is shown in in the lower panel of Figure \ref{fig:ad-cot-test}.
-This result is highly consistent with the model for spiral winding proposed by \citet{2019arXiv190910291P} {\bf with a $\cot{\phi_\mathrm{min}} = $ {15\degree} } and can be interpreted as evidence that spirals are formed through local disc perturbations, and are primarily governed by local forces.
+This result is highly consistent with the model for spiral winding proposed by \citet{2019arXiv190910291P} {\bf with $\cot{\phi_\mathrm{min}} = $ {15\degree} } and can be interpreted as evidence that spirals are formed through local disc perturbations, and are primarily governed by local forces.
 

### 2021-02-27 sections/summary.tex
deleted unused plots (still available in data repo). Made some small changes to numbers to reflect most recent results

@@
 This is not an unfair assumption given the amount of human effort that went into obtaining spiral arm measurements (more so than any other pitch angle measurement method, with each galaxy receiving at least 30 human classifications).
-The galaxy sample used is not selected to be representative of the general spiral population, but is comparable in size to those used in other similar studies (\citealt{2013MNRAS.436.1074S}, \citealt{2019ApJ...871..194Y}, \citealt{2019arXiv190910291P}) and covers a range of masses ($9.45 < \log(M_* / M_\odot) < 11.05$ and spiral types.
+The galaxy sample used is not selected to be representative of the general spiral population, but is comparable in size to those used in other similar studies (\citealt{2013MNRAS.436.1074S}, \citealt{2019ApJ...871..194Y}, \citealt{2019arXiv190910291P}) and covers a range of masses ($9.45 < \log(M_* / M_\odot) < 11.05$) and spiral types.
 

### 2021-03-31 referee_response1.md
Starting edits for 2nd report.

--- 2021-02-25 referee_response.md
@@
 They also say "In general, logarithmic spirals are good approximations of the shape of galactic spiral arms".
-In the face of these comments, which show that the assumption of logarithmic spirals is made despite the more general applicability of the method of Considere & Athanassoula (1988) we think there is evidence in the literature that the assumption of logarithmic spirals is widespread, even if it's clear the 2DFFT method can in principle discover radial variation in phi.
+In the face of these comments, which show that the assumption of logarithmic spirals is made despite the more general applicability of the method of Considere & Athanassoula (1988), we think there is evidence in the literature that the assumption of logarithmic spirals is widespread, even if the 2DFFT method can in principle discover radial variation in phi.
 
@@
 In general sample sizes are small.
-We note this as an interesting followup question for larger samples, including potentially a large sample run through Galaxy Builder.
+We note this as an interesting followup question for larger samples, including potentially a larger sample run through a second phase of Galaxy Builder.
 
@@
 
- In this set of 98 galaxies, we find that 93% have the number of arms the same within +/-1 (53% have exactly the same number of arms).
+ We added a line to point the reader to Lingard et al.
+(2020) where this is discussed, and also add a comment that in this set of 98 galaxies, we find that 93% have the number of arms the same within +/-1 (53% have exactly the same number of arms).
 
@@
 
-We agree. Our statistical analysis incorporates this (which is one of the reasons we don't find anything significant). No changes have been made (or asked for).
+We agree. Our statistical analysis incorporates this (which is one of the reasons we don't find anything significant). We have made this clearer in the text.
 
@@
 When we looked into this we found this needed to be updated to 129 galaxies with 247 spiral arms which is closer to two arms per galaxy (1.9).
-We are unsure of the reason for this discrepancy, it may have been as simple as a typo.
+We are unsure of the reason for this discrepancy.
+All plots and numberical result have been updated.
 
@@
 
- We have found that in some cases, while volunteers drew two arms, the clustering/aggregation code was unable to cluster them, so the final method reported a single arm.
+ We have found that in some cases, while most volunteers drew two arms, the clustering/aggregation code was unable to cluster them, so the final method reported a single arm.
 This happened in 35 cases (the majority of the 1 arm galaxies), and is a consequence of clustering noisy data.
@@
 
-This isn't really in the scope of this paper.
+We think addressing this, while interesting, is beyond the scope of this paper.
 The Hart et al.
-(2017) result really showed that pitch angle depended on disc mass more than bulge size (but more massive discs tend to have larger bulges.
+(2017) result really showed that pitch angle depended on disc mass more than bulge size (but more massive discs tend to have larger bulges).
 In the interest of keeping to the point of this paper, we have removed that citation, but also note that these correlation results can depend on both how bulge size and pitch angles are measured.
@@
 We agree we cannot make any strong claims regarding Manifold theory, however it is still valid that we find no link between pitch angle and our measure of bar strength, which we feel should be reported.
-Edits have been made to the text.
+Edits have been made to the text to note the limitations.
 
@@
 That's a specific observation that a more detailed model based on the simple model of Pringle & Dobbs (2019) would have to explain.
-We have made edits to the text to make this point.
+We have made edits to the text to make this point, which we think is a really nice addition to the paper.
 
@@
 
-We assume no selection bias in galaxy builder results, which is not necessarily valid, however the requirement for a lower limit of pitch angle is needed for the test.
+We assume no selection bias in Galaxy Builder results, which is not necessarily valid, however the requirement for a lower limit of pitch angle is needed for the test and we have added to the explanation of why in the text (also see comments above).
 
@@
 
-Following this feedback, we have added (see at end of the README) a description of the contents of each of the Python Notebooks.
+Following this feedback, we have additionally added (see at end of the README) a description of the contents of each of the Python Notebooks.
 

### 2021-03-31 referee_response2.md
Starting edits for 2nd report.

# Referee Response, second round.

> Major comments:

> 1. In the Introduction, the authors change the three criteria of the ordering in the Hubble classification sequence. However, Sandage (2005) repeats these several times: the size of the central bulge, the degree of resolution into condensations, and the openness of the spiral arms.

> The authors change the "degree of resolution into condensations" to "the degree of resolution of the spiral arms (how 'patchy' they are)". Yet the word patchy is never used in this context, and for flocculent galaxies (where there are no well described spiral arms), such as NGC 2841 (type SAT3 in the Revised Shapley Ames Catalog), an ordering is nonetheless given, in coherence with all three criteria.

> Please refrain from reformulating the Hubble classification to a scheme based on your own criteria: e.g. the prototype flocculent spiral galaxy NGC 2841 has condensations in its disc, but no clear spiral arms.

OK. We have made edits to use the recommended language from Sandage (2005) and refrained from adding any context to help modern readers (I know this is snark - which I will edit later).

> 2. In this context, it is striking that the authors now choose NGC 4689 as a example of a flocculent galaxy, even though spiral structure is visible, in particular the eastern arm. True, Elmegreen & Elmegreen (1987) list it as a class 3 spiral, i.e. in the same category as NGC 2841, but the spiral arms are clearly there in the 3.6 micron Spitzer image. I suggest using NGC 2841 as a good example for a flocculent galaxy, and note that for NGC 5055 there is a paper by Michele Thornley -  1996ApJ...469L..45T - pointing out that in the K'-band there is spiral structure seen, even though it is also listed as a prototype flocculent spiral in Elmegreen & Elmegreen (1987).

We have edited to use NGC 2841 as the example of a flocculent galaxy ** TO DO - I put the SDSS image of it in the drive.  **

> 3. The authors write on page 2 left side line 38: Spiral tightness is also observed to correlate with rotation curve shape (Seigar et al. 2005), with rising rotation curves creating more open spiral structure.

> I think you mean that galaxies with rising rotation curves have more open spiral structure. I don't think rotation curves are creating anything. More likely, the associated lower shear allows more open spirals.

We have edited this sentence as suggested.

> 4. In section 3.2.2 the authors write: One of the predictions of Manifold theory is that pitch angle increases with bar strength (Athanassoula et al. 2009b).

> Please make this section self-contained: ... pitch angle increases with bar strength (Athanassoula et al. 2009b), as evaluated at the Lagragian L_1 and L_2 points. The bar strength evaluated that manner is very specific, and in Athanassoula et al. 2009b it is shown that a correlation with Q_bar is much less strong. Since you use here p_bar as a proxy for bar strength, it is only remotely correlated with the bar strength defined in Athanassoula et al. 2009b. True, you admit that your test is not very convincing, but the interested reader should not have to consult explicitly the Athanassoula et al. 2009b paper to find out why.

We have edited this section as suggested.

> 5. In the conclusion section, the formulation of the conclusions on section 3.2.2. are much stronger than justified: you say that you cannot completely rule out the manifold theory, even though you admit that the proxy you use is not convincing. Since hurried readers could read this conclusion section without reading the bulk of the paper, you leave them with the impression that you almost rule out the Manifold theory, while this is not the case, since your proxy for bar strength is very poor, and your sample is small. Please harmonize your conclusions with the reality you describe in the main section.

> 6. In the conclusions, page 11 left side lines 48-49, you suggest that the absence of galaxies with low pitch angles is not a selection effect. How does this square with the data of Kennicutt (1981), where nearly all Sa galaxies, as classified in the RSA catalogue, have pitch angles less than 10 degrees? Are you sure that you are not missing the tightly wound early type spiral galaxies?

> I did a cursory inspection of the early type spirals in the Yu & Ho (2018) sample, and find that most of the early type spirals (S0/a, Sa and Sab) are actually barred spirals when checking the classification in Hyperleda and displaying them in the legacysurvey viewer. Perhaps the authors can check the distribution of the results of Kennicutt (1981) in a diagram such as Figure 9, to see whether his sample is in agreement with the ideas of Pringle & Dobbs (2019)?

> Minor comments and typos:

Thank you for noting these minor typos and other errors. We have corrected these, and some other ones we noted in editing.

> 1. page 2 right side line 50: disk - please use English spelling throughout: disc.

> 2. page 3 left side line 40: we further separate the galaxies ...

> 3. page 4 left side line 21: and report the pitch angle

> 4. page 4 left side line 53: which results in a sample favouring (3 times "which" is too much)

> 5. page 4 left side line 60: of these classifications

> 6. page 6 right side line 19: morphological

> 7. page 10 right side line 27: you write "pitch angle calculation; obtaining" but the ; should disappear

### 2021-03-31 sections/abstract.tex
Starting edits for 2nd report.

@@
 We present a new approach to deal with the large variations in pitch angle between different arms in a single galaxy, which obtains full posterior distributions on parameters.
-We make use of our pitch angles to examine previously reported links between bulge and bar strength and pitch angle, finding no correlation in our data {\bf (with a caveat that we use observational proxies for both bulge size and bar strength which differ from other work)}.
-We test a recent model for spiral arm winding, which predicts uniformity of the cotangent of pitch angle between some unknown upper and lower limits, {\bf finding our observations are consistent with} this model of transient and recurrent spiral pitch angle {\bf as long as the pitch angle at which most winding spirals dissipate or disappear is larger than 10$^\circ$.}
+We make use of our pitch angles to examine previously reported links between bulge and bar strength and pitch angle, finding no correlation in our data  (with a caveat that we use observational proxies for both bulge size and bar strength which differ from other work).
+We test a recent model for spiral arm winding, which predicts uniformity of the cotangent of pitch angle between some unknown upper and lower limits, finding our observations are consistent with this model of transient and recurrent spiral pitch angle as long as the pitch angle at which most winding spirals dissipate or disappear is larger than 10$^\circ$.

### 2021-03-31 sections/introduction.tex
Starting edits for 2nd report.

@@
 \citealt{2014PASA...31...35D}).
-Spirals are as diverse as the theories proposed to govern their evolution, from the quintessential pair of well-defined arcs of the grand design spiral, to the fragmented arm segments of the flocculent spiral, to the disjointed multi-armed spiral
-\sout{. These variations on structure account for $18\%$, $50\%$ and $32\%$ of the population respectively}
+Spirals are as diverse as the theories proposed to govern their evolution, from the quintessential pair of well-defined arcs of the grand design spiral, to the fragmented arm segments of the flocculent spiral, to the disjointed multi-armed spiral.
 (\citealt{2011ApJ...737...32E}; examples of each type are shown in Figure \ref{fig:spiral-galaxy-types}).
-The Hubble classification scheme \citep{1926ApJ....64..321H} and its revisions and expansions \citep{1961hag..book.....S,1991rc3..book.....D} contain detailed variations of different types of spiral galaxy, divided by the presence of a bar and ordered by \textbf{the degree of resolution of the spiral arms (how ``patchy” they are)}, how tightly wound they are and the prominence of a central bulge.
-\textbf{Building on this, \citet{1982MNRAS.201.1021E} found that flocculent spirals are more prevalent in unbarred, isolated galaxies.
-The presence of a bar, a binary companion or group membership result in a higher fraction of observed grand design spiral patterns.}
+The Hubble classification scheme \citep{1926ApJ....64..321H} and its revisions and expansions \citep{1961hag..book.....S,1991rc3..book.....D} contain detailed variations of different types of spiral galaxy, divided by the presence of a bar and ordered by the \textbf{openness of the spiral arms, the} degree of resolution \textbf{into condensations} and the prominence of a central bulge.
+Building on this, \citet{1982MNRAS.201.1021E} found that flocculent spirals are more prevalent in unbarred, isolated galaxies.
+The presence of a bar, a binary companion or group membership result in a higher fraction of observed grand design spiral patterns.
 
@@
   \caption{Examples of the different types of spiral galaxy present in the sky.
-The left column shows the \textbf{grand design spiral NGC 5248}, the middle shows the \textbf{many-armed spiral NGC-3184} and the right shows the \textbf{flocculent spiral NGC-4689}.
+The left column shows the grand design spiral NGC 5248.
+the middle shows the many-armed spiral NGC-3184 and the right shows the flocculent spiral NGC {\bf 2841}.
 Images were taken with the Sloan Digital Sky Survey Telescope.}
@@
 
-\textbf{
 Whatever kind of spiral is present in a disc galaxy, there is plenty of evidence to suggest that they have a significant role on the overall evolution of that galaxy.
@@
 Studies of spiral morphology have also found interesting correlations with other galactic properties, such as a correlation between the tightness of spiral arms and central mass concentration (\citealt{2019ApJ...871..194Y}, though neither \citealt{2017MNRAS.472.2263H} nor \citealt{2019MNRAS.487.1808M} found such a relation in large samples).
-Spiral tightness is also observed to correlate with with rotation curve shape (\citealt{2005MNRAS.359.1065S}), with rising rotation curves creating more open spiral structure.
+Spiral tightness is also observed to correlate with with rotation curve shape (\citealt{2005MNRAS.359.1065S}), with \textbf{galaxies with rising rotation curves having} more open spiral structure.
 These predictions and observations provide compelling reasons for continued investigation of the underlying rules and dynamics of spiral structure, as doing so is essential for understanding the secular evolution of disc galaxies.
-}
 
@@
 One of the fundamental assumptions of early work on spiral formation mechanisms (primarily QSDW) was that the disc of a galaxy, if unstable to spiral perturbations, would create a stable, static wave which would exist unchanged for many rotational periods \citep{1964ApJ...140..646L}.
-The motivation for static waves with small numbers of arms \sout{(with a preference for $m=2$)} was primarily observational: most disc galaxies observed at the time showed spiral structure with low spiral arm numbers, suggesting that spirals exist for a long time or are continually rebuilt.
-\textbf{This, in combination with theoretical arguments about the ``winding problem", motivated the original static density waves of \citet{1964ApJ...140..646L}, to which swing amplification was added by \citet{Toomre1981}  to provide a way to counteract the short lifetime of stellar density waves.}
+The motivation for static waves with small numbers of arms was primarily observational: most disc galaxies observed at the time showed spiral structure with low spiral arm numbers, suggesting that spirals exist for a long time or are continually rebuilt.
+This, in combination with theoretical arguments about the ``winding problem", motivated the original static density waves of \citet{1964ApJ...140..646L}, to which swing amplification was added by \citet{Toomre1981}  to provide a way to counteract the short lifetime of stellar density waves.
 
@@
 Many other galactic components may correlate with spiral morphology, including bulge fraction (\citealt{1975A&A....44..363Y}, \citealt{2013MNRAS.436.1074S}, \citealt{2019MNRAS.487.1808M}) and black hole mass (\citealt{2008ApJ...678L..93S}, \citealt{2017MNRAS.471.2187D}, \citealt{2019MS&E..571a2118A}).
-Larger bulges and more massive central black holes have both \textbf{been observed to} correlate with more tightly wound spiral arms.
-\textbf{We can also test this with the data presented in this paper.}
+Larger bulges and more massive central black holes have both been observed to correlate with more tightly wound spiral arms.
+We can also test this with the data presented in this paper.
 
-\textbf{This paper is structured as follows: In Section \ref{section:method} we introduce methods to measure galaxy pitch angle, and present our sample, and our Bayseian hierarchical modeling method making use of \textit{Galaxy Builder} to estimate galaxy and population pitch angles.
-Section \ref{section:constraints-on-galaxy-phi} presents our general constraints on pitch angles in our section, while} Section \ref{section:morphology-comparison} examines the correlation between pitch angle and bulge size implied by the Hubble sequence, and pitch angle and bar strength implied by Manifold theory.
+This paper is structured as follows: In Section \ref{section:method} we introduce methods to measure galaxy pitch angle, and present our sample, and our Bayseian hierarchical modeling method making use of \textit{Galaxy Builder} to estimate galaxy and population pitch angles.
+Section \ref{section:constraints-on-galaxy-phi} presents our general constraints on pitch angles in our section, while Section \ref{section:morphology-comparison} examines the correlation between pitch angle and bulge size implied by the Hubble sequence, and pitch angle and bar strength implied by Manifold theory.
 Section \ref{section:spiral_winding} investigates spiral arm winding using the test derived by \cite{2019arXiv190910291P} (uniformity of galaxy pitch angle in $\cot\;\phi$).
-\textbf{We provide a summary and conclusions in Section \ref{section:summary}.}  Where necessary, we make use of $H_0 = 70\ \text{km}\ \text{s}^{-1}\ \text{Mpc}^{-1}$.
+We provide a summary and conclusions in Section \ref{section:summary}.
+Where necessary, we make use of $H_0 = 70\ \text{km}\ \text{s}^{-1}\ \text{Mpc}^{-1}$.

### 2021-03-31 sections/method.tex
Starting edits for 2nd report.

--- 2021-02-26 sections/method.tex
@@
 where $\phi$ is the arm's pitch angle, $A$ is an amplitude coefficient and $\theta$ is the polar coordinate.
-\textbf{Different arms in a galaxy could have different values of $\phi$, however for each arm, $\phi$ is assumed to be constant with radius}.
+Different arms in a galaxy could have different values of $\phi$, however for each arm, $\phi$ is assumed to be constant with radius.
 One method used to obtain a pitch angle of a galaxy is to fit logarithmic spirals to individually identified arm segments and take the weighted mean of their pitch angles (which often vary by upwards of $10^\circ$, \citealt{2014ApJ...790...87D}).
@@
 
-Fourier analysis in one- and two-dimensions (as performed by \citealt{2019arXiv190804246D}, \citealt{2012ApJS..199...33D}, \citealt{2018MNRAS.474.2594M}, \textbf{and dating back to the seminal work of \citealt{ConsidereAthanassoula1988}}) is another widely used method of computationally obtaining galaxy pitch angles.
+Fourier analysis in one- and two-dimensions (as performed by \citealt{2019arXiv190804246D}, \citealt{2012ApJS..199...33D}, \citealt{2018MNRAS.474.2594M}, and dating back to the seminal work of \citealt{ConsidereAthanassoula1988}) is another widely used method of computationally obtaining galaxy pitch angles.
 Two-dimensional Fourier methods generally decompose a deprojected image of a galaxy into a superposition of logarithmic spirals between inner and outer annuli \citep{2012ApJS..199...33D} and reports the pitch angle with the highest amplitude as the galaxy's pitch angle.
@@
 It is unclear how the variation between pitch angles of individual arms impacts this measurement.
-\textbf{We note that while this method is able to model non-logarithmic spirals -- as a sum of logarithmic spirals with differing pitch angles, most applications use models which assume that the pitch angle is constant with radius, in some cases picking regions of a galaxy in which this is true - e.g.
-see Section 4.3.2 of \citealt{2012ApJS..199...33D}.}
+We note that while this method is able to model non-logarithmic spirals -- as a sum of logarithmic spirals with differing pitch angles, most applications use models which assume that the pitch angle is constant with radius, in some cases picking regions of a galaxy in which this is true - e.g.
+see Section 4.3.2 of \citealt{2012ApJS..199...33D}.
 
@@
 A histogram of stellar masses for our subset can be see in Section \ref{section:constraints-on-galaxy-phi}, where variation with stellar mass, to check the impact of this limited mass range, is also investigated.
-{\bf For the reader's convenience we also reproduce Figure 4 from \citet{2020arXiv200610450L} here (see Figure \ref{fig:stellarmass}) which shows the redshift and stellar mass distribution of our analysis sample compared to the full \textit{stellar mass-complete sample} from \citet{2017MNRAS.472.2263H}.
-Our choice (see \citet{2020arXiv200610450L} for details) to prefer lower redshift galaxies for {\it Galaxy Builder} analysis is clear in the mass distribution which which results in a sample which favours galaxies $9.5 < \log(M_* / M_\odot) < 10.0$, and includes a smaller number of spirals with masses up to  $\log(M_* / M_\odot) = 11.05$}
+For the reader's convenience we also reproduce Figure 4 from \citet{2020arXiv200610450L} here (see Figure \ref{fig:stellarmass}) which shows the redshift and stellar mass distribution of our analysis sample compared to the full \textit{stellar mass-complete sample} from \citet{2017MNRAS.472.2263H}.
+Our choice (see \citet{2020arXiv200610450L} for details) to prefer lower redshift galaxies for {\it Galaxy Builder} analysis is clear in the mass distribution which which results in a sample which favours galaxies $9.5 < \log(M_* / M_\odot) < 10.0$, and includes a smaller number of spirals with masses up to  $\log(M_* / M_\odot) = 11.05$.
 
@@
 Some galaxies in \citet{2020arXiv200610450L} were shown to volunteers a second time in a repeat validation subset to create a second aggregate model used to test internal consistency.
-{\bf Section 3.2 of \citet{2020arXiv200610450L} presents a comparison if these classifications to investigate volunteer consistency.
-We can see that arm number is highly reliable within $\pm1$ (93\% galaxies have arm number counts which agree in this range; 53\% have exactly identical arm number counts).} In this work, we combine the 30 classifications of galaxies in this validation subset with the 30 original classifications.
+Section 3.2 of \citet{2020arXiv200610450L} presents a comparison if these classifications to investigate volunteer consistency.
+We can see that arm number is highly reliable within $\pm1$ (93\% galaxies have arm number counts which agree in this range; 53\% have exactly identical arm number counts).
+In this work, we combine the 30 classifications of galaxies in this validation subset with the 30 original classifications.
 Clustering of drawn spiral arms and cleaning of points was then performed as detailed in \citet{2020arXiv200610450L}.
-We remove any galaxies for which no spiral arms were identified, resulting in a hierarchical data structure of {\bf 129 galaxies, 247 spiral arms and 238,433 points.
+We remove any galaxies for which no spiral arms were identified, resulting in a hierarchical data structure of 129 galaxies, 247 spiral arms and 238,433 points.
 This breaks down further to 68 galaxies (53\% of the sample) having two arms identified (meaning that they were marked by enough users to cluster into an arm), 19 (15\%) with three such arms, 4 (3\%) with four, and the remainder (38 or 29\%) with a single identified arm.
@@
 Given this, we do not recommend using these statistics to make general conclusions about the number of arms per galaxy.
-However, as our hierarchical model incorporates the uncertainty involved with missing spiral arms, the results related to pitch angle should not be significantly affected.}
+However, as our hierarchical model incorporates the uncertainty involved with missing spiral arms, the results related to pitch angle should not be significantly affected.
 
@@
 Logarithmic spirals have the desirable properties of a constant pitch angle and a small number of free parameters.
-\textbf{For this first analysis of the \textit{Galaxy Builder} models we choose to} make use of it here without an explicit comparison to other models.
+For this first analysis of the \textit{Galaxy Builder} models we choose to make use of it here without an explicit comparison to other models.
 A simple visual inspection of the fitted logarithmic spirals suggests that it is an appropriate model, however, a comparison of a logarithmic spiral profile to other spiral forms (i.e.
@@
 
-\textbf{As suggested by spiral formation models which correlate a galaxy wide pitch angle with galaxy wide properties,} we will assume that a given galaxy has some preferred value for arm pitch angle, $\phi_\mathrm{gal}$, and that the pitch angles of spiral arms in that galaxy, $\phi_\mathrm{arm}$, are constant with radius (giving logarithmic spirals) and drawn from a normal distribution centred on $\phi_\mathrm{gal}$, with some spread $\sigma_\mathrm{gal}$ common to all galaxies.
+As suggested by spiral formation models which correlate a galaxy wide pitch angle with galaxy wide properties, we will assume that a given galaxy has some preferred value for arm pitch angle, $\phi_\mathrm{gal}$, and that the pitch angles of spiral arms in that galaxy, $\phi_\mathrm{arm}$, are constant with radius (giving logarithmic spirals) and drawn from a normal distribution centred on $\phi_\mathrm{gal}$, with some spread $\sigma_\mathrm{gal}$ common to all galaxies.
 We truncate the normal distribution of spiral arm pitch angles in a single galaxy between the physical limits of {0\degree} (a ring) and {90\degree} (a ``spoke''), giving
@@
 
-\textbf{The choice to assume all galaxies show the same inter-arm variation in pitch angle (represented by a common value of $\sigma_\mathrm{gal}$ across all galaxies) was motivated by our small sample size and the low number of arms measured per galaxy.
+The choice to assume all galaxies show the same inter-arm variation in pitch angle (represented by a common value of $\sigma_\mathrm{gal}$ across all galaxies) was motivated by our small sample size and the low number of arms measured per galaxy.
 With this sample size we do not find, nor expect to be sensitive to variations in this parameter.
@@
 In a very detailed study of four very nearby spirals, \citet{HonigRead2015} conclude there are large variations of pitch angles  between spirals, and among arms in a given spiral, but made no comments as to if the variation was consistent with being constant.
-Further investigation of this issue in a larger \textit{Galaxy Builder} sample would be an interesting follow-up project.}
+Further investigation of this issue in a larger \textit{Galaxy Builder} sample would be an interesting follow-up project.
 
@@
 
-\textbf{These are conservative priors, which are not expected to have significant impact on the results.} The inverse gamma distribution is used to aid the convergence of the Hamiltonian Monte Carlo (HMC) algorithm used (discussed later).
+These are conservative priors, which are not expected to have significant impact on the results.
+The inverse gamma distribution is used to aid the convergence of the Hamiltonian Monte Carlo (HMC) algorithm used (discussed later).
 The Cauchy distribution is equivalent to the Student's t-distribution with one degree of freedom, and was chosen due to its fatter tails than the normal distribution.

### 2021-03-31 sections/results.tex
Starting edits for 2nd report.

--- 2021-02-25 sections/results.tex
@@
 The spread of arm pitch angle from the mean galaxy pitch angle can be seen in Figure \ref{fig:arm-pa-spread}, with points colour-coded by the number of arms measured for a galaxy.
-We see a slight drop in the expectation values of galaxy pitch angle ($E[\phi_\mathrm{gal}]$) compared to the expectation of arm pitch angles ($E[\phi_\mathrm{arm}]$) at small galaxy pitch angles, {\bf which is caused by a combination of the} truncation of $\phi_\mathrm{gal}$ at {0\degree} {\bf and the large spread (so the mean value differs from the mode as the distribution is highly skewed).}
+We see a slight drop in the expectation values of galaxy pitch angle ($E[\phi_\mathrm{gal}]$) compared to the expectation of arm pitch angles ($E[\phi_\mathrm{arm}]$) at small galaxy pitch angles, which is caused by a combination of the truncation of $\phi_\mathrm{gal}$ at {0\degree} and the large spread (so the mean value differs from the mode as the distribution is highly skewed).
 
@@
 
-\textbf{In Figure \ref{fig:stellar-mass-phigal} we present the stellar mass distribution of the sample, and investigate how the global galaxy pitch angle depends on this parameter.
+In Figure \ref{fig:stellar-mass-phigal} we present the stellar mass distribution of the sample, and investigate how the global galaxy pitch angle depends on this parameter.
 The majority of our sample has stellar masses $9.5 < \log(M_* / M_\odot) < 10.0$, and galaxy average pitch angles 10$^\circ$--20$^\circ$.
@@
 We observe no significant trend of the global pitch angle with mass, although there is a hint that more massive spirals may on average have more tightly wound arms.
-}
 
@@
 
-Morphological classification commonly links bulge size to spiral tightness, and such a link is implied by the Hubble Sequence (\citealt{2005ARA&A..43..581S}, \citealt{2009MNRAS.393.1531G}, \citealt{2013seg..book..155B}), \textbf{although small bulge Sa galaxies have been noted for decades (e.g.
-for a review see \citet{2005ARA&A..43..581S}; this is also noted in \citealt{2019MNRAS.487.1808M})}.
+Morphological classification commonly links bulge size to spiral tightness, and such a link is implied by the Hubble Sequence (\citealt{2005ARA&A..43..581S}, \citealt{2009MNRAS.393.1531G}, \citealt{2013seg..book..155B}), although small bulge Sa galaxies have been noted for decades (e.g.
+for a review see \citet{2005ARA&A..43..581S}; this is also noted in \citealt{2019MNRAS.487.1808M}).
 Some recent studies have indeed reported a link between measured spiral galaxy pitch angle and bulge size (i.e.
 \citealt{2019ApJ...873...85D}), while others have not found any significant correlation \citep{2019MNRAS.487.1808M}.
-\textbf{The differing results of this may depend on the details of how both the bulge size, and spiral pitch angles are measured and suggest further investigation is needed.} We investigate this relationship here using a measure of bulge prominence from Galaxy Zoo 2, as Equation 3 in \citet{2019MNRAS.487.1808M}:
+The differing results of this may depend on the details of how both the bulge size, and spiral pitch angles are measured and suggest further investigation is needed.
+We investigate this relationship here using a measure of bulge prominence from Galaxy Zoo 2, as Equation 3 in \citet{2019MNRAS.487.1808M}:
 
@@
 
-One of the predictions of Manifold theory is that pitch angle increases with bar strength \citep{2009MNRAS.400.1706A}.
-\textbf{In this work we do not have any direct measurement of bar strength; nor are there significant numbers of strongly barred galaxies in our sample.
-However in an attempt to} investigate this relationship in our data, we make use of Galaxy Zoo 2's bar fraction ($p_\mathrm{bar}$), which has been demonstrated to be a good measure of bar length \citep{Willett2013:1308.3496v2} and bar strength \citep{2012MNRAS.423.1485S,2012MNRAS.424.2180M,2018MNRAS.473.4731K} and therefore a good measure of the torque applied on the disc gas.
+One of the predictions of Manifold theory is that pitch angle increases with bar strength \textbf{as evaluated at the Lagrangian L1 and L2 points}  \citep{2009MNRAS.400.1706A}.
+In this work we do not have any \textbf{similar} measurement of bar strength \textbf{and we note that \citet{2009MNRAS.400.1706A} find a much weaker correlation with other measures of bar strength; we also do not have} significant numbers of strongly barred galaxies in our sample.
+However in an attempt to investigate this relationship in our data, we make use of Galaxy Zoo 2's bar fraction ($p_\mathrm{bar}$), which has been demonstrated to be a good measure of bar length \citep{Willett2013:1308.3496v2} and bar strength \citep{2012MNRAS.423.1485S,2012MNRAS.424.2180M,2018MNRAS.473.4731K} and therefore a good measure of the torque applied on the disc gas.
 
@@
 Confidence intervals are shown, with moving rightwards indicating more confidence in rejecting the null hypothesis that the compared values were drawn from the same parent distribution.
-We cannot reject the null hypotheis at the 1\% level for any of the tests conducted, meaning there is no evidence in this sample that bulge size or bar strength impacts pitch angle.}
+We cannot reject the null hypothesis at the 1\% level for any of the tests conducted, meaning there is no evidence in this sample that bulge size or bar strength impacts pitch angle.}
   \label{fig:ad-morphology-test}
@@
 
-The fact that we do not find any link between \textbf{our measure of bar strength (based on the prominance of the bar in Galaxy Zoo 2) and pitch angle is suggestive that there is actually no link between bar strength and pitch angle, which would exclude Manifold theory as the primary mechanism driving the evolution of the spirals in our sample.
-However since we use only a proxy for bar strength, which has not been well tested, this is not conclusive.}
+The fact that we do not find any link between our measure of bar strength (based on the prominence of the bar in Galaxy Zoo 2) and pitch angle is suggestive that there is actually no link between bar strength and pitch angle, which would exclude Manifold theory as the primary mechanism driving the evolution of the spirals in our sample.
+However since we use only a proxy for bar strength, which has not been well tested, this is not conclusive.
 

### 2021-04-01 referee_response2.md
Completed most of the text changes.

@@
 # Referee Response, second round.
+
+We thank the referee for these additional comments. As before, major changes have been hi-lighted in bold.
 
@@
 
-OK. We have made edits to use the recommended language from Sandage (2005) and refrained from adding any context to help modern readers (I know this is snark - which I will edit later).
+We have made edits to use the recommended language.
 
@@
 
+We have further toned down the conclusion on Manifold theory and believe it now includes most of the caveats given in the main section.
+
 > 6.
@@
 Are you sure that you are not missing the tightly wound early type spiral galaxies?
+
+Our sample selection is quite different to Kennicutt (1981) which may be one factor.
+In addition a colleague compared Kennicutt (1981) pitch angles to those measured using the Spacfire software, and found that Sparcfire measured more open arms in general for the early type spirals.
+We have not run the Kennicutt sample through Galaxy Builder, but will make sure to do that for the next phase of the project.
 
@@
 
-Thank you for noting these minor typos and other errors. We have corrected these, and some other ones we noted in editing.
+Thank you for noting these minor typos and other errors. We have corrected these, and a handful of other typos we noted in editing.
 
 > 1. page 2 right side line 50: disk - please use English spelling throughout: disc.
-
 > 2. page 3 left side line 40: we further separate the galaxies ...
-
 > 3. page 4 left side line 21: and report the pitch angle
-
 > 4. page 4 left side line 53: which results in a sample favouring (3 times "which" is too much)
-
 > 5. page 4 left side line 60: of these classifications
-
 > 6. page 6 right side line 19: morphological
-
 > 7. page 10 right side line 27: you write "pitch angle calculation; obtaining" but the ; should disappear

### 2021-04-01 sections/introduction.tex
Completed most of the text changes.

@@
 ``wave amplification by stimulated emission of radiation'', \citealt{1976ApJ...205..363M}; swing amplification, \citealt{1965MNRAS.130..125G}), but do not require the idealistic disc conditions required for the formation and maintenance of a stationary wave.
-The pitch angles of these transient spiral arms will decrease due to the differential rotation of the disk, with the density of the arm peaking at some critical pitch angle, before dissipating to be reformed.
+The pitch angles of these transient spiral arms will decrease due to the differential rotation of the disc, with the density of the arm peaking at some critical pitch angle, before dissipating to be reformed.
 
@@
 
-Using Galaxy Zoo 2 data \citep{Willett2013:1308.3496v2} we further separate of galaxies by the presence and strength of a stellar bar.
+Using Galaxy Zoo 2 data \citep{Willett2013:1308.3496v2} we further separate the galaxies by the presence and strength of a stellar bar.
 This will allow us to test simulations of gas in barred galaxies, which often demonstrate that bars can drive long-term spiral evolution \citep{2008A&A...489..115R}, or boost transient spiral structure \citep{2012MNRAS.426..167G}.
@@
 
-This paper is structured as follows: In Section \ref{section:method} we introduce methods to measure galaxy pitch angle, and present our sample, and our Bayseian hierarchical modeling method making use of \textit{Galaxy Builder} to estimate galaxy and population pitch angles.
+This paper is structured as follows: In Section \ref{section:method} we introduce methods to measure galaxy pitch angle, and present our sample, and our Bayesian hierarchical modelling method making use of \textit{Galaxy Builder} to estimate galaxy and population pitch angles.
 Section \ref{section:constraints-on-galaxy-phi} presents our general constraints on pitch angles in our section, while Section \ref{section:morphology-comparison} examines the correlation between pitch angle and bulge size implied by the Hubble sequence, and pitch angle and bar strength implied by Manifold theory.

### 2021-04-01 sections/method.tex
Completed most of the text changes.

@@
 Fourier analysis in one- and two-dimensions (as performed by \citealt{2019arXiv190804246D}, \citealt{2012ApJS..199...33D}, \citealt{2018MNRAS.474.2594M}, and dating back to the seminal work of \citealt{ConsidereAthanassoula1988}) is another widely used method of computationally obtaining galaxy pitch angles.
-Two-dimensional Fourier methods generally decompose a deprojected image of a galaxy into a superposition of logarithmic spirals between inner and outer annuli \citep{2012ApJS..199...33D} and reports the pitch angle with the highest amplitude as the galaxy's pitch angle.
+Two-dimensional Fourier methods generally decompose a deprojected image of a galaxy into a superposition of logarithmic spirals between inner and outer annuli \citep{2012ApJS..199...33D} and report the pitch angle with the highest amplitude as the galaxy's pitch angle.
 \citet{2020MNRAS.493.3854H} combined Fourier analysis of spiral galaxies with a visual tracing of spiral arms, successfully eliminating observed bias in a sample of toy images of galaxies.
@@
 For the reader's convenience we also reproduce Figure 4 from \citet{2020arXiv200610450L} here (see Figure \ref{fig:stellarmass}) which shows the redshift and stellar mass distribution of our analysis sample compared to the full \textit{stellar mass-complete sample} from \citet{2017MNRAS.472.2263H}.
-Our choice (see \citet{2020arXiv200610450L} for details) to prefer lower redshift galaxies for {\it Galaxy Builder} analysis is clear in the mass distribution which which results in a sample which favours galaxies $9.5 < \log(M_* / M_\odot) < 10.0$, and includes a smaller number of spirals with masses up to  $\log(M_* / M_\odot) = 11.05$.
+Our choice (see \citet{2020arXiv200610450L} for details) to prefer lower redshift galaxies for {\it Galaxy Builder} analysis is clear in the mass distribution which results in a sample favouring galaxies $9.5 < \log(M_* / M_\odot) < 10.0$, and includes a smaller number of spirals with masses up to  $\log(M_* / M_\odot) = 11.05$.
 
 \begin{figure*}
-  \includegraphics[width=17.7cm]{plots/stellarmass.png}
+  \includegraphics[width=17.7cm]{plots/stellar_mass_selection_plot.pdf}
   \caption{A plot of redshift against stellar mass for the \textit{stellar mass-complete sample} from \citet{2017MNRAS.472.2263H}; the subset we use for analysis in Galaxy Builder are shown in red.
@@
 Some galaxies in \citet{2020arXiv200610450L} were shown to volunteers a second time in a repeat validation subset to create a second aggregate model used to test internal consistency.
-Section 3.2 of \citet{2020arXiv200610450L} presents a comparison if these classifications to investigate volunteer consistency.
+Section 3.2 of \citet{2020arXiv200610450L} presents a comparison of these classifications to investigate volunteer consistency.
 We can see that arm number is highly reliable within $\pm1$ (93\% galaxies have arm number counts which agree in this range; 53\% have exactly identical arm number counts).
@@
 
-Spiral arm points are deprojected to a face-on orientation using the disk inclination and position angle obtained through photometric model fitting performed in \citet{2020arXiv200610450L}.
+Spiral arm points are deprojected to a face-on orientation using the disc inclination and position angle obtained through photometric model fitting performed in \citet{2020arXiv200610450L}.
 Arms are individually corrected to all have the same chirality (a pitch angle greater than or equal to zero) using the logarithmic spiral fit in \citet{2020arXiv200610450L}.

### 2021-04-01 sections/results.tex
Completed most of the text changes.

@@
 
-We make use of Galaxy Zoo 2 data for morhpological comparison.
+We make use of Galaxy Zoo 2 data for morphological comparison.
 Two of the galaxies in our sample could not be matched to Galaxy Zoo 2 data, and as such have been dropped from this comparison (leaving 127 galaxies).
@@
 
-One of the predictions of Manifold theory is that pitch angle increases with bar strength \textbf{as evaluated at the Lagrangian L1 and L2 points}  \citep{2009MNRAS.400.1706A}.
-In this work we do not have any \textbf{similar} measurement of bar strength \textbf{and we note that \citet{2009MNRAS.400.1706A} find a much weaker correlation with other measures of bar strength; we also do not have} significant numbers of strongly barred galaxies in our sample.
+One of the predictions of Manifold theory is that pitch angle increases with bar strength \textbf{as evaluated by the Quadrupole moment, $Q$ at the Lagrangian $L_1$ point (a value that differs from typical ``bar strength'', which is this value averaged over all radii in the bar; \citealt{2009MNRAS.400.1706A})}.
+In this work we do not have any \textbf{similar} measurement of bar strength \textbf{and we note that \citet{2009MNRAS.400.1706A} caution that other measures of bar strength may not show this relation; we also do not have} significant numbers of strongly barred galaxies in our sample.
 However in an attempt to investigate this relationship in our data, we make use of Galaxy Zoo 2's bar fraction ($p_\mathrm{bar}$), which has been demonstrated to be a good measure of bar length \citep{Willett2013:1308.3496v2} and bar strength \citep{2012MNRAS.423.1485S,2012MNRAS.424.2180M,2018MNRAS.473.4731K} and therefore a good measure of the torque applied on the disc gas.
@@
 If $\Omega_p$ instead varies with radius we would expect $\cot{\phi}$ to be uniformly distributed between $\cot{\phi_\mathrm{max}}$ and $\cot{\phi_\mathrm{min}}$.
-{\bf The model presented in \citet{2019arXiv190910291P} does not give any physical justification for what $\cot{\phi_\mathrm{max}}$ and $\cot{\phi_\mathrm{min}}$ should be.
-}
+The model presented in \citet{2019arXiv190910291P} does not give any physical justification for what $\cot{\phi_\mathrm{max}}$ and $\cot{\phi_\mathrm{min}}$ should be.
 
-To test this theory, \citet{2019arXiv190910291P} used a Kolmogorov-Smirnov test to examine whether {\bf a sample of 113} galaxies with measured pitch angles was likely to have been drawn from a distribution uniform in its cotangent.
-Pitch angles were measured {\bf by \citet{2019ApJ...871..194Y}} using discrete Fourier transformations in one- and two-dimensions, and as such do not account for inter-arm variations.
-They {\bf conclude the model works within limits of} $\cot{\phi} \in [1.00, 4.75]$ (roughly $11.9^\circ < \phi < 45.0^\circ$), motivated by examination of the data.
+To test this theory, \citet{2019arXiv190910291P} used a Kolmogorov-Smirnov test to examine whether a sample of 113 galaxies with measured pitch angles was likely to have been drawn from a distribution uniform in its cotangent.
+Pitch angles were measured by \citet{2019ApJ...871..194Y} using discrete Fourier transformations in one- and two-dimensions, and as such do not account for inter-arm variations.
+They conclude the model works within limits of $\cot{\phi} \in [1.00, 4.75]$ (roughly $11.9^\circ < \phi < 45.0^\circ$), motivated by examination of the data.
 
@@
 We will make use of the marginalized Anderson-Darling test described above, and examine winding on a per-arm basis, as well as a per-galaxy basis.
-Observation of the distribution of arm pitch angles in our sample (Figure \ref{fig:pa-cot-distributions}) suggests {\bf they are close to uniform in cotangent} within limits of $15^\circ < \phi < 50.0^\circ$.
+Observation of the distribution of arm pitch angles in our sample (Figure \ref{fig:pa-cot-distributions}) suggests they are close to uniform in cotangent within limits of $15^\circ < \phi < 50.0^\circ$.
 
@@
 
-This result suggests that \sout{we cannot rule out} {\bf our data are consistent with} a cot-uniform source distribution for galaxy pitch angle, but the large uncertainty in $\phi_\mathrm{gal}$ makes it difficult to make any conclusive statements.
+This result suggests that our data are consistent with a cot-uniform source distribution for galaxy pitch angle, but the large uncertainty in $\phi_\mathrm{gal}$ makes it difficult to make any conclusive statements.
 This result is also highly sensitive to the lower limit of $\phi$: decreasing it to {10\degree} results in us rejecting the cot-uniform model at greater than the 0.1\% level for 96\% of the posterior samples.
-{\bf We can conclude from this that the \citet{2019arXiv190910291P} cot-uniform model is an adequate fit to the data, as long as the minimum pitch angle, $\cot{\phi_\mathrm{min}}$, at which the majority of winding dissipate or disappear is  $\phi_\mathrm{min} > $ {10\degree}, and more confidently $\phi_\mathrm{min} = $ {15\degree}.
-We reiterate that there is no prediction in \citet{2019arXiv190910291P} as to what this minimum pitch angle should be, so our observation constrains the allowed range.} \sout{As we have no information available on the selection biases present for classification of extremely loose or tight spiral arms in \textit{Galaxy Builder}, we choose to keep the less strict limit of {15\degree}.}
+We can conclude from this that the \citet{2019arXiv190910291P} cot-uniform model is an adequate fit to the data, as long as the minimum pitch angle, $\cot{\phi_\mathrm{min}}$, at which the majority of winding dissipate or disappear is  $\phi_\mathrm{min} > $ {10\degree}, and more confidently $\phi_\mathrm{min} = $ {15\degree}.
+We reiterate that there is no prediction in \citet{2019arXiv190910291P} as to what this minimum pitch angle should be, so our observation constrains the allowed range.
 
@@
 The resulting distribution of Anderson-Darling statistics is shown in in the lower panel of Figure \ref{fig:ad-cot-test}.
-This result is highly consistent with the model for spiral winding proposed by \citet{2019arXiv190910291P} {\bf with $\cot{\phi_\mathrm{min}} = $ {15\degree} } and can be interpreted as evidence that spirals are formed through local disc perturbations, and are primarily governed by local forces.
+This result is highly consistent with the model for spiral winding proposed by \citet{2019arXiv190910291P} with $\cot{\phi_\mathrm{min}} = $ {15\degree} and can be interpreted as evidence that spirals are formed through local disc perturbations, and are primarily governed by local forces.
 
@@
 In this instance, we would not be able to reject the null hypothesis at the 1\% level for either $\phi_\mathrm{gal}$ or $\phi_\mathrm{arm}$, meaning our sample is consistent with a cot uniform distribution.
-{\bf The larger error in $\phi_\mathrm{gal}$  means that this result is more significant for $\phi_\mathrm{arm}$, which is also physically motivated, as arms can wind independently.}}
+The larger error in $\phi_\mathrm{gal}$  means that this result is more significant for $\phi_\mathrm{arm}$, which is also physically motivated, as arms can wind independently.}
   \label{fig:ad-cot-test}

### 2021-04-01 sections/summary.tex
Completed most of the text changes.

@@
 This paper presents a new Bayesian approach to estimate galaxy pitch angle, making use of citizen science results to measure spiral arms through photometric modelling.
-We introduce an adaptation of the Anderson-Darling test, which we name the \textit{marginalized Anderson-Darling test}, to incorporate full Bayesian posterior probabilities and utilize this test to investigate theories governing spiral formation and evolution.
+We introduce an adaptation of the Anderson-Darling test, which we name the \textit{marginalized Anderson-Darling test}, to incorporate full Bayesian posterior probabilities and use this test to investigate theories governing spiral formation and evolution.
 
-The hierarchical Bayesian approach implemented in this paper allows a more thorough examination of pitch angle than length-weighted pitch angle calculation; obtaining posterior distributions of measured parameters.
-It better accounts for the large variations observed in inter-arm pitch angle than Fourier analysis, which assumes all arms in a given \textbf{symmetric} mode have the same pitch angle.
+The hierarchical Bayesian approach implemented in this paper allows a more thorough examination of pitch angle than length-weighted pitch angle calculation obtaining posterior distributions of measured parameters.
+It better accounts for the large variations observed in inter-arm pitch angle than Fourier analysis, which assumes all arms in a given symmetric mode have the same pitch angle.
 In this work, we find that the mean inter-arm difference in pitch angle is $11.0^\circ\pm 0.9^\circ$.
@@
 
-We do not find any link between bar strength and pitch angle \textbf{in our sample.
-However, rather that a direct measure of bar strength, we make use of an available parameter which correlates with bar strength, so at best this observation} suggests that the primary mechanism driving the evolution of the spirals in our sample is not Manifold theory (see Section \ref{section:morphology-comparison-bar}), \textbf{but cannot completely rule it out.}
+We do not find any link between {\bf our measure of} bar strength and pitch angle in our sample.
+However, rather that a direct measure of bar strength, we make use of an available parameter which correlates with bar strength, so at best this observation {\bf is suggestive} that the primary mechanism driving the evolution of the spirals in our sample {\bf may not be} Manifold theory (see Section \ref{section:morphology-comparison-bar}), \textbf{however, since this is not the measure of bar strength predicted to correlate with pitch angle by \citet{2009MNRAS.400.1706A}, and those authors caution that the details of the bar strength measure can wash out the predicted correlation, this is not strong evidence against Manifold theory models}.
 
 Our results are consistent with spiral winding of the form described by \citet{2019arXiv190910291P}, in which spiral arms are transient and recurrent, evolve through mechanisms such as swing-amplification \citep{1965MNRAS.130..125G} and which wind up over time.
-\textbf{This model predict a distribution of pitch angles that is uniform in cotangent space across some range.
+This model predict a distribution of pitch angles that is uniform in cotangent space across some range.
 No prediction is provided as to what that range should be.
-Our data are consistent with this model, if the minimum pitch angle is $\phi_\mathrm{min} = $ {15\degree}, but rule it out if the minimum pitch angle is $\phi_\mathrm{min} = $ {10\degree}}.
+Our data are consistent with this model, if the minimum pitch angle is $\phi_\mathrm{min} = $ {15\degree}, but rule it out if the minimum pitch angle is $\phi_\mathrm{min} = $ {10\degree}.
 The assumptions of this model of spiral winding are highly simplistic, and it leaves many unanswered questions: what determines the limits on $\phi$?
 Is the spiral arm equally apparent at all pitch angles, or is a selection effect present?
-\textbf{Our observations suggest that any further development of this model needs to predict that the minimum pitch angle, $\phi_\mathrm{min}>${10\degree}.} This result is also not evidence against QSDW, as our distribution of pitch angles may be dictated by other factors such as disk shear.
+Our observations suggest that any further development of this model needs to predict that the minimum pitch angle, $\phi_\mathrm{min}>${10\degree}.This result is also not evidence against QSDW, as our distribution of pitch angles may be dictated by other factors such as disc shear.
 
-In this work, we assume that spiral arms are equally likely to be identified and recovered at all pitch angles, \textbf{which suggests the absence of galaxies at low pitch angles is not a selection effect}.
+In this work, we assume that spiral arms are equally likely to be identified and recovered at all pitch angles, which suggests the absence of galaxies at low pitch angles is not a selection effect.
 This is not an unfair assumption given the amount of human effort that went into obtaining spiral arm measurements (more so than any other pitch angle measurement method, with each galaxy receiving at least 30 human classifications).
@@
 
-\textbf{We have presented evidence that} the methodology proposed here is a robust solution to the problems facing investigation of spiral morphology, namely that of reliably identifying spiral arms, and properly accounting for the spread in pitch angles of arms within a galaxy.
-This is one of the largest samples for which this test has been done \textbf{and is scaleable to larger samples}; such a sample would make possible further comparisons, such as splitting galaxies into spiral type (grand design / many-armed / flocculent), examining the differences between populations, \textbf{investigating if the interarm spread depends on other galaxy properties.}
+We have presented evidence that the methodology proposed here is a robust solution to the problems facing investigation of spiral morphology, namely that of reliably identifying spiral arms, and properly accounting for the spread in pitch angles of arms within a galaxy.
+This is one of the largest samples for which this test has been done and is scaleable to larger samples; such a sample would make possible further comparisons, such as splitting galaxies into spiral type (grand design / many-armed / flocculent), examining the differences between populations, investigating if the interarm spread depends on other galaxy properties.
 

### 2021-04-01 referee_response2.md
Added pitch angle files, and a thought about sample selection

@@
 Our sample selection is quite different to Kennicutt (1981) which may be one factor.
+It seems possibly that very tightly wound spirals may be more likely to be misidentified as smooth galaxies, and therefore not make it into the spiral sample in Galaxy Zoo at all.
 In addition a colleague compared Kennicutt (1981) pitch angles to those measured using the Spacfire software, and found that Sparcfire measured more open arms in general for the early type spirals.
-We have not run the Kennicutt sample through Galaxy Builder, but will make sure to do that for the next phase of the project.
+We have not run the Kennicutt sample through Galaxy Builder, so we cannot reliably comment if that would be an effect here, but will make sure to do that for the next phase of the project.
 

### 2021-04-05 referee_response2.md
More tweaks on sample selection comments
Final edits from KLM (maybe)

@@
 
-We have edited to use NGC 2841 as the example of a flocculent galaxy ** TO DO - I put the SDSS image of it in the drive.  **
+We have edited to use NGC 2841 as the example of a flocculent galaxy.
 
@@
 
-Our sample selection is quite different to Kennicutt (1981) which may be one factor.
-It seems possibly that very tightly wound spirals may be more likely to be misidentified as smooth galaxies, and therefore not make it into the spiral sample in Galaxy Zoo at all.
-In addition a colleague compared Kennicutt (1981) pitch angles to those measured using the Spacfire software, and found that Sparcfire measured more open arms in general for the early type spirals.
+Our sample selection is quite different to Kennicutt (1981) which we suspect is the primary factor.
+For example the most distant of the Kennicutt galaxies is at a similar redshift the lower limit of our sample.
+This means the mass range covered will be very different too, with the Kennicutt sample likely weighted to lower mass galaxies.
+Another factor, is that we have found that very tightly wound spirals may be misidentified as smooth galaxies at larger distances, and therefore not make it into the spiral sample in Galaxy Zoo at all.
+In addition a co-author (BK) compared Kennicutt (1981) pitch angles to those measured using the Spacfire software, and found that Sparcfire measured more open arms in general for the early type spirals.
 We have not run the Kennicutt sample through Galaxy Builder, so we cannot reliably comment if that would be an effect here, but will make sure to do that for the next phase of the project.
+
+ We have added some additional lines about sample selection.
 
@@
 Perhaps the authors can check the distribution of the results of Kennicutt (1981) in a diagram such as Figure 9, to see whether his sample is in agreement with the ideas of Pringle & Dobbs (2019)?
+
+ We did this, and the Kennicutt sample is peaked at lower pitch angle, so the agreement is not as good.
+We have added a note about this to the paper, including comments on sample selection, and we include the plot for the referee.
 

### 2021-04-05 sections/summary.tex
Final edits from KLM (maybe)

@@
 We do not find any link between {\bf our measure of} bar strength and pitch angle in our sample.
-However, rather that a direct measure of bar strength, we make use of an available parameter which correlates with bar strength, so at best this observation {\bf is suggestive} that the primary mechanism driving the evolution of the spirals in our sample {\bf may not be} Manifold theory (see Section \ref{section:morphology-comparison-bar}), \textbf{however, since this is not the measure of bar strength predicted to correlate with pitch angle by \citet{2009MNRAS.400.1706A}, and those authors caution that the details of the bar strength measure can wash out the predicted correlation, this is not strong evidence against Manifold theory models}.
+However, rather that a direct measure of bar strength, we make use of an available parameter which correlates with bar strength, so at best this observation {\bf is suggestive} that the primary mechanism driving the evolution of the spirals in our sample {\bf may not be} Manifold theory (see Section \ref{section:morphology-comparison-bar}).
+\textbf{Since this is not the measure of bar strength predicted to correlate with pitch angle by \citet{2009MNRAS.400.1706A}, and those authors caution that the details of the bar strength measure can wash out the predicted correlation, this is not strong evidence against Manifold theory models}.
 
@@
 
-In this work, we assume that spiral arms are equally likely to be identified and recovered at all pitch angles, which suggests the absence of galaxies at low pitch angles is not a selection effect.
+In this work, we assume that spiral arms are equally likely to be identified and recovered at all pitch angles, which suggests the absence of galaxies at low pitch angles is not {\bf due to an inability of us to measure such arms}.
 This is not an unfair assumption given the amount of human effort that went into obtaining spiral arm measurements (more so than any other pitch angle measurement method, with each galaxy receiving at least 30 human classifications).
-The galaxy sample used is not selected to be representative of the general spiral population, but is comparable in size to those used in other similar studies (\citealt{2013MNRAS.436.1074S}, \citealt{2019ApJ...871..194Y}, \citealt{2019arXiv190910291P}) and covers a range of masses ($9.45 < \log(M_* / M_\odot) < 11.05$) and spiral types.
+The galaxy sample used is {\bf a random subset of a volume limited sample (see Figure \ref{fig:stellarmass}), and} is comparable in size to those used in other similar studies (\citealt{2013MNRAS.436.1074S}, \citealt{2019ApJ...871..194Y}, \citealt{2019arXiv190910291P}).
+The sample covers a range of masses ($9.45 < \log(M_* / M_\odot) < 11.05$) and spiral types, {\bf however it is possible that tightly wound spirals are preferentially missed by the pre-selection from {\it Galaxy Zoo}, if they are less obviously identified as having spirals at these distances ($0.02<z<0.055$).
+The \citet{1981AJ.....86.1847K} sample of 113 much more nearby spirals (all at $z<0.019$ and most at $z<0.009$), includes several with arms at much lower pitch angles, but very few arms which are loosely wound (none with $\cot \phi > 31^\circ$), meaning it does not match the same $\cot \phi$ constant model well, although as at the time the sample was set by data availability, it is unclear how conclusive this is.
+In a future version of {\it Galaxy Builder} we intend to include this sample as a comparison set.}
 

### 2021-04-05 sections/summary.tex
Added note that K81 comment on their own sample incompleteness (missing open arms)

@@
 The sample covers a range of masses ($9.45 < \log(M_* / M_\odot) < 11.05$) and spiral types, {\bf however it is possible that tightly wound spirals are preferentially missed by the pre-selection from {\it Galaxy Zoo}, if they are less obviously identified as having spirals at these distances ($0.02<z<0.055$).
-The \citet{1981AJ.....86.1847K} sample of 113 much more nearby spirals (all at $z<0.019$ and most at $z<0.009$), includes several with arms at much lower pitch angles, but very few arms which are loosely wound (none with $\cot \phi > 31^\circ$), meaning it does not match the same $\cot \phi$ constant model well, although as at the time the sample was set by data availability, it is unclear how conclusive this is.
+The \citet{1981AJ.....86.1847K} sample of 113 much more nearby spirals (all at $z<0.019$ and most at $z<0.009$), includes several with arms at much lower pitch angles, but very few arms which are loosely wound (none with $\cot \phi > 31^\circ$), meaning it does not match the same $\cot \phi$ constant model well, although as at the time the sample was set by data availability, it is unclear how conclusive this is, and \citet{1981AJ.....86.1847K} note the incompleteness of their sample for open armed spirals.
 In a future version of {\it Galaxy Builder} we intend to include this sample as a comparison set.}

### 2021-04-13 sections/introduction.tex
changes flocculent galaxy used to match referee's suggestion
Removes highlighting ready for final submission

@@
 (\citealt{2011ApJ...737...32E}; examples of each type are shown in Figure \ref{fig:spiral-galaxy-types}).
-The Hubble classification scheme \citep{1926ApJ....64..321H} and its revisions and expansions \citep{1961hag..book.....S,1991rc3..book.....D} contain detailed variations of different types of spiral galaxy, divided by the presence of a bar and ordered by the \textbf{openness of the spiral arms, the} degree of resolution \textbf{into condensations} and the prominence of a central bulge.
+The Hubble classification scheme \citep{1926ApJ....64..321H} and its revisions and expansions \citep{1961hag..book.....S,1991rc3..book.....D} contain detailed variations of different types of spiral galaxy, divided by the presence of a bar and ordered by the openness of the spiral arms, the degree of resolution into condensations and the prominence of a central bulge.
 Building on this, \citet{1982MNRAS.201.1021E} found that flocculent spirals are more prevalent in unbarred, isolated galaxies.
@@
 \begin{figure*}
-  \includegraphics[width=15cm]{plots/galaxy_types.jpg}
+  \includegraphics[width=15cm]{plots/galaxy_types_remade.jpg}
   \caption{Examples of the different types of spiral galaxy present in the sky.
 The left column shows the grand design spiral NGC 5248.
-the middle shows the many-armed spiral NGC-3184 and the right shows the flocculent spiral NGC {\bf 2841}.
+the middle shows the many-armed spiral NGC-3184 and the right shows the flocculent spiral NGC 2841.
 Images were taken with the Sloan Digital Sky Survey Telescope.}
@@
 Studies of spiral morphology have also found interesting correlations with other galactic properties, such as a correlation between the tightness of spiral arms and central mass concentration (\citealt{2019ApJ...871..194Y}, though neither \citealt{2017MNRAS.472.2263H} nor \citealt{2019MNRAS.487.1808M} found such a relation in large samples).
-Spiral tightness is also observed to correlate with with rotation curve shape (\citealt{2005MNRAS.359.1065S}), with \textbf{galaxies with rising rotation curves having} more open spiral structure.
+Spiral tightness is also observed to correlate with with rotation curve shape (\citealt{2005MNRAS.359.1065S}), with galaxies with rising rotation curves having more open spiral structure.
 These predictions and observations provide compelling reasons for continued investigation of the underlying rules and dynamics of spiral structure, as doing so is essential for understanding the secular evolution of disc galaxies.

### 2021-04-13 sections/results.tex
Removes highlighting ready for final submission

@@
 
-One of the predictions of Manifold theory is that pitch angle increases with bar strength \textbf{as evaluated by the Quadrupole moment, $Q$ at the Lagrangian $L_1$ point (a value that differs from typical ``bar strength'', which is this value averaged over all radii in the bar; \citealt{2009MNRAS.400.1706A})}.
-In this work we do not have any \textbf{similar} measurement of bar strength \textbf{and we note that \citet{2009MNRAS.400.1706A} caution that other measures of bar strength may not show this relation; we also do not have} significant numbers of strongly barred galaxies in our sample.
+One of the predictions of Manifold theory is that pitch angle increases with bar strength as evaluated by the Quadrupole moment, $Q$ at the Lagrangian $L_1$ point (a value that differs from typical ``bar strength'', which is this value averaged over all radii in the bar; \citealt{2009MNRAS.400.1706A}).
+In this work we do not have any similar measurement of bar strength and we note that \citet{2009MNRAS.400.1706A} caution that other measures of bar strength may not show this relation; we also do not have significant numbers of strongly barred galaxies in our sample.
 However in an attempt to investigate this relationship in our data, we make use of Galaxy Zoo 2's bar fraction ($p_\mathrm{bar}$), which has been demonstrated to be a good measure of bar length \citep{Willett2013:1308.3496v2} and bar strength \citep{2012MNRAS.423.1485S,2012MNRAS.424.2180M,2018MNRAS.473.4731K} and therefore a good measure of the torque applied on the disc gas.

### 2021-04-13 sections/summary.tex
Removes highlighting ready for final submission

@@
 
-We do not find any link between {\bf our measure of} bar strength and pitch angle in our sample.
-However, rather that a direct measure of bar strength, we make use of an available parameter which correlates with bar strength, so at best this observation {\bf is suggestive} that the primary mechanism driving the evolution of the spirals in our sample {\bf may not be} Manifold theory (see Section \ref{section:morphology-comparison-bar}).
-\textbf{Since this is not the measure of bar strength predicted to correlate with pitch angle by \citet{2009MNRAS.400.1706A}, and those authors caution that the details of the bar strength measure can wash out the predicted correlation, this is not strong evidence against Manifold theory models}.
+We do not find any link between our measure of bar strength and pitch angle in our sample.
+However, rather that a direct measure of bar strength, we make use of an available parameter which correlates with bar strength, so at best this observation is suggestive that the primary mechanism driving the evolution of the spirals in our sample may not be Manifold theory (see Section \ref{section:morphology-comparison-bar}).
+Since this is not the measure of bar strength predicted to correlate with pitch angle by \citet{2009MNRAS.400.1706A}, and those authors caution that the details of the bar strength measure can wash out the predicted correlation, this is not strong evidence against Manifold theory models.
 
@@
 
-In this work, we assume that spiral arms are equally likely to be identified and recovered at all pitch angles, which suggests the absence of galaxies at low pitch angles is not {\bf due to an inability of us to measure such arms}.
+In this work, we assume that spiral arms are equally likely to be identified and recovered at all pitch angles, which suggests the absence of galaxies at low pitch angles is not due to an inability of us to measure such arms.
 This is not an unfair assumption given the amount of human effort that went into obtaining spiral arm measurements (more so than any other pitch angle measurement method, with each galaxy receiving at least 30 human classifications).
-The galaxy sample used is {\bf a random subset of a volume limited sample (see Figure \ref{fig:stellarmass}), and} is comparable in size to those used in other similar studies (\citealt{2013MNRAS.436.1074S}, \citealt{2019ApJ...871..194Y}, \citealt{2019arXiv190910291P}).
-The sample covers a range of masses ($9.45 < \log(M_* / M_\odot) < 11.05$) and spiral types, {\bf however it is possible that tightly wound spirals are preferentially missed by the pre-selection from {\it Galaxy Zoo}, if they are less obviously identified as having spirals at these distances ($0.02<z<0.055$).
+The galaxy sample used is a random subset of a volume limited sample (see Figure \ref{fig:stellarmass}), and is comparable in size to those used in other similar studies (\citealt{2013MNRAS.436.1074S}, \citealt{2019ApJ...871..194Y}, \citealt{2019arXiv190910291P}).
+The sample covers a range of masses ($9.45 < \log(M_* / M_\odot) < 11.05$) and spiral types, however it is possible that tightly wound spirals are preferentially missed by the pre-selection from \textit{Galaxy Zoo}, if they are less obviously identified as having spirals at these distances ($0.02<z<0.055$).
 The \citet{1981AJ.....86.1847K} sample of 113 much more nearby spirals (all at $z<0.019$ and most at $z<0.009$), includes several with arms at much lower pitch angles, but very few arms which are loosely wound (none with $\cot \phi > 31^\circ$), meaning it does not match the same $\cot \phi$ constant model well, although as at the time the sample was set by data availability, it is unclear how conclusive this is, and \citet{1981AJ.....86.1847K} note the incompleteness of their sample for open armed spirals.
-In a future version of {\it Galaxy Builder} we intend to include this sample as a comparison set.}
+In a future version of {\it Galaxy Builder} we intend to include this sample as a comparison set.
 
